### 2018-11-16 README.Rmd
Set up main repository files

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fortebaseline

This repository contains the source code and raw manuscript files for the FoRTE baseline ED2 runs manuscript.

## Installation

You can install the development version of fortebaseline from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("ashiklom/fortebaseline")
```

### 2018-11-16 analysis/paper/paper.Rmd
Set up main repository files

---
title: "FoRTE baseline"
author:
  - Alexey N. Shiklomanov
  - Jeff Atkins
  - Ben Bond-Lamberty
  - Christopher M. Gough
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/ecology.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(fortebaseline) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

# Introduction

Here is a citation [@bond-lamberty_2015_moderate] and another citation [@clm45_note;@norman_1979_modelling].
Another citation [@bond-lamberty_2015_moderate].

# Background

```{r demo, eval = FALSE}
1 + 1
```

# Methods

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2018-11-16 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(fortebaseline)
+
+test_check("fortebaseline")

### 2018-11-20 analysis/paper/paper.Rmd
Add some notes to paper draft

@@
 
-# Background
+# Methods
 
-```{r demo, eval = FALSE}
-1 + 1
-```
+## Site description
 
-# Methods
+- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (/Acer rubrum/), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
+- Average age -- 95 years (2013)
+- NEP in UMBS tower is 1.58 MG C ha-1 yr-1 (0.80-1.98) (1999-2006), but with landscape variation (REF?)
+- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923
+- Present day composition typical of upper Great Lakes region
+- Previously, the site of the FASET experiment [@gough_2013_sustained; @gough_2008_multi]
+
+## Model description: ED2
+
+The Ecosystem Demography Model, version 2 (ED2) [@medvigy_2009_mechanistic].
+
+Summary of common features (can move into supplement if necessary):
+
+- Cohort-based scaling approach
+- Photosynthesis / stomatal conductance
+- Allometries
+
+In this study, we run a factorial experiment of ED2 with the following features enabled or disabled:
+
+- Radiative transfer
+  - Two-stream model
+    - Same general approach as in CLM 4.5 [@clm45_note], but rather than solving for just two layers (sunlit and shaded), solve for each canopy layer.
+    - Set up the two-stream equations for each layer as a linear system and solve the entire set of equations numerically
+    - Similar to approaches used in other models: JULES [@mercado_2007_improving], ORCHIDEE-CAN [@naudts_2015_vertically], CLM(SPA) [@bonan_2014_modeling]
+  - Multiple scatter model **TODO**
 

### 2019-01-29 analysis/paper/paper.Rmd
Add some notes to paper.Rmd

@@
 
-Here is a citation [@bond-lamberty_2015_moderate] and another citation [@clm45_note;@norman_1979_modelling].
-Another citation [@bond-lamberty_2015_moderate].
+Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
+Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
 

### 2019-01-29 analysis/paper/paper.Rmd
PAPER: Brain dump in introduction

@@
 
+Earth's climate is changing, and its terrestrial ecosystems are likely to be affected (REF: IPCC).
+Many of these ecosystems will be affected through persistent changes in their environment, for instance through elevated atmospheric CO~2~ concentrations and temperature, changing precipitation regimes (REF).
+At the same time, climate change is likely change the frequency and severity of disturbances [@dale_2001_climate], and the impacts of these changes may be more acute.
+
+<!-- Add something positive about each of these? Then, the section about models is more compelling. -->
+In trying to anticipate future changes to terrestrial ecosystems, ecologists have four general techniques.
+One approach is to look to the past to see how ecosystems responded to similar changes (REFs).
+This is challenging for two reasons:
+For one, we often have to look very far back to find similar climate changes, and the techniques we have for determining the state of ecosystems in the distant past have limited detail (REF).
+More importantly, there are few (if any) times where the climate has changed as rapidly as it is now, and it is unclear whether slower changes in the past will trigger responses similar to the faster changes happening now (REF).
+A second approach is space-for-time substitution, whereby we use existing climatic gradients through space (e.g. with latitude or elevation) to infer the corresponding changes through time (REF).
+However, many processes that drive spatial variation in ecosystems play out over hundreds or thousands of years (e.g.
+evolution), far longer than policy-relevant timescales of years to decades and far slower than the current pace of climate change (REF).
+A third approach is to conduct experimental manipulations, where individuals are subjected to controlled changes in environmental conditions [@medlyn_2015_using].
+Because they are able to isolate specific changes by controlling for most other factors, experiments are very useful for testing mechanisms.
+Unfortunately, resource constraints limit the spatial and temporal extent of experiments, and emerging evidence from (the relatively few) long-term experiments suggests that short-term results cannot always be extrapolated from longer-term results [@reich_2018_unexpected].
+
+The fourth approach is to use computer models based on existing understanding of ecosystem function.
+Models unconstrained by data are little more than thought experiments with limited value to decision-making.
+However, models provide a useful tool for confronting hypotheses with observations from a variety of sources, including the three described above [@dietze_2013_improving].
+
+Model design involves an invariable trade-off between complexity and predictive uncertainty.
+The more processes are represented in a model (and the more parameters are used to represent those processes), the higher the uncertainty in the "true" value of each parameter, and the higher the uncertainty in model predictions (especially out of sample) (REF -- information theory, e.g.
+AIC?).
+On the other hand, simplistic models often fail to match important properties of the observations.
+A key challenge in model development is to find the balance between these two extremes.
+
+Big leaf models (e.g. CLM) fail to capture vegetation dynamics (REF).
+This is due to their inability to represent demographic processes, which are important (REF).
+An emerging class of models can simulate vegetation demographic processes, including competition for resources within a patch, recruitment, and mortality [@fisher_2017_vegetation].
+Individually-based models are one possibility (REFs -- Shugart, Ensheng Wang, etc.).
+However, they are complicated (REF).
+An alternative are models of intermediate complexity ("middle ground"), which use approximations.
+One such approximation is the "Ecosystem Demography" approach, whereby groups of individuals ("cohorts") of similar species, size, and structure are modeled together, and the statistical properties of these cohorts are modeled rather than the characteristics of each individual [@moorcroft_2001_method].
+
+One realization of this is the Ecosystem Demography (v2) model [@moorcroft_2001_method; @medvigy_2011_predicting; @medvigy_2009_mechanistic].
+This model has been used to simulate (examples...).
+
 Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
 Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
+
+<!--
+TODO: Also discuss the other ED modules.
+-->
+
+This manuscript addresses the following research questions:
+(Q1) To what extent can the ED2 model reproduce 100 years of observed succession (or observed _anything_) at UMBS starting from bare ground?
+(Q2) How does the inclusion of various processes affect the accuracy and predictive uncertainty of ED2 predictions?
+(Q3) What drives the predictive uncertainty of the ED2 model at UMBS, and what additional observations are necessary to improve predictions / constrain uncertainty?
+
+Based on previous results [@fisher_2015_taking], we predict ED2 will over-predict dominance of early-successional canopy species (and under-predict understory productivity) under its default configuration.
+Enabling dynamic N cycling and a finite canopy radius model will alleviate these issues, but will significantly increase the parameteric uncertainty due to strong sensitivity to N-cycling and canopy gap fraction parameters, respectively.
+Constraining these parameters should be a high priority
 

### 2019-01-29 analysis/paper/paper.Rmd
PAPER: Start going through ED modules.

@@
 
+## ED modules
+
+(NOTE: Not a real section; just brainstorming notes for now)
+
+- Big leaf model (no demographics) (`IBIGLEAF`)
+    - Simplistic case
+    - Only one PFT per patch (by definition). Can ask: Does it get the right one?
+    - Compare fluxes against other cases.
+- Branches (`IBRANCH_THERMO`)
+    - Options:
+        - (0) No branches in energy balance or biophysics
+        - (1; default) Branches in energy balance, but not biophysics
+        - (2) Branches in both
+    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
+- Temperature variation of physiology (`IPHYSIOL`)
+    - Options:
+        - (0) Arrhenius function, `Gamma*` based on Foley et al. (1996) (TODO: REF)
+        - (1) Arrhenius, `Gamma*` based on Michaelis-Mentel coefficients for CO2 and O2, as in Farquhar et al. (1980) (TODO: REF)
+        - (2; default) Q10 equations and parameters of Collatz et al. (1991), with MM coefficients. High/low temperature correction in Moorcroft et al. (2001)
+        - (3) As (2), but `Gamma*` as in Farquhar et al. (1980) and CLM.
+    - Need to do some exploratory analysis here. Grab the relevant equations and coefficients and plot them (outside of ED) to compare behavior.
+        - Possible angle: More important at temperature extremes?
+    - Consult @rogers_2016_roadmap for inspiration
+- Finite canopy radius (`CROWN_MOD`)
+    - Options:
+        - (0; default): Complete shading
+        - (1): Finite canopy radius (Dietze 2008)
+    - See current hypotheses
+- Soil decomposition (`DECOMP_SCHEME`)
+    - Options:
+        - (0; default): Exponential
+        - (1) Lloyd and Taylor (1994)
+    - (1) requires additional parameters, so should increase parameteric uncertainty.
+    - Hypothesis: Affects overall GPP/R~E~ balance, but not successional dynamics.
+- Water limitation of photosynthesis (`H2O_PLANT_LIM`)
+    - Options:
+        - (0) No limitation.
+        - (1; default): Supply is total soil water above wilting point integrated across all levels in rooting zone.
+        - (2): Supply is total soil water at field capacity minus wilting point, scaled by "wilting factor":
+          ```
+          # k -- Soil layer
+          # z -- Layer depth
+          # psi -- Matric potential (for layer, or at wp or fc)
+          # wp -- wilting point
+          # fc -- field capacity
+          (psi(k) - (H - z(k)) - psi_wp) / (psi_fc - psi_wp)
+          ```
+    - Sensitivity should depend on site climate and soil conditions.
+    - UMBS are pretty wet sites, so shouldn't expect a huge effect.
+- Structural growth (`ISTRUCT_GROWTH_SCHEME`)
+    - Options:
+        - (0; default) Use all `bstorage` (NSC?)
+        - (1) Reserve `bstorage` to re-flush canopy and fine roots once before calculating structural growth.
+    - (1) is a more conservative strategy, so should reduce the growth and relative dominance of early-successional PFTs.
+In particular, growth rate should decline more with size (because more LAI and fine roots means more reserves needed).
+- Stomatal conductance model (`ISTOMATA_SCHEME`)
+    - Options:
+        - (0, default): Leuning
+        - (1) Katul optimization model (see Xu et al. 2016 New Phyt.)
+    - Need to read more about this...
+- Trait plasticity (`TRAIT_PLASTICITY_SCHEME`)
+    - Options:
+        - (0, default): Static traits
+        - (1) Vm0 (SLA) decrease (increase) with shading, based on Lloyd et al. (2010 Biogeosciences). Updated yearly.
+        - (2) Same as (1), but updated monthly
+        - (-1) Same as 1, but adjust SLA using height
+        - (-2) Same as (2), but adjust SLA using height
+    - If trait plasticity is optimal, this should reduce (?) the relative dominance of early-successional PFTs.
+    - Probably introduces considerable predictive uncertainty, given new parameters for plasticity models.
+- Other configs that ED has, but that we will probably ignore:
+    - Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs. Worth checking.
+    - Grass (`IGRASS`) -- I'm not running with any grass PFTs.
+    - Phenology (`IPHEN_SCHEME`) -- Differences affect grasses and tropical species. Conifers are always evergreen, and hardwoods are always cold-deciduous (need to check that schemes are the same).
+    - Seed dispersal (`REPRO_SCHEME`) -- Only seems to matter for multi-site runs, with some toggles for tropical (drought-deciduous) species.
+    - Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
+
 # Methods

### 2019-01-30 analysis/paper/paper.Rmd
PAPER: Finish the list of ED2 schemes

@@
     - See current hypotheses
-- Soil decomposition (`DECOMP_SCHEME`)
+- Soil decomposition temperature sensitivity (`DECOMP_SCHEME`)
     - Options:
@@
     - Probably introduces considerable predictive uncertainty, given new parameters for plasticity models.
+- Carbon balance contribution to mortality (`IDDMORT_SCHEME`)
+    - Options:
+        - (0, default): Carbon balance in terms of fluxes only
+        - (1) Carbon balance is offset by storage pool (i.e. carbon storage is an additional buffer against mortality)
+    - Additional protection from (1) should reduce mortality for all trees, so will probably reinforce dominance of early successional trees?
+    - But likely to interact with other schemes, especially structural growth, which determines size of NSC reserves
+- Carbon stress scheme (`CBR_SCHEME`)
+    - Options:
+        - (0, default): Single stress, with maximum carbon balance determined at full sun and no moisture
+        - (1) Co-limitation of light and moisture; calculate `cb/cb_lightmax` and `cb/cb_moistmax`, and weight based on `DDMORT_CONST` (tunable global[?] parameter)
+        - (2) "Leibig style"; limitation based on whether soil or light are more important (`cb/max(cb_lightmax, cb_moistmax)`)
+    - No idea about this one. Need to do more reading.
+- Quantum yield model (`QUANTUM_EFFICIENCY_T`)
+    - Options:
+        - (0, default): Constant quantum efficiency
+        - (1): Quantum efficiency varies with temperature, based on Ehleringer (1978) (TODO: Get ref)
+    - Uncertainty should increase with (1) assuming parameters are tunable. 
+    - Depends on where UMBS conditions fall within the range of temperatures. Good candidate for exploratory analysis outside of ED.
+- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
+    - Options:
+        - (0, default): No limitation
+        - (1) Limitation based on internal model (TODO: what is it?)
+    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
+- Nitrogen limitation of decomposition (`N_PLANT_LIM`)
+    - Options:
+        - (0, default): No limitation
+        - (1) Limitation based on internal model (TODO: what is it?)
+    - Same as above
+- Canopy roughness model (`ICANTURB`)
+    - Options:
+        - (0): Based on Leuning et al. (1995; TODO REF). Wind computed using "similarity theory" for top cohort, then extinguished with cumulative LAI.
+        - (1, default) "Default ED2.1 scheme" (TODO: 0? Figure it out!), but using "zero-plane" displacement height.
+        - (2) Massman (1997; TODO) using constant drag and no sheltering factor
+        - (3) Also Massman, but with varying drag and sheltering within canopy
+        - (4) Like 0, but calculate ground conductance following CLM (eq. 5.98-5.100)
+    - No idea, except that (3) will have higher uncertainty than (2). Need to read more about this.
+- Similarity theory model (for canopy roughness) (`ISFCLYRM`)
+    - Computes `u*`, `T*`, etc.
+    - Options:
+        - (1) BRAMS default, based on Louis (1979 TODO). 
+        - (2) Oncley and Dudhia (1995 TODO), based on MM5
+        - (3) Beljaars and Holtslag (1991 TODO); Like (2), but alternative method for stable case that has more mixing
+        - (4, default) CLM (2004; TODO: Check that this is same as 2013?). Like (2) and (3), but with special functions to deal with very stable cases.
+    - Again, no idea. Need to read more about micro-meteorology and how it affects vegetation processes.
+- Ground to canopy conductance (`IED_GRNDVAP`)
+    - Options:
+        - (0, default): Modified Lee Pielke (1992 TODO), with field capacity, but with beta factor without square (as in Noilhan and Planton 1989 TODO). Like original ED2 and LEAF-3.
+        - (1-4) Mahfouf and Noilhan (1991 TODO), #1-4
+        - (5) Combination of (1) (alpha) and (2) (soil resistance)
+    - No idea. Again, need to do some micro-met reading.
+- Percolation and infiltration (`IPERCOL`)
+    - Options:
+        - (0) Constant soil conductivity. Temporary surface water (only exists if top soil layer is saturated) exceeding 1:9 liquid:ice ratio shed via percolation.
+        - (1, default) Constant soil conductivity, with percolation model from Anderson (1976 NOAA technical report NWS 19 TODO).
+Temporary surface water can exist after heavy rain event even if soil doesn't saturate.
+        - (2) Like (1), but soil conductivity decreases with depth even under constant soil moisture
+    - Presumably, not a huge impact at UMBS because water shouldn't be an issue.
+    - See also `RUNOFF_TIME` (e-folding lifetime of temporary surface water)
+- Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)
+    - Options:
+        - (0, default) No treefall disturbance
+        - (> 0) Disturbance rate, in 1/years. Will create new patches.
+        - (< 0) Also rate, but will be added to mortality (kill plants, but no new patches)
+    - Reference value for this is 0.014 trees year^-1^ (note that because of ED's cohort approximation, number of individuals is continuous, not discrete)
+    - See also `TIME2CANOPY`, minimum age for tree-fall disturbance.
+- Growth respiration (`GROWTH_RESP_SCHEME`)
+    - Options:
+        - (0, default) Growth respiration is a tax on GPP, at PFT-specific rate. Treated as aboveground wood to canopy-airspace flux.
+        - (1) Growth respiration calculated as in (0), but split into leaf, sapwood-a, sapwood-b, and root sources proportionally to biomass.
+No effect on overall C budget, but greater within-ecosystem flux resolution.
+    - These should produce identical results, and if they don't, that means there is a bug!
+    - If output is identical, we should set to (1), since we can always calculate total respiration otherwise.
+- Storage respiration (`STORAGE_RESP_SCHEME`) -- same as growth
+    - Note that this whole system is a _hack_, because there is no real "storage respiration" process.
+In PEcAn, we hard-code the relevant parameters to zero.
+An interesting question is whether we may be able to get better results with this hack enabled -- i.e.
+if it compensates a problem in some other process.
 - Other configs that ED has, but that we will probably ignore:
@@
     - Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
+    - Fire disturbance (`INCLUDE_FIRE`) -- We know there weren't any fires at UMBS during this time, and more generally are interested in stability.
+        - This could potentially be fun to play with in a later paper about disturbance regimes. Potentially, with Jackie Matthes?
+    - Anthropogenic disturbances (`IANTH_DISTURB`) -- We are currently looking at stable state, without disturbances.
+But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.
 

### 2019-01-30 analysis/paper/paper.Rmd
PAPER: Sort ED modules based on paper priority

@@
 
-(NOTE: Not a real section; just brainstorming notes for now)
+NOTE: Not a real section; just brainstorming notes for now
+Details about these sub-models will probably makes their way into the methods and/or appendices.
+
+### High priority
+
+These are switches that I understand (more-or-less), and that I think will be most relevant to FoRTe.
+
+- Finite canopy radius (`CROWN_MOD`)
+    - Options:
+        - (0; default): Complete shading
+        - (1): Finite canopy radius (Dietze 2008)
+    - See current hypotheses
+- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
+    - Options:
+        - (0, default): No limitation
+        - (1) Limitation based on internal model (TODO: what is it?)
+    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
+- Nitrogen limitation of decomposition (`N_PLANT_LIM`)
+    - Options:
+        - (0, default): No limitation
+        - (1) Limitation based on internal model (TODO: what is it?)
+    - Same as above
+
+### Medium priority
+
+These are switches that may be relevant to FoRTE, but may be harder to pull off.
 
@@
     - Compare fluxes against other cases.
-- Branches (`IBRANCH_THERMO`)
-    - Options:
-        - (0) No branches in energy balance or biophysics
-        - (1; default) Branches in energy balance, but not biophysics
-        - (2) Branches in both
-    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
 - Temperature variation of physiology (`IPHYSIOL`)
@@
     - Consult @rogers_2016_roadmap for inspiration
-- Finite canopy radius (`CROWN_MOD`)
-    - Options:
-        - (0; default): Complete shading
-        - (1): Finite canopy radius (Dietze 2008)
-    - See current hypotheses
 - Soil decomposition temperature sensitivity (`DECOMP_SCHEME`)
@@
     - UMBS are pretty wet sites, so shouldn't expect a huge effect.
+- Quantum yield model (`QUANTUM_EFFICIENCY_T`)
+    - Options:
+        - (0, default): Constant quantum efficiency
+        - (1): Quantum efficiency varies with temperature, based on Ehleringer (1978) (TODO: Get ref)
+    - Uncertainty should increase with (1) assuming parameters are tunable. 
+    - Depends on where UMBS conditions fall within the range of temperatures. Good candidate for exploratory analysis outside of ED.
 - Structural growth (`ISTRUCT_GROWTH_SCHEME`)
@@
     - But likely to interact with other schemes, especially structural growth, which determines size of NSC reserves
+- Percolation and infiltration (`IPERCOL`)
+    - Options:
+        - (0) Constant soil conductivity. Temporary surface water (only exists if top soil layer is saturated) exceeding 1:9 liquid:ice ratio shed via percolation.
+        - (1, default) Constant soil conductivity, with percolation model from Anderson (1976 NOAA technical report NWS 19 TODO).
+Temporary surface water can exist after heavy rain event even if soil doesn't saturate.
+        - (2) Like (1), but soil conductivity decreases with depth even under constant soil moisture
+    - Presumably, not a huge impact at UMBS because water shouldn't be an issue.
+    - See also `RUNOFF_TIME` (e-folding lifetime of temporary surface water)
 - Carbon stress scheme (`CBR_SCHEME`)
@@
     - No idea about this one. Need to do more reading.
-- Quantum yield model (`QUANTUM_EFFICIENCY_T`)
+
+### Lower priority
+
+These are processes that I think will not have huge impacts, are farther-removed from the FoRTE project goals, or are things I simply don't understand very well (particularly all the micro-meteorology stuff).
+
+- Branches (`IBRANCH_THERMO`)
     - Options:
-        - (0, default): Constant quantum efficiency
-        - (1): Quantum efficiency varies with temperature, based on Ehleringer (1978) (TODO: Get ref)
-    - Uncertainty should increase with (1) assuming parameters are tunable. 
-    - Depends on where UMBS conditions fall within the range of temperatures. Good candidate for exploratory analysis outside of ED.
-- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
-    - Options:
-        - (0, default): No limitation
-        - (1) Limitation based on internal model (TODO: what is it?)
-    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
-- Nitrogen limitation of decomposition (`N_PLANT_LIM`)
-    - Options:
-        - (0, default): No limitation
-        - (1) Limitation based on internal model (TODO: what is it?)
-    - Same as above
+        - (0) No branches in energy balance or biophysics
+        - (1; default) Branches in energy balance, but not biophysics
+        - (2) Branches in both
+    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
 - Canopy roughness model (`ICANTURB`)
@@
     - No idea. Again, need to do some micro-met reading.
-- Percolation and infiltration (`IPERCOL`)
-    - Options:
-        - (0) Constant soil conductivity. Temporary surface water (only exists if top soil layer is saturated) exceeding 1:9 liquid:ice ratio shed via percolation.
-        - (1, default) Constant soil conductivity, with percolation model from Anderson (1976 NOAA technical report NWS 19 TODO).
-Temporary surface water can exist after heavy rain event even if soil doesn't saturate.
-        - (2) Like (1), but soil conductivity decreases with depth even under constant soil moisture
-    - Presumably, not a huge impact at UMBS because water shouldn't be an issue.
-    - See also `RUNOFF_TIME` (e-folding lifetime of temporary surface water)
 - Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)
@@
 if it compensates a problem in some other process.
-- Other configs that ED has, but that we will probably ignore:
-    - Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs. Worth checking.
+
+### Almost-certainly omit
+
+These configurations are almost certainly irrelevant to the goals of this paper or will not apply at UMBS (e.g. grass or tropical species).
+Mostly for reference.
+
+- Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs (TODO: Check that this is the case).
 - Grass (`IGRASS`) -- I'm not running with any grass PFTs.

### 2019-01-31 analysis/paper/ed-notes.Rmd
PAPER: Start `ed-notes` doc. Update references, fix typos.

---
title: ED2 model notes
author: Alexey Shiklomanov
---

# ED2 switches

## Finite canopy radius (`CROWN_MOD`)

In `src/dynamics/radiate_driver.f90:L550`:
If disabled, set canopy crown area for all patches to 1.
If enabled, set finite crown areas for each cohort.

Crown area used in some of the following places:

- Canopy structural dynamics (`dynamics/canopy_struct_dynamics.f90:L470`)
- Plant hydrology (`dynamics/plant_hydro.f90:L145,204,240,447,626`)

Crown area comes from @dietze_2008_capturing.
Crown area calculated by allometry in `utils/allometry.f90:L633` (inside subroutine `area_indices`), as `min(1, nplant * dbh2ca(dbh, hite, sla, pft))`.
`dbh2ca` is defined in `utils/allometry.f90:L373` as:

```
ca = biCa[pft] * dbh ^ b2Ca[pft]
```

But, note that crown area is restricted such that "Local LAI / Crown area" is never less than one (i.e you cannot have more crown area than leaf area index -- your maximum crown area is the leaf area index):

```
loclai = sla * size2bl(dbh, hite, pft)
ca = min(loclai, ca)
```

where `size2bl` is the leaf biomass allometry (so leaf biomass; therefore, `sla * size2bl` gives LAI).

Crown area set or calculated in:

- `dynamics/strcutral_growth.f90:L669,929`

## Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)

## Nitrogen limitation of decomposition (`N_DECOMP_LIM`)

### 2019-01-31 analysis/paper/paper.Rmd
PAPER: Start `ed-notes` doc. Update references, fix typos.

@@
     - See current hypotheses
+    - This model is based on @dietze_2008_capturing
 - Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
@@
     - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
-- Nitrogen limitation of decomposition (`N_PLANT_LIM`)
+- Nitrogen limitation of decomposition (`N_DECOMP_LIM`)
     - Options:

### 2019-01-31 analysis/paper/ed-notes.Rmd
PAPER: Add notes on finite canopy radius in ED

@@
 
-In `src/dynamics/radiate_driver.f90:L550`:
-If disabled, set canopy crown area for all patches to 1.
-If enabled, set finite crown areas for each cohort.
-
-Crown area used in some of the following places:
-
-- Canopy structural dynamics (`dynamics/canopy_struct_dynamics.f90:L470`)
-- Plant hydrology (`dynamics/plant_hydro.f90:L145,204,240,447,626`)
+### Calculating crown area
 
@@
 ```
-ca = biCa[pft] * dbh ^ b2Ca[pft]
+ca = b1Ca[pft] * dbh ^ b2Ca[pft]
 ```
 
-But, note that crown area is restricted such that "Local LAI / Crown area" is never less than one (i.e you cannot have more crown area than leaf area index -- your maximum crown area is the leaf area index):
+Note that crown area is restricted such that "Local LAI / Crown area" is never less than one (i.e you cannot have more crown area than leaf area index -- your maximum crown area is the leaf area index):
 
@@
 
-Crown area set or calculated in:
+The allometry coefficients `b1Ca` and `b2Ca` have the following defaults (from `init/ed_params.f90:L3055`):
 
-- `dynamics/strcutral_growth.f90:L669,929`
+- `b1Ca`
+    - 5:13 (C3 grass and temperate trees) -- `2.490154` (constant)
+    - All others -- `exp(-1.853) * exp(b1Ht(pft)) ** 1.888`; log-linear relationship with height allometry
+- `b2Ca`
+    - 5:13 -- `0.8068806` (constant)
+    - All others -- `b2Ht(pft) * 1.888`; linear relationship with height allometry
+
+Note that only tropical PFTs are affected by the allometry model (`IALLOM`); coefficient adjustments are wrapped in `if (is_tropical(pft))`.
+
+If two cohorts are fused, their crown areas are summed, up to a maximum of 1 (in `utils/fuse_fiss_utils.f90:L2932`).
+
+### Radiative transfer
+
+In `src/dynamics/radiate_driver.f90:L550`:
+If disabled, set canopy crown area for all patches to 1.
+If enabled, set finite crown areas for each cohort.
+
+Crown area array (for each cohort) is passed to the various radiative transfer models (see `ICANRAD` parameter).
+Inside the RTMs, it is the `cai` variable.
+
+The following sections focus on the contributions of finite canopy radius to radiative transfer calculations.
+For the full descriptions, consult the "Canopy radiative transfer" module section.
+
+#### Two-stream
+
+In the two-stream model, crown area is used to correct `mu_bar` (average inverse optical depth per unit leaf and stem area) for crown area (resulting variable `mu`):
+
+```
+mu = -etai / log(1 - cai + cai * exp(-etai / (cai * mu_bar)))
+```
+
+where `etai` is the total area index (TAI = LAI * clumping factor + WAI) of the layer.
+
+(If `cai = 1`, this simplifies to just `mu_bar`).
+
+In the longwave (thermal), crown area is also used as a linear scaling of the blackbody emissivity.
+
+In the shortwave two-stream model, the crown area correction also happens for direct radiation (`mu0`):
+
+```
+mu0 = -etai / log(1 - cai + cai * exp((-proj_area * etai) / (cai * czen)))
+```
+
+where `czen` is the cosine of the solar zenith angle (radians) and `proj_area` is the projected area, calculated as a function of leaf orientation angle (`orient`):
+
+```
+phi1 = 0.5 - orient * (0.633 + 0.33 * orient)
+phi2 = 0.877 * (1 - 2 * phi1)
+proj_area = phi1 + phi2 * czen
+```
+
+These calculations follow CLM 4.5 [@clm45_note] (section 3.1, eq. 3.3), which in turn is based on @sellers_1985_canopy.
+This is described in more detail in the canopy radiative transfer section.
+
+#### Multiple-scatter
+
+In the multiple-scatter model [@zhao_2005_multiple; @zhao_2006_modeling], crown area is used calculate the exposed total area index (TAI) of each layer (`locetai`):
+
+```
+locetai = etai / cai
+```
+
+Crown area is also used to adjust the layer transmittance coefficient for beam radiation (`tau_beam`):
+
+```
+lambda = proj_area / mu
+tau_beam = (1 - cai) + cai * exp(-lambda * locetai)
+```
+
+A similar correction is applied to diffuse radiation, but the contribution from each hemisphere has to be integrated.
+This uses the analytical solution for that integral.
+
+```
+d1 = phi1 * locetai
+d2 = phi2 * locetai
+tau_diff = (1 - cai) - cai * exp(-d1 * d2) * (d1^2 * exp(d1) * eifun(-d1) + (d1 - 1))
+```
+
+where `eifun` is the exponential integral function.
+
+### Plant hydraulics
+
+Source: `dynamcs/plant_hydro.f90`.
+
+Subroutine `plant_hydro_driver`:
+Crown area for each cohort calculated as `crown_area / nplant`, then passed to `calc_plant_water_flux` (as variable `crown_area`).
+
+Calculate the root area index (RAI) at each depth:
+
+```
+RAI = broot_d * SRA * root_frac / (4 * crown_area)
+```
+
+where SRA is the PFT specific root area parameter, and `root_frac` is the fraction of roots in that layer (function of PFT parameter `root_beta` and depth).
+
+Calculate the soil-root water conductance at each depth:
+
+```
+gw_cond = (4 * crown_area) * soil_cond * sqrt(RAI) / (pi * dslz)
+```
+
+where `soil_cond` is soil conductivity at that depth and `dslz` is the soil layer thickness.
+This is based on @katul_2003_relationship.
+
+### Canopy turbulence
+
+In file `dynamics/canopy_struct_dynamics.f90`, subroutine `canopy_turbulence`.
+
+Crown area used in calculation of the wind profile extinction coefficients.
+
+```
+#-- dynamics/canopy_struct_dynamics.f90:L470 --
+extinct_full = crown_area * exp(-0.25 * lai / crown_area) + (1 - crown_area)
+extinct_full = crown_area * exp(-0.5 * lai / crown_area) + (1 - crown_area)
+veg_wind = max(ugbmin, uh[i - 1] * extinct_half)
+uh[i] = uh[i - 1] * extinct_full
+```
+
+Note that there is a copy of the same calculation performed in double precision (subroutine `canopy_turbulence8`) later in the file.
+
+Crown area is also used to calculate the open canopy fraction of each cohort by subroutine `update_patch_derived_props` (file `utils/update_derived_props.f90:L43`).
+That subroutine is in turn called by subroutine `structural_growth_eq_0` (file `dynamics/structural_growth.f90`).
+This open canopy fraction is used to calculate the canopy roughness (`rough`).
+
+```
+rough = snow_rough * snowfac_can +
+    (soil_rough * opencan_frac + veg_rough * (1 - opencan_frac)) *
+    (1 - snowfac_can)
+```
+
+Open canopy fraction is also used to calculate the net ground conductance (`ggnet`):
+
+```
+ggnet = ggbare * ggveg / (ggveg + ggbare * (1 - opencan_frac))
+```
 

### 2019-01-31 analysis/paper/paper.Rmd
PAPER: Add notes on finite canopy radius in ED

@@
     - See current hypotheses
-    - This model is based on @dietze_2008_capturing
+    - This model is based on crown area allometries from @dietze_2008_capturing
 - Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
@@
     - Compare fluxes against other cases.
+- Canopy radiative transfer model (`ICANRAD`)
+    - Options:
+        - (0, deprecated?) Original two-stream model (Medvigy 2006)
+        - (1) Multiple-scattering model [@zhao_2006_modeling; @zhao_2005_multiple]
+        - (2, default) Updated two-stream model [@liou_2002]
+    - In theory, multiple-scattering model is more detailed and should give "better" results. Open question as to how that affects predictions.
+    - Good candidate for exploratory analysis.
+I already implemented the two-stream model in R -- could be a good idea to implement the multi-scatter model as well and compare their radiation profiles.
+- Allometry (`IALLOM`)
+    - Options:
+        - (0, default): Original ED2
+        - (1)
+            - Structural biomass parameterized so total AGB similar to @baker_2004_variation, eq. 2. `Balive` similar to ED2.
+            - Root depth configured so root depth is 5m for canopy trees and 0.5 m for grasses/seedlings
+            - Crown area as in @poorter_2006_architecture, with maximum crown area
+        - (2) Like (1), but...
+            - Height to DBH allometry from @poorter_2006_architecture
+            - `Balive` has extra allometric equations (see @cole_2006_allometric and @calvo-alvarado_2008_allometric)
+        - (3) Like (2), but tropical leaf biomass estimated from @lescure_1983_la, and PFT differences in leaf biomass based on SLA rather than wood density.
+Also, split leaf allometry into small (sapling) and large (tree) classes, as in @cole_2006_allometric
 - Temperature variation of physiology (`IPHYSIOL`)
@@
         - (0) Arrhenius function, `Gamma*` based on Foley et al. (1996) (TODO: REF)
-        - (1) Arrhenius, `Gamma*` based on Michaelis-Mentel coefficients for CO2 and O2, as in Farquhar et al. (1980) (TODO: REF)
-        - (2; default) Q10 equations and parameters of Collatz et al. (1991), with MM coefficients. High/low temperature correction in Moorcroft et al. (2001)
-        - (3) As (2), but `Gamma*` as in Farquhar et al. (1980) and CLM.
+        - (1) Arrhenius, `Gamma*` based on Michaelis-Mentel coefficients for CO2 and O2, as in @farquhar_1980_biochemical
+        - (2; default) Q10 equations and parameters of @collatz_1991_physiological, with MM coefficients. High/low temperature correction in @moorcroft_2001_method
+        - (3) As (2), but `Gamma*` as in @farquhar_1980_biochemical and CLM.
     - Need to do some exploratory analysis here. Grab the relevant equations and coefficients and plot them (outside of ED) to compare behavior.
@@
 
-- Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs (TODO: Check that this is the case).
 - Grass (`IGRASS`) -- I'm not running with any grass PFTs.
@@
 But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.
+
+### Additional references
+
+@katul_2003_relationship
 

### 2019-01-31 analysis/paper/ed-notes.Rmd
PAPER: Add description of N limitation to photosynthesis

@@
 
+### Calculating N supply and demand for live tissue growth
+
+Basically, N limitation places a limit on the fraction of stomata that are open (due to N limitation; `cpatch%fsn`).
+
+Calculating changes in live biomass (subroutines `dbalive_dt` and `dbalive_dt_eq_0` in `dynamics/growth_balive.f90`).
+
+If plant limitation enabled, call subroutine `potential_N_uptake` to see what would happen if stomata were open.
+This sets the value of potential N uptake (`N_uptake_pot`).
+
+If no limitation, set fraction of open stomata (from N limitation; `fsn`) to 1.
+Otherwise, calculate N limitation factor.
+
+```
+#- dynamics/growth_balive.f90:L293
+nitrogen_supply = plant_N_supply_scale * broot * mineralized_soil_N
+fsn = nitrogen_supply / (nitrogen_supply + N_uptake_pot)
+```
+
+### Initialization from near bare ground
+
+Subroutines `near_bare_ground_init` and `near_barea_ground_big_leaf_init` (in `init/ed_nbg_init.f90`):
+
+If there is N limitation, initialize as follows:
+
+- Fast soil C (`fast_soil_C`): 0.2
+- Slow soil C (`slow_soil_C`): 0.01
+- Structural soil C (`structural_soil_C`): 10.0
+- Structural soil lignin (`structural_soil_L`): 10.0
+- Mineralized soil N (`mineralized_soil_N`): 1.0
+- Fast soil N (`fast_soil_N`): 1.0
+
+If no N limitation, set all of these to zero.
+
 ## Nitrogen limitation of decomposition (`N_DECOMP_LIM`)

### 2019-01-31 analysis/paper/ed-notes.Rmd
PAPER: Add description of N decomposition limitation scheme

@@
 ## Nitrogen limitation of decomposition (`N_DECOMP_LIM`)
+
+Calculation of N immobilization factor (subroutine `resp_f_decomp`, in `dynamics/soil_respiration.f90`).
+
+N decomposition limitation affects the decomposition fraction (`csite%f_decomp`).
+If no limitation, this is set to 1.
+
+If limitation, calculate similarly to plant (supply relative to demand):
+
+```
+#-- dynamics/soil_respiration.f90:L226
+Lc = exp(-3 * structural_soil_L / structural_soil_C)
+
+N_immobilization_demand = A_decomp * Lc * decay_rate_stsc * structural_soil_C *
+  ((1 - r_stsc) / c2n_slow - 1 / c2n_structural)
+
+f_decomp = N_immobil_supply_scale * mineralized_soil_N / 
+  (N_immobilization_demand + N_immobil_supply_scale * mineralized_soil_N)
+```
+
+Whether N decomposition limitation is enabled also affects decay rate of slow carbon pool if using `DECOMP_SCHEME = 2`.
+With N limitation, decay rate is 100.2 year^-1^; otherwise, 0.2 year^-1^.

### 2019-02-12 analysis/paper/exploratory.Rmd
Add exploratory analysis Rmd

---
title: Exploration of ED outputs
author: Alexey Shiklomanov
output:
  html_document:
    self_contained: true
---

# Setup

```{r setup, cache = FALSE}
library(tidyverse)
library(fst)

requireNamespace("udunits2", quietly = TRUE)
requireNamespace("knitr", quietly = TRUE)
requireNamespace("data.table", quietly = TRUE)
requireNamespace("here", quietly = TRUE)

knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE)
```

This code has two requjred input datasets.
The first -- `ed_in` -- contains the half-hourly ED2 outputs by variable and ensemble member.
It is huge (~137 million rows).
It is generated by the `analysis/scratch/read_pecan_output.R` script.

```{r read-ed-in, cache = FALSE}
ed_in_file <- here::here("analysis/data/derived-data/ed-ensemble-out.fst")
ed_in <- read_fst(ed_in_file) %>% as_tibble()
ed_in
```

The second -- `params` -- contains ED2 input parameters for each ensemble member.
It is generated by the `analysis/scratch/read_trait_data.R` script.

```{r read-params, cache = FALSE}
params_file <- here::here("analysis/data/derived-data/ed-params.fst")
params <- read_fst(params_file) %>% as_tibble()
params
```

Both are currently tied to a specific workflow ID (99000000066).
Eventually, I will modify this to use `drake` to improve reproducibility.
The data are stored in `fst` format for fast IO.

Most of the parameters are not currently varied by PEcAn, so there isn't much use in looking at them in the context of an ensemble analysis.
The following code removes any parameters that are constant across ensemble members.

```{r subset-params}
params_sub <- params %>%
  semi_join(
    params %>%
      group_by(pft, variable) %>%
      filter(sd(value) > 0) %>%
      ungroup() %>%
      distinct(variable)
  ) %>%
  rename(
    parameter = variable,
    parameter_value = value
  )
```

Since we will be using these parameters as dependent variables, let's create a form of these that is wide.

```{r params-wide}
params_wide <- spread(params_sub, parameter, parameter_value)
params_wide
```

To make the ED2 output more manageable, let's calculate the annual total GPP and NPP (and convert them to more sensible units), the max annual LAI, and the max annual water table depth.
(We'll use `data.table` to make this a bit faster.)

```{r ed-summary}
.data.table.aware <- TRUE
data.table::setDT(ed_in)

ed_wide <- data.table::dcast(ed_in, time + run_id ~ variable)
ed_wide <- ed_wide[, year := lubridate::year(time)]

ed_summary_wide <- ed_wide[, .(
  GPP = udunits2::ud.convert(sum(gpp), "kg m-2", "Mg ha-1") * 60 * 30,
  NPP = udunits2::ud.convert(sum(npp), "kg m-2", "Mg ha-1") * 60 * 30,
  LAI = max(lai),
  water_table = min(watertable)
), by = .(run_id, year)]

ed_summary <- data.table::melt(
  ed_summary_wide,
  id.vars = c("run_id", "year")
) %>% as_tibble()
```

Now, let's have a look at the results.
For reference, I also include values from @hardiman_2013_maintaining in red (min-max as dashed lines, mean as solid line).

```{r ts-plot}
hardiman <- tribble(
  ~variable, ~low, ~mean, ~hi,
  "LAI", 1.8, 4.14, 6.56,
  "NPP", 1.68, 3.11, 7.26
)

ggplot(ed_summary) +
  aes(x = year, y = value, group = run_id) +
  geom_line(alpha = 0.2) +
  geom_hline(aes(yintercept = low), data = hardiman,
             color = "red", linetype = "dashed") +
  geom_hline(aes(yintercept = mean), data = hardiman,
             color = "red") +
  geom_hline(aes(yintercept = hi), data = hardiman,
             color = "red", linetype = "dashed") +
  facet_wrap(~variable, scales = "free_y")
```

Compute the average across the time series (excluding establishment years 1902-1912), and join with parameter values.

```{r time-average}
time_average <- ed_summary %>%
  filter(year > 1912) %>%
  group_by(variable, run_id) %>%
  summarize(value = mean(value)) %>%
  left_join(params_sub)
```

Calculate linear regression between parameter values and time-series averages.
Note that there because parameters are PFT-specific, I have to fit regressions for each PFT parameter (so 4 regressions per variable-parameter combination).

```{r linmod-calc}
linmod <- time_average %>%
  group_by(variable, pft, parameter) %>%
  nest() %>%
  mutate(
    fit = suppressWarnings(map(
      data,
      possibly(lm, NULL),
      formula = value ~ parameter_value
    )),
    failed_fit = map_lgl(fit, is.null),
  ) %>%
  filter(!failed_fit) %>%
  mutate(
    r2 = map2_dbl(
      fit,
      data,
      ~suppressWarnings(modelr::rsquare(.x, .y))
    ),
    slope = map_dbl(fit, ~coefficients(.x)[[2]])
  )
```

Now, let's look at the top predictors:

```{r linmod-output}
linmod %>%
  select(variable, pft, parameter, r2, slope) %>%
  arrange(desc(r2)) %>%
  print(n = 20)
```

### 2019-02-13 analysis/paper/exploratory.Rmd
Add LAI by PFT ensemble analysis

@@
 
+# Summary of output
+
 Now, let's have a look at the results.
@@
 ```
+
+# LAI, by ensemble member
+
+Load the LAI output (generated by script `analysis/scratch/read_ed_lai.R`).
+
+```{r load-lai}
+lai_file <- here::here("analysis/data/derived-data/ed-lai-output.fst")
+all_lai <- fst::read_fst(lai_file)
+```
+
+Calculate annual maximum and plot.
+
+```{r lai-ensemble-plot}
+all_lai %>%
+  group_by(year = lubridate::floor_date(month, "year"),
+           pft, ensemble) %>%
+  summarize(lai = max(lai)) %>%
+  ungroup() %>%
+  filter(pft != "Late conifer") %>%
+  ggplot() +
+  aes(x = year, y = lai, color = pft) +
+  geom_line() +
+  facet_wrap(~ensemble)
+```

### 2019-02-13 analysis/scratch/read_ed_ensemble.R

@@ -109,16 +109,7 @@ for (i in seq_along(years)) {
 #   - Loop through first 
 
 ##################################################
-library(magrittr)
-library(data.table)
-library(ggplot2)
 
-workflow_id <- 99000000066
-outdir <- file.path("analysis", "data", "model_output", workflow_id)
-runs <- list.files(outdir)
-
-run <- runs[[1]]
-rundir <- file.path(outdir, run)
 
 read_variable <- function(nc, varname) {
   var_dimnames <- purrr::map_chr(nc[[c("var", varname, "dim")]], "name")

### 2019-02-13 analysis/scratch/read_ed_lai.R

@@ -1,0 +1,54 @@
+library(tidyverse)
+library(data.table)
+library(furrr)
+
+workflow_id <- 99000000066
+outdir <- file.path("analysis", "data", "model_output", workflow_id)
+runs <- list.files(outdir)
+
+get_monthly_lai <- function(rundir, year) {
+  filename <- file.path(
+    rundir,
+    glue::glue("analysis-T-{year}-00-00-000000-g01.h5")
+  )
+  nc <- ncdf4::nc_open(filename)
+  lai_raw <- ncdf4::ncvar_get(
+    nc,
+    "LAI_PY",
+    start = c(6, 1, 1, 1),
+    count = c(6, -1, -1, -1) # Only include PFTs 6, (7), 8, 9, 10, 11
+  )[-2,,] # Drop PFT 7
+
+  pft_levels <- c("Early hardwood", "Mid hardwood", "Late hardwood",
+                  "Pine", "Late conifer")
+  pftnames <- factor(pft_levels[c(4, 5, 1:3)], pft_levels)
+  basedate <- ISOdate(year, 1, 1, 0, 0, 0, "UTC")
+
+  data.table::melt(lai_raw) %>%
+    tibble::as_tibble() %>%
+    dplyr::rename(ipft = Var1, cohort = Var2, timestep = Var3) %>%
+    dplyr::mutate(
+      pft = pftnames[ipft],
+      dt = basedate + lubridate::dhours((timestep - 1) / 2),
+      month = lubridate::floor_date(dt, "month")
+    ) %>%
+    dplyr::group_by(month, pft, cohort) %>%
+    dplyr::summarize(lai = mean(value)) %>%
+    dplyr::ungroup() %>%
+    dplyr::group_by(month, pft) %>%
+    dplyr::summarize(lai = sum(lai)) %>%
+    dplyr::ungroup()
+}
+
+get_ensemble_lai <- function(rundir, years = seq(1902, 1990)) {
+  ensemble <- basename(rundir)
+  future_map_dfr(years, get_monthly_lai, rundir = rundir) %>%
+    dplyr::mutate(ensemble = !!ensemble)
+}
+
+plan(list(tweak(multiprocess, workers = 2),
+          tweak(multiprocess, workers = 4)))
+all_lai <- future_map_dfr(file.path(outdir, runs), possibly(get_ensemble_lai, NULL),
+                          .progress = TRUE)
+
+fst::write_fst(all_lai, "analysis/data/derived-data/ed-lai-output.fst")

### 2019-03-22 analysis/paper/paper.Rmd
Publish current version of paper on GitHub

@@
   echo = FALSE,
-  comment = "#>",
-  fig.path = "../figures/"
+  comment = "#>"
+  ## fig.path = "../figures/"
 )
 
-library(fortebaseline) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+## library(fortebaseline) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
 ```
@@
 
-```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
+```{r, include = FALSE}
+lai_10yr_cap <- paste(
+  "ED2 leaf area index (LAI) predictions by plant functional type."
+)
+```
+
+```{r lai_10yr, fig.cap = lai_10yr_cap, fig.dim = c(10, 7)}
+readd(results_plot)
 ```

### 2019-03-22 analysis/drake.R

@@ -1,3 +1,4 @@
+#!/usr/bin/env Rscript
 library(drake)
 library(fortebaseline)
 library(ggplot2)
@@ -21,7 +22,6 @@ import::from("future", "plan", "availableCores", .into = "")
 # end imports
 
 expose_imports("fortebaseline")
-theme_set(theme_cowplot())
 
 created_since <- "2019-03-20"
 
@@ -92,6 +92,7 @@ plan <- drake_plan(
     labs(y = "Leaf area index", color = "PFT") +
     scale_x_datetime(date_breaks = "2 years",
                      date_labels = "%Y") +
+    cowplot::theme_cowplot() +
     theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
           axis.title.x = element_blank()),
   results_plot_png = save_plot(file_out(!!(
@@ -99,11 +100,11 @@ plan <- drake_plan(
       dir_create() %>%
       path("lai_10year.png")
   )), results_plot, base_width = 10, base_height = 7),
-  paper = rmarkdown::render(
+  paper = target(
+    rmarkdown::render(
     file_in(!!(here("analysis", "paper", "paper.Rmd"))),
-    "github_document",
-    output_dir = "_rendered_output"
-  )
+    "github_document"
+  ), trigger = trigger(condition = !file.exists("analysis/paper/paper.md")) )
 )
 
 future::plan(future.callr::callr)

### 2019-03-22 publish.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+set -e
+
+# Run drake to compile everything
+./analysis/drake.R
+
+# Copy the output
+cd analysis/paper
+cp -rf paper.md paper_files _rendered_output/
+
+# Force-push the results up to GitHub
+cd _rendered_output
+git checkout rendered
+git add .
+git commit --amend -m "Rendered output"
+git push --force

### 2019-03-25 analysis/paper/ed-notes.Rmd
Move old ED module notes from paper to notes

@@
 With N limitation, decay rate is 100.2 year^-1^; otherwise, 0.2 year^-1^.
+
+# Old notes
+
+## ED modules
+
+NOTE: Not a real section; just brainstorming notes for now
+Details about these sub-models will probably makes their way into the methods and/or appendices.
+
+### High priority
+
+These are switches that I understand (more-or-less), and that I think will be most relevant to FoRTe.
+
+- Finite canopy radius (`CROWN_MOD`)
+    - Options:
+        - (0; default): Complete shading
+        - (1): Finite canopy radius (Dietze 2008)
+    - See current hypotheses
+    - This model is based on crown area allometries from @dietze_2008_capturing
+- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
+    - Options:
+        - (0, default): No limitation
+        - (1) Limitation based on internal model (TODO: what is it?)
+    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
+- Nitrogen limitation of decomposition (`N_DECOMP_LIM`)
+    - Options:
+        - (0, default): No limitation
+        - (1) Limitation based on internal model (TODO: what is it?)
+    - Same as above
+
+### Medium priority
+
+These are switches that may be relevant to FoRTE, but may be harder to pull off.
+
+- Big leaf model (no demographics) (`IBIGLEAF`)
+    - Simplistic case
+    - Only one PFT per patch (by definition). Can ask: Does it get the right one?
+    - Compare fluxes against other cases.
+- Canopy radiative transfer model (`ICANRAD`)
+    - Options:
+        - (0, deprecated?) Original two-stream model (Medvigy 2006)
+        - (1) Multiple-scattering model [@zhao_2006_modeling; @zhao_2005_multiple]
+        - (2, default) Updated two-stream model [@liou_2002]
+    - In theory, multiple-scattering model is more detailed and should give "better" results. Open question as to how that affects predictions.
+    - Good candidate for exploratory analysis.
+I already implemented the two-stream model in R -- could be a good idea to implement the multi-scatter model as well and compare their radiation profiles.
+- Allometry (`IALLOM`)
+    - Options:
+        - (0, default): Original ED2
+        - (1)
+            - Structural biomass parameterized so total AGB similar to @baker_2004_variation, eq. 2. `Balive` similar to ED2.
+            - Root depth configured so root depth is 5m for canopy trees and 0.5 m for grasses/seedlings
+            - Crown area as in @poorter_2006_architecture, with maximum crown area
+        - (2) Like (1), but...
+            - Height to DBH allometry from @poorter_2006_architecture
+            - `Balive` has extra allometric equations (see @cole_2006_allometric and @calvo-alvarado_2008_allometric)
+        - (3) Like (2), but tropical leaf biomass estimated from @lescure_1983_la, and PFT differences in leaf biomass based on SLA rather than wood density.
+Also, split leaf allometry into small (sapling) and large (tree) classes, as in @cole_2006_allometric
+- Temperature variation of physiology (`IPHYSIOL`)
+    - Options:
+        - (0) Arrhenius function, `Gamma*` based on Foley et al. (1996) (TODO: REF)
+        - (1) Arrhenius, `Gamma*` based on Michaelis-Mentel coefficients for CO2 and O2, as in @farquhar_1980_biochemical
+        - (2; default) Q10 equations and parameters of @collatz_1991_physiological, with MM coefficients. High/low temperature correction in @moorcroft_2001_method
+        - (3) As (2), but `Gamma*` as in @farquhar_1980_biochemical and CLM.
+    - Need to do some exploratory analysis here. Grab the relevant equations and coefficients and plot them (outside of ED) to compare behavior.
+        - Possible angle: More important at temperature extremes?
+    - Consult @rogers_2016_roadmap for inspiration
+- Soil decomposition temperature sensitivity (`DECOMP_SCHEME`)
+    - Options:
+        - (0; default): Exponential
+        - (1) Lloyd and Taylor (1994)
+    - (1) requires additional parameters, so should increase parameteric uncertainty.
+    - Hypothesis: Affects overall GPP/R~E~ balance, but not successional dynamics.
+- Water limitation of photosynthesis (`H2O_PLANT_LIM`)
+    - Options:
+        - (0) No limitation.
+        - (1; default): Supply is total soil water above wilting point integrated across all levels in rooting zone.
+        - (2): Supply is total soil water at field capacity minus wilting point, scaled by "wilting factor":
+          ```
+          # k -- Soil layer
+          # z -- Layer depth
+          # psi -- Matric potential (for layer, or at wp or fc)
+          # wp -- wilting point
+          # fc -- field capacity
+          (psi(k) - (H - z(k)) - psi_wp) / (psi_fc - psi_wp)
+          ```
+    - Sensitivity should depend on site climate and soil conditions.
+    - UMBS are pretty wet sites, so shouldn't expect a huge effect.
+- Quantum yield model (`QUANTUM_EFFICIENCY_T`)
+    - Options:
+        - (0, default): Constant quantum efficiency
+        - (1): Quantum efficiency varies with temperature, based on Ehleringer (1978) (TODO: Get ref)
+    - Uncertainty should increase with (1) assuming parameters are tunable. 
+    - Depends on where UMBS conditions fall within the range of temperatures. Good candidate for exploratory analysis outside of ED.
+- Structural growth (`ISTRUCT_GROWTH_SCHEME`)
+    - Options:
+        - (0; default) Use all `bstorage` (NSC?)
+        - (1) Reserve `bstorage` to re-flush canopy and fine roots once before calculating structural growth.
+    - (1) is a more conservative strategy, so should reduce the growth and relative dominance of early-successional PFTs.
+In particular, growth rate should decline more with size (because more LAI and fine roots means more reserves needed).
+- Stomatal conductance model (`ISTOMATA_SCHEME`)
+    - Options:
+        - (0, default): Leuning
+        - (1) Katul optimization model (see Xu et al. 2016 New Phyt.)
+    - Need to read more about this...
+- Trait plasticity (`TRAIT_PLASTICITY_SCHEME`)
+    - Options:
+        - (0, default): Static traits
+        - (1) Vm0 (SLA) decrease (increase) with shading, based on Lloyd et al. (2010 Biogeosciences). Updated yearly.
+        - (2) Same as (1), but updated monthly
+        - (-1) Same as 1, but adjust SLA using height
+        - (-2) Same as (2), but adjust SLA using height
+    - If trait plasticity is optimal, this should reduce (?) the relative dominance of early-successional PFTs.
+    - Probably introduces considerable predictive uncertainty, given new parameters for plasticity models.
+- Carbon balance contribution to mortality (`IDDMORT_SCHEME`)
+    - Options:
+        - (0, default): Carbon balance in terms of fluxes only
+        - (1) Carbon balance is offset by storage pool (i.e. carbon storage is an additional buffer against mortality)
+    - Additional protection from (1) should reduce mortality for all trees, so will probably reinforce dominance of early successional trees?
+    - But likely to interact with other schemes, especially structural growth, which determines size of NSC reserves
+- Percolation and infiltration (`IPERCOL`)
+    - Options:
+        - (0) Constant soil conductivity. Temporary surface water (only exists if top soil layer is saturated) exceeding 1:9 liquid:ice ratio shed via percolation.
+        - (1, default) Constant soil conductivity, with percolation model from Anderson (1976 NOAA technical report NWS 19 TODO).
+Temporary surface water can exist after heavy rain event even if soil doesn't saturate.
+        - (2) Like (1), but soil conductivity decreases with depth even under constant soil moisture
+    - Presumably, not a huge impact at UMBS because water shouldn't be an issue.
+    - See also `RUNOFF_TIME` (e-folding lifetime of temporary surface water)
+- Carbon stress scheme (`CBR_SCHEME`)
+    - Options:
+        - (0, default): Single stress, with maximum carbon balance determined at full sun and no moisture
+        - (1) Co-limitation of light and moisture; calculate `cb/cb_lightmax` and `cb/cb_moistmax`, and weight based on `DDMORT_CONST` (tunable global[?] parameter)
+        - (2) "Leibig style"; limitation based on whether soil or light are more important (`cb/max(cb_lightmax, cb_moistmax)`)
+    - No idea about this one. Need to do more reading.
+
+### Lower priority
+
+These are processes that I think will not have huge impacts, are farther-removed from the FoRTE project goals, or are things I simply don't understand very well (particularly all the micro-meteorology stuff).
+
+- Branches (`IBRANCH_THERMO`)
+    - Options:
+        - (0) No branches in energy balance or biophysics
+        - (1; default) Branches in energy balance, but not biophysics
+        - (2) Branches in both
+    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
+- Canopy roughness model (`ICANTURB`)
+    - Options:
+        - (0): Based on Leuning et al. (1995; TODO REF). Wind computed using "similarity theory" for top cohort, then extinguished with cumulative LAI.
+        - (1, default) "Default ED2.1 scheme" (TODO: 0? Figure it out!), but using "zero-plane" displacement height.
+        - (2) Massman (1997; TODO) using constant drag and no sheltering factor
+        - (3) Also Massman, but with varying drag and sheltering within canopy
+        - (4) Like 0, but calculate ground conductance following CLM (eq. 5.98-5.100)
+    - No idea, except that (3) will have higher uncertainty than (2). Need to read more about this.
+- Similarity theory model (for canopy roughness) (`ISFCLYRM`)
+    - Computes `u*`, `T*`, etc.
+    - Options:
+        - (1) BRAMS default, based on Louis (1979 TODO). 
+        - (2) Oncley and Dudhia (1995 TODO), based on MM5
+        - (3) Beljaars and Holtslag (1991 TODO); Like (2), but alternative method for stable case that has more mixing
+        - (4, default) CLM (2004; TODO: Check that this is same as 2013?). Like (2) and (3), but with special functions to deal with very stable cases.
+    - Again, no idea. Need to read more about micro-meteorology and how it affects vegetation processes.
+- Ground to canopy conductance (`IED_GRNDVAP`)
+    - Options:
+        - (0, default): Modified Lee Pielke (1992 TODO), with field capacity, but with beta factor without square (as in Noilhan and Planton 1989 TODO). Like original ED2 and LEAF-3.
+        - (1-4) Mahfouf and Noilhan (1991 TODO), #1-4
+        - (5) Combination of (1) (alpha) and (2) (soil resistance)
+    - No idea. Again, need to do some micro-met reading.
+- Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)
+    - Options:
+        - (0, default) No treefall disturbance
+        - (> 0) Disturbance rate, in 1/years. Will create new patches.
+        - (< 0) Also rate, but will be added to mortality (kill plants, but no new patches)
+    - Reference value for this is 0.014 trees year^-1^ (note that because of ED's cohort approximation, number of individuals is continuous, not discrete)
+    - See also `TIME2CANOPY`, minimum age for tree-fall disturbance.
+- Growth respiration (`GROWTH_RESP_SCHEME`)
+    - Options:
+        - (0, default) Growth respiration is a tax on GPP, at PFT-specific rate. Treated as aboveground wood to canopy-airspace flux.
+        - (1) Growth respiration calculated as in (0), but split into leaf, sapwood-a, sapwood-b, and root sources proportionally to biomass.
+No effect on overall C budget, but greater within-ecosystem flux resolution.
+    - These should produce identical results, and if they don't, that means there is a bug!
+    - If output is identical, we should set to (1), since we can always calculate total respiration otherwise.
+- Storage respiration (`STORAGE_RESP_SCHEME`) -- same as growth
+    - Note that this whole system is a _hack_, because there is no real "storage respiration" process.
+In PEcAn, we hard-code the relevant parameters to zero.
+An interesting question is whether we may be able to get better results with this hack enabled -- i.e.
+if it compensates a problem in some other process.
+
+### Almost-certainly omit
+
+These configurations are almost certainly irrelevant to the goals of this paper or will not apply at UMBS (e.g. grass or tropical species).
+Mostly for reference.
+
+- Grass (`IGRASS`) -- I'm not running with any grass PFTs.
+- Phenology (`IPHEN_SCHEME`) -- Differences affect grasses and tropical species. Conifers are always evergreen, and hardwoods are always cold-deciduous (need to check that schemes are the same).
+- Seed dispersal (`REPRO_SCHEME`) -- Only seems to matter for multi-site runs, with some toggles for tropical (drought-deciduous) species.
+- Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
+- Fire disturbance (`INCLUDE_FIRE`) -- We know there weren't any fires at UMBS during this time, and more generally are interested in stability.
+    - This could potentially be fun to play with in a later paper about disturbance regimes. Potentially, with Jackie Matthes?
+- Anthropogenic disturbances (`IANTH_DISTURB`) -- We are currently looking at stable state, without disturbances.
+But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.

### 2019-03-25 analysis/paper/paper.Rmd
Move old ED module notes from paper to notes

--- 2019-01-31 analysis/paper/paper.Rmd
@@
 
-## ED modules
+### Additional references
 
-NOTE: Not a real section; just brainstorming notes for now
-Details about these sub-models will probably makes their way into the methods and/or appendices.
-
-### High priority
-
-These are switches that I understand (more-or-less), and that I think will be most relevant to FoRTe.
-
-- Finite canopy radius (`CROWN_MOD`)
-    - Options:
-        - (0; default): Complete shading
-        - (1): Finite canopy radius (Dietze 2008)
-    - See current hypotheses
-    - This model is based on @dietze_2008_capturing
-- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
-    - Options:
-        - (0, default): No limitation
-        - (1) Limitation based on internal model (TODO: what is it?)
-    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
-- Nitrogen limitation of decomposition (`N_DECOMP_LIM`)
-    - Options:
-        - (0, default): No limitation
-        - (1) Limitation based on internal model (TODO: what is it?)
-    - Same as above
-
-### Medium priority
-
-These are switches that may be relevant to FoRTE, but may be harder to pull off.
-
-- Big leaf model (no demographics) (`IBIGLEAF`)
-    - Simplistic case
-    - Only one PFT per patch (by definition). Can ask: Does it get the right one?
-    - Compare fluxes against other cases.
-- Temperature variation of physiology (`IPHYSIOL`)
-    - Options:
-        - (0) Arrhenius function, `Gamma*` based on Foley et al. (1996) (TODO: REF)
-        - (1) Arrhenius, `Gamma*` based on Michaelis-Mentel coefficients for CO2 and O2, as in Farquhar et al. (1980) (TODO: REF)
-        - (2; default) Q10 equations and parameters of Collatz et al. (1991), with MM coefficients. High/low temperature correction in Moorcroft et al. (2001)
-        - (3) As (2), but `Gamma*` as in Farquhar et al. (1980) and CLM.
-    - Need to do some exploratory analysis here. Grab the relevant equations and coefficients and plot them (outside of ED) to compare behavior.
-        - Possible angle: More important at temperature extremes?
-    - Consult @rogers_2016_roadmap for inspiration
-- Soil decomposition temperature sensitivity (`DECOMP_SCHEME`)
-    - Options:
-        - (0; default): Exponential
-        - (1) Lloyd and Taylor (1994)
-    - (1) requires additional parameters, so should increase parameteric uncertainty.
-    - Hypothesis: Affects overall GPP/R~E~ balance, but not successional dynamics.
-- Water limitation of photosynthesis (`H2O_PLANT_LIM`)
-    - Options:
-        - (0) No limitation.
-        - (1; default): Supply is total soil water above wilting point integrated across all levels in rooting zone.
-        - (2): Supply is total soil water at field capacity minus wilting point, scaled by "wilting factor":
-          ```
-          # k -- Soil layer
-          # z -- Layer depth
-          # psi -- Matric potential (for layer, or at wp or fc)
-          # wp -- wilting point
-          # fc -- field capacity
-          (psi(k) - (H - z(k)) - psi_wp) / (psi_fc - psi_wp)
-          ```
-    - Sensitivity should depend on site climate and soil conditions.
-    - UMBS are pretty wet sites, so shouldn't expect a huge effect.
-- Quantum yield model (`QUANTUM_EFFICIENCY_T`)
-    - Options:
-        - (0, default): Constant quantum efficiency
-        - (1): Quantum efficiency varies with temperature, based on Ehleringer (1978) (TODO: Get ref)
-    - Uncertainty should increase with (1) assuming parameters are tunable. 
-    - Depends on where UMBS conditions fall within the range of temperatures. Good candidate for exploratory analysis outside of ED.
-- Structural growth (`ISTRUCT_GROWTH_SCHEME`)
-    - Options:
-        - (0; default) Use all `bstorage` (NSC?)
-        - (1) Reserve `bstorage` to re-flush canopy and fine roots once before calculating structural growth.
-    - (1) is a more conservative strategy, so should reduce the growth and relative dominance of early-successional PFTs.
-In particular, growth rate should decline more with size (because more LAI and fine roots means more reserves needed).
-- Stomatal conductance model (`ISTOMATA_SCHEME`)
-    - Options:
-        - (0, default): Leuning
-        - (1) Katul optimization model (see Xu et al. 2016 New Phyt.)
-    - Need to read more about this...
-- Trait plasticity (`TRAIT_PLASTICITY_SCHEME`)
-    - Options:
-        - (0, default): Static traits
-        - (1) Vm0 (SLA) decrease (increase) with shading, based on Lloyd et al. (2010 Biogeosciences). Updated yearly.
-        - (2) Same as (1), but updated monthly
-        - (-1) Same as 1, but adjust SLA using height
-        - (-2) Same as (2), but adjust SLA using height
-    - If trait plasticity is optimal, this should reduce (?) the relative dominance of early-successional PFTs.
-    - Probably introduces considerable predictive uncertainty, given new parameters for plasticity models.
-- Carbon balance contribution to mortality (`IDDMORT_SCHEME`)
-    - Options:
-        - (0, default): Carbon balance in terms of fluxes only
-        - (1) Carbon balance is offset by storage pool (i.e. carbon storage is an additional buffer against mortality)
-    - Additional protection from (1) should reduce mortality for all trees, so will probably reinforce dominance of early successional trees?
-    - But likely to interact with other schemes, especially structural growth, which determines size of NSC reserves
-- Percolation and infiltration (`IPERCOL`)
-    - Options:
-        - (0) Constant soil conductivity. Temporary surface water (only exists if top soil layer is saturated) exceeding 1:9 liquid:ice ratio shed via percolation.
-        - (1, default) Constant soil conductivity, with percolation model from Anderson (1976 NOAA technical report NWS 19 TODO).
-Temporary surface water can exist after heavy rain event even if soil doesn't saturate.
-        - (2) Like (1), but soil conductivity decreases with depth even under constant soil moisture
-    - Presumably, not a huge impact at UMBS because water shouldn't be an issue.
-    - See also `RUNOFF_TIME` (e-folding lifetime of temporary surface water)
-- Carbon stress scheme (`CBR_SCHEME`)
-    - Options:
-        - (0, default): Single stress, with maximum carbon balance determined at full sun and no moisture
-        - (1) Co-limitation of light and moisture; calculate `cb/cb_lightmax` and `cb/cb_moistmax`, and weight based on `DDMORT_CONST` (tunable global[?] parameter)
-        - (2) "Leibig style"; limitation based on whether soil or light are more important (`cb/max(cb_lightmax, cb_moistmax)`)
-    - No idea about this one. Need to do more reading.
-
-### Lower priority
-
-These are processes that I think will not have huge impacts, are farther-removed from the FoRTE project goals, or are things I simply don't understand very well (particularly all the micro-meteorology stuff).
-
-- Branches (`IBRANCH_THERMO`)
-    - Options:
-        - (0) No branches in energy balance or biophysics
-        - (1; default) Branches in energy balance, but not biophysics
-        - (2) Branches in both
-    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
-- Canopy roughness model (`ICANTURB`)
-    - Options:
-        - (0): Based on Leuning et al. (1995; TODO REF). Wind computed using "similarity theory" for top cohort, then extinguished with cumulative LAI.
-        - (1, default) "Default ED2.1 scheme" (TODO: 0? Figure it out!), but using "zero-plane" displacement height.
-        - (2) Massman (1997; TODO) using constant drag and no sheltering factor
-        - (3) Also Massman, but with varying drag and sheltering within canopy
-        - (4) Like 0, but calculate ground conductance following CLM (eq. 5.98-5.100)
-    - No idea, except that (3) will have higher uncertainty than (2). Need to read more about this.
-- Similarity theory model (for canopy roughness) (`ISFCLYRM`)
-    - Computes `u*`, `T*`, etc.
-    - Options:
-        - (1) BRAMS default, based on Louis (1979 TODO). 
-        - (2) Oncley and Dudhia (1995 TODO), based on MM5
-        - (3) Beljaars and Holtslag (1991 TODO); Like (2), but alternative method for stable case that has more mixing
-        - (4, default) CLM (2004; TODO: Check that this is same as 2013?). Like (2) and (3), but with special functions to deal with very stable cases.
-    - Again, no idea. Need to read more about micro-meteorology and how it affects vegetation processes.
-- Ground to canopy conductance (`IED_GRNDVAP`)
-    - Options:
-        - (0, default): Modified Lee Pielke (1992 TODO), with field capacity, but with beta factor without square (as in Noilhan and Planton 1989 TODO). Like original ED2 and LEAF-3.
-        - (1-4) Mahfouf and Noilhan (1991 TODO), #1-4
-        - (5) Combination of (1) (alpha) and (2) (soil resistance)
-    - No idea. Again, need to do some micro-met reading.
-- Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)
-    - Options:
-        - (0, default) No treefall disturbance
-        - (> 0) Disturbance rate, in 1/years. Will create new patches.
-        - (< 0) Also rate, but will be added to mortality (kill plants, but no new patches)
-    - Reference value for this is 0.014 trees year^-1^ (note that because of ED's cohort approximation, number of individuals is continuous, not discrete)
-    - See also `TIME2CANOPY`, minimum age for tree-fall disturbance.
-- Growth respiration (`GROWTH_RESP_SCHEME`)
-    - Options:
-        - (0, default) Growth respiration is a tax on GPP, at PFT-specific rate. Treated as aboveground wood to canopy-airspace flux.
-        - (1) Growth respiration calculated as in (0), but split into leaf, sapwood-a, sapwood-b, and root sources proportionally to biomass.
-No effect on overall C budget, but greater within-ecosystem flux resolution.
-    - These should produce identical results, and if they don't, that means there is a bug!
-    - If output is identical, we should set to (1), since we can always calculate total respiration otherwise.
-- Storage respiration (`STORAGE_RESP_SCHEME`) -- same as growth
-    - Note that this whole system is a _hack_, because there is no real "storage respiration" process.
-In PEcAn, we hard-code the relevant parameters to zero.
-An interesting question is whether we may be able to get better results with this hack enabled -- i.e.
-if it compensates a problem in some other process.
-
-### Almost-certainly omit
-
-These configurations are almost certainly irrelevant to the goals of this paper or will not apply at UMBS (e.g. grass or tropical species).
-Mostly for reference.
-
-- Allometry (`IALLOM`) -- Seems to mostly affect tropical PFTs (TODO: Check that this is the case).
-- Grass (`IGRASS`) -- I'm not running with any grass PFTs.
-- Phenology (`IPHEN_SCHEME`) -- Differences affect grasses and tropical species. Conifers are always evergreen, and hardwoods are always cold-deciduous (need to check that schemes are the same).
-- Seed dispersal (`REPRO_SCHEME`) -- Only seems to matter for multi-site runs, with some toggles for tropical (drought-deciduous) species.
-- Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
-- Fire disturbance (`INCLUDE_FIRE`) -- We know there weren't any fires at UMBS during this time, and more generally are interested in stability.
-    - This could potentially be fun to play with in a later paper about disturbance regimes. Potentially, with Jackie Matthes?
-- Anthropogenic disturbances (`IANTH_DISTURB`) -- We are currently looking at stable state, without disturbances.
-But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.
+@katul_2003_relationship
 

### 2019-03-25 analysis/paper/paper.Rmd
[paper] Add minor notes about RTM and crown model

@@
 In trying to anticipate future changes to terrestrial ecosystems, ecologists have four general techniques.
-One approach is to look to the past to see how ecosystems responded to similar changes (REFs).
+One approach is to look to the past to see how ecosystems responded to similar changes (REFs -- e.g. PALEON?).
 This is challenging for two reasons:
@@
 - Radiative transfer
-  - Two-stream model
+  - Two-stream model (ED default)
     - Same general approach as in CLM 4.5 [@clm45_note], but rather than solving for just two layers (sunlit and shaded), solve for each canopy layer.
@@
     - Similar to approaches used in other models: JULES [@mercado_2007_improving], ORCHIDEE-CAN [@naudts_2015_vertically], CLM(SPA) [@bonan_2014_modeling]
-  - Multiple scatter model **TODO**
+  - Multiple scatter model
+    - Based on @zhao_2005_multiple
+- Crown model
+  - Infinite (ED2 default)
+    - Marginally taller cohorts completely shade out shorter ones
+  - Finite
+    - Cohorts have a finite crown area, so multiple cohorts have access to the same light level
+    - Based on crown area allometries from @dietze_2008_changing
+- Trait plasticity
 

### 2019-03-25 analysis/paper/paper.Rmd
[paper] Add more discussion of ED submodels to intro

@@
 This model has been used to simulate (examples...).
+Previous uncertainty analyses with ED:
+[@dietze_2014_quantitative]...
+More recently, [@raczka_2018_what]...
 
-Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
-Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
+Light availability is a first order control on photosynthesis, and competition for light is a key process in driving competitive dynamics and community succession.
+How light is distributed through a canopy is therefore critically important to accurately simulating community dynamics.
+Radiative transfer models are used to simulate light proliferation through a canopy.
+Detailed, 3D representations [@widlowski_2014_abstract; @widlowski_2007_third; @malenovsky_2008_influence] have prohibitively high computational costs for use in dynamic vegetation models.
+Therefore, these models have had to rely on computationally-efficient approximations.
+One common approach is the "two-stream approximation", originally developed for atmospheric radiative transfer [@meador_1980_two] and later adapted for plant canopies [@dickinson_1983_land; @sellers_1985_canopy].
+<!-- TODO: Further discussion? -->
+Although popular, the two-stream approach has some important limitations (REFS, details...).
+An alternative approach is the "multiple scattering" model of @zhao_2005_multiple.
+Accounting for multiple scattering enhances canopy light absorption, particularly in near-infrared wavelengths [@zhao_2005_multiple], therefore reducing the amount of light that makes it into the understory (TODO confirm).
+<!-- TODO: Further discussion -->
+
+Competition is also strongly determined by canopy structure and how it is represented (REF).
+In other words, how much does canopy height matter compared to gappiness and horizontal competition?
+C.f. [@hogan_2018_fast; @dietze_2008_changing].
+
+Trait plasticity with light level also plays a role -- i.e. shaded leaves adapt to lower light levels.
+C.f. [@keenan_2016_global; @lloyd_2010_optimisation; @niinemets_2016_within].
 
@@
 Enabling dynamic N cycling and a finite canopy radius model will alleviate these issues, but will significantly increase the parameteric uncertainty due to strong sensitivity to N-cycling and canopy gap fraction parameters, respectively.
-Constraining these parameters should be a high priority
+Constraining these parameters should be a high priority...
 
@@
 
+## Future directions: Disturbance response modeling
+
+Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
+Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
+
 # Conclusion

### 2019-03-25 analysis/paper/paper.Rmd
[paper] Start working on submodel descriptions

@@
 - Trait plasticity
+  - None (ED default) -- Trait values are the same at all canopy heights and light levels
+  - Plastic
+    - V~cmax~(0) decreases with shading, and SLA increases with shading, based on @lloyd_2010_optimisation
 
@@
 
+# Appendix 1: Submodel descriptions
+
+## Radiative transfer: Definitions
+
+- $\omega$ -- Clumping factor
+- $LAI$ -- Leaf area index
+- TAI -- Total area index (leaf and woody area)
+- $LAI_e$ / $TAI_e$ -- "Effective" area indices (accounting for clumping)
+
+## Two-stream RTM
+
+<!-- TODO -->
+
+## Multiple scatter RTM
+
+$$ LAI_e = \omega LAI $$
+
+$$ TAI_e = LAI_e + WAI $$
+
+$$ TAI_{l} = \frac{TAI_e}{CAI} $$
+
+Leaf ($w_l$) and wood ($w_w$) weights:
+
+$$ w_l = \frac{LAI_e}{TAI_e} $$
+$$ w_w = 1 - w_l $$
+
+Projected area, based on coefficients $\phi_1$ and $\phi_2$.
+
+$$ a_{proj} = \phi_1 + \phi_2 \mu $$
+
+$$ \lambda = \frac{a_{proj}}{\mu} $$
+
 ##### pagebreak

### 2019-03-26 analysis/paper/paper.Rmd
Add radiation profile analysis

@@
 
-```{r, include = FALSE}
+```{r captions, include = FALSE}
 lai_10yr_cap <- paste(
@@
 )
+
+rad_barplot_cap <- paste(
+  "Comparison of radiation profiles under two-stream and multiple scatter",
+  "canopy radiative transfer models."
+)
 ```
+
+Different model configurations produce significantly different ecological trajectories.
 
@@
 readd(results_plot)
+```
+
+```{r rad_prof, fig.cap = rad_barplot_cap, fig.dim = c(5, 7)}
+readd(rad_lineplot)
 ```

### 2019-03-26 R/download_history_file.R

@@ -1,0 +1,25 @@
+#' Download a history output file for a given workflow
+#'
+#' @template workflow_id
+#' @param year Output year
+#' @param month Output month
+#' @param day Output day
+#' @param hour Output hour
+#' @param outdir Target directory for storing files. Default is
+#'   `analysis/data/model_output/histfiles`
+#' @return Full path to downloaded file
+#' @author Alexey Shiklomanov
+#' @export
+download_history_file <- function(workflow_id, year, month, day, hour,
+                                  outdir = file.path("analysis", "data",
+                                                     "model_output", "histfiles")) {
+  filename <- sprintf("history-S-%04d-%02d-%02d-%02d0000-g01.h5",
+                      year, month, day, hour)
+  prun_url <- purrr::insistently(pecanapi::run_url, purrr::rate_backoff(max_times = 10))
+  url <- prun_url(workflow_id, filename, port = 7999)
+  full_outdir <- file.path(outdir, workflow_id)
+  dir.create(full_outdir, showWarnings = FALSE, recursive = TRUE)
+  outfile <- file.path(full_outdir, filename)
+  download.file(url, outfile)
+  outfile
+}

### 2019-03-26 R/rad_profile.R

@@ -1,0 +1,38 @@
+#' Read the radiation profile from a history file and return as a tidy
+#' `data.frame`
+#'
+#' @param file Full path to history file
+#' @return `data.frame` with columns `workflow_id`, `datetime`,
+#'   `cohort` (integer - tallest first), and one column per radiation flux.
+#' @author Alexey Shiklomanov
+#' @export
+read_tidy_rad_profile <- function(file) {
+  nc <- ncdf4::nc_open(file)
+  rad_prof <- ncdf4::ncvar_get(nc, "RAD_PROFILE_CO")
+  trad_prof <- t(rad_prof)
+  colnames(trad_prof) <- c(
+    "PAR_beam_down", "PAR_beam_up",
+    "PAR_diff_down", "PAR_diff_up",
+    "NIR_beam_down", "NIR_beam_up",
+    "NIR_diff_down", "NIR_diff_up",
+    "TIR_diff_down", "TIR_diff_up"
+  )
+  file_workflow <- stringr::str_extract(
+    dirname(file),
+    "9900[[:digit:]]+"
+  ) %>% as.numeric()
+  stopifnot(length(file_workflow) == 1, !is.na(file_workflow))
+  file_datetime <- stringr::str_extract(
+    basename(file),
+    "([[:digit:]]+-)+[[:digit:]]+"
+  ) %>% lubridate::ymd_hms()
+  stopifnot(length(file_datetime) == 1, inherits(file_datetime, "POSIXct"))
+  tibble::as_tibble(trad_prof) %>%
+    dplyr::mutate(
+      workflow_id = file_workflow,
+      datetime = file_datetime,
+      cohort = dplyr::row_number()
+    ) %>%
+    dplyr::select(workflow_id, datetime, cohort,
+                  dplyr::everything())
+}

### 2019-03-26 analysis/drake.R

@@ -1,12 +1,15 @@
 #!/usr/bin/env Rscript
 library(drake)
-library(fortebaseline)
 library(ggplot2)
+
+## library(fortebaseline)
+devtools::load_all(here::here(), attach_testthat = FALSE)
+expose_imports("fortebaseline")
 
 # begin imports
 import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
-             "if_else", "group_by_at", "bind_rows", .into = "")
+             "if_else", "group_by_at", "bind_rows", "summarize_all", .into = "")
 import::from("tidyr", "unnest", .into = "")
 import::from("tibble", "as_tibble", .into = "")
 import::from("here", "here", .into = "")
@@ -21,32 +24,43 @@ import::from("future.callr", "callr", .into = "")
 import::from("future", "plan", "availableCores", .into = "")
 # end imports
 
-expose_imports("fortebaseline")
-
 created_since <- "2019-03-20"
 
-workflow_df <- tbl(default_connection(), "workflows") %>%
-  filter(start_date == "1902-06-01",
-         end_date == "1912-12-31",
-         created_at > created_since) %>%
-  select(workflow_id = id, start_date, end_date, notes) %>%
-  collect() %>%
-  mutate(configs_df = map(notes, parse_notes),
-         workflow_id = as.numeric(workflow_id)) %>%
-  unnest(configs_df) %>%
-  select(-notes)
+workflow_df <- tryCatch(
+  readd(workflow_df_drake),
+  error = function(e) {
+    tbl(default_connection(), "workflows") %>%
+      filter(start_date == "1902-06-01",
+             end_date == "1912-12-31",
+             created_at > created_since) %>%
+      select(workflow_id = id, start_date, end_date, notes) %>%
+      collect() %>%
+      mutate(configs_df = map(notes, parse_notes),
+             workflow_id = as.numeric(workflow_id)) %>%
+      unnest(configs_df) %>%
+      select(-notes)
+  }
+)
 
-workflows_years <- expand.grid(
-  workflow_id = as.numeric(pull(workflow_df, workflow_id)),
-  year = seq(1902, 1912)
-) %>%
-  as_tibble() %>%
-  mutate(startmonth = case_when(year == 1902 ~ 6, TRUE ~ 1))
+workflows_years <- tryCatch(
[…]

### 2019-03-26 analysis/scratch/misc.R

@@ -310,3 +310,79 @@ rowMeans(h_light)
 url_t <- "http://localhost:7999/thredds/dodsC/outputs/PEcAn_99000000080/out/99000000048/analysis-T-1903-00-00-000000-g01.h5"
 hf_t <- ncdf4::nc_open(url_t)
 light <- ncdf4::ncvar_get(hf_t, "FMEAN_RAD_PRO")
+
+##################################################
+wf_sub <- workflow_df %>%
+  filter(!(n_limit_soil & !n_limit_ps)) %>%
+  select(-dplyr::ends_with("_date")) %>%
+  dplyr::arrange(workflow_id)
+
+ids <- wf_sub %>%
+  filter(!trait_plasticity, n_limit_ps, !n_limit_soil)
+# 88, 90, 104, 106
+
+wfid <- 99000000088
+## outfile <- pecanapi::run_dap(wfid, "analysis-D-1910-07-05-000000-g01.h5")
+hist_dates <- as_tibble(expand.grid(time = seq(0, 18, 6), day = seq(1, 10)))
+files <- with(hist_dates, sprintf("history-S-1910-07-%02d-%02d0000-g01.h5", day, time))
+
+outdir <- file.path("analysis/data/model_output/histfiles/", wfid)
+dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
+for (f in files) {
+  message("Processing file: ", f)
+  url <- pecanapi::run_url(wfid, f)
+  outfile <- file.path(outdir, f)
+  if (!file.exists(outfile)) download.file(url, outfile)
+  ## nc <- ncdf4::nc_open(url)
+  ## l[[f]] <- ncdf4::ncvar_get(nc, "RAD_PROFILE_CO")
+  ## ncdf4::nc_close(nc)
+}
+
+result <- purrr::map(
+  files,
+  function(x) {
+    path <- pecanapi::run_dap(wfid, x)
+    ## path <- file.path(outdir, x)
+    nc <- ncdf4::nc_open(path)
+    on.exit(ncdf4::nc_close(nc), add = TRUE)
+    ncdf4::ncvar_get(nc, "RAD_PROFILE_CO")
+  }
+)
+names(result) <- files
+
+tidy_rad_profile <- function(rad_prof) {
+  trad_prof <- t(rad_prof)
+  colnames(trad_prof) <- c(
+    "PAR_beam_down", "PAR_beam_up",
+    "PAR_diff_down", "PAR_diff_up",
[…]

### 2019-03-26 analysis/paper/paper.Rmd
Flesh out paper outline

@@
 ```
+
+# Outline
+
+- **Introduction**
+    - Why we need models?
+        - Forests are important, for C sequestration, other ecosystem services
+        - Policy decisions related to forests depend on future projections
+        - We need models to make future projections
+    - What kinds of models are out there?
+        - Traditional: "Big leaf" models
+            - Needed to be simple to run on old computers
+            - Don't need as much data
+            - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
+        - Individual-based models
+            - Much more detailed, but harder to run, and need much more data.
+            - Trade-off between complexity and accuracy.
+        - Cohort-based models like ED2 as a middle ground?
+    - How much complexity do we need to accurately model ecological succession?
+        - Lots of existing work on parameteric uncertainty (e.g. Dietze, Raczka, others).
+        - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
+    - What are the processes that drive succession of ecological communities?
+        - Competition for light
+        - Competition for nutrients
+        - Competition for water?
+        - Others?
+        - Brief review of community/ecosystem ecology of the Upper Midwest
+    - Guiding questions:
+        - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
+        - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
+        - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+    - Study overview:
+        - ED with factorial combination of submodels
+        - Ensemble analysis to look at parameter uncertainty
+- **Methods**
+    - Site description
+    - ED2 model description
+        - Common features
+        - Submodel descriptions (in brief; more details in appendices)
+            - Two-stream RTM
+            - Multiple-scatter RTM
+            - Finite canopy radius model
+            - Trait plasticity scheme
+            - N limitation (?)
+    - Data sources
+        - TRY data
+        - Other data? FIA? Remote sensing?
+    - Modeling workflow (PEcAn, etc.)
+    - Data analysis
+        - Structural vs. parametric uncertainty
+        - Comparison against "reality"
+- **Results**
+    - Summary:
+        - LAI time series with uncertainties, faceted by model structure
+    - Structural uncertainty details
+        - Fig: RTM comparison (Radiation profile)
+        - Fig: Crown model comparison
+        - Fig: Trait plasticity?
+    - Uncertainty analysis
+        - Which parameters contribute the most uncertainty?
+        - Structural vs. parameteric uncertainty
+- **Discussion**
+    - (Generally, mirrors results section)
+    - Overview of major findings
+    - Ecological implications of structural uncertainty
+    - Implications of uncertainty analysis for future data collection and model development
+    - Future directions: Predicting disturbance response
 
@@
 # Methods
+
+## General outline
 

### 2019-03-27 analysis/paper/paper.Rmd
Minor outline revision

@@
 
-Different model configurations produce significantly different ecological trajectories.
-
 ```{r lai_10yr, fig.cap = lai_10yr_cap, fig.dim = c(10, 7)}
@@
 ```
+`r lai_10yr_cap`
+
+Different model configurations produce significantly different ecological trajectories.
 
@@
 ```
+`r rad_barplot_cap`
 

### 2019-03-27 analysis/paper/paper.Rmd
Expand paper outline

--- 2019-03-26 analysis/paper/paper.Rmd
@@
 ## library(fortebaseline) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+n <- 0
+mdcap <- function(text) {
+  n <<- n + 1
+  out <- sprintf("**Figure %d**: %s", n, text)
+  if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
+    out
+  } else {
+    invisible(NULL)
+  }
+}
 ```
+
+# Outline
+
+- **Introduction**
+    - Why we need models?
+        - Forests are important, for C sequestration, other ecosystem services
+        - Policy decisions related to forests depend on future projections
+        - We need models to make future projections
+    - What kinds of models are out there?
+        - Traditional: "Big leaf" models
+            - Needed to be simple to run on old computers
+            - Don't need as much data
+            - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
+        - Individual-based models
+            - Much more detailed, but harder to run, and need much more data.
+            - Trade-off between complexity and accuracy.
+        - Cohort-based models like ED2 as a middle ground?
+    - How much complexity do we need to accurately model ecological succession?
+        - Lots of existing work on parameteric uncertainty (e.g. Dietze, Raczka, others).
+        - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
+    - What are the processes that drive succession of ecological communities?
+        - Competition for light
+        - Competition for nutrients
+        - Competition for water?
+        - Others?
+        - Brief review of community/ecosystem ecology of the Upper Midwest
+    - Guiding questions:
+        - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
+        - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
+        - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+    - Study overview:
+        - ED with factorial combination of submodels
+        - Ensemble analysis to look at parameter uncertainty
+- **Methods**
+    - Site description
+    - ED2 model description
+        - Common features
+        - Submodel descriptions (in brief; more details in appendices)
+            - Two-stream RTM
+            - Multiple-scatter RTM
+            - Finite canopy radius model
+            - Trait plasticity scheme
+            - N limitation (?)
+    - Data sources
+        - TRY data
+        - Other data? FIA? Remote sensing?
+    - Modeling workflow (PEcAn, etc.)
+    - Data analysis
+        - Structural vs. parametric uncertainty
+        - Comparison against "reality"
+- **Results**
+    - Summary:
+        - LAI time series with uncertainties, faceted by model structure
+    - Structural uncertainty details
+        - Understory light environment and how it affects species matters
+        - The multiple-scatter
+            - Fig: RTM comparison (Radiation profile)
+            - Fig: Crown model comparison
+            - Fig: Trait plasticity?
+    - Uncertainty analysis
+        - Which parameters contribute the most uncertainty?
+        - Structural vs. parameteric uncertainty
+- **Discussion**
+    - Ecological implications of structural uncertainty
+        - Predictions about ecological succession are very sensitive to the understory light environment.
+            - Need a good representation of canopy structure (finite canopy radius)
+            - How you model canopy radiative transfer matters! Small differences in the model can give significant results.
+        - But, trait plasticity can provide an alternative mechanism for success in the understory.
+            - Important modeling implication for all the emerging research on intra-specific trait variability.
+            - This combination of discrete functional types with plastic traits is closer to reality than purely trait-based ("eco-evolutionary?") models or completely discretized models with static traits on the other.
+            - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
+    - Implications of uncertainty analysis for future data collection and model development
+        - Parameter uncertainty is still a _huge_ source of overall model predictive uncertainty. Different parameters can give _qualitatively_ different ecological outcomes.
+        - Looking beyond aggregate variables (e.g. net primary productivity, total LAI) is essential to model evaluation and calibration. It's possible to generate a relatively?
+        - (Expected) Subsurface parameters (e.g. leaf-root ratios) have high sensitivity, but are poorly constrained.
+        - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
+    - Future directions: Predicting disturbance response
 
@@
 
+## General outline
+
 ## Site description
@@
 
+`r mdcap(lai_10yr_cap)`
+
+Different model configurations produce significantly different ecological trajectories.
+
 ```{r rad_prof, fig.cap = rad_barplot_cap, fig.dim = c(5, 7)}
@@
 ```
+
+`r mdcap(rad_barplot_cap)`
 

### 2019-03-27 analysis/paper/paper.Rmd
Add old LAI parameter uncertainty analysis to paper

@@
     - Future directions: Predicting disturbance response
+        - FoRTE project -- Forecasting a disturbance.
+            - Compare ecological resilience/memory between model and reality
+
+Journal ideas:
+
+- _Global Change Biology_ -- 8000 word limit may be OK, especially if many methodological details are punted to the appendices.
+- _Ecological Monographs_ -- To do this topic justice may require a lot of words and figures. It may be hard to squeeze all of this information into a regular research paper.
 
@@
 
-### Additional references
-
-@katul_2003_relationship
-
 # Methods
-
-## General outline
 
@@
 
-The Ecosystem Demography Model, version 2 (ED2) [@medvigy_2009_mechanistic].
+The Ecosystem Demography Model, version 2.X (ED2) [@medvigy_2009_mechanistic].
 
@@
 - Photosynthesis / stomatal conductance
+    - I think I need to cite @katul_2003_relationship somewhere in here
 - Allometries
@@
 
+## Data sources
+
+TRY data [@kattge_2011_try].
+
+Other data from FoRTE? FIA? Remote sensing?
+
+## Modeling workflow
+
+PEcAn parameter sensitivity and uncertainty analysis [@lebauer_2013_facilitating].
+Other PEcAn references: [@dietze_2013_improving].
+
+{TODO: Conceptual diagram of workflow}
+
 # Results
 
-```{r captions, include = FALSE}
+```{r captions}
 lai_10yr_cap <- paste(
@@
   "canopy radiative transfer models."
+)
+
+annual_lai_workflows_cap <- paste(
+  "Max annual leaf area index predictions by plant functional type",
+  "across different parameter combinations."
 )
@@
 
+## Structural uncertainty
+
 ```{r rad_prof, fig.cap = rad_barplot_cap, fig.dim = c(5, 7)}
@@
 `r mdcap(rad_barplot_cap)`
+
+Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation.
+
+## Parameter uncertainty
+
+{TODO: PEcAn variance decomposition analysis, by configuration.}
+
+```{r annual_lai_workflows, fig.cap = annual_lai_workflows_cap, fig.dim = c(10, 8)}
+readd(old_lai_plot)
+```
+
+`r mdcap(annual_lai_workflows_cap)`
+
+Different parameter combinations produce
 

### 2019-03-27 analysis/drake.R

@@ -9,12 +9,13 @@ expose_imports("fortebaseline")
 # begin imports
 import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
-             "if_else", "group_by_at", "bind_rows", "summarize_all", .into = "")
+             "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
+             .into = "")
 import::from("tidyr", "unnest", .into = "")
 import::from("tibble", "as_tibble", .into = "")
 import::from("here", "here", .into = "")
 import::from("fs", "dir_create", "path", .into = "")
-import::from("fst", "write_fst", .into = "")
+import::from("fst", "write_fst", "read_fst", .into = "")
 import::from("forcats", "fct_relabel", .into = "")
 import::from("cowplot", "save_plot", "theme_cowplot", .into = "")
 import::from("furrr", "future_pmap_dfr", .into = "")
@@ -23,6 +24,7 @@ import::from("purrr", "map", "possibly", "discard", "safely",
              "negate", "map_dfr", .into = "")
 import::from("future.callr", "callr", .into = "")
 import::from("future", "plan", "availableCores", .into = "")
+import::from("lubridate", "floor_date", .into = "")
 # end imports
 
 created_since <- "2019-03-20"
@@ -191,7 +193,28 @@ plan <- drake_plan(
   ),
   cohort_data = cohort_raw_list %>%
     discard(~is.null(.x[["result"]])) %>%
-    map_dfr("result")
+    map_dfr("result"),
+  # Old parameter sensitivity analysis
+  old_lai_summary = file_in(!!file.path("analysis", "data",
+                                        "derived-data", "ed-lai-output.fst")) %>%
+    read_fst() %>%
+    as_tibble() %>%
+    group_by(year = floor_date(month, "year"),
+             pft, run_id = ensemble) %>%
+    summarize(max_lai = max(lai)) %>%
+    ungroup(),
+  old_lai_plot = old_lai_summary %>%
+    filter(pft != "Late conifer") %>%
+    ggplot() +
+    aes(x = year, y = max_lai, color = pft) +
+    geom_line() +
+    facet_wrap(~run_id) +
+    labs(y = "Max annual leaf area index",
+         color = "PFT") +
+    theme_cowplot() +
+    theme(axis.title.x = element_blank(),
+          axis.text.x = element_text(angle = 90, vjust = 0.5),
[…]

### 2019-03-27 analysis/paper/paper.Rmd
Add another parameter means plot

@@
 )
-
-annual_lai_workflows_cap <- paste(
-  "Max annual leaf area index predictions by plant functional type",
-  "across different parameter combinations."
-)
 ```
@@
 
+```{r captions_uncertainty}
+annual_means_cap <- paste(
+  "Total annual gross (GPP) and net (NPP) primary productivity,",
+  "and maximum annual leaf area index (LAI)",
+  "from ED2 parameter ensemble simulations.",
+  "Red lines are mean (solid) and min/max (dashed) values",
+  "from Hardiman et al. (2013)."
+)
+
+annual_lai_workflows_cap <- paste(
+  "Max annual leaf area index predictions by plant functional type",
+  "across different parameter combinations."
+)
+```
+
 {TODO: PEcAn variance decomposition analysis, by configuration.}
+
+```{r annual_means, fig.cap = annual_means_cap, fig.dim = c(5, 7)}
+readd(old_ensemble_plot)
+```
+
+`r mdcap(annual_means_cap)`
 
@@
 
-Different parameter combinations produce
+Different parameter combinations produce qualitatively different successional outcomes.
 

### 2019-03-27 analysis/drake.R

@@ -11,12 +11,12 @@ import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
              "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
              .into = "")
-import::from("tidyr", "unnest", .into = "")
-import::from("tibble", "as_tibble", .into = "")
+import::from("tidyr", "unnest", "spread", .into = "")
+import::from("tibble", "as_tibble", "tribble", .into = "")
 import::from("here", "here", .into = "")
 import::from("fs", "dir_create", "path", .into = "")
 import::from("fst", "write_fst", "read_fst", .into = "")
-import::from("forcats", "fct_relabel", .into = "")
+import::from("forcats", "fct_relabel", "lvls_revalue", .into = "")
 import::from("cowplot", "save_plot", "theme_cowplot", .into = "")
 import::from("furrr", "future_pmap_dfr", .into = "")
 import::from("magrittr", "%>%", .into = "")
@@ -121,7 +121,7 @@ plan <- drake_plan(
   )), results_plot, base_width = 10, base_height = 7),
   paper = target(
     rmarkdown::render(
-      file_in(!!(here("analysis", "paper", "paper.Rmd"))),
+      knitr_in(!!(here("analysis", "paper", "paper.Rmd"))),
       "github_document"
     )),
   # Representative workflows for two-stream and multiple scatter
@@ -214,7 +214,50 @@ plan <- drake_plan(
     theme_cowplot() +
     theme(axis.title.x = element_blank(),
           axis.text.x = element_text(angle = 90, vjust = 0.5),
-          legend.position = "bottom")
+          legend.position = "bottom"),
+  old_ensemble = file_in(!!file.path("analysis", "data",
+                                     "derived-data", "ed-ensemble-out.fst")) %>%
+    read_fst() %>%
+    as_tibble() %>%
+    spread(variable, value) %>%
+    group_by(run_id, year = floor_date(time, "year")) %>%
+    summarize(
+      npp = udunits2::ud.convert(sum(npp), "kg m-2", "Mg ha-1") * 60 * 30,
+      gpp = udunits2::ud.convert(sum(gpp), "kg m-2", "Mg ha-1") * 60 * 30,
+      lai = max(lai)
+    ) %>%
+    ungroup() %>%
[…]

### 2019-03-27 analysis/scratch/misc.R

@@ -415,3 +415,33 @@ cohort_data %>%
 readd(workflow_df) %>%
   filter(workflow_id == 99000000089) %>%
   dplyr::glimpse()
+
+##################################################
+# Old ED ensemble output
+ed_out <- read_fst("analysis/data/derived-data/ed-ensemble-out.fst") %>%
+  as_tibble()
+
+old_ensemble_summary <- ed_out %>%
+  spread(variable, value) %>%
+  group_by(run_id, year = lubridate::floor_date(time, "year")) %>%
+  summarize(npp = sum(npp), gpp = sum(gpp), lai = max(lai)) %>%
+  ungroup() %>%
+  gather(variable, value, npp, gpp, lai)
+
+old_ensemble_summary %>%
+  mutate(variable = factor(variable, c("gpp", "npp", "lai")) %>%
+           lvls_revalue(c("GPP", "NPP", "LAI"))) %>%
+  ggplot() +
+  aes(x = year, y = value, group = run_id) +
+  geom_line(alpha = 0.2) +
+  facet_wrap(vars(variable), scales = "free") +
+  theme_cowplot() +
+  theme(axis.title.x = element_blank())
+
+old_ensemble_summary %>%
+  filter(year == floor_date(year, "10 years")) %>%
+  ggplot() +
+  aes(x = factor(strftime(year, "%Y")), y = value) +
+  geom_violin() +
+  ## geom_jitter() +
+  facet_wrap(vars(variable), scales = "free")

### 2019-03-27 analysis/paper/paper.Rmd
Order LAI ensemble plots by NPP

@@
   "Max annual leaf area index predictions by plant functional type",
-  "across different parameter combinations."
+  "across different parameter combinations.",
+  "Facets are arranged in order of increasing mean annual NPP",
+  "after 1940."
 )

### 2019-03-27 analysis/drake.R

@@ -10,13 +10,14 @@ expose_imports("fortebaseline")
 import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
              "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
-             .into = "")
+             "arrange", .into = "")
 import::from("tidyr", "unnest", "spread", .into = "")
 import::from("tibble", "as_tibble", "tribble", .into = "")
 import::from("here", "here", .into = "")
 import::from("fs", "dir_create", "path", .into = "")
 import::from("fst", "write_fst", "read_fst", .into = "")
-import::from("forcats", "fct_relabel", "lvls_revalue", .into = "")
+import::from("forcats", "fct_relabel", "lvls_revalue", "fct_inorder",
+             .into = "")
 import::from("cowplot", "save_plot", "theme_cowplot", .into = "")
 import::from("furrr", "future_pmap_dfr", .into = "")
 import::from("magrittr", "%>%", .into = "")
@@ -208,7 +209,7 @@ plan <- drake_plan(
     ggplot() +
     aes(x = year, y = max_lai, color = pft) +
     geom_line() +
-    facet_wrap(~run_id) +
+    facet_wrap(vars(factor(run_id, old_npp_max[["run_id"]]))) +
     labs(y = "Max annual leaf area index",
          color = "PFT") +
     theme_cowplot() +
@@ -228,6 +229,12 @@ plan <- drake_plan(
     ) %>%
     ungroup() %>%
     tidyr::gather(variable, value, npp, gpp, lai),
+  old_npp_max = old_ensemble %>%
+    filter(year > "1959-12-31", variable == "npp") %>%
+    group_by(run_id) %>%
+    summarize(npp = mean(value)) %>%
+    ungroup() %>%
+    arrange(npp),
   hardiman = tribble(
     ~variable, ~low, ~mean, ~hi,
     "LAI", 1.8, 4.14, 6.56,

### 2019-04-15 analysis/paper/paper.Rmd
Add some notes from Longo ED2 GMD papers

@@
 
+[@longo_2019_ed1] provide a description of the most recent version of ED (2.2), and [@longo_2019_ed2] provide an evaluation.
+However, these studies only consider the default configuration of ED2.
+In fact, ED2 provides several alternative structures for some of its submodels.
+
 Light availability is a first order control on photosynthesis, and competition for light is a key process in driving competitive dynamics and community succession.
@@
 
-The Ecosystem Demography Model, version 2.X (ED2) [@medvigy_2009_mechanistic].
-
-Summary of common features (can move into supplement if necessary):
+The Ecosystem Demography Model, version 2.2 (ED2).
+Default configuration is described in [@longo_2019_ed1].
+A brief summary of common features is as follows:
 
@@
     - Similar to approaches used in other models: JULES [@mercado_2007_improving], ORCHIDEE-CAN [@naudts_2015_vertically], CLM(SPA) [@bonan_2014_modeling]
+    - For details, see Supplements S9 and S10 in [@longo_2019_ed1].
   - Multiple scatter model
@@
 
+## Comparison against other studies
+
+In two tropical sites, ED2 did a good job simulating observed top-of-canopy outgoing shortwave radiation, as well as radiation profiles [@longo_2019_ed2].
+Completely closed canopy (i.e. no finite canopy radius) did not lead to an excessively dark understory; in fact, understory was slightly brighter than observed.
+
 ## Future directions: Disturbance response modeling

### 2019-04-17 analysis/paper/paper.Rmd
Add more notes and references to paper
Many updates for reading ensemble output, plus some paper notes

@@
 
+[@rollinson_2017_emergent] -- ED2 paleoclimate
+[@miller_2015_alteration] -- ED2 at Duke FACE
+[@dekauwe_2013_forest] -- ED2 at Duke and Oak Ridge FACE
+
 [@longo_2019_ed1] provide a description of the most recent version of ED (2.2), and [@longo_2019_ed2] provide an evaluation.
@@
 
-Light availability is a first order control on photosynthesis, and competition for light is a key process in driving competitive dynamics and community succession.
+Light availability is a first order control on photosynthesis, and competition for light is a key process in driving competitive dynamics and community succession [@bazzaz_1979_physiological].
 How light is distributed through a canopy is therefore critically important to accurately simulating community dynamics.
@@
     - V~cmax~(0) decreases with shading, and SLA increases with shading, based on @lloyd_2010_optimisation
+    - V~cmax~(0) decline with shade tolerance is supported by classic ecological theory on succession [@bazzaz_1979_physiological]
 
@@
 {TODO: Conceptual diagram of workflow}
+
+## Analysis of results
+
+Photosynthesis / respiration ratio (i.e.
+C balance) in light and shade can be used as an indication of succession -- expect early successional species to have higher ratio in light and late successional species to have higher ratio in shade [@bazzaz_1979_physiological].
 

### 2019-04-17 R/pft.R

@@ -1,0 +1,16 @@
+#' Set PFT from ED integer number
+#'
+#' @param i Vector of PFT integer numbers
+#' @return Factor of PFT names
+#' @author Alexey Shiklomanov
+#' @export
+set_pft <- function(i) {
+  stopifnot(all(i %in% c(6, 9, 10, 11)))
+  pfts <- c(
+    rep(NA_character_, 5),
+    "Pine", NA_character_, NA_character_,
+    "Early hardwood", "Mid hardwood", "Late hardwood"
+  )
+  ipfts <- pfts[i]
+  factor(ipfts, pfts[c(9:11, 6)])
+}

### 2019-04-17 R/read_configxml.R

@@ -1,0 +1,25 @@
+#' Read ED2 parameter file (`config.xml`) into a `data.frame`
+#'
+#' @param path Path to run directory containing XML config file
+#' @param filename Name of config file. Default = `config.xml`
+#' @return Wide `data.frame` containing parameters for each PFT.
+#' @author Alexey Shiklomanov
+#' @importFrom magrittr "%>%"
+#' @export
+read_configxml <- function(path, filename = "config.xml") {
+  config_file <- fs::path(path, filename)
+  stopifnot(fs::file_exists(config_file))
+  xml <- xml2::read_xml(config_file) %>% xml2::as_list()
+  x <- out$leaf_respiration_rate_m2
+  xml[["config"]] %>%
+    purrr::map_dfr(as_tibble) %>%
+    dplyr::mutate_all(purrr::simplify) %>%
+    ## Missing parameters are set to `NULL`, which `simplify` can't
+    ## handle. Convert to NA and then re-simplify only those columns.
+    dplyr::mutate_if(is.list, ~purrr::map_if(.x, is.null, ~NA_character_)) %>%
+    dplyr::mutate_if(is.list, purrr::simplify) %>%
+    dplyr::mutate_all(readr::parse_guess) %>%
+    dplyr::mutate_if(is.character, readr::parse_guess) %>%
+    dplyr::mutate(pft = set_pft(num)) %>%
+    dplyr::select(pft, everything(), -num)
+}

### 2019-04-17 R/workflow_run_matrix.R

@@ -1,0 +1,54 @@
+#' Create a `data.frame` of workflow and run IDs and directories.
+#' Useful as an object over which to iterate.
+#'
+#' @param path Path to local workflows. Default is
+#'   `analysis/data/model_output/workflows`.
+#' @param since Date of earliest workflow to use. Default =
+#'   `1970-01-01`.
+#' @param upto Date of latest workflow to use. Default = `2050-01-01`.
+#' @param type Type of output for paths. Either `out` for output or
+#'   `run` for run configuration.
+#' @return `data.frame` of workflow and run IDs and directory paths
+#' @author Alexey Shiklomanov
+#' @importFrom magrittr %>%
+#' @export
+workflow_run_matrix <- function(path = fs::path("analysis", "data",
+                                                "model_output", "workflows"),
+                                since = "1970-01-01",
+                                upto = "2050-01-01",
+                                type = c("out", "run")) {
+  type <- match.arg(type)
+  if (!inherits(since, "POSIX")) {
+    since <- tryCatch(
+      as.POSIXct(since),
+      error = function(e) stop(glue::glue(
+        "Unable to parse `{since}` as date. ",
+        "Failed with the following error:\n",
+        conditionMessage(e)
+      ))
+    )
+  }
+  stopifnot(fs::dir_exists(path))
+
+  dir_info <- fs::dir_info(path) %>%
+    dplyr::filter(modification_time >= since)
+  stopifnot(nrow(dir_info) > 0)
+
+  dir_info %>%
+    dplyr::select(path) %>%
+    dplyr::transmute(
+      workflow_id = as.numeric(stringr::str_extract(path, "[[:digit:]]+")),
+      run_paths = purrr::map(path, fs::path, type) %>%
+        purrr::map(fs::dir_ls) %>%
+        purrr::map(as.character)
+    ) %>%
+    tidyr::unnest() %>%
+    dplyr::mutate(
+      run_id = fs::path_file(run_paths) %>%
+        stringr::str_extract("[[:digit:]]+") %>%
+        as.numeric()
+    ) %>%
+    dplyr::select(workflow_id, run_id,
+                  path = run_paths) %>%
+    dplyr::filter(!is.na(run_id))
+}

### 2019-04-17 analysis/scratch/misc.R

@@ -581,3 +581,117 @@ result %>%
 ##   group_by(workflow_id, pft, runs) %>%
 ##   summarize_all(mean) %>%
 ##   gather(variable, value, )
+
+##################################################
+# MonetDB
+library(purrr)
+library(magrittr)
+month_file <- "analysis/data/model_output/workflows/PEcAn_99000000112/out/99000000072/analysis-E-1902-06-00-000000-g01.h5"
+mnc <- ncdf4::nc_open(month_file)
+mnc_vardims <- map(mnc[["var"]], "dim") %>% map(~map_chr(.x, "name"))
+mnc_vardims_p <- map_chr(mnc_vardims, paste, collapse = " ") %>%
+  discard(., grepl("phony_dim", names(.))) %>%
+  map_chr(~gsub("phony_dim_", "", .))
+
+writeLines(sprintf("\"%s\",", names(which(mnc_vardims_p == "0"))))
+
+cohorts <- ncdf4::ncvar_get(mnc, "NCOHORTS_GLOBAL")
+slz <- ncdf4::ncvar_get(mnc, "SLZ")
+
+i2f <- function(i, l) {
+  structure(as.integer(i), levels = levels, class = "factor")
+}
+
+setpft <- function(i) {
+  stopifnot(all(i %in% c(6, 9, 10, 11)))
+  pfts <- c(
+    rep(NA_character_, 5),
+    "Pine", NA_character_, NA_character_,
+    "Early hardwood", "Mid hardwood", "Late hardwood"
+  )
+  ipfts <- pfts[i]
+  factor(ipfts, pfts[c(9:11, 6)])
+}
+
+cohort_vars <- tibble::tribble(
+  ~hdf_varname, ~variable, ~unit,
+  "AGB_CO", "Aboveground biomass", "kgC ~ plant^{-1}",
+  "BALIVE", "Live tissue biomass", "kgC ~ plant^{-1}",
+  "BA_CO", "Basal area", "cm^2",
+  "BDEAD", "Stuctural wood biomass", "kgC ~ plant^{-1}",
+  "BLEAF", "Leaf biomass", "kgC ~ plant^{-1}",
+  "BROOT", "Root biomass", "kgC ~ plant^{-1}",
+  "BSAPWOODA", "Aboveground sapwood biomass", "kgC ~ plant^{-1}",
[…]

### 2019-04-17 analysis/scratch/read_10year_ensemble.R

@@ -1,21 +1,91 @@
 library(tidyverse)
 stopifnot(
   requireNamespace("future"),
-  requireNamespace("furrr")
+  requireNamespace("furrr"),
+  requireNamespace("here")
 )
 
-read_month <- function(file) {
-  nc <- ncdf4::nc_open(file)
-  on.exit(ncdf4::nc_close(nc), add = TRUE)
-  v <- nc[["var"]]
-  vnames <- purrr::map_chr(v, "name")
-  vdims <- purrr::map(v, "dim")
-  i <- which(lengths(vdims) == 1 &
-               purrr::map_chr(vdims, list(1, "name")) == "phony_dim_0")
-  purrr::map_dfc(names(i), ncdf4::ncvar_get, nc = nc) %>%
-    setNames(names(i)) %>%
+monthly_mean_vars <- readLines(
+  here::here("analysis", "data",
+             "derived-data", "monthly_mean_vars.txt")
+)
+
+monthly_pft_vars <- readLines(
+  here::here("analysis", "data",
+             "derived-data", "monthly_pft_vars.txt")
+)
+
+monthly_cohort_vars <- readLines(
+  here::here("analysis", "data",
+             "derived-data", "monthly_cohort_vars.txt")
+)
+
+monthly_cb_vars <- readLines(
+  here::here("analysis", "data",
+             "derived-data", "monthly_cb_vars.txt")
+)
+
+
+setpft <- function(i) {
+  stopifnot(all(i %in% c(6, 9, 10, 11)))
+  pfts <- c(
+    rep(NA_character_, 5),
+    "Pine", NA_character_, NA_character_,
+    "Early hardwood", "Mid hardwood", "Late hardwood"
+  )
+  ipfts <- pfts[i]
+  factor(ipfts, pfts[c(9:11, 6)])
+}
+
+read_month <- function(file, pkg = c("RNetCDF", "ncdf4")) {
+  pkg <- match.arg(pkg)
+  stopifnot(requireNamespace(pkg, quietly = TRUE))
+
+  if (pkg == "ncdf4") {
+    nc <- ncdf4::nc_open(file)
+    on.exit(ncdf4::nc_close(nc), add = TRUE)
+    getnc <- purrr::partial(ncdf4::ncvar_get, nc = nc)
+    allnc <- -1
+  } else if (pkg == "RNetCDF") {
+    nc <- RNetCDF::open.nc(file)
[…]

### 2019-04-17 analysis/scratch/read_ed_params.R

@@ -1,51 +1,36 @@
 library(tidyverse)
-library(fst)
+library(fortebaseline)
 
-workflow_id <- 99000000066
-outdir <- file.path("analysis", "data", "model_output", workflow_id)
-runs <- list.files(outdir)
+requireNamespace("fst", quietly = TRUE)
 
-pft_dict <- tribble(
-  ~pft, ~num,
-  "Early hardwood", 9,
-  "Mid hardwood", 10,
-  "Late hardwood", 11,
-  "Pine", 6
-) %>%
-  mutate(pft = fct_inorder(pft))
+runs <- workflow_run_matrix(type = "run")
 
-read_run_configs <- function(run_id) {
-  import::from("magrittr", "%>%", .into = "")
-  import::from("pecanapi", "output_url", .into = "")
-  import::from("XML", "xmlToList", "xmlParse", .into = "")
-  import::from("tidyr", "gather", .into = "")
-  import::from("dplyr", "bind_rows", "mutate_all", "mutate",
-               "left_join", "select", .into = "")
+path <- runs$path[[1]]
 
-  config_file <- output_url(
-    workflow_id,
-    file.path("run", run_id, "config.xml")
-  )
-  raw_xml <- xmlToList(xmlParse(config_file))
+configs <- runs %>%
+  mutate(config = map(path, possibly(read_configxml, NULL))) %>%
+  filter(map_lgl(config, negate(is.null))) %>%
+  unnest()
 
-  bind_rows(raw_xml) %>%
-    mutate_all(as.numeric) %>%
-    mutate(run_id = !!run_id) %>%
-    gather(variable, value, -num, -run_id) %>%
-    left_join(pft_dict, by = "num") %>%
-    select(run_id, pft, variable, value)
-}
+fst::write_fst(configs, "analysis/data/derived-data/ed-ensemble-params.fst")
 
-pft_configs <- map_dfr(runs, read_run_configs)
+meta_vars <- configs %>%
+  select(-workflow_id, -path) %>%
+  group_by(pft) %>%
+  summarize_all(~length(unique(.x))) %>%
+  gather(variable, count, -pft, -run_id) %>%
+  filter(count > 1) %>%
+  distinct(variable) %>%
+  pull()
 
-## pft_configs %>%
-##   group_by(variable, pft) %>%
-##   filter(sd(value) > 0) %>%
-##   ungroup(variable, pft) %>%
-##   ggplot() +
[…]

### 2019-04-18 analysis/paper/paper.Rmd
Add some paper notes about UMBS composition

@@
 - Previously, the site of the FASET experiment [@gough_2013_sustained; @gough_2008_multi]
+- UMBS is 87% upland and 13% wetland [@bergen_2007_observing]. We focus on the upland ecosystem, which has the following characteristics (from @bergen_2007_observing):
+    - 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, 1.2% lowland glacial lake
+    - 60.9% aspen-dominated, 16.6% northern hardwood (/Acer saccharum/, /Acer rubrum/, /Fagus grandifolia/, /Tilia americana/ (basswood), /Betula alleghaniensis/ (yellow birch), /Fraxinus americana/ (white ash), /Tsuga canadensis/), 13.3% upland conifer (/Pinus strobus/, /Pinus resinosa/)
+    - 1.6% northern red oak (/Quercus rubra/)
+    - 60.9% transitioning from aspen to next succession, of which: 55.8% to northern hardwood, 34.7% to upland conifer, 3.8% to northern red oak, 2.6% to northern red oak-red maple, 3.1% to upland conifer-northern hardwood
 

### 2019-04-18 analysis/paper/paper.Rmd
Paper: Add notes about PFT definitions

@@
 
+Plant functional type definitions:
+- Early successional hardwood
+    - _Betula papyrifera_ (white/paper birch)
+    - _Populus grandidentata_ (bigtooth aspen)
+    - ?? _Populus tremuloides_ (trembling aspen)
+    - ?? _Prunus pensylvatica_ (fire cherry)
+- Mid successional hardwood
+    - _Quercus rubra_ (red oak)
+    - _Acer rubrum_ (red maple)
+    - ?? _Betula alleghaniensis_ (yellow birch)
+    - ?? _Tilia americana_ (basswood)
+    - ?? _Fraxinus americana_ (white ash)
+- Late successional hardwood
+    - _Acer saccharum_ (sugar maple)
+    - _Fagus grandifolia_ (American beech)
+    - ?? _Ostyra virginiana_ (hop-hornbeam)
+    - ?? _Acer pensylvaticum_ (striped maple)
+- Pine
+    - _Pinus strobus_ (white pine)
+    - ?? _Pinus resinosa_
+
 {TODO: Conceptual diagram of workflow}

### 2019-04-22 analysis/paper/chunks.Rmd
Update introduction outline. Move old intro to chunks.Rmd

--- 2019-04-17 analysis/paper/paper.Rmd
@@
----
-title: "FoRTE baseline"
-author:
-  - Alexey N. Shiklomanov
-  - Jeff Atkins
-  - Ben Bond-Lamberty
-  - Christopher M. Gough
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: 
-    bookdown::html_document2:
-      fig_caption: yes
-      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
-bibliography: references.bib
-csl: "../templates/ecology.csl" # Insert path for the bib-style
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
----
-
-```{r, setup, echo = FALSE}
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  warning = FALSE,
-  message = FALSE,
-  echo = FALSE,
-  comment = "#>"
-  ## fig.path = "../figures/"
-)
-
-## library(fortebaseline) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
-n <- 0
-mdcap <- function(text) {
-  n <<- n + 1
-  out <- sprintf("**Figure %d**: %s", n, text)
-  if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
-    out
-  } else {
-    invisible(NULL)
-  }
-}
-```
-
-# Outline
-
-- **Introduction**
-    - Why we need models?
-        - Forests are important, for C sequestration, other ecosystem services
-        - Policy decisions related to forests depend on future projections
-        - We need models to make future projections
-    - What kinds of models are out there?
-        - Traditional: "Big leaf" models
-            - Needed to be simple to run on old computers
-            - Don't need as much data
-            - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
-        - Individual-based models
-            - Much more detailed, but harder to run, and need much more data.
-            - Trade-off between complexity and accuracy.
-        - Cohort-based models like ED2 as a middle ground?
-    - How much complexity do we need to accurately model ecological succession?
-        - Lots of existing work on parameteric uncertainty (e.g. Dietze, Raczka, others).
-        - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
-    - What are the processes that drive succession of ecological communities?
-        - Competition for light
-        - Competition for nutrients
-        - Competition for water?
-        - Others?
-        - Brief review of community/ecosystem ecology of the Upper Midwest
-    - Guiding questions:
-        - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
-        - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
-        - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-    - Study overview:
-        - ED with factorial combination of submodels
-        - Ensemble analysis to look at parameter uncertainty
-- **Methods**
-    - Site description
-    - ED2 model description
-        - Common features
-        - Submodel descriptions (in brief; more details in appendices)
-            - Two-stream RTM
-            - Multiple-scatter RTM
-            - Finite canopy radius model
-            - Trait plasticity scheme
-            - N limitation (?)
-    - Data sources
-        - TRY data
-        - Other data? FIA? Remote sensing?
-    - Modeling workflow (PEcAn, etc.)
-    - Data analysis
-        - Structural vs. parametric uncertainty
-        - Comparison against "reality"
-- **Results**
-    - Summary:
-        - LAI time series with uncertainties, faceted by model structure
-    - Structural uncertainty details
-        - Understory light environment and how it affects species matters
-        - The multiple-scatter
-            - Fig: RTM comparison (Radiation profile)
-            - Fig: Crown model comparison
-            - Fig: Trait plasticity?
-    - Uncertainty analysis
-        - Which parameters contribute the most uncertainty?
-        - Structural vs. parameteric uncertainty
-- **Discussion**
-    - Ecological implications of structural uncertainty
-        - Predictions about ecological succession are very sensitive to the understory light environment.
-            - Need a good representation of canopy structure (finite canopy radius)
-            - How you model canopy radiative transfer matters! Small differences in the model can give significant results.
-        - But, trait plasticity can provide an alternative mechanism for success in the understory.
-            - Important modeling implication for all the emerging research on intra-specific trait variability.
-            - This combination of discrete functional types with plastic traits is closer to reality than purely trait-based ("eco-evolutionary?") models or completely discretized models with static traits on the other.
-            - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
-    - Implications of uncertainty analysis for future data collection and model development
-        - Parameter uncertainty is still a _huge_ source of overall model predictive uncertainty. Different parameters can give _qualitatively_ different ecological outcomes.
-        - Looking beyond aggregate variables (e.g. net primary productivity, total LAI) is essential to model evaluation and calibration. It's possible to generate a relatively?
-        - (Expected) Subsurface parameters (e.g. leaf-root ratios) have high sensitivity, but are poorly constrained.
-        - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
-    - Future directions: Predicting disturbance response
-        - FoRTE project -- Forecasting a disturbance.
-            - Compare ecological resilience/memory between model and reality
-
-Journal ideas:
-
-- _Global Change Biology_ -- 8000 word limit may be OK, especially if many methodological details are punted to the appendices.
-- _Ecological Monographs_ -- To do this topic justice may require a lot of words and figures. It may be hard to squeeze all of this information into a regular research paper.
-
 # Introduction
@@
 Constraining these parameters should be a high priority...
-
-# Methods
-
-## Site description
-
-- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (/Acer rubrum/), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
-- Average age -- 95 years (2013)
-- NEP in UMBS tower is 1.58 MG C ha-1 yr-1 (0.80-1.98) (1999-2006), but with landscape variation (REF?)
-- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923
-- Present day composition typical of upper Great Lakes region
-- Previously, the site of the FASET experiment [@gough_2013_sustained; @gough_2008_multi]
-
-## Model description: ED2
-
-The Ecosystem Demography Model, version 2.2 (ED2).
-Default configuration is described in [@longo_2019_ed1].
-A brief summary of common features is as follows:
-
-- Cohort-based scaling approach
-- Photosynthesis / stomatal conductance
-    - I think I need to cite @katul_2003_relationship somewhere in here
-- Allometries
-
-In this study, we run a factorial experiment of ED2 with the following features enabled or disabled:
-
-- Radiative transfer
-  - Two-stream model (ED default)
-    - Same general approach as in CLM 4.5 [@clm45_note], but rather than solving for just two layers (sunlit and shaded), solve for each canopy layer.
-    - Set up the two-stream equations for each layer as a linear system and solve the entire set of equations numerically
-    - Similar to approaches used in other models: JULES [@mercado_2007_improving], ORCHIDEE-CAN [@naudts_2015_vertically], CLM(SPA) [@bonan_2014_modeling]
-    - For details, see Supplements S9 and S10 in [@longo_2019_ed1].
-  - Multiple scatter model
-    - Based on @zhao_2005_multiple
-- Crown model
-  - Infinite (ED2 default)
-    - Marginally taller cohorts completely shade out shorter ones
-  - Finite
-    - Cohorts have a finite crown area, so multiple cohorts have access to the same light level
-    - Based on crown area allometries from @dietze_2008_changing
-- Trait plasticity
-  - None (ED default) -- Trait values are the same at all canopy heights and light levels
-  - Plastic
-    - V~cmax~(0) decreases with shading, and SLA increases with shading, based on @lloyd_2010_optimisation
-    - V~cmax~(0) decline with shade tolerance is supported by classic ecological theory on succession [@bazzaz_1979_physiological]
-
-## Data sources
-
-TRY data [@kattge_2011_try].
-
-Other data from FoRTE? FIA? Remote sensing?
-
-## Modeling workflow
-
-PEcAn parameter sensitivity and uncertainty analysis [@lebauer_2013_facilitating].
-Other PEcAn references: [@dietze_2013_improving].
-
-{TODO: Conceptual diagram of workflow}
-
-## Analysis of results
-
-Photosynthesis / respiration ratio (i.e.
-C balance) in light and shade can be used as an indication of succession -- expect early successional species to have higher ratio in light and late successional species to have higher ratio in shade [@bazzaz_1979_physiological].
-
-# Results
-
-```{r captions}
-lai_10yr_cap <- paste(
-  "ED2 leaf area index (LAI) predictions by plant functional type."
-)
-
-rad_barplot_cap <- paste(
-  "Comparison of radiation profiles under two-stream and multiple scatter",
-  "canopy radiative transfer models."
-)
-```
-
-```{r lai_10yr, fig.cap = lai_10yr_cap, fig.dim = c(10, 7)}
-readd(results_plot)
-```
-
-`r mdcap(lai_10yr_cap)`
-
-Different model configurations produce significantly different ecological trajectories.
-
-## Structural uncertainty
-
-```{r rad_prof, fig.cap = rad_barplot_cap, fig.dim = c(5, 7)}
-readd(rad_lineplot)
-```
-
-`r mdcap(rad_barplot_cap)`
-
-Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation.
-
-## Parameter uncertainty
-
-```{r captions_uncertainty}
-annual_means_cap <- paste(
-  "Total annual gross (GPP) and net (NPP) primary productivity,",
-  "and maximum annual leaf area index (LAI)",
-  "from ED2 parameter ensemble simulations.",
-  "Red lines are mean (solid) and min/max (dashed) values",
-  "from Hardiman et al. (2013)."
-)
-
-annual_lai_workflows_cap <- paste(
-  "Max annual leaf area index predictions by plant functional type",
-  "across different parameter combinations.",
-  "Facets are arranged in order of increasing mean annual NPP",
-  "after 1940."
-)
-```
-
-{TODO: PEcAn variance decomposition analysis, by configuration.}
-
-```{r annual_means, fig.cap = annual_means_cap, fig.dim = c(5, 7)}
-readd(old_ensemble_plot)
-```
-
-`r mdcap(annual_means_cap)`
-
-```{r annual_lai_workflows, fig.cap = annual_lai_workflows_cap, fig.dim = c(10, 8)}
-readd(old_lai_plot)
-```
-
-`r mdcap(annual_lai_workflows_cap)`
-
-Different parameter combinations produce qualitatively different successional outcomes.
-
-# Discussion
-
-## Comparison against other studies
-
-In two tropical sites, ED2 did a good job simulating observed top-of-canopy outgoing shortwave radiation, as well as radiation profiles [@longo_2019_ed2].
-Completely closed canopy (i.e. no finite canopy radius) did not lead to an excessively dark understory; in fact, understory was slightly brighter than observed.
-
-## Future directions: Disturbance response modeling
-
-Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
-Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
-
-# Conclusion
-
-# Acknowledgements
-
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-##### pagebreak
-
-# References 
-<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-<div id="refs"></div>
-
-# Appendix 1: Submodel descriptions
-
-## Radiative transfer: Definitions
-
-- $\omega$ -- Clumping factor
-- $LAI$ -- Leaf area index
-- TAI -- Total area index (leaf and woody area)
-- $LAI_e$ / $TAI_e$ -- "Effective" area indices (accounting for clumping)
-
-## Two-stream RTM
-
-<!-- TODO -->
-
-## Multiple scatter RTM
-
-$$ LAI_e = \omega LAI $$
-
-$$ TAI_e = LAI_e + WAI $$
-
-$$ TAI_{l} = \frac{TAI_e}{CAI} $$
-
-Leaf ($w_l$) and wood ($w_w$) weights:
-
-$$ w_l = \frac{LAI_e}{TAI_e} $$
-$$ w_w = 1 - w_l $$
-
-Projected area, based on coefficients $\phi_1$ and $\phi_2$.
-
-$$ a_{proj} = \phi_1 + \phi_2 \mu $$
-
-$$ \lambda = \frac{a_{proj}}{\mu} $$
-
-##### pagebreak
-
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r}
-# what commit is this file at? You may need to change the path value
-# if your Rmd is not in analysis/paper/
-git2r::repository("../..")
-```

### 2019-04-22 analysis/paper/paper.Rmd
Update introduction outline. Move old intro to chunks.Rmd

@@
 - **Introduction**
-    - Why we need models?
-        - Forests are important, for C sequestration, other ecosystem services
-        - Policy decisions related to forests depend on future projections
-        - We need models to make future projections
-    - What kinds of models are out there?
-        - Traditional: "Big leaf" models
-            - Needed to be simple to run on old computers
-            - Don't need as much data
-            - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
-        - Individual-based models
-            - Much more detailed, but harder to run, and need much more data.
-            - Trade-off between complexity and accuracy.
-        - Cohort-based models like ED2 as a middle ground?
-    - How much complexity do we need to accurately model ecological succession?
-        - Lots of existing work on parameteric uncertainty (e.g. Dietze, Raczka, others).
-        - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
-    - What are the processes that drive succession of ecological communities?
-        - Competition for light
-        - Competition for nutrients
-        - Competition for water?
-        - Others?
-        - Brief review of community/ecosystem ecology of the Upper Midwest
-    - Guiding questions:
-        - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
-        - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
-        - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-    - Study overview:
-        - ED with factorial combination of submodels
-        - Ensemble analysis to look at parameter uncertainty
 - **Methods**
@@
 
-Earth's climate is changing, and its terrestrial ecosystems are likely to be affected (REF: IPCC).
-Many of these ecosystems will be affected through persistent changes in their environment, for instance through elevated atmospheric CO~2~ concentrations and temperature, changing precipitation regimes (REF).
-At the same time, climate change is likely change the frequency and severity of disturbances [@dale_2001_climate], and the impacts of these changes may be more acute.
-
-<!-- Add something positive about each of these? Then, the section about models is more compelling. -->
-In trying to anticipate future changes to terrestrial ecosystems, ecologists have four general techniques.
-One approach is to look to the past to see how ecosystems responded to similar changes (REFs -- e.g. PALEON?).
-This is challenging for two reasons:
-For one, we often have to look very far back to find similar climate changes, and the techniques we have for determining the state of ecosystems in the distant past have limited detail (REF).
-More importantly, there are few (if any) times where the climate has changed as rapidly as it is now, and it is unclear whether slower changes in the past will trigger responses similar to the faster changes happening now (REF).
-A second approach is space-for-time substitution, whereby we use existing climatic gradients through space (e.g. with latitude or elevation) to infer the corresponding changes through time (REF).
-However, many processes that drive spatial variation in ecosystems play out over hundreds or thousands of years (e.g.
-evolution), far longer than policy-relevant timescales of years to decades and far slower than the current pace of climate change (REF).
-A third approach is to conduct experimental manipulations, where individuals are subjected to controlled changes in environmental conditions [@medlyn_2015_using].
-Because they are able to isolate specific changes by controlling for most other factors, experiments are very useful for testing mechanisms.
-Unfortunately, resource constraints limit the spatial and temporal extent of experiments, and emerging evidence from (the relatively few) long-term experiments suggests that short-term results cannot always be extrapolated from longer-term results [@reich_2018_unexpected].
-
-The fourth approach is to use computer models based on existing understanding of ecosystem function.
-Models unconstrained by data are little more than thought experiments with limited value to decision-making.
-However, models provide a useful tool for confronting hypotheses with observations from a variety of sources, including the three described above [@dietze_2013_improving].
-
-Model design involves an invariable trade-off between complexity and predictive uncertainty.
-The more processes are represented in a model (and the more parameters are used to represent those processes), the higher the uncertainty in the "true" value of each parameter, and the higher the uncertainty in model predictions (especially out of sample) (REF -- information theory, e.g.
-AIC?).
-On the other hand, simplistic models often fail to match important properties of the observations.
-A key challenge in model development is to find the balance between these two extremes.
-
-Big leaf models (e.g. CLM) fail to capture vegetation dynamics (REF).
-This is due to their inability to represent demographic processes, which are important (REF).
-An emerging class of models can simulate vegetation demographic processes, including competition for resources within a patch, recruitment, and mortality [@fisher_2017_vegetation].
-Individually-based models are one possibility (REFs -- Shugart, Ensheng Wang, etc.).
-However, they are complicated (REF).
-An alternative are models of intermediate complexity ("middle ground"), which use approximations.
-One such approximation is the "Ecosystem Demography" approach, whereby groups of individuals ("cohorts") of similar species, size, and structure are modeled together, and the statistical properties of these cohorts are modeled rather than the characteristics of each individual [@moorcroft_2001_method].
-
-One realization of this is the Ecosystem Demography (v2) model [@moorcroft_2001_method; @medvigy_2011_predicting; @medvigy_2009_mechanistic].
-This model has been used to simulate (examples...).
-Previous uncertainty analyses with ED:
-[@dietze_2014_quantitative]...
-More recently, [@raczka_2018_what]...
-
-[@rollinson_2017_emergent] -- ED2 paleoclimate
-[@miller_2015_alteration] -- ED2 at Duke FACE
-[@dekauwe_2013_forest] -- ED2 at Duke and Oak Ridge FACE
-
-[@longo_2019_ed1] provide a description of the most recent version of ED (2.2), and [@longo_2019_ed2] provide an evaluation.
-However, these studies only consider the default configuration of ED2.
-In fact, ED2 provides several alternative structures for some of its submodels.
-
-Light availability is a first order control on photosynthesis, and competition for light is a key process in driving competitive dynamics and community succession [@bazzaz_1979_physiological].
-How light is distributed through a canopy is therefore critically important to accurately simulating community dynamics.
-Radiative transfer models are used to simulate light proliferation through a canopy.
-Detailed, 3D representations [@widlowski_2014_abstract; @widlowski_2007_third; @malenovsky_2008_influence] have prohibitively high computational costs for use in dynamic vegetation models.
-Therefore, these models have had to rely on computationally-efficient approximations.
-One common approach is the "two-stream approximation", originally developed for atmospheric radiative transfer [@meador_1980_two] and later adapted for plant canopies [@dickinson_1983_land; @sellers_1985_canopy].
-<!-- TODO: Further discussion? -->
-Although popular, the two-stream approach has some important limitations (REFS, details...).
-An alternative approach is the "multiple scattering" model of @zhao_2005_multiple.
-Accounting for multiple scattering enhances canopy light absorption, particularly in near-infrared wavelengths [@zhao_2005_multiple], therefore reducing the amount of light that makes it into the understory (TODO confirm).
-<!-- TODO: Further discussion -->
-
-Competition is also strongly determined by canopy structure and how it is represented (REF).
-In other words, how much does canopy height matter compared to gappiness and horizontal competition?
-C.f. [@hogan_2018_fast; @dietze_2008_changing].
-
-Trait plasticity with light level also plays a role -- i.e. shaded leaves adapt to lower light levels.
-C.f. [@keenan_2016_global; @lloyd_2010_optimisation; @niinemets_2016_within].
-
-<!--
-TODO: Also discuss the other ED modules.
--->
-
-This manuscript addresses the following research questions:
-(Q1) To what extent can the ED2 model reproduce 100 years of observed succession (or observed _anything_) at UMBS starting from bare ground?
-(Q2) How does the inclusion of various processes affect the accuracy and predictive uncertainty of ED2 predictions?
-(Q3) What drives the predictive uncertainty of the ED2 model at UMBS, and what additional observations are necessary to improve predictions / constrain uncertainty?
-
-Based on previous results [@fisher_2015_taking], we predict ED2 will over-predict dominance of early-successional canopy species (and under-predict understory productivity) under its default configuration.
-Enabling dynamic N cycling and a finite canopy radius model will alleviate these issues, but will significantly increase the parameteric uncertainty due to strong sensitivity to N-cycling and canopy gap fraction parameters, respectively.
-Constraining these parameters should be a high priority...
+- Why we need models?
+    - Forests are important, for C sequestration, other ecosystem services
+    - Policy decisions related to forests depend on future projections
+    - We need models to make future projections
+- What kinds of models are out there?
+    - Traditional: "Big leaf" models
+        - Needed to be simple to run on old computers
+        - Don't need as much data
+        - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
+    - Individual-based models
+        - Much more detailed, but harder to run, and need much more data.
+        - Trade-off between complexity and accuracy.
+- Cohort-based models like ED2 as a middle ground?
+    - Core ED concept [@moorcroft_2001_method]
+    - ED2 developments -- [@medvigy_2009_mechanistic; @medvigy_2011_predicting]
+    - Most recent version -- [@longo_2019_ed1]
+    - Validation/applications of the ED model:
+        - [@rollinson_2017_emergent] -- ED2 paleoclimate
+        - [@miller_2015_alteration] -- ED2 at Duke FACE
+        - [@dekauwe_2013_forest] -- ED2 at Duke and Oak Ridge FACE
+        - [@longo_2019_ed2] -- Model evaluation in Tropics
+- How much complexity do we need to accurately model ecological dynamics?
+    - Lots of existing work on parameteric uncertainty
+        - [@dietze_2014_quantitative]
+        - [@raczka_2018_what]
+    - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
+- What are the processes that drive succession of ecological communities?
+    - Competition for light [@bazzaz_1979_physiological]
+    - Others? Leaning towards omitting because factorial combination becomes too large.
+        - Competition for water?
+        - Competition for nutrients? (ED's N model isn't very good. Alternatives are being developed, but not in ED mainline yet.)
+- Brief review of community/ecosystem ecology of the Upper Midwest
+    - UMBS is a useful site for testing these theories because of its natural history.
+    - Around 1900, mature forest dominated by white (*Pinus strobus*) and red pine (*P. resinosa*), hemlock (*Tsuga canadensis*), and sugar maple (*Acer saccharum*).
+    - But, repeat harvest and burning between 1880 and 1920 [@kilburn_1960_effects] led to replacement by early successional aspen (*Populus grandidentata* and *P.
+tremuloides*), birch (*Betula papyrifera*), and cherry (*Prunus pensylvatica*) [@gates_1930_aspen; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging; @bergen_2007_observing].
+    - Such aspen-dominated secondary forests are sustained by background levels of disturbance, particularly fire [@gates_1930_aspen]
+    - However, due to aggressive fire suppression in the US, these forests are now undergoing succession, primarily to mid/late-successional hardwood (typified by sugar maple and American beech, *Fagus grandifolia*) or upland conifer (typified by red and white pine) [@bergen_2007_observing].
+    - The combination of stand-replacing disturbance ~100 years ago and suppressed disturbance since then makes this an ideal site for testing a vegetation model run from bare ground without having to prescribe disturbance.
+- Guiding questions:
+    - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
+    - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
+    - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+- Study overview:
+    - ED with factorial combination of submodels related to radiative transfer formulation (two-stream vs.
+multiple scatter), horizontal competition (finite canopy radius vs.
+complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
+    - Ensemble analysis to look at parameter uncertainty
 

### 2019-04-22 analysis/paper/paper.Rmd
Standardize italics character

@@
     - UMBS is a useful site for testing these theories because of its natural history.
-    - Around 1900, mature forest dominated by white (*Pinus strobus*) and red pine (*P. resinosa*), hemlock (*Tsuga canadensis*), and sugar maple (*Acer saccharum*).
-    - But, repeat harvest and burning between 1880 and 1920 [@kilburn_1960_effects] led to replacement by early successional aspen (*Populus grandidentata* and *P.
-tremuloides*), birch (*Betula papyrifera*), and cherry (*Prunus pensylvatica*) [@gates_1930_aspen; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging; @bergen_2007_observing].
+    - Around 1900, mature forest dominated by white (_Pinus strobus_) and red pine (_Pinus resinosa_), hemlock (_Tsuga canadensis_), and sugar maple (_Acer saccharum_).
+    - But, repeat harvest and burning between 1880 and 1920 [@kilburn_1960_effects] led to replacement by early successional aspen (_Populus grandidentata_ and _Populus tremuloides_), birch (_Betula papyrifera_), and cherry (_Prunus pensylvatica_) [@gates_1930_aspen; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging; @bergen_2007_observing].
     - Such aspen-dominated secondary forests are sustained by background levels of disturbance, particularly fire [@gates_1930_aspen]
-    - However, due to aggressive fire suppression in the US, these forests are now undergoing succession, primarily to mid/late-successional hardwood (typified by sugar maple and American beech, *Fagus grandifolia*) or upland conifer (typified by red and white pine) [@bergen_2007_observing].
+    - However, due to aggressive fire suppression in the US, these forests are now undergoing succession, primarily to mid/late-successional hardwood (typified by sugar maple and American beech, _Fagus grandifolia_) or upland conifer (typified by red and white pine) [@bergen_2007_observing].
     - The combination of stand-replacing disturbance ~100 years ago and suppressed disturbance since then makes this an ideal site for testing a vegetation model run from bare ground without having to prescribe disturbance.
@@
 - Average age -- 95 years (2013)
-- NEP in UMBS tower is 1.58 MG C ha-1 yr-1 (0.80-1.98) (1999-2006), but with landscape variation (REF?)
-- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923
+- NEP in UMBS tower is 1.58 MG C ha^-1^ yr^-1^ (0.80-1.98) (1999-2006), but with landscape variation (REF?)
+- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923 (REF)
 - Present day composition typical of upper Great Lakes region
@@
     - 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, 1.2% lowland glacial lake
-    - 60.9% aspen-dominated, 16.6% northern hardwood (/Acer saccharum/, /Acer rubrum/, /Fagus grandifolia/, /Tilia americana/ (basswood), /Betula alleghaniensis/ (yellow birch), /Fraxinus americana/ (white ash), /Tsuga canadensis/), 13.3% upland conifer (/Pinus strobus/, /Pinus resinosa/)
-    - 1.6% northern red oak (/Quercus rubra/)
+    - 60.9% aspen-dominated, 16.6% northern hardwood (_Acer saccharum_, _Acer rubrum_, _Fagus grandifolia_, _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_), 13.3% upland conifer (_Pinus strobus_, _Pinus resinosa_)
+    - 1.6% northern red oak (_Quercus rubra_)
     - 60.9% transitioning from aspen to next succession, of which: 55.8% to northern hardwood, 34.7% to upland conifer, 3.8% to northern red oak, 2.6% to northern red oak-red maple, 3.1% to upland conifer-northern hardwood

### 2019-04-22 analysis/paper/paper.Rmd
Minor cleanup of methods outline

@@
 
-- **Introduction**
-- **Methods**
-    - Site description
-    - ED2 model description
-        - Common features
-        - Submodel descriptions (in brief; more details in appendices)
-            - Two-stream RTM
-            - Multiple-scatter RTM
-            - Finite canopy radius model
-            - Trait plasticity scheme
-            - N limitation (?)
-    - Data sources
-        - TRY data
-        - Other data? FIA? Remote sensing?
-    - Modeling workflow (PEcAn, etc.)
-    - Data analysis
-        - Structural vs. parametric uncertainty
-        - Comparison against "reality"
 - **Results**
@@
 
-## Model description: ED2
+## ED2 model description
 
@@
 Default configuration is described in [@longo_2019_ed1].
-A brief summary of common features is as follows:
-
-- Cohort-based scaling approach
-- Photosynthesis / stomatal conductance
-    - I think I need to cite @katul_2003_relationship somewhere in here
-- Allometries
 
@@
 
+BETYdb [@lebauer_2013_facilitating].
+
 TRY data [@kattge_2011_try].
 
-Other data from FoRTE? FIA? Remote sensing?
+Other FoRTE data?
 
@@
 
-{TODO: Conceptual diagram of workflow}
+For each combination of ED submodels, we ran 10 ED ensemble simulations.
 
@@
 
-Photosynthesis / respiration ratio (i.e.
-C balance) in light and shade can be used as an indication of succession -- expect early successional species to have higher ratio in light and late successional species to have higher ratio in shade [@bazzaz_1979_physiological].
+- Structural vs. parameteric uncertainty
+- Comparison against observations
+- Others...
 

### 2019-04-22 analysis/paper/paper.Rmd
Add parameter distributions to paper.Rmd

@@
 Plant functional type definitions:
-- Early successional hardwood
+
+- **Early successional hardwood**
     - _Betula papyrifera_ (white/paper birch)
@@
     - ?? _Prunus pensylvatica_ (fire cherry)
-- Mid successional hardwood
+- **Mid successional hardwood**
     - _Quercus rubra_ (red oak)
@@
     - ?? _Fraxinus americana_ (white ash)
-- Late successional hardwood
+- **Late successional hardwood**
     - _Acer saccharum_ (sugar maple)
@@
     - ?? _Acer pensylvaticum_ (striped maple)
-- Pine
+- **Pine**
     - _Pinus strobus_ (white pine)
     - ?? _Pinus resinosa_
+
+```{r methods_captions}
+param_dist_cap <- paste(
+  "Input parameter distributions."
+)
+```
+
+```{r param_dist, fig.cap = param_dist_cap, fig.dim = c(13, 10)}
+readd(param_dist_gg)
+```
 
@@
 ```{r annual_means, fig.cap = annual_means_cap, fig.dim = c(5, 7)}
-readd(old_ensemble_plot)
+## readd(old_ensemble_plot)
 ```

### 2019-04-22 analysis/drake.R

@@ -10,7 +10,7 @@ expose_imports("fortebaseline")
 import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
              "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
-             "arrange", .into = "")
+             "arrange", "distinct", "one_of", .into = "")
 import::from("tidyr", "unnest", "spread", .into = "")
 import::from("tibble", "as_tibble", "tribble", .into = "")
 import::from("here", "here", .into = "")
@@ -209,62 +209,86 @@ plan <- drake_plan(
     ggplot() +
     aes(x = year, y = max_lai, color = pft) +
     geom_line() +
-    facet_wrap(vars(factor(run_id, old_npp_max[["run_id"]]))) +
+    ## facet_wrap(vars(factor(run_id, old_npp_max[["run_id"]]))) +
+    facet_wrap(vars(factor(run_id))) +
     labs(y = "Max annual leaf area index",
          color = "PFT") +
     theme_cowplot() +
     theme(axis.title.x = element_blank(),
           axis.text.x = element_text(angle = 90, vjust = 0.5),
           legend.position = "bottom"),
-  old_ensemble = file_in(!!file.path("analysis", "data",
-                                     "derived-data", "ed-ensemble-out.fst")) %>%
-    read_fst() %>%
-    as_tibble() %>%
-    spread(variable, value) %>%
-    group_by(run_id, year = floor_date(time, "year")) %>%
-    summarize(
-      npp = udunits2::ud.convert(sum(npp), "kg m-2", "Mg ha-1") * 60 * 30,
-      gpp = udunits2::ud.convert(sum(gpp), "kg m-2", "Mg ha-1") * 60 * 30,
-      lai = max(lai)
-    ) %>%
-    ungroup() %>%
-    tidyr::gather(variable, value, npp, gpp, lai),
-  old_npp_max = old_ensemble %>%
-    filter(year > "1959-12-31", variable == "npp") %>%
-    group_by(run_id) %>%
-    summarize(npp = mean(value)) %>%
-    ungroup() %>%
-    arrange(npp),
-  hardiman = tribble(
-    ~variable, ~low, ~mean, ~hi,
-    "LAI", 1.8, 4.14, 6.56,
[…]

### 2019-04-22 analysis/scratch/plot_params.R

@@ -1,34 +1,0 @@
-library(tidyverse)
-library(fst)
-
-configs <- read_fst(here::here("analysis/data/derived-data/ed-ensemble-params.fst")) %>%
-  as_tibble()
-
-meta_vars <- configs %>%
-  select(-workflow_id, -path) %>%
-  group_by(pft) %>%
-  summarize_all(~length(unique(.x))) %>%
-  gather(variable, count, -pft, -run_id) %>%
-  filter(count > 1) %>%
-  distinct(variable) %>%
-  pull()
-
-meta_vars
-
-params <- configs %>%
-  select(workflow_id, run_id, pft, !!!meta_vars)
-
-params_long <- params %>%
-  gather(param, value, -workflow_id, -run_id, -pft)
-
-ggplot(params_long) +
-  aes(x = pft, y = value, color = pft, fill = pft) +
-  geom_violin(color = "black", alpha = 0.5) +
-  geom_jitter() +
-  facet_wrap(vars(param), scales = "free_y")
-
-##################################################
-trait.mcmc <- ma_result
-trait <- names(trait.mcmc)[[1]]
-
-??fitdistr

### 2019-04-23 analysis/paper/chunks.Rmd
Minor cleanup of methods outline
Move results and discussion outline into sections

@@
+**Journal ideas**
+
+- _Global Change Biology_ -- 8000 word limit may be OK, especially if many methodological details are punted to the appendices.
+- _Ecological Monographs_ -- To do this topic justice may require a lot of words and figures. It may be hard to squeeze all of this information into a regular research paper.
+
 # Introduction
@@
 Constraining these parameters should be a high priority...
+
+# Methods
+
+Photosynthesis / respiration ratio (i.e.
+C balance) in light and shade can be used as an indication of succession -- expect early successional species to have higher ratio in light and late successional species to have higher ratio in shade [@bazzaz_1979_physiological].
+
+# Discussion
+
+## Comparison against other studies
+
+In two tropical sites, ED2 did a good job simulating observed top-of-canopy outgoing shortwave radiation, as well as radiation profiles [@longo_2019_ed2].
+Completely closed canopy (i.e. no finite canopy radius) did not lead to an excessively dark understory; in fact, understory was slightly brighter than observed.
+
+## Future directions: Disturbance response modeling
+
+Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
+Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].

### 2019-04-23 analysis/paper/paper.Rmd
Move results and discussion outline into sections

@@
 ```
-
-# Outline
-
-- **Results**
-    - Summary:
-        - LAI time series with uncertainties, faceted by model structure
-    - Structural uncertainty details
-        - Understory light environment and how it affects species matters
-        - The multiple-scatter
-            - Fig: RTM comparison (Radiation profile)
-            - Fig: Crown model comparison
-            - Fig: Trait plasticity?
-    - Uncertainty analysis
-        - Which parameters contribute the most uncertainty?
-        - Structural vs. parameteric uncertainty
-- **Discussion**
-    - Ecological implications of structural uncertainty
-        - Predictions about ecological succession are very sensitive to the understory light environment.
-            - Need a good representation of canopy structure (finite canopy radius)
-            - How you model canopy radiative transfer matters! Small differences in the model can give significant results.
-        - But, trait plasticity can provide an alternative mechanism for success in the understory.
-            - Important modeling implication for all the emerging research on intra-specific trait variability.
-            - This combination of discrete functional types with plastic traits is closer to reality than purely trait-based ("eco-evolutionary?") models or completely discretized models with static traits on the other.
-            - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
-    - Implications of uncertainty analysis for future data collection and model development
-        - Parameter uncertainty is still a _huge_ source of overall model predictive uncertainty. Different parameters can give _qualitatively_ different ecological outcomes.
-        - Looking beyond aggregate variables (e.g. net primary productivity, total LAI) is essential to model evaluation and calibration. It's possible to generate a relatively?
-        - (Expected) Subsurface parameters (e.g. leaf-root ratios) have high sensitivity, but are poorly constrained.
-        - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
-    - Future directions: Predicting disturbance response
-        - FoRTE project -- Forecasting a disturbance.
-            - Compare ecological resilience/memory between model and reality
-
-Journal ideas:
-
-- _Global Change Biology_ -- 8000 word limit may be OK, especially if many methodological details are punted to the appendices.
-- _Ecological Monographs_ -- To do this topic justice may require a lot of words and figures. It may be hard to squeeze all of this information into a regular research paper.
 
@@
 
+- Summary:
+    - LAI time series with uncertainties, faceted by model structure
+- Structural uncertainty details
+    - Understory light environment and how it affects species matters
+    - The multiple-scatter
+        - Fig: RTM comparison (Radiation profile)
+        - Fig: Crown model comparison
+        - Fig: Trait plasticity?
+- Uncertainty analysis
+    - Which parameters contribute the most uncertainty?
+    - Structural vs. parameteric uncertainty
+
 ```{r captions}
@@
 
-## Comparison against other studies
+- Ecological implications of structural uncertainty
+    - Predictions about ecological succession are very sensitive to the understory light environment.
+        - Need a good representation of canopy structure (finite canopy radius)
+        - How you model canopy radiative transfer matters! Small differences in the model can give significant results.
+    - But, trait plasticity can provide an alternative mechanism for success in the understory.
+        - Important modeling implication for all the emerging research on intra-specific trait variability.
+        - This combination of discrete functional types with plastic traits is closer to reality than purely trait-based ("eco-evolutionary?") models or completely discretized models with static traits on the other.
+        - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
+- Implications of uncertainty analysis for future data collection and model development
+    - Parameter uncertainty is still a _huge_ source of overall model predictive uncertainty. Different parameters can give _qualitatively_ different ecological outcomes.
+    - Looking beyond aggregate variables (e.g. net primary productivity, total LAI) is essential to model evaluation and calibration. It's possible to generate a relatively?
+    - (Expected) Subsurface parameters (e.g. leaf-root ratios) have high sensitivity, but are poorly constrained.
+    - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
+- Future directions: Predicting disturbance response
+    - FoRTE project -- Forecasting a disturbance.
+        - Compare ecological resilience/memory between model and reality
 
-In two tropical sites, ED2 did a good job simulating observed top-of-canopy outgoing shortwave radiation, as well as radiation profiles [@longo_2019_ed2].
-Completely closed canopy (i.e. no finite canopy radius) did not lead to an excessively dark understory; in fact, understory was slightly brighter than observed.
+# Conclusions
 
-## Future directions: Disturbance response modeling
-
-Net ecosystem productivity (NEP) declines immediately after disturbance, but recovers in 10-20 years [@amiro_2010_ecosystem].
-Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
-
-# Conclusion
+<!-- TODO -->
 
 # Acknowledgements
+
+This project funded by NSF grant.
+Cyberinfrastructure provided by Pacific Northwest National Laboratory (PNNL).
+Data from University of Michigan Biological Station (UMBS).
+Data from TRY (TODO: Specific TRY statement).
 

### 2019-04-23 analysis/paper/paper.Rmd
Add meta-analysis result figure

@@
 param_dist_cap <- paste(
-  "Input parameter distributions."
+  "Input parameter distributions from PEcAn trait meta-analysis."
 )
@@
 - Summary:
-    - LAI time series with uncertainties, faceted by model structure
+    - Time series of aggregate variables (GPP, NPP, AGB, Total LAI), with error bars by structure
 - Structural uncertainty details

### 2019-04-23 R/pft.R

@@ -1,3 +1,30 @@
+#' PFT information
+#'
+#' @param col Column to return. If `NULL` (default), return `data.frame`
+#' @return If `is.null(col)`, `data.frame` containing PEcAn PFT name
+#'   (`bety_name`), formatted name (`pft`), short name (`shortname`),
+#'   ED PFT number (`num`), and color (`color`).
+#' @author Alexey Shiklomanov
+#' @return
+#' @export
+pfts <- function(col = NULL) {
+  pfts <- data.frame(
+    bety_name = paste0("umbs.", c(
+      "early_hardwood",
+      "mid_hardwood",
+      "late_hardwood",
+      "northern_pine"
+    )),
+    pft = c("Early hardwood", "Mid hardwood", "Late hardwood", "Pine"),
+    shortname = c("Early", "Mid", "Late", "Pine"),
+    num = c(6, 9, 10, 11),
+    color = ggsci::pal_startrek()(4),
+    stringsAsFactors = FALSE
+  )
+  if (is.null(col)) return(pfts)
+  pfts[[col]]
+}
+
 #' Set PFT from ED integer number
 #'
 #' @param i Vector of PFT integer numbers
@@ -12,5 +39,5 @@ set_pft <- function(i) {
     "Early hardwood", "Mid hardwood", "Late hardwood"
   )
   ipfts <- pfts[i]
-  factor(ipfts, pfts[c(9:11, 6)])
+  factor(ipfts, pfts("pft"))
 }

### 2019-04-23 analysis/drake.R

@@ -1,7 +1,6 @@
 #!/usr/bin/env Rscript
 library(drake)
 library(ggplot2)
-
 ## library(fortebaseline)
 devtools::load_all(here::here(), attach_testthat = FALSE)
 expose_imports("fortebaseline")
@@ -59,6 +58,41 @@ workflows_years <- tryCatch(
 )
 
 plan <- drake_plan(
+  ##############
+  ## Figure 1: Parameter distributions ##
+  ##############
+  ma_posterior = readRDS(file_in(
+    !!here::here("analysis", "data", "derived-data", "meta-analysis.rds")
+  )) %>% tidy_posterior(),
+  param_dist_gg = ma_posterior %>%
+    mutate(pft = factor(pft, pfts("pft"))) %>%
+    unnest(draws) %>%
+    ggplot() +
+    aes(x = pft, y = draws, fill = pft) +
+    geom_violin() +
+    facet_wrap(vars(trait), scales = "free_y") +
+    scale_fill_manual(values = pfts("color")) +
+    labs(x = "PFT", fill = "PFT") +
+    theme_cowplot() +
+    theme(axis.title.y = element_blank(),
+          axis.text.x = element_blank()),
+  ##############
+  ## Figure 2 ##
+  ##############
+  #####################
+  ## Old drake stuff ##
+  #####################
+  run_params = read_fst(file_in(
+    !!here("analysis", "data", "derived-data", "ed-ensemble-params.fst")
+  )) %>% as_tibble(),
+  meta_vars = run_params %>%
+    select(-workflow_id, -path) %>%
+    group_by(pft) %>%
+    summarize_all(~length(unique(.x))) %>%
+    tidyr::gather(variable, count, -pft, -run_id) %>%
+    filter(count > 1) %>%
+    distinct(variable) %>%
+    pull(),
   workflow_df_drake = workflow_df,
   workflows_years_drake = workflows_years,
   lai_raw = target(
@@ -266,29 +300,6 @@ plan <- drake_plan(
   ##   ) +
   ##   theme_cowplot() +
   ##   theme(axis.title = element_blank()),
-  run_params = read_fst(file_in(
-    !!here("analysis", "data", "derived-data", "ed-ensemble-params.fst")
-  )) %>% as_tibble(),
-  meta_vars = run_params %>%
-    select(-workflow_id, -path) %>%
-    group_by(pft) %>%
-    summarize_all(~length(unique(.x))) %>%
[…]

### 2019-04-23 analysis/scratch/meta_analysis.R

@@ -1,36 +1,0 @@
-library(tidyverse)
-library(RPostgres)
-library(fortebaseline)
-
-con <- dbConnect(
-  Postgres(),
-  user = "bety",
-  password = "bety",
-  host = "localhost",
-  port = 7990
-)
-
-pfts <- paste0("umbs.", c(
-  "early_hardwood",
-  "mid_hardwood",
-  "late_hardwood",
-  "northern_pine"
-))
-
-PEcAn.logger::logger.setLevel("WARN")
-outfile <- here::here("analysis", "data", "raw-data", "meta-analysis.rds")
-if (!file.exists(outfile)) {
-  ma_results <- map(pfts, pecan_ma_pft, con = con) %>% setNames(pfts)
-  saveRDS(ma_results, outfile)
-} else {
-  ma_results <- readRDS(outfile)
-}
-
-posterior <- tidy_posterior(ma_results)
-
-trait_draws %>%
-  mutate(pft = factor(pft, pfts) %>% fct_relabel(~gsub("umbs.", "", .x))) %>%
-  ggplot() +
-  aes(x = pft, y = Mean, ymin = lo, ymax = hi) +
-  geom_pointrange() +
-  facet_wrap(vars(trait), scales = "free_y")

### 2019-04-23 analysis/scripts/run_meta_analysis.R

@@ -1,0 +1,18 @@
+library(tidyverse)
+library(RPostgres)
+library(fortebaseline)
+
+con <- dbConnect(
+  Postgres(),
+  user = "bety",
+  password = "bety",
+  host = "localhost",
+  port = 7990
+)
+
+PEcAn.logger::logger.setLevel("WARN")
+outfile <- here::here("analysis", "data", "derived-data", "meta-analysis.rds")
+pfts <- pfts()
+ma_results <- map(pfts[["bety_name"]], pecan_ma_pft, con = con) %>%
+  setNames(ma_results, pfts[["pft"]])
+saveRDS(ma_results, outfile)

### 2019-04-23 analysis/paper/paper.Rmd
Clean up summary time series output

@@
 
-- Summary:
+## Summary
+
 - Time series of aggregate variables (GPP, NPP, AGB, Total LAI), with error bars by structure
-- Structural uncertainty details
+
+```{r summary-captions}
+summary_ts_cap <- paste(
+  "ED2 plot-level predictions by model configuration."
+)
+```
+
+```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 7)}
+readd(summary_ts_plot)
+```
+
+## Structural uncertainty details
+
 - Understory light environment and how it affects species matters
@@
     - Fig: Trait plasticity?
-- Uncertainty analysis
+
+## Uncertainty analysis
+
 - Which parameters contribute the most uncertainty?

### 2019-04-23 R/workflow_info.R

@@ -17,7 +17,7 @@ workflow_structures <- function(...) {
     ) %>%
     dplyr::select(workflow_id, notes) %>%
     tidyr::unnest(notes) %>%
-    transmute(
+    dplyr::transmute(
       workflow_id = workflow_id,
       crown = fct_inorder(if_else(crown_model, "finite", "closed")),
       rtm = fct_inorder(if_else(multiple_scatter, "multi-scatter", "two-stream")),

### 2019-04-23 analysis/drake.R

@@ -9,7 +9,8 @@ expose_imports("fortebaseline")
 import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
              "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
-             "arrange", "distinct", "one_of", .into = "")
+             "arrange", "distinct", "one_of", "everything", "summarize_at",
+             "starts_with", .into = "")
 import::from("tidyr", "unnest", "spread", .into = "")
 import::from("tibble", "as_tibble", "tribble", .into = "")
 import::from("here", "here", .into = "")
@@ -24,7 +25,8 @@ import::from("purrr", "map", "possibly", "discard", "safely",
              "negate", "map_dfr", .into = "")
 import::from("future.callr", "callr", .into = "")
 import::from("future", "plan", "availableCores", .into = "")
-import::from("lubridate", "floor_date", .into = "")
+import::from("lubridate", "floor_date", "month", "%m+%", "period", 
+    .into = "")
 # end imports
 
 created_since <- "2019-03-20"
@@ -58,15 +60,17 @@ workflows_years <- tryCatch(
 )
 
 plan <- drake_plan(
-  ##############
+  #######################################
   ## Figure 1: Parameter distributions ##
-  ##############
+  #######################################
   ma_posterior = readRDS(file_in(
     !!here::here("analysis", "data", "derived-data", "meta-analysis.rds")
   )) %>% tidy_posterior(),
   param_dist_gg = ma_posterior %>%
     mutate(pft = factor(pft, pfts("pft"))) %>%
     unnest(draws) %>%
+    filter(draws < quantile(draws, 0.975),
+           draws > quantile(draws, 0.025)) %>%
     ggplot() +
     aes(x = pft, y = draws, fill = pft) +
     geom_violin() +
@@ -76,9 +80,58 @@ plan <- drake_plan(
     theme_cowplot() +
     theme(axis.title.y = element_blank(),
           axis.text.x = element_blank()),
-  ##############
-  ## Figure 2 ##
-  ##############
+  ########################################
+  ## Figure 2: Aggregate variables plot ##
+  ########################################
+  raw_monthly_output = read_fst(file_in(
[…]

### 2019-04-23 analysis/scratch/plot_monthly.R

@@ -1,14 +1,59 @@
-library(fortebaseline)
+## library(fortebaseline)
 library(tidyverse)
+library(lubridate)
+devtools::load_all(here::here(), attach_testthat = FALSE)
 stopifnot(requireNamespace("fst", quietly = TRUE))
 
+raw_monthly_output %>%
+  select(dplyr::matches("agb"))
+
+raw_monthly_output %>%
+  filter(date == tail(date, 1), runs == unique(runs)[[3]]) %>%
+  select(-(workflow_id:date))
+
+  select(pft, nplant, nplant_py, mmean_lai_py, lai_co) %>%
+  mutate(lai = lai_co * nplant,
+         lai_sum = cumsum(lai))
+
 f <- "analysis/data/derived-data/monthly-ensemble-output.fst"
+
+raw_monthly_output %>%
+  as_tibble() %>%
+  mutate(workflow_id = as.numeric(gsub("PEcAn_", "", workflows))) %>%
+  select(workflow_id, everything()) %>%
+  left_join(workflow_structures(), by = "workflow_id")
 
 result <- fst::read_fst(f) %>%
   as_tibble() %>%
   mutate(workflow_id = as.numeric(gsub("PEcAn_", "", workflows))) %>%
   select(workflow_id, everything()) %>%
   left_join(workflow_structures(), by = "workflow_id")
+
+monthly_subset %>%
+  mutate(
+    year = floor_date(date, "years") %m+% lubridate::period(6, "months"),
+    month = month(date)
+  ) %>%
+  filter(month %in% 7:9) %>%
+  gather(variable, value, GPP:AGB) %>%
+  mutate(variable = fct_inorder(variable)) %>%
+  group_by(variable, year, add = TRUE) %>%
+  summarize(
+    mean = mean(value),
+    lo = quantile(value, 0.2),
+    hi = quantile(value, 0.8)
+  ) %>%
+  filter(!is.na(variable)) %>%
+  mutate(model = interaction(crown, rtm, traits)) %>%
+  ggplot() +
+  aes(x = year, y = mean, ymin = lo, ymax = hi, fill = model, color = model) +
+  geom_ribbon(alpha = 0.5) +
+  facet_wrap(vars(variable), scales = "free_y") +
+  scale_color_brewer(palette = "Paired") +
[…]

### 2019-04-23 analysis/paper/paper.Rmd
Add "bigfacet" code and plot

@@
 
-- Time series of aggregate variables (GPP, NPP, AGB, Total LAI), with error bars by structure
+Time series of aggregate variables (GPP, NPP, AGB, Total LAI), with error bars by structure
 
@@
 )
+
+bigfacet_cap <- paste(
+  "ED2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
+)
 ```
@@
 readd(summary_ts_plot)
+```
+
+`r mdcap(summary_ts_cap)`
+
+Big facet plot (exploratory) of LAI by PFT for all model combinations
+
+```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 12)}
+readd(bigfacet_plot)
 ```

### 2019-04-23 analysis/drake.R

@@ -10,7 +10,7 @@ import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "pull", "case_when", "rename", "ungroup", "group_by", "left_join",
              "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
              "arrange", "distinct", "one_of", "everything", "summarize_at",
-             "starts_with", .into = "")
+             "starts_with", "row_number", "transmute", .into = "")
 import::from("tidyr", "unnest", "spread", .into = "")
 import::from("tibble", "as_tibble", "tribble", .into = "")
 import::from("here", "here", .into = "")
@@ -132,6 +132,33 @@ plan <- drake_plan(
     scale_fill_brewer(palette = "Paired") +
     theme_cowplot() +
     theme(axis.title.y = element_blank()),
+  bigfacet_plot = monthly_means %>%
+    ungroup() %>%
+    select(crown, rtm, traits, date, pft, runs,
+           mmean_lai_py) %>%
+    filter(month(date) == 7) %>%
+    left_join(
+      # Order runs in order of increasing max Pine LAI
+      monthly_means %>%
+        ungroup() %>%
+        select(workflow_id, runs, pft, mmean_lai_py) %>%
+        filter(pft == "Pine") %>%
+        group_by(workflow_id, runs) %>%
+        summarize(pine_lai = max(mmean_lai_py)) %>%
+        mutate(run_i = rank(pine_lai)),
+      by = "runs") %>%
+    mutate(model = interaction(crown, rtm, traits, sep = "\n")) %>%
+    ggplot() +
+    aes(x = date, y = mmean_lai_py, color = pft) +
+    geom_line() +
+    facet_grid(vars(run_i), vars(model), drop = TRUE) +
+    labs(y = "Leaf area index", color = "PFT") +
+    scale_color_manual(values = pfts("color")) +
+    scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
+    theme_cowplot() +
+    theme(strip.text.y = element_blank(),
+          axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+          axis.title.x = element_blank()),
   #####################
   ## Old drake stuff ##
   #####################

### 2019-04-23 analysis/scratch/misc.R

@@ -720,3 +720,36 @@ tnc <- ncdf4::nc_open(t_file)
 # 3 - DBH class
 # 4 - PFT
 
+monthly_means %>%
+  ungroup() %>%
+  select(workflow_id, runs, pft, mmean_lai_py) %>%
+  filter(pft == "Pine") %>%
+  group_by(workflow_id, runs) %>%
+  summarize(pine_lai = max(mmean_lai_py)) %>%
+  mutate(run_i = rank(pine_lai)) %>%
+  arrange(workflow_id, pine_lai) %>%
+  print(n = 30)
+
+##################################################
+## bigfacet_plot = 
+monthly_means %>%
+  ungroup() %>%
+  select(crown, rtm, traits, date, pft, runs,
+         mmean_lai_py) %>%
+  filter(month(date) == 7) %>%
+  group_by(crown, rtm, traits, date, pft) %>%
+  mutate(run_i = row_number()) %>%
+  select(-runs) %>%
+  ungroup() %>%
+  mutate(model = interaction(crown, rtm, traits)) %>%
+  ggplot() +
+  aes(x = date, y = mmean_lai_py, color = pft) +
+  geom_line() +
+  facet_grid(vars(run_i), vars(model)) +
+  labs(y = "Leaf area index", color = "PFT") +
+  scale_color_manual(values = pfts("color")) +
+  theme_cowplot() +
+  theme(axis.title.y = element_blank(),
+        strip.text.y = element_blank())
+  
+  

### 2019-04-23 analysis/paper/paper.Rmd
Use facets (not colors) for summary plot

@@
 
-```{r summary-captions}
+```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 7)}
 summary_ts_cap <- paste(
-  "ED2 plot-level predictions by model configuration."
+  "ED2 plot-level ensemble predictions of",
+  "gross (GPP) and net (NPP) primary productivity,",
+  "total leaf area index (LAI), and aboveground biomass (AGB)",
+  "by model configuration.",
+  "Red lines are mean (dashed) and min/max (dotted) values",
+  "from Hardiman et al. (2013)."
 )
-
-bigfacet_cap <- paste(
-  "ED2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
-)
-```
-
-```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 7)}
 readd(summary_ts_plot)
@@
 ```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 12)}
+bigfacet_cap <- paste(
+  "ED2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
+)
 readd(bigfacet_plot)
 ```
+
+`r mdcap(bigfacet_cap)`
 
@@
 
-```{r captions}
-lai_10yr_cap <- paste(
-  "ED2 leaf area index (LAI) predictions by plant functional type."
-)
-
-rad_barplot_cap <- paste(
-  "Comparison of radiation profiles under two-stream and multiple scatter",
-  "canopy radiative transfer models."
-)
-```
-
-```{r lai_10yr, fig.cap = lai_10yr_cap, fig.dim = c(10, 7)}
-readd(results_plot)
-```
-
-`r mdcap(lai_10yr_cap)`
-
 Different model configurations produce significantly different ecological trajectories.
@@
 ```{r rad_prof, fig.cap = rad_barplot_cap, fig.dim = c(5, 7)}
+rad_barplot_cap <- paste(
+  "Comparison of radiation profiles under two-stream and multiple scatter",
+  "canopy radiative transfer models."
+)
 readd(rad_lineplot)
@@
 
-```{r captions_uncertainty}
-annual_means_cap <- paste(
-  "Total annual gross (GPP) and net (NPP) primary productivity,",
-  "and maximum annual leaf area index (LAI)",
-  "from ED2 parameter ensemble simulations.",
-  "Red lines are mean (solid) and min/max (dashed) values",
-  "from Hardiman et al. (2013)."
-)
-
-annual_lai_workflows_cap <- paste(
-  "Max annual leaf area index predictions by plant functional type",
-  "across different parameter combinations.",
-  "Facets are arranged in order of increasing mean annual NPP",
-  "after 1940."
-)
-```
-
 {TODO: PEcAn variance decomposition analysis, by configuration.}
-
-```{r annual_means, fig.cap = annual_means_cap, fig.dim = c(5, 7)}
-## readd(old_ensemble_plot)
-```
-
-`r mdcap(annual_means_cap)`
-
-```{r annual_lai_workflows, fig.cap = annual_lai_workflows_cap, fig.dim = c(10, 8)}
-readd(old_lai_plot)
-```
-
-`r mdcap(annual_lai_workflows_cap)`
 

### 2019-04-23 analysis/drake.R

@@ -91,6 +91,11 @@ plan <- drake_plan(
     select(workflow_id, everything()) %>%
     select(-workflows) %>%
     left_join(workflow_structures(), by = "workflow_id"),
+  hardiman = tribble(
+    ~variable, ~low, ~mean, ~hi,
+    "LAI", 1.8, 4.14, 6.56,
+    "NPP", 1.68, 3.11, 7.26
+  ) %>% mutate(variable = factor(variable, c("GPP", "NPP", "LAI", "AGB"))),
   monthly_means = raw_monthly_output %>%
     group_by(workflow_id, crown, rtm, traits, date, runs, pft) %>%
     select(starts_with("mmean"), agb_py) %>%
@@ -103,6 +108,8 @@ plan <- drake_plan(
       LAI_pft = mmean_lai_py,
       AGB_pft = agb_py
     ) %>%
+    group_by(crown, rtm, traits, date, runs) %>%
+    # Aggregate PFTs
     summarize(
       GPP = mean(GPP),
       NPP = mean(NPP),
@@ -116,22 +123,34 @@ plan <- drake_plan(
     filter(month %in% 7:9) %>%
     tidyr::gather(variable, value, GPP:AGB) %>%
     mutate(variable = fct_inorder(variable)) %>%
-    group_by(variable, year, add = TRUE) %>%
+    # Aggregate runs
+    group_by(crown, rtm, traits, variable, year) %>%
     summarize(
       mean = mean(value),
       lo = quantile(value, 0.2),
       hi = quantile(value, 0.8)
     ) %>%
     filter(!is.na(variable)) %>%
-    mutate(model = interaction(crown, rtm, traits)) %>%
+    mutate(model = interaction(crown, rtm, traits, sep = "\n")) %>%
     ggplot() +
-    aes(x = year, y = mean, ymin = lo, ymax = hi, fill = model, color = model) +
-    geom_ribbon(alpha = 0.5) +
-    facet_wrap(vars(variable), scales = "free_y") +
+    aes(x = year, y = mean, ymin = lo, ymax = hi) +
+    geom_line() +
+    geom_ribbon(alpha = 0.35) +
+    geom_hline(aes(yintercept = mean), data = hardiman, linetype = "dashed", color = "red") +
+    geom_hline(aes(yintercept = low), data = hardiman, linetype = "dotted", color = "red") +
+    geom_hline(aes(yintercept = hi), data = hardiman, linetype = "dotted", color = "red") +
[…]

### 2019-04-24 analysis/paper/paper.Rmd
Add draft of PEcAn-like parameter sensitivity analysis

@@
 
-Time series of aggregate variables (GPP, NPP, AGB, Total LAI), with error bars by structure
-
 ```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 7)}
@@
 
-Big facet plot (exploratory) of LAI by PFT for all model combinations
+Take-away: Parameter uncertainty is as large as, or even larger, than structural uncertainty.
 
@@
 `r mdcap(bigfacet_cap)`
+
+Take-away: Different parameter combinations produce not just quantitatively different bulk ecosystem properties (e.g.
+NPP, total LAI), but also qualitatively different successional/ecological outcomes.
 
@@
 
+```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(20, 15)}
+uncertainty_analysis_cap <- paste(
+  "PEcAn-liked parameter sensitivity and uncertainty analysis, by model type."
+)
+readd(sensitivity_plot)
+```
+
+`r mdcap(uncertainty_analysis_cap)`
+
 ## Structural uncertainty
@@
 Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation.
-
-## Parameter uncertainty
-
-{TODO: PEcAn variance decomposition analysis, by configuration.}
-
-Different parameter combinations produce qualitatively different successional outcomes.
 

### 2019-04-24 R/sensitivity_analysis.R

@@ -1,0 +1,35 @@
+#' One-at-a-time PEcAn sensitivity analysis
+#' 
+#' NOTE: Fitting all at once results in too few degrees of freedom.
+#' Need to fit one at a time.
+#' @param y Response variable
+#' @param x Dependent variable
+#' @param alpha Fraction of median around which to calculate
+#' derivative for sensitivity. Default = 0.01
+#' @return `tibble` of sensitivity analysis results, with columns
+#' `cv` (Coefficient of variation for `x`), `sensitivity`
+#' (unnormalized sensitivity of `y` to `x` at median), `elasticity`
+#' (normalized `sensitivity`), and `pvar` (prediction variance).
+#' @export
+sensitivity_analysis <- function(y, x, alpha = 0.01) {
+  stopifnot(is.numeric(y), is.numeric(x))
+  fit <- mgcv::gam(y ~ s(x))
+  xmedian <- median(x)
+  ymedian <- median(y)
+  # CV -- Normalized parameter variance
+  cv <- var(x) / xmedian
+  # Sensitivity -- derivative at the median
+  xpm <- xmedian * (1 + alpha * c(1, -1))
+  predpm <- predict(fit, data.frame(x = xpm))
+  sensitivity <- diff(predpm) / diff(xpm)
+  # Elasticity -- normalized sensitivity
+  elasticity <- sensitivity / (ymedian / xmedian)
+  # Prediction variance
+  pvar <- var(predict(fit))
+  tibble::tibble(
+    cv = cv,
+    sensitivity = sensitivity,
+    elasticity = elasticity,
+    pvar = pvar
+  )
+}

### 2019-04-24 analysis/drake.R

@@ -11,7 +11,7 @@ import::from("dplyr", "tbl", "filter", "select", "collect", "mutate",
              "if_else", "group_by_at", "bind_rows", "summarize_all", "summarize",
              "arrange", "distinct", "one_of", "everything", "summarize_at",
              "starts_with", "row_number", "transmute", .into = "")
-import::from("tidyr", "unnest", "spread", .into = "")
+import::from("tidyr", "unnest", "spread", "unite", "separate", .into = "")
 import::from("tibble", "as_tibble", "tribble", .into = "")
 import::from("here", "here", .into = "")
 import::from("fs", "dir_create", "path", .into = "")
@@ -22,49 +22,54 @@ import::from("cowplot", "save_plot", "theme_cowplot", .into = "")
 import::from("furrr", "future_pmap_dfr", .into = "")
 import::from("magrittr", "%>%", .into = "")
 import::from("purrr", "map", "possibly", "discard", "safely",
-             "negate", "map_dfr", .into = "")
+             "negate", "map_dfr", "map_lgl", .into = "")
 import::from("future.callr", "callr", .into = "")
 import::from("future", "plan", "availableCores", .into = "")
 import::from("lubridate", "floor_date", "month", "%m+%", "period", 
     .into = "")
 # end imports
 
-created_since <- "2019-03-20"
+## created_since <- "2019-03-20"
 
-workflow_df <- tryCatch(
-  readd(workflow_df_drake),
-  error = function(e) {
-    tbl(default_connection(), "workflows") %>%
-      filter(start_date == "1902-06-01",
-             end_date == "1912-12-31",
-             created_at > created_since) %>%
-      select(workflow_id = id, start_date, end_date, notes) %>%
-      collect() %>%
-      mutate(configs_df = map(notes, parse_notes),
-             workflow_id = as.numeric(workflow_id)) %>%
-      unnest(configs_df) %>%
-      select(-notes)
-  }
-)
+## workflow_df <- tryCatch(
+##   readd(workflow_df_drake),
+##   error = function(e) {
+##     tbl(default_connection(), "workflows") %>%
+##       filter(start_date == "1902-06-01",
+##              end_date == "1912-12-31",
+##              created_at > created_since) %>%
[…]

### 2019-04-24 analysis/scratch/misc.R

@@ -727,25 +727,37 @@ devtools::load_all(here::here(), attach_testthat = FALSE)
 
 drake::loadd(run_params, monthly_means_site)
 
-run_params_wide <- run_params %>%
-  mutate(pft = lvls_revalue(pft, pfts("shortname"))) %>%
-  gather(param, value, -workflow_id, -run_id, -pft) %>%
-  unite(pft_param, pft, param, sep = "..") %>%
-  spread(pft_param, value)
+d <- sensitivity_inputs %>%
+  ungroup() %>%
+  filter(crown == crown[[1]], rtm == rtm[[1]], traits == traits[[1]])
 
-sensitivity_inputs <- monthly_means_site %>%
-  filter(date > "1910-01-01", month %in% 7:9) %>%
-  group_by(crown, rtm, traits, run_id = as.numeric(runs)) %>%
-  summarize_at(vars(GPP:AGB), mean) %>%
-  left_join(run_params_wide, by = "run_id")
+y <- pull(d, GPP)
+x <- pull(d, Early..SLA)
 
-# NOTE: Fitting all at once results in too few degrees of freedom.
-# Need to fit one at a time.
-sensitivity_analysis <- function(y, x) {
-  stopifnot(is.numeric(y), is.data.frame(x))
-  dat <- cbind(x, y = y)
-  rhs <- sprintf("s(%s)", names(x)) %>% paste(collapse = " + ")
-  form <- sprintf("y ~ %s", rhs)
-  fit <- mgcv::gam(as.formula(form), data = dat)
-  head(dat)
-}
+ssout %>%
+  mutate(model = interaction(crown, rtm, traits, sep = "\n"),
+         elasticity = pmax(elasticity, -200)) %>%
+  separate(xvar, c("pft", "trait"), sep = "\\.\\.") %>%
+  ggplot() +
+  aes(x = pft, y = trait, fill = elasticity) +
+  geom_tile() +
+  facet_grid(vars(yvar), vars(model)) +
+  scale_fill_gradient2() +
+  labs(x = "PFT", y = "Trait") +
+  cowplot::theme_cowplot() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
+        axis.text.y = element_text(size = rel(0.7)))
+
+ssout %>%
+  mutate(model = interaction(crown, rtm, traits)) %>%
+  separate(xvar, c("pft", "trait"), sep = "\\.\\.") %>%
+  ggplot() +
[…]

### 2019-04-24 analysis/paper/paper.Rmd
Remove CV from sensitivity plot, and tweak paper

@@
 
-- Which parameters contribute the most uncertainty?
-- Structural vs. parameteric uncertainty
-
-Different model configurations produce significantly different ecological trajectories.
-
-```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(20, 15)}
+```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(15, 15)}
 uncertainty_analysis_cap <- paste(
-  "PEcAn-liked parameter sensitivity and uncertainty analysis, by model type."
+  "PEcAn-liked parameter sensitivity and uncertainty analysis, by model type.",
+  "Elasticity (a) is the normalized sensitivity of the model to a fixed change in the parameter.",
+  "pvar (b) is the partial variance, which describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
+  "Input parameter distributions are shown in Figure XXX."
 )
@@
 `r mdcap(uncertainty_analysis_cap)`
+
+Take-away: Model parameter uncertainty is sensitive to model configuration.
+For example, turning on the finite canopy radius with the two-stream canopy RTM increases sensitivity across many parameters.
+
+**TODO**: Structural vs. parameteric uncertainty
 

### 2019-04-24 analysis/drake.R

@@ -232,10 +232,10 @@ plan <- drake_plan(
       labs(x = "PFT", y = "Trait") +
       theme_cowplot() +
       theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
-            axis.text.y = element_text(size = rel(0.5))),
-    transform = map(YYY = c(cv, elasticity, pvar),
-                    SCALE = c(scale_fill_gradient(low = "white", high = "red"),
-                              scale_fill_gradient2(),
+            axis.text.y = element_text(size = rel(0.6)),
+            legend.position = "bottom"),
+    transform = map(YYY = c(elasticity, pvar),
+                    SCALE = c(scale_fill_gradient2(),
                               scale_fill_continuous(low = "white", high = "blue")))
   ),
   sensitivity_plot = target(

### 2019-04-29 analysis/paper/paper.Rmd
Add parameter vs. structural uncertainty figure

@@
 
+```{r within-vs-across, fig.cap = within_vs_across_cap, fig.dim = c(6, 6)}
+within_vs_across_cap <- paste(
+  "Comparison of parameteric uncertainty within ensembles (colored bars, colored by model type)",
+  "and \"structural\" uncertainty (variance in ensemble means; black bar) by output variable.",
+  "Output variables are expressed as growing season averages for all years after 1910."
+)
+readd(within_across_plot)
+```
+
+Take-away: Overall, structural uncertainty (variance in ensemble means across structures) is comparable or lower than parameter uncertainty (uncertainty within ensembles for a given structure).
+
+`r mdcap within_vs_across_cap`
+
 ```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(15, 15)}
@@
 For example, turning on the finite canopy radius with the two-stream canopy RTM increases sensitivity across many parameters.
-
-**TODO**: Structural vs. parameteric uncertainty
 

### 2019-04-29 analysis/drake.R

@@ -46,6 +46,11 @@ plan <- drake_plan(
       knitr_in(!!(here("analysis", "paper", "paper.Rmd"))),
       "github_document"
     )),
+  # Common elements
+  model_scale = c(RColorBrewer::brewer.pal(
+    n = nrow(both_uncertainty) - 1,
+    name = "Paired"
+  ), "black") %>% setNames(both_uncertainty[["model"]]),
   #######################################
   ## Figure 1: Parameter distributions ##
   #######################################
@@ -128,8 +133,8 @@ plan <- drake_plan(
     geom_hline(aes(yintercept = low), data = hardiman, linetype = "dotted", color = "red") +
     geom_hline(aes(yintercept = hi), data = hardiman, linetype = "dotted", color = "red") +
     facet_grid(vars(variable), vars(model), scales = "free_y", switch = "y") +
-    scale_color_brewer(palette = "Paired") +
-    scale_fill_brewer(palette = "Paired") +
+    scale_color_manual(values = unname(head(model_scale, -1))) +
+    scale_fill_manual(values = unname(head(model_scale, -1))) +
     theme_cowplot() +
     theme(axis.title.y = element_blank(),
           axis.title.x = element_blank(),
@@ -187,10 +192,11 @@ plan <- drake_plan(
     tidyr::gather(param, value, -workflow_id, -run_id, -pft) %>%
     unite(pft_param, pft, param, sep = "..") %>%
     spread(pft_param, value),
-  sensitivity_inputs = monthly_means_site %>%
+  growing_season_averages = monthly_means_site %>%
     filter(date > "1910-01-01", month %in% 7:9) %>%
     group_by(crown, rtm, traits, run_id = as.numeric(runs)) %>%
-    summarize_at(vars(GPP:AGB), mean) %>%
+    summarize_at(vars(GPP:AGB), mean),
+  sensitivity_inputs = growing_season_averages %>%
     left_join(run_params_wide, by = "run_id"),
   sensitivity_results = sensitivity_inputs %>%
     ungroup() %>%
@@ -223,7 +229,34 @@ plan <- drake_plan(
   sensitivity_plot = target(
     cowplot::plot_grid(sensitivity_plot_piece),
     transform = combine(sensitivity_plot_piece)
-  )
+  ),
+  #######################################
+  ## Within vs. across model structure ##
+  #######################################
+  parameter_uncertainty = growing_season_averages %>%
[…]

### 2019-04-29 analysis/scratch/misc.R

@@ -758,6 +758,3 @@ ssout %>%
   scale_fill_manual(values = pfts("color")) +
   theme(axis.title.x = element_blank(),
         axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
-  
-
-

### 2019-04-29 analysis/paper/paper.Rmd
[paper] Add takeaways for within vs. across figure

@@
 Take-away: Overall, structural uncertainty (variance in ensemble means across structures) is comparable or lower than parameter uncertainty (uncertainty within ensembles for a given structure).
+Parameter uncertainty varies significantly across model structures.
+Enabling trait plasticity increases variance in predictions of AGB.
+Enabling the finite canopy radius increases the variance in LAI predictions.
+GPP and NPP have the highest variance by far for the combination of finite canopy radius and two-stream radiative transfer.
 

### 2019-04-30 analysis/paper/paper.Rmd
[paper] Fix within vs across caption bug
[paper] Minor cleanup of outline

@@
 
-## Structural uncertainty details
-
-- Understory light environment and how it affects species matters
-- The multiple-scatter
-    - Fig: RTM comparison (Radiation profile)
-    - Fig: Crown model comparison
-    - Fig: Trait plasticity?
-
 ## Uncertainty analysis
@@
 
+`r mdcap(within_vs_across_cap)`
+
 Take-away: Overall, structural uncertainty (variance in ensemble means across structures) is comparable or lower than parameter uncertainty (uncertainty within ensembles for a given structure).
@@
 GPP and NPP have the highest variance by far for the combination of finite canopy radius and two-stream radiative transfer.
-
-`r mdcap within_vs_across_cap`
 
@@
 
-## Structural uncertainty
+## Details about model structures
 
@@
 )
-readd(rad_lineplot)
 ```
@@
 
-Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation.
+<!-- Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation. -->
+
+(TODO)
 

### 2019-04-30 analysis/paper/paper.Rmd
[paper] Add parameter table (and remove trailing whitespace)

@@
 
+CRUNCEP met.
+
 Other FoRTE data?
@@
 
+| Parameter_name          | Description                                          | Unit                          |
+|-------------------------|------------------------------------------------------|-------------------------------|
+| `c2n_fineroot`          | C:N ratio in fine roots                              | unitless (mass ratio)         |
+| `c2n_leaf`              | C:N ratio in leaves                                  | unitless (mass ratio)         |
+| `f_labile`              | Fraction of litter that goes to labile (fast) C pool | unitless (0-1)                |
+| `fineroot2leaf` / `q`   | Ratio of fine root to leaf biomass                   | unitless (mass ratio)         |
+| `growth_resp_factor`    | Fraction of daily C gain lost to growth respiration  | unitless (0-1)                |
+| `leaf_width`            | Mean leaf width (for boundary layer conductance)     | m                             |
+| `mort2`                 | C balance ratio at which mortality rapidly increases | unitless (mass ratio)         |
+| `nonlocal_dispersal`    | "Proportion of dispersal that is global"             | unitless (0-1)                |
+| `quantum_efficiency`    | Farquhar model parameter (TODO)                      | mol CO~2~ (mol photons)^-1^   |
+| `r_fract`               | Fraction of C storage to seed reproduction           | unitless                      |
+| `root_respiration_rate` | Root respiration rate at reference soil temperature  | μmol CO~2~ (kg fine root)^-1^ |
+| `root_turnover_rate`    | Rate of fine rootloss                                | year^-1^                      |
+| `SLA`                   | Specific leaf area                                   | m^2^ kg^-1^ C                 |
+| `seedling_mortality`    | Proportion of seed that dies and goes to litter pool | unitless (0-1)                |
+| `stomatal_slope`        | Slope between _A_ and stomatal conductance (Leuning) | unitless                      |
+| `Vcmax`                 | Maximum rate of CO~2~ carboxylation at 15 °C         | μmol m^-2^ s^-1^              |
+| `Vm_low_temp`           | Minimum temperature for photosynthesis               | °C                            |
+| `water_conductance`     | "Water availability factor"                          | m^-2^ a^-1^ (kg C root)^-1^   |
+
+Table: Parameter descriptions
+
 ```{r methods_captions}
@@
 ```
+
+`r mdcap(param_dist_cap)`
 

### 2019-04-30 analysis/paper/paper.Rmd
[paper] Add notes about FoRTE data from Chris
[paper] Re-order and tweak parameter table

--- 2019-04-30 analysis/paper/paper.Rmd
@@
 
+CRUNCEP met.
+
 Other FoRTE data?
+- LAI trends: 1997-2014
+- Aboveground NPP data: 1999-2017
 
@@
 
+| Parameter name (ED name)            | Description                                          | Unit                          |
+|-------------------------------------|------------------------------------------------------|-------------------------------|
+| `c2n_fineroot`                      | C:N ratio in fine roots                              | unitless (mass ratio)         |
+| `c2n_leaf`                          | C:N ratio in leaves                                  | unitless (mass ratio)         |
+| `f_labile`                          | Fraction of litter that goes to labile (fast) C pool | unitless (0-1)                |
+| `fineroot2leaf` (`q`)               | Ratio of fine root to leaf biomass                   | unitless (mass ratio)         |
+| `growth_resp_factor`                | Fraction of daily C gain lost to growth respiration  | unitless (0-1)                |
+| `leaf_turnover_rate`                | Temperature dependent rate of leaf loss              | year^-1^                      |
+| `Rd0`                               | Leaf dark respiration at 15 °C                       | ???                           |
+| `dark_respiration_factor` #         | Ratio of leaf respiration to Vcmax                   | ???                           |
+| `root_respiration_rate` (`_factor`) | Root respiration rate at 15 °C                       | μmol CO~2~ (kg fine root)^-1^ |
+| `root_turnover_rate`                | Temperature dependent rate of fine root loss         | year^-1^                      |
+| `leaf_width`                        | Mean leaf width (for boundary layer conductance)     | m                             |
+| `mort2`                             | C balance ratio at which mortality rapidly increases | unitless (mass ratio)         |
+| `nonlocal_dispersal`                | "Proportion of dispersal that is global"             | unitless (0-1)                |
+| `quantum_efficiency`                | Farquhar model parameter (TODO)                      | mol CO~2~ (mol photons)^-1^   |
+| `r_fract`                           | Fraction of C storage to seed reproduction           | unitless                      |
+| `SLA`                               | Specific leaf area                                   | m^2^ kg^-1^ C                 |
+| `seedling_mortality`                | Proportion of seed that dies and goes to litter pool | unitless (0-1)                |
+| `stomatal_slope`                    | Slope between _A_ and stomatal conductance (Leuning) | unitless                      |
+| `Vcmax` (`Vm0`)                     | Maximum rate of CO~2~ carboxylation at 15 °C         | μmol m^-2^ s^-1^              |
+| `Vm_low_temp`                       | Minimum temperature for photosynthesis               | °C                            |
+| `water_conductance`                 | "Water availability factor"                          | m^-2^ a^-1^ (kg C root)^-1^   |
+
+Table: Parameter descriptions
+
 ```{r methods_captions}

### 2019-04-30 analysis/paper/paper.Rmd
[paper] Add ED2 model wiki link. Minor fixes.

@@
 
-- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (/Acer rubrum/), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
+- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (_Acer rubrum_), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
 - Average age -- 95 years (2013)
@@
 
-Table: Parameter descriptions
+Table: Parameter descriptions. For additional information, see the [ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters).
 

### 2019-04-30 analysis/paper/paper.Rmd
Clean up parameter table

@@
 | Parameter name (ED name)            | Description                                            | Unit                          |
-|-------------------------------------|------------------------------------------------------|-------------------------------|
+|-------------------------------------|--------------------------------------------------------|-------------------------------|
 | `c2n_fineroot`                      | C:N ratio in fine roots                                | unitless (mass ratio)         |
@@
 | `growth_resp_factor`                | Fraction of daily C gain lost to growth respiration    | unitless (0-1)                |
-| `leaf_turnover_rate`                | Temperature dependent rate of leaf loss              | year^-1^                      |
+| `leaf_turnover_rate`                | Temperature dependent rate of leaf loss (conifer only) | year^-1^                      |
 | `Rd0`                               | Leaf dark respiration at 15 °C                         | ???                           |

### 2019-05-02 analysis/paper/paper.Rmd
[paper] Label parameters based on appearance in Raczka paper

--- 2019-04-30 analysis/paper/paper.Rmd
@@
 | Parameter name (ED name)             | Description                                            | Unit                         |
-|-------------------------------------|------------------------------------------------------|-------------------------------|
-| `c2n_fineroot`                      | C:N ratio in fine roots                              | unitless (mass ratio)         |
-| `c2n_leaf`                          | C:N ratio in leaves                                  | unitless (mass ratio)         |
-| `f_labile`                          | Fraction of litter that goes to labile (fast) C pool | unitless (0-1)                |
-| `fineroot2leaf` (`q`)               | Ratio of fine root to leaf biomass                   | unitless (mass ratio)         |
-| `growth_resp_factor`                | Fraction of daily C gain lost to growth respiration  | unitless (0-1)                |
-| `leaf_turnover_rate`                | Temperature dependent rate of leaf loss              | year^-1^                      |
-| `Rd0`                               | Leaf dark respiration at 15 °C                       | ???                           |
-| `dark_respiration_factor` #         | Ratio of leaf respiration to Vcmax                   | ???                           |
-| `root_respiration_rate` (`_factor`) | Root respiration rate at 15 °C                       | μmol CO~2~ (kg fine root)^-1^ |
-| `root_turnover_rate`                | Temperature dependent rate of fine root loss         | year^-1^                      |
-| `leaf_width`                        | Mean leaf width (for boundary layer conductance)     | m                             |
-| `mort2`                             | C balance ratio at which mortality rapidly increases | unitless (mass ratio)         |
-| `nonlocal_dispersal`                | "Proportion of dispersal that is global"             | unitless (0-1)                |
-| `quantum_efficiency`                | Farquhar model parameter (TODO)                      | mol CO~2~ (mol photons)^-1^   |
-| `r_fract`                           | Fraction of C storage to seed reproduction           | unitless                      |
-| `SLA`                               | Specific leaf area                                   | m^2^ kg^-1^ C                 |
-| `seedling_mortality`                | Proportion of seed that dies and goes to litter pool | unitless (0-1)                |
-| `stomatal_slope`                    | Slope between _A_ and stomatal conductance (Leuning) | unitless                      |
-| `Vcmax` (`Vm0`)                     | Maximum rate of CO~2~ carboxylation at 15 °C         | μmol m^-2^ s^-1^              |
-| `Vm_low_temp`                       | Minimum temperature for photosynthesis               | °C                            |
-| `water_conductance`                 | "Water availability factor"                          | m^-2^ a^-1^ (kg C root)^-1^   |
+|--------------------------------------|--------------------------------------------------------|------------------------------|
+| `c2n_fineroot` #                     | C:N ratio in fine roots                                | unitless (mass ratio)        |
+| `c2n_leaf` #                         | C:N ratio in leaves                                    | unitless (mass ratio)        |
+| `f_labile` *                         | Fraction of litter that goes to labile (fast) C pool   | unitless (0-1)               |
+| `fineroot2leaf` (`q`) *              | Ratio of fine root to leaf biomass                     | unitless (mass ratio)        |
+| `growth_resp_factor` *               | Fraction of daily C gain lost to growth respiration    | unitless (0-1)               |
+| `leaf_turnover_rate` #               | Temperature dependent rate of leaf loss (conifer only) | year^-1^                     |
+| `Rd0`  *                             | Leaf dark respiration at 15 °C                         | ???                          |
+| `dark_respiration_factor` *          | Ratio of leaf respiration to Vcmax                     | ???                          |
+| `root_respiration_rate` (`_factor`)* | Root respiration rate at 15 °C                         | μmol CO~2~ (kg fineroot)^-1^ |
+| `root_turnover_rate` *               | Temperature dependent rate of fine root loss           | year^-1^                     |
+| `leaf_width` +                       | Mean leaf width (for boundary layer conductance)       | m                            |
+| `mort2` *                            | C balance ratio at which mortality rapidly increases   | unitless (mass ratio)        |
+| `mort3` +%                           | Density-independent (background) mortality rate        | year^-1^                     |
+| `nonlocal_dispersal` #               | "Proportion of dispersal that is global"               | unitless (0-1)               |
+| `quantum_efficiency` *               | Farquhar model parameter (TODO)                        | mol CO~2~ (mol photons)^-1^  |
+| `r_fract` *                          | Fraction of C storage to seed reproduction             | unitless                     |
+| `SLA` *                              | Specific leaf area                                     | m^2^ kg^-1^ C                |
+| `seedling_mortality` #               | Proportion of seed that dies and goes to litter pool   | unitless (0-1)               |
+| `stomatal_slope` *                   | Slope between _A_ and stomatal conductance (Leuning)   | unitless                     |
+| `Vcmax` (`Vm0`) *                    | Maximum rate of CO~2~ carboxylation at 15 °C           | μmol m^-2^ s^-1^             |
+| `Vm_low_temp` #                      | Minimum temperature for photosynthesis                 | °C                           |
+| `water_conductance` *                | "Water availability factor"                            | m^-2^ a^-1^ (kg C root)^-1^  |
+| `minimum_height` +%                  | Minimum height for plant reproduction                  | m                            |
+| `leaf_reflect/trans_vis/nir` #%      | Leaf reflectance/transmittance in visible/NIR          | unitless (0-1)               |
 
 Table: Parameter descriptions. For additional information, see the [ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters).
+**\*** indicates parameters common to this study and @raczka_2018_what.
+**+** indicates parameters that are in @raczka_2018_what but not this study.
+**#** indicates parameters that are in this study but not @raczka_2018_what.
+**%** indicates parameters that have not yet been included in the runs.
 

### 2019-05-08 analysis/paper/paper.Rmd
Use CSV parameter table in paper and elsewhere

--- 2019-04-30 analysis/paper/paper.Rmd
@@
     - _Populus grandidentata_ (bigtooth aspen)
-    - ?? _Populus tremuloides_ (trembling aspen)
-    - ?? _Prunus pensylvatica_ (fire cherry)
+    - _Populus tremuloides_ (trembling aspen)
 - **Mid successional hardwood**
@@
     - _Acer rubrum_ (red maple)
-    - ?? _Betula alleghaniensis_ (yellow birch)
-    - ?? _Tilia americana_ (basswood)
-    - ?? _Fraxinus americana_ (white ash)
+    - _Acer pensylvaticum_ (striped maple)
 - **Late successional hardwood**
@@
     - _Fagus grandifolia_ (American beech)
-    - ?? _Ostyra virginiana_ (hop-hornbeam)
-    - ?? _Acer pensylvaticum_ (striped maple)
 - **Pine**
     - _Pinus strobus_ (white pine)
-    - ?? _Pinus resinosa_
+- _Omitted candidates_
+    - Early hardwood
+        - _Prunus pensylvatica_ (fire cherry)
+    - Mid hardwood
+        - _Betula alleghaniensis_ (yellow birch)
+        - _Tilia americana_ (basswood)
+        - _Fraxinus americana_ (white ash)
+    - Late hardwood
+        - _Ostyra virginiana_ (hop-hornbeam)
+    - Pine
+        - _Pinus resinosa_ (red pine)
 
-| Parameter name (ED name)            | Description                                          | Unit                          |
-|-------------------------------------|------------------------------------------------------|-------------------------------|
-| `c2n_fineroot`                      | C:N ratio in fine roots                              | unitless (mass ratio)         |
-| `c2n_leaf`                          | C:N ratio in leaves                                  | unitless (mass ratio)         |
-| `f_labile`                          | Fraction of litter that goes to labile (fast) C pool | unitless (0-1)                |
-| `fineroot2leaf` (`q`)               | Ratio of fine root to leaf biomass                   | unitless (mass ratio)         |
-| `growth_resp_factor`                | Fraction of daily C gain lost to growth respiration  | unitless (0-1)                |
-| `leaf_turnover_rate`                | Temperature dependent rate of leaf loss              | year^-1^                      |
-| `Rd0`                               | Leaf dark respiration at 15 °C                       | ???                           |
-| `dark_respiration_factor` #         | Ratio of leaf respiration to Vcmax                   | ???                           |
-| `root_respiration_rate` (`_factor`) | Root respiration rate at 15 °C                       | μmol CO~2~ (kg fine root)^-1^ |
-| `root_turnover_rate`                | Temperature dependent rate of fine root loss         | year^-1^                      |
-| `leaf_width`                        | Mean leaf width (for boundary layer conductance)     | m                             |
-| `mort2`                             | C balance ratio at which mortality rapidly increases | unitless (mass ratio)         |
-| `nonlocal_dispersal`                | "Proportion of dispersal that is global"             | unitless (0-1)                |
-| `quantum_efficiency`                | Farquhar model parameter (TODO)                      | mol CO~2~ (mol photons)^-1^   |
-| `r_fract`                           | Fraction of C storage to seed reproduction           | unitless                      |
-| `SLA`                               | Specific leaf area                                   | m^2^ kg^-1^ C                 |
-| `seedling_mortality`                | Proportion of seed that dies and goes to litter pool | unitless (0-1)                |
-| `stomatal_slope`                    | Slope between _A_ and stomatal conductance (Leuning) | unitless                      |
-| `Vcmax` (`Vm0`)                     | Maximum rate of CO~2~ carboxylation at 15 °C         | μmol m^-2^ s^-1^              |
-| `Vm_low_temp`                       | Minimum temperature for photosynthesis               | °C                            |
-| `water_conductance`                 | "Water availability factor"                          | m^-2^ a^-1^ (kg C root)^-1^   |
-
-Table: Parameter descriptions. For additional information, see the [ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters).
+```{r param-table}
+tbl <- readd(param_table)
+tbl <- subset(tbl, select = -unit_parsed)
+tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
+knitr::kable(tbl, escape = FALSE, caption = paste0(
+  "**Table**: Parameter descriptions. ",
+  "For additional information, see the ",
+  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
+))
+```
 

### 2019-05-08 analysis/drake.R

@@ -23,6 +23,7 @@ library(tidyr)
 library(forcats)
 library(purrr)
 library(lubridate)
+library(readr)
 # Parallel execution
 library(future)
 library(future.callr)
@@ -60,6 +61,8 @@ plan <- drake_plan(
     n = nrow(both_uncertainty) - 1,
     name = "Paired"
   ), "black") %>% setNames(both_uncertainty[["model"]]),
+  param_table = file_in(!!(path(data_dir, "derived-data", "parameter-table.csv"))) %>%
+    read_csv(),
   #######################################
   ## Figure 1: Parameter distributions ##
   #######################################

### 2019-05-29 analysis/paper/paper.Rmd
Add laptop changes

@@
         - [@raczka_2018_what]
+        - [@pandit_2018_optimizing]
     - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
@@
 
+- Comparisons against other studies
+    - Parameter sensitivity
+        - [@pandit_2018_optimizing] -- Most important parameters for shrubs were SLA, Vcmax, stomatal slope, root:leaf ratio, and root turonver rate (slightly less important: leaf turnover rate, growth respiration, cuticular conductance; much less important: water conductance; basically irrelevant: storage turnover, leaf width)
 - Ecological implications of structural uncertainty

### 2019-05-29 analysis/scratch/misc.R

@@ -1,5 +1,5 @@
 # Read CRUNCEP wind inputs. Is the same wind time series looped
-# from 1902 to 1948? 
+# from 1902 to 1948?
 winds <- list()
 files <- glue::glue("http://localhost:7999/thredds/dodsC/dbfiles/CRUNCEP_site_1-33/CRUNCEP.{1903:1910}.nc")
 for (yr in 1903:1951) {
@@ -29,7 +29,7 @@ ggplot(wind_df) +
 # Yes, yes it is.
 ############################################################
 # Is read.output actually loading everything it can from the ED output?
-# No! 
+# No!
 
 filename <- "http://localhost:7999/thredds/dodsC/outputs/PEcAn_99000000032/out/99000000030/analysis-T-1932-00-00-000000-g01.h5"
 hf <- ncdf4::nc_open(filename)
@@ -98,7 +98,7 @@ submit_workflow(settings)
 watch_workflow(workflow_id)
 
 ##################################################
-# Related to reading ED 
+# Related to reading ED
 ##################################################
 ed_summary %>%
   filter(variable == "nee") %>%
@@ -114,7 +114,7 @@ lai <- ncdf4::ncvar_get(nc, "LAI")
 ## nc <- ncdf4::nc_open(run_dap(workflow_id, "1902.nc", run1))
 
 ## catalog_url <- pecanapi:::thredds_fs_url()
-## raw_catalog <- 
+## raw_catalog <-
 
 ## con <- DBI::dbConnect(
 ##   RPostgres::Postgres(),
@@ -673,7 +673,7 @@ cohort_vars <- tibble::tribble(
 )
 
 x <- tibble::tibble(
-  date = 
+  date =
 )
 
   map_dfc(cohort_vars[["hdf_varname"]], ncdf4::ncvar_get, nc = mnc) %>%
@@ -688,7 +688,7 @@ for (i in seq_along(x)) {
 # 1 - Site (patch?)
 # 2 - DBH class
 # 3 - PFT
-# 4 - Last 12 months, plus current month 
+# 4 - Last 12 months, plus current month
 # 5 - Disturbance type
 # 6 - Soil layer?
 # 7 - Height class?
@@ -830,7 +830,7 @@ tbl(con, "traits")
 
 dat <- PEcAn.DB::get.trait.data(
   pfts = list(
-    
+
   )
   pfts = c("umbs.early_hardwood", "umbs.mid_hardwood", "umbs.late_hardwood",
            "umbs.northern_pine"),
[…]

### 2019-06-05 analysis/paper/paper.Rmd
Draft first two paragraphs of intro

@@
 # Introduction
+
+Temperate forests in the Upper Midwest are important.
+{C sequestration}.
+{Other ecosystem services}.
+However, the fate of these forests is highly uncertain.
+The natural ecological succession of these forests has been significantly altered by the interactions of a significantly different and rapidly changing climate, and novel disturbance regimes under which some natural disturbances (e.g.
+fire) are suppressed while others (e.g.
+harvest; pests and pathogens) may be becoming more frequent.
+<!-- TODO refs -->
+Therefore, predictions about the future of these forests based on past analogs may not be reliable.
+
+Instead, more reliable predictions can be obtained by using dynamic vegetation models that explicitly represent process involved in forest growth and mortality.
+Vegetation models fall broadly along the following spectrum of complexity.
+On one side are relatively simple "Big leaf" models---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples?
+-->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
+<!-- TODO refs -->
+These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models, though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
+<!-- TODO refs -->
+On the other end of the spectrum are "individual-based" models (a.k.a.
+"gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer].
+<!-- TODO other refs?
+-->
+Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Others? -->
+Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
+Moreover, these models' predictions of species composition and ecosystem structure are relevant outside of carbon cycle science, particularly for forest management and wildlife conservation.
+<!-- TODO refs -->
+However, these benefits of individual-based models come at the cost of much higher data and computational requirements, which effectively restrict their application to individual, well-studied sites.
+<!-- TODO refs.
+Also, is this actually true?
+-->
 

### 2019-06-05 analysis/paper/paper.Rmd
Add paragraph on ED to introduction

@@
 -->
+
+A key challenge of vegetation modeling research over the last several decades has been the development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
+One such approach is the "ecosystem demography" approach, which uses a system of partial differential equations to simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method].
+<!-- TODO Run this description by Mike?
+-->
+The original "Ecosystem Demography" (ED) model [@moorcroft_2001_method] was enhanced by @medvigy_2009_mechanistic (as ED2) with fully-coupled leaf physiology, canopy radiative transfer, and micrometeorology, and each of these components have been further refined over the last decade [culminating in ED2.2; @longo_2019_ed1].
+Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:
+- paleoecological studies [@rollinson_2017_emergent];
+- free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration];
+- estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
 

### 2019-06-05 analysis/paper/paper.Rmd
Add paragraph introducing uncertainty

@@
 <!-- TODO refs -->
+{UMBS ecological background/natural history?} 
 Therefore, predictions about the future of these forests based on past analogs may not be reliable.
@@
 - estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
+- vegetation-climate feedbacks {Swann, etc} <!-- TODO -->
+
+Vegetation model projections are inherently uncertain, and this uncertainty stems from several sources.
+_Driver_ uncertainty refers to uncertainty in data about processes that are essential to the system in question but are not represented by the model (e.g. weather and climate for vegetation models).
+_Initial condition_ uncertainty arises from the fact that models have to start somewhere, and the exact conditions at the place and time simulations begin are uncertain.
+_Process_ or _structural_ uncertainty arises because vegetation models necessarily only represent a subset of all processes involved in plant biology, and that the process that are included can usually be represented in multiple different ways.
+Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
+The focus of this study is on the interaction between parametric and structural uncertainty; driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
+<!-- TODO This whole paragraph is awkward -->
 

### 2019-06-05 analysis/paper/paper.Rmd
Add ED2 uncertainty text to intro

@@
 _Initial condition_ uncertainty arises from the fact that models have to start somewhere, and the exact conditions at the place and time simulations begin are uncertain.
-_Process_ or _structural_ uncertainty arises because vegetation models necessarily only represent a subset of all processes involved in plant biology, and that the process that are included can usually be represented in multiple different ways.
+_Process_ or _structural_ uncertainty arises because vegetation models necessarily only represent a subset of all processes involved in plant biology, and that the process that are included can usually be represented in multiple different ways (i.e.
+which processes are included, and how they are represented).
 Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
-The focus of this study is on the interaction between parametric and structural uncertainty; driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
 <!-- TODO This whole paragraph is awkward -->
+
+Several recent studies have looked at parameter uncertainty in ED2.
+@dietze_2014_quantitative evaluated which ED2 parameters contributed the most to uncertainties in ED2 simulations of productivity across multiple biomes in North America at 1, 5, and 10 year time scales.
+They found that the parameters contributing the most to model uncertainty across all biomes were those related to growth respiration, mortality, stomatal conductance, and water uptake.
+However, they also found that despite some common patterns, parameter uncertainty varied with biome.
+More recently, a similar analysis evaluated parameter uncertainty over 100 year time scale at the Willow Creek Ameriflux site, and found that the most important parameters to model uncertainty were quantum efficiency of photosynthesis, leaf respiration, and soil-plant water transfer [@raczka_2018_what].
+Another recent study assessed parameter uncertainty in ED2 as part of its overall goal of parameterizing the model for sagebrush [@pandit_2018_optimizing].
+They found that the most important parameters were specific leaf area, the maximum rate of carbon fixation (V~cmax~), slope of stomatal conductance, the fine-root to leaf carbon ratio, and the fine root turnover rate.
+Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large (>20%) impact on predictions of net primary productivity [@viskari_2019_influence].
+The variability in results across these studies demonstrates that parameter contributions to ED2 uncertainty are highly variable, and that additional parameter uncertainty studies are needed.
+
+While parameter uncertainty ED2 has been previously evaluated, less work has been done on structural uncertainty.
+Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization (REF - Medlyn?), high-latitude warming (Rogers and Serbin?), {others...}.
+<!-- TODO -->
+These differences have been attributed to differences in model representations of key processes, including canopy radiative transfer [@fisher_2017_vegetation] and photosynthesis [@rogers_2016_roadmap].
+<!-- TODO More here?
+-->
+However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways.
+In this study, we evaluate structural uncertainty more precisely/directly by selectively toggling and switching individual processes/model components in ED2 while keeping everything else constant.
+
+The focus of this study is on the interaction between parametric and structural uncertainty in the ED2 model.
+Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
 

### 2019-06-05 analysis/paper/paper.Rmd
Add final introduction paragraph

@@
 Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
+Our study is organized around the following guiding questions:
+(1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
+(2) What is the cost of considering these processes, in terms of additional parameteric uncertainty?
+(3) What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+To answer these questions, we ran ED2 ensemble simulationes with a factorial combination of submodels related to radiative transfer formulation (two-stream vs.
+multiple scatter), horizontal competition (finite canopy radius vs.
+complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
+We then used a sensitivity and variance decomposition analysis to evaluate the contribution of parameter uncertainty for each model configuration.
+
+## Old introduction outline
 

### 2019-06-06 analysis/paper/paper.Rmd
Update references
Render ED2 list in intro correctly

@@
 -->
-Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Others? -->
+Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
 Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
@@
 Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:
+
 - paleoecological studies [@rollinson_2017_emergent];

### 2019-06-06 analysis/paper/paper.Rmd
Highlight key ideas in introduction

@@
 
-Temperate forests in the Upper Midwest are important.
+**Temperate forests in the Upper Midwest are important.**
 {C sequestration}.
@@
 {UMBS ecological background/natural history?} 
-Therefore, predictions about the future of these forests based on past analogs may not be reliable.
+**Therefore, predictions about the future of these forests based on past analogs may not be reliable.**
 
-Instead, more reliable predictions can be obtained by using dynamic vegetation models that explicitly represent process involved in forest growth and mortality.
+**Instead, more reliable predictions can be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
 Vegetation models fall broadly along the following spectrum of complexity.
-On one side are relatively simple "Big leaf" models---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples?
+On one side are relatively simple **"Big leaf" models**---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples?
 -->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
@@
 <!-- TODO refs -->
-On the other end of the spectrum are "individual-based" models (a.k.a.
+On the other end of the spectrum are **"individual-based" models** (a.k.a.
 "gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer].
@@
 Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
-Moreover, these models' predictions of species composition and ecosystem structure are relevant outside of carbon cycle science, particularly for forest management and wildlife conservation.
+Moreover, these models' **predictions of species composition and ecosystem structure are relevant outside of carbon cycle science**, particularly for forest management and wildlife conservation.
 <!-- TODO refs -->
-However, these benefits of individual-based models come at the cost of much higher data and computational requirements, which effectively restrict their application to individual, well-studied sites.
+However, these **benefits of individual-based models come at the cost of much higher data and computational requirements**, which effectively restrict their application to individual, well-studied sites.
 <!-- TODO refs.
@@
 
-A key challenge of vegetation modeling research over the last several decades has been the development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
-One such approach is the "ecosystem demography" approach, which uses a system of partial differential equations to simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method].
+A key challenge of vegetation modeling research over the last several decades has been the **development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual** [@purves_2008_predictive].
+One such approach is the **"ecosystem demography" approach**, which uses a system of partial differential equations to simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method].
 <!-- TODO Run this description by Mike?
@@
 
-Vegetation model projections are inherently uncertain, and this uncertainty stems from several sources.
+**Vegetation model projections are inherently uncertain, and this uncertainty stems from several sources.**
 _Driver_ uncertainty refers to uncertainty in data about processes that are essential to the system in question but are not represented by the model (e.g. weather and climate for vegetation models).
@@
 
-Several recent studies have looked at parameter uncertainty in ED2.
+**Several recent studies have looked at parameter uncertainty in ED2.**
 @dietze_2014_quantitative evaluated which ED2 parameters contributed the most to uncertainties in ED2 simulations of productivity across multiple biomes in North America at 1, 5, and 10 year time scales.
@@
 
-While parameter uncertainty ED2 has been previously evaluated, less work has been done on structural uncertainty.
+**While parameter uncertainty ED2 has been previously evaluated, less work has been done on structural uncertainty**.
 Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization (REF - Medlyn?), high-latitude warming (Rogers and Serbin?), {others...}.
@@
 -->
-However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways.
+**However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways.**
 In this study, we evaluate structural uncertainty more precisely/directly by selectively toggling and switching individual processes/model components in ED2 while keeping everything else constant.
 
-The focus of this study is on the interaction between parametric and structural uncertainty in the ED2 model.
+**The focus of this study is on the interaction between parametric and structural uncertainty in the ED2 model.**
 Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).

### 2019-06-07 analysis/paper/paper.Rmd
Move old intro outline to bottom of paper
First draft of methods site description

@@
 
-## Old introduction outline
-
-- Why we need models?
-    - Forests are important, for C sequestration, other ecosystem services
-    - Policy decisions related to forests depend on future projections
-    - We need models to make future projections
-- What kinds of models are out there?
-    - Traditional: "Big leaf" models
-        - Needed to be simple to run on old computers
-        - Don't need as much data
-        - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
-    - Individual-based models
-        - Much more detailed, but harder to run, and need much more data.
-        - Trade-off between complexity and accuracy.
-- Cohort-based models like ED2 as a middle ground?
-    - Core ED concept [@moorcroft_2001_method]
-    - ED2 developments -- [@medvigy_2009_mechanistic; @medvigy_2011_predicting]
-    - Most recent version -- [@longo_2019_ed1]
-    - Validation/applications of the ED model:
-        - [@rollinson_2017_emergent] -- ED2 paleoclimate
-        - [@miller_2015_alteration] -- ED2 at Duke FACE
-        - [@dekauwe_2013_forest] -- ED2 at Duke and Oak Ridge FACE
-        - [@longo_2019_ed2] -- Model evaluation in Tropics
-- How much complexity do we need to accurately model ecological dynamics?
-    - Lots of existing work on parameteric uncertainty
-        - [@dietze_2014_quantitative]
-        - [@raczka_2018_what]
-        - [@pandit_2018_optimizing]
-    - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
-- What are the processes that drive succession of ecological communities?
-    - Competition for light [@bazzaz_1979_physiological]
-    - Others? Leaning towards omitting because factorial combination becomes too large.
-        - Competition for water?
-        - Competition for nutrients? (ED's N model isn't very good. Alternatives are being developed, but not in ED mainline yet.)
-- Brief review of community/ecosystem ecology of the Upper Midwest
-    - UMBS is a useful site for testing these theories because of its natural history.
-    - Around 1900, mature forest dominated by white (_Pinus strobus_) and red pine (_Pinus resinosa_), hemlock (_Tsuga canadensis_), and sugar maple (_Acer saccharum_).
-    - But, repeat harvest and burning between 1880 and 1920 [@kilburn_1960_effects] led to replacement by early successional aspen (_Populus grandidentata_ and _Populus tremuloides_), birch (_Betula papyrifera_), and cherry (_Prunus pensylvatica_) [@gates_1930_aspen; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging; @bergen_2007_observing].
-    - Such aspen-dominated secondary forests are sustained by background levels of disturbance, particularly fire [@gates_1930_aspen]
-    - However, due to aggressive fire suppression in the US, these forests are now undergoing succession, primarily to mid/late-successional hardwood (typified by sugar maple and American beech, _Fagus grandifolia_) or upland conifer (typified by red and white pine) [@bergen_2007_observing].
-    - The combination of stand-replacing disturbance ~100 years ago and suppressed disturbance since then makes this an ideal site for testing a vegetation model run from bare ground without having to prescribe disturbance.
-- Guiding questions:
-    - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
-    - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
-    - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-- Study overview:
-    - ED with factorial combination of submodels related to radiative transfer formulation (two-stream vs.
-multiple scatter), horizontal competition (finite canopy radius vs.
-complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
-    - Ensemble analysis to look at parameter uncertainty
-
 # Methods
@@
 
-- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (_Acer rubrum_), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
-- Average age -- 95 years (2013)
-- NEP in UMBS tower is 1.58 MG C ha^-1^ yr^-1^ (0.80-1.98) (1999-2006), but with landscape variation (REF?)
-- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923 (REF)
-- Present day composition typical of upper Great Lakes region
-- Previously, the site of the FASET experiment [@gough_2013_sustained; @gough_2008_multi]
-- UMBS is 87% upland and 13% wetland [@bergen_2007_observing]. We focus on the upland ecosystem, which has the following characteristics (from @bergen_2007_observing):
-    - 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, 1.2% lowland glacial lake
-    - 60.9% aspen-dominated, 16.6% northern hardwood (_Acer saccharum_, _Acer rubrum_, _Fagus grandifolia_, _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_), 13.3% upland conifer (_Pinus strobus_, _Pinus resinosa_)
-    - 1.6% northern red oak (_Quercus rubra_)
-    - 60.9% transitioning from aspen to next succession, of which: 55.8% to northern hardwood, 34.7% to upland conifer, 3.8% to northern red oak, 2.6% to northern red oak-red maple, 3.1% to upland conifer-northern hardwood
+We performed this study at the University of Michigan Biological Station (Ameriflux site, "UMBS-dist") ({coords}), <!-- TODO --> located in Northern Lower Michigan, USA.
+The area surrounding the research station is 87% well-drained upland forest and 13% wetland [@bergen_2007_observing]; the focus of this study is on the former.
+The landscape geography of the UMBS upland forest is 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, and 1.2% lowland glacial lake [@bergen_2007_observing].
+Most of the UMBS upland forest canopy is dominated by temperate deciduous early-successional species, most importantly _Populus grandidentata_ (bigtooth aspen) and, to a lesser extent, _Betula papyrifera_ (paper birch), with _Acer saccharum_ (sugar maple), _Acer rubrum_ (red maple), _Fagus grandifolia_ (American beech), _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_ (eastern hemlock), _Quercus rubra_ (northern red oak), _Pinus strobus_ (white pine), and _Pinus resinosa_ (red pine) existing in various fractions in the understory (and, in patches, in the canopy).
+This composition is a legacy of the site's disturbance history:
+The site was intensively logged in the late 1800s and early 1900s, and experienced regularly recurring fires until the mid-1920s, at which point a regime of active fire suppression started that has persisted to the present day.
+<!-- TODO ref -->
+As a result, the average stand age in 2013 was 95 years. <!-- TODO ref -->
+The majority of the forest that is aspen-dominated is undergoing succession to "northern hardwood" (maple, beech, basswood, birch, ash, hemlock), "upland conifer" (red and white pine), or "northern red-oak" ecotypes [@bergen_2007_observing].
 
-## ED2 model description
+## Ecosystem Demography Model (version 2.2)
 
@@
 
+# Old outline
+
+## Introduction
+
+- Why we need models?
+    - Forests are important, for C sequestration, other ecosystem services
+    - Policy decisions related to forests depend on future projections
+    - We need models to make future projections
+- What kinds of models are out there?
+    - Traditional: "Big leaf" models
+        - Needed to be simple to run on old computers
+        - Don't need as much data
+        - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
+    - Individual-based models
+        - Much more detailed, but harder to run, and need much more data.
+        - Trade-off between complexity and accuracy.
+- Cohort-based models like ED2 as a middle ground?
+    - Core ED concept [@moorcroft_2001_method]
+    - ED2 developments -- [@medvigy_2009_mechanistic; @medvigy_2011_predicting]
+    - Most recent version -- [@longo_2019_ed1]
+    - Validation/applications of the ED model:
+        - [@rollinson_2017_emergent] -- ED2 paleoclimate
+        - [@miller_2015_alteration] -- ED2 at Duke FACE
+        - [@dekauwe_2013_forest] -- ED2 at Duke and Oak Ridge FACE
+        - [@longo_2019_ed2] -- Model evaluation in Tropics
+- How much complexity do we need to accurately model ecological dynamics?
+    - Lots of existing work on parameteric uncertainty
+        - [@dietze_2014_quantitative]
+        - [@raczka_2018_what]
+        - [@pandit_2018_optimizing]
+    - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
+- What are the processes that drive succession of ecological communities?
+    - Competition for light [@bazzaz_1979_physiological]
+    - Others? Leaning towards omitting because factorial combination becomes too large.
+        - Competition for water?
+        - Competition for nutrients? (ED's N model isn't very good. Alternatives are being developed, but not in ED mainline yet.)
+- Brief review of community/ecosystem ecology of the Upper Midwest
+    - UMBS is a useful site for testing these theories because of its natural history.
+    - Around 1900, mature forest dominated by white (_Pinus strobus_) and red pine (_Pinus resinosa_), hemlock (_Tsuga canadensis_), and sugar maple (_Acer saccharum_).
+    - But, repeat harvest and burning between 1880 and 1920 [@kilburn_1960_effects] led to replacement by early successional aspen (_Populus grandidentata_ and _Populus tremuloides_), birch (_Betula papyrifera_), and cherry (_Prunus pensylvatica_) [@gates_1930_aspen; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging; @bergen_2007_observing].
+    - Such aspen-dominated secondary forests are sustained by background levels of disturbance, particularly fire [@gates_1930_aspen]
+    - However, due to aggressive fire suppression in the US, these forests are now undergoing succession, primarily to mid/late-successional hardwood (typified by sugar maple and American beech, _Fagus grandifolia_) or upland conifer (typified by red and white pine) [@bergen_2007_observing].
+    - The combination of stand-replacing disturbance ~100 years ago and suppressed disturbance since then makes this an ideal site for testing a vegetation model run from bare ground without having to prescribe disturbance.
+- Guiding questions:
+    - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
+    - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
+    - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+- Study overview:
+    - ED with factorial combination of submodels related to radiative transfer formulation (two-stream vs.
+multiple scatter), horizontal competition (finite canopy radius vs.
+complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
+    - Ensemble analysis to look at parameter uncertainty
+
 $$ \lambda = \frac{a_{proj}}{\mu} $$
+
+## Methods
+
+### Site description
+
+- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (_Acer rubrum_), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
+- Average age -- 95 years (2013)
+- NEP in UMBS tower is 1.58 MG C ha^-1^ yr^-1^ (0.80-1.98) (1999-2006), but with landscape variation (REF?)
+- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923 (REF)
+- Present day composition typical of upper Great Lakes region
+- Previously, the site of the FASET experiment [@gough_2013_sustained; @gough_2008_multi]
+- UMBS is 87% upland and 13% wetland [@bergen_2007_observing]. We focus on the upland ecosystem, which has the following characteristics (from @bergen_2007_observing):
+    - 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, 1.2% lowland glacial lake
+    - 60.9% aspen-dominated, 16.6% northern hardwood (_Acer saccharum_, _Acer rubrum_, _Fagus grandifolia_, _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_), 13.3% upland conifer (_Pinus strobus_, _Pinus resinosa_)
+    - 1.6% northern red oak (_Quercus rubra_)
+    - 60.9% transitioning from aspen to next succession, of which: 55.8% to northern hardwood, 34.7% to upland conifer, 3.8% to northern red oak, 2.6% to northern red oak-red maple, 3.1% to upland conifer-northern hardwood
 

### 2019-06-07 analysis/paper/paper.Rmd
First pass at ED2 model overview

@@
 
-The Ecosystem Demography Model, version 2.2 (ED2).
-Default configuration is described in [@longo_2019_ed1].
+For this study, we used the Ecosystem Demography Model, version 2.2 (ED2).
+Below, we provide only a high-level description of key model features and focus in detail only on submodels and configurations specific to this study.
+A full description of the default configuration of the model is provided by @longo_2019_ed1.
+
+ED2 represents both spatial and temporal variability hierarchically.
+The largest spatial unit is the "polygon", within which all elements share the same meteorological forcings.
+Polygons are divided into multiple "sites", each site with its own set of abiotic characteristics such as soil texture and topography.
+Sites are divided into discrete "patches" according to time since last disturbance.
+Finally, each patch is composed of an arbitrary (and dynamic) number of "cohorts", representing the mean behavior of groups of plant individuals of similar size and composition.
+Temporal variability in ED2 is similarly hierarchical.
+Thermodynamics, including energy, water, and eddy (e.g. CO~2~) fluxes, are resolved on the order of seconds.
+Biophysics, including photosynthesis, respiration, and radiative transfer, and the evaluation of energy, water, and CO2 budgets, are resolved on the order of minutes.
+Phenology and plant carbon balance are evaluated daily.
+Cohort dynamics, including structural growth, mortality, and reproduction, are evaluated monthly.
+Finally, patch dynamics (e.g. annual disturbances) are evaluated annually.
+In this study, we consider only a single patch (and, therefore, a single site and polygon) starting from a "near bare ground" condition.
+
+<!-- TODO -->
+{Brief summary of thermodynamics?}
+{Brief summary of water cycle?}
+
+ED2 represents the carbon flux for each cohort as the net sum of contributions from turbulent mixing, leaf photosynthesis and respiration, (fine) root respiration, growth and maintenance respiration, and heterotrophic respiration.
+In addition, each cohort also has a virtual carbon balance that links fast processes like photosynthesis and respiration to slow processes like allocation (to growth, storage, and reproduction) and mortality.
+
+The official ED2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained  https://github.com/ashiklom/ED2/tree/b048950971e91699b78fc566df133caf977fda32.
 

### 2019-06-07 analysis/paper/paper.Rmd
Add very rough first pass at descriptions of parameters

@@
 
+## ED2 submodels
+
 In this study, we run a factorial experiment of ED2 with the following features enabled or disabled:
-
 - Radiative transfer
@@
 
+## ED2 parameters
+
+The full list of parameters in ED2 is large, with over 100 parameters per plant functional type.
+A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
+Instead, we selected a subset of parameters that were identified as important by previous ED2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
+
+Three parameters are related to leaf-level physiology:
+Following the enzyme-kinetic model of @farquhar_1980_biochemical, the rate of photosynthesis is the minimum of light-limited and enzyme-limited reactions.
+The former are controlled by the _quantum efficiency_ parameter---maximum efficiency with which absorbed photosynthetically active radiation is converted to CO~2~.
+The latter are controlled by _V~cmax~_, the maximum rate of carbon fixation by Rubisco.
+The water demand of photosynthesis is controlled by the _stomatal slope_, the sensitivity of stomatal conductance of CO~2~ as a function of CO~2~ concentration and humidity at the leaf surface [@leuning_1995_critical; @ball_1987_model].
+Three more parameters correspond to the respiration rates of leaves, roots, and "growth maintenance".
+Two more parameters control carbon allocation:
+One is the ratio of fine root to leaf biomass (`fineroot2leaf`, or `q`), and another is the ratio of "storage" carbon allocated to reproduction (`r_fract`).
+Three parameters control various aspects of adult tree mortality, namely the density-independent mortality rate (`mort3`), and the time scale (`mort1`) and critical carbon balance (`mort2`) of mortality from carbon starvation.
+An additional parameter controls seedling mortality.
+_Specific leaf area_ (`SLA`) is used to convert leaf biomass to leaf area index, which in turn is used in a variety of calculations related to canopy radiative transfer and micrometeorology.
+Two additional parameters describe canopy structure in ways critical to the canopy radiative transfer:
+Canopy _clumping factor_ describes how evenly leaf area is distributed in horizontal space (1 being perfectly evenly; 0 being a "black hole" where all leaves are concentrated in a single point); and
+leaf _orientation factor_ describes the average distribution of leaf angles (-1 being perfectly vertical, 1 being perfectly horizontal, and 0 being random).
+Four parameters control leaf optical properties, namely the fractions of light reflected or transmitted in visible and near-infrared wavelengths (`leaf_(reflect|trans)_(vis|nir)`).
+Three other parameters:
+
+- Fraction of litter carbon that enters the labile (fast) soil carbon pool (`f_labile`)
+- Minimum plant height for reproduction (`minimum_height`)
+- Soil-plant water conductance, or "water availability factor" (`water_conductance`)
+
 ## Data sources
@@
 - Aboveground NPP data: 1999-2017
+
+Priors on leaf optical properties from Shiklomanov dissertation (chapter 3).
 

### 2019-06-07 analysis/paper/paper.Rmd
First draft of text on submodels

@@
 
-In this study, we run a factorial experiment of ED2 with the following features enabled or disabled:
-- Radiative transfer
+In this study, we ran a factorial combination of the following ED2 configurations: (1) two-stream vs.
+multiple-scatter canopy radiative transfer models; (2) infinite vs.
+finite crown area (a.k.a.
+complete vs.
+partial shading); and (3) static vs.
+light-plastic traits.
+
+Both canopy radiative transfer models in ED2 resolve the full vertical radiation profile within a patch as a function of canopy structure (leaf and wood area indices, crown area, leaf angle distribution) and incident solar radiation.
+The two models differ in the mathematical approximations of the underlying differential radiative transfer equations.
+
 - Two-stream model (ED default)
@@
     - Based on @zhao_2005_multiple
-- Crown model
-  - Infinite (ED2 default)
-    - Marginally taller cohorts completely shade out shorter ones
-  - Finite
-    - Cohorts have a finite crown area, so multiple cohorts have access to the same light level
-    - Based on crown area allometries from @dietze_2008_changing
-- Trait plasticity
-  - None (ED default) -- Trait values are the same at all canopy heights and light levels
-  - Plastic
-    - V~cmax~(0) decreases with shading, and SLA increases with shading, based on @lloyd_2010_optimisation
-    - V~cmax~(0) decline with shade tolerance is supported by classic ecological theory on succession [@bazzaz_1979_physiological]
+
+The crown area submodel in ED2 determines the nature of competition for light between cohorts.
+In the default configuration ("infinite crown area", or "complete shading"), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
+This means that taller cohorts always receive more incoming radiation than shorter cohorts, even when the height difference is small.
+Put another way, this means that there is no horizontal competition, only vertical.
+This has been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous canopies [@fisher_2015_taking].
+In the alternate configuration ("finite crown area", or "partial shading"), canopies take up only a fraction of the available horizontal area, meaning that multiple cohorts of similar height can receive the same level of light.
+The horizontal area of crowns is determined by allometric equations from @dietze_2008_changing.
+
+The third submodel we evaluated was trait plasticity.
+In the default configuration, all cohorts of a given plant functional type will have the same parameters, regardless of environmental conditions.
+This ignores the globally-documented intraspecific trait variability as a function of light level [@niinemets_2010_review; @keenan_2016_global].
+<!-- TODO Also, classical ecological succession theory bazzaz_1979_physiological?
+-->
+In the alternate configuration, as light level decreases (trees become more shaded), specific leaf area increases and V~cmax~ decreases, following empirical relationships from the tropics [@lloyd_2010_optimisation].
 

### 2019-06-07 analysis/paper/paper.Rmd
More details on modeling workflow, other methods

@@
 
-We performed this study at the University of Michigan Biological Station (Ameriflux site, "UMBS-dist") ({coords}), <!-- TODO --> located in Northern Lower Michigan, USA.
+We performed this study at the University of Michigan Biological Station (Ameriflux site US-UMd, 45.5625$^{\circ}$, -84.6975$^{\circ}$), located in Northern Lower Michigan, USA.
 The area surrounding the research station is 87% well-drained upland forest and 13% wetland [@bergen_2007_observing]; the focus of this study is on the former.
@@
 
-## Data sources
+By default, ED2 supports 17 different plant functional types, which divide plant species according to photosynthetic pathway (C3 vs.
+C4), growth form (grass vs.
+tree), leaf phenology habit (deciduous vs.
+evergreen), biome (e.g.
+temperate vs.
+tropical), and successional status (e.g.
+early, mid, late).
+However, we limited our simulations to the four plant functional types that have any appreciable presence at UMBS: Early, mid, and late temperate deciduous trees and pines.
+The species comprising these plant functional types are shown in Table X.
 
-BETYdb [@lebauer_2013_facilitating].
+Table: Plant functional type-species mappings
 
-TRY data [@kattge_2011_try].
+| Plant functional type | Species                 |
+|-----------------------|-------------------------|
+| Early hardwood        | _Betula papyrifera_     |
+|                       | _Populus grandidentata_ |
+|                       | _Populus tremuloides_   |
+| Mid hardwood          | _Quercus rubra_         |
+|                       | _Acer rubrum_           |
+|                       | _Acer pensylvaticum_    |
+| Late hardwood         | _Acer saccharum_        |
+|                       | _Fagus grandifolia_     |
+| Pine                  | _Pinus strobus_         |
 
-CRUNCEP met.
-
-Other FoRTE data?
-- LAI trends: 1997-2014
-- Aboveground NPP data: 1999-2017
-
-Priors on leaf optical properties from Shiklomanov dissertation (chapter 3).
-
-## Modeling workflow
-
-PEcAn parameter sensitivity and uncertainty analysis [@lebauer_2013_facilitating].
-Other PEcAn references: [@dietze_2013_improving].
-
-Plant functional type definitions:
-
-- **Early successional hardwood**
-    - _Betula papyrifera_ (white/paper birch)
-    - _Populus grandidentata_ (bigtooth aspen)
-    - _Populus tremuloides_ (trembling aspen)
-- **Mid successional hardwood**
-    - _Quercus rubra_ (red oak)
-    - _Acer rubrum_ (red maple)
-    - _Acer pensylvaticum_ (striped maple)
-- **Late successional hardwood**
-    - _Acer saccharum_ (sugar maple)
-    - _Fagus grandifolia_ (American beech)
-- **Pine**
-    - _Pinus strobus_ (white pine)
-- _Omitted candidates_
-    - Early hardwood
-        - _Prunus pensylvatica_ (fire cherry)
-    - Mid hardwood
-        - _Betula alleghaniensis_ (yellow birch)
-        - _Tilia americana_ (basswood)
-        - _Fraxinus americana_ (white ash)
-    - Late hardwood
-        - _Ostyra virginiana_ (hop-hornbeam)
-    - Pine
-        - _Pinus resinosa_ (red pine)
-
-```{r param-table}
-tbl <- readd(param_table)
-tbl <- subset(tbl, select = -unit_parsed)
-tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-knitr::kable(tbl, escape = FALSE, caption = paste0(
-  "**Table**: Parameter descriptions. ",
-  "For additional information, see the ",
-  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
-))
-```
+For each plant functional type, we generated a distribution of parameter values via a trait-meta analysis [@lebauer_2013_facilitating; see also @dietze_2014_quantitative; @raczka_2018_what].
+Prior distributions for this meta-analysis were the same as those used in @dietze_2014_quantitative, @raczka_2018_what, and @viskari_2019_influence.
+Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
+The resulting parameter distributions are shown in Figure X.
+<!-- TODO more on this? -->
+All parameters not described in this section were set to their ED2 PFT-specific defaults.
 
@@
 
-For each combination of ED submodels, we ran 10 ED ensemble simulations.
+## Modeling workflow
+
+For each factorial combination of model configurations/submodels (described above), we ran 100 ensemble members from 1901 to 2000.
+Each ensemble member was initialized from ED2's "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
+Driving meteorological data was 6-hourly CRU-NCEP <!-- TODO ref.
+More here?
+--> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric].
+Soil texture was set to 92% sand, 7% silt, and 1% clay, per site-level observations in [@gough_2010_wood].
+The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
+The entire model execution workflow was conducted using the Predictive Ecosystem Analyzer (PEcAn, www.pecanproject.org).
 
@@
 
-- Structural vs. parameteric uncertainty
-- Comparison against observations
-- Others...
+- Overall summary:
+    - Overall parameter uncertainty -- Time-series plots of total annual GPP and NPP, and maximum LAI and AGB, averaged across ensemble members
+    - Impact of parameter uncertainty on ecological trajectories -- plot of max annual LAI by PFT
+    - Both compared against observations of recently observed LAI and NPP at UMBS (e.g. FASET, FoRTE)
+- Uncertainty analysis:
+    - Parameter vs. structural uncertainty -- Variance in mean annual GPP from 19XX to 19YY across ensembles within model structure, and compare against 
+    - Breakdown of parameter uncertainty -- PEcAn sensitivity and uncertainty analysis [@lebauer_2013_facilitating; @dietze_2014_quantitative]
+- Details of structural uncertainty:
+    - Vertical profiles of radiation, traits?
 
@@
 
+### PFT definitions
+
+- **Early successional hardwood**
+    - _Betula papyrifera_ (white/paper birch)
+    - _Populus grandidentata_ (bigtooth aspen)
+    - _Populus tremuloides_ (trembling aspen)
+- **Mid successional hardwood**
+    - _Quercus rubra_ (red oak)
+    - _Acer rubrum_ (red maple)
+    - _Acer pensylvaticum_ (striped maple)
+- **Late successional hardwood**
+    - _Acer saccharum_ (sugar maple)
+    - _Fagus grandifolia_ (American beech)
+- **Pine**
+    - _Pinus strobus_ (white pine)
+- _Omitted candidates_
+    - Early hardwood
+        - _Prunus pensylvatica_ (fire cherry)
+    - Mid hardwood
+        - _Betula alleghaniensis_ (yellow birch)
+        - _Tilia americana_ (basswood)
+        - _Fraxinus americana_ (white ash)
+    - Late hardwood
+        - _Ostyra virginiana_ (hop-hornbeam)
+    - Pine
+        - _Pinus resinosa_ (red pine)
+
+```{r param-table}
+tbl <- readd(param_table)
+tbl <- subset(tbl, select = -unit_parsed)
+tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
+knitr::kable(tbl, escape = FALSE, caption = paste0(
+  "**Table**: Parameter descriptions. ",
+  "For additional information, see the ",
+  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
+))
+```
+
 ##### pagebreak

### 2019-06-19 analysis/paper/paper.Rmd
Update `drake.R` for new outputs and figures

@@
 )
-readd(bigfacet_plot)
+readd(lai_pft_plot)
 ```

### 2019-06-19 analysis/drake.R

@@ -16,32 +16,32 @@ if (getRversion() >= "3.6") {
           immediate. = TRUE)
 }
 
-library(drake)
-# Data processing
-library(dplyr)
-library(tidyr)
-library(forcats)
-library(purrr)
-library(lubridate)
-library(readr)
-# Parallel execution
-library(future)
-library(future.callr)
-library(furrr)
-# Files and file paths
-library(fs)
-library(fst)
-library(here)
-# Plotting
-library(ggplot2)
-library(cowplot)
+suppressPackageStartupMessages({
+  library(drake)
+  # Data processing
+  library(dplyr)
+  library(tidyr)
+  library(forcats)
+  library(purrr)
+  library(lubridate)
+  library(readr)
+  # Parallel execution
+  library(future)
+  library(future.callr)
+  library(furrr)
+  # Files and file paths
+  library(fs)
+  library(fst)
+  library(here)
+  # Plotting
+  library(ggplot2)
+  library(cowplot)
+
+  devtools::load_all(here::here(), attach_testthat = FALSE)
+  expose_imports("fortebaseline")
+})
 
 cmdargs <- commandArgs(trailingOnly = TRUE)
-
-## library(fortebaseline)
-
-devtools::load_all(here::here(), attach_testthat = FALSE)
-expose_imports("fortebaseline")
 
 # Set common directories and paths
 analysis_dir <- dir_create(here("analysis"))
@@ -51,24 +51,88 @@ download_dir <- dir_create(path(data_dir, "retrieved"))
 monthly_output_file <- path(download_dir, "monthly-ensemble-output.fst")
 meta_analysis_file <- path(download_dir, "meta-analysis.rds")
 ensemble_params_file <- path(download_dir, "ed-ensemble-params.fst")
+cohort_file <- path(download_dir, "cohort_output.fst")
 
 plan <- drake_plan(
+  #########################################
   # Common elements
+  #########################################
   paper = target(
     rmarkdown::render(
       knitr_in(!!(path(analysis_dir, "paper", "paper.Rmd"))),
       .format
     ),
     transform = map(.format = c("html_document"))),
-  model_scale = c(RColorBrewer::brewer.pal(
-    n = nrow(both_uncertainty) - 1,
-    name = "Paired"
-  ), "black") %>% setNames(both_uncertainty[["model"]]),
   param_table = file_in(!!(path(data_dir, "derived-data", "parameter-table.csv"))) %>%
     read_csv(),
[…]

### 2019-06-21 analysis/paper/paper.Rmd
[paper] Move ED2 paragraph to methods

@@
 -->
-
-A key challenge of vegetation modeling research over the last several decades has been the **development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual** [@purves_2008_predictive].
-One such approach is the **"ecosystem demography" approach**, which uses a system of partial differential equations to simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method].
-<!-- TODO Run this description by Mike?
--->
-The original "Ecosystem Demography" (ED) model [@moorcroft_2001_method] was enhanced by @medvigy_2009_mechanistic (as ED2) with fully-coupled leaf physiology, canopy radiative transfer, and micrometeorology, and each of these components have been further refined over the last decade [culminating in ED2.2; @longo_2019_ed1].
-Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:
-
-- paleoecological studies [@rollinson_2017_emergent];
-- free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration];
-- estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
-- vegetation-climate feedbacks {Swann, etc} <!-- TODO -->
+Therefore, a key objective of vegetation modeling research over the last several decades has been the development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
+In this study, we focus on the Ecosystem Demography model [ED2.2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which constitutes an example of such an approach.
 
@@
 A full description of the default configuration of the model is provided by @longo_2019_ed1.
+
+The key feature of the "ecosystem demography" approach underlying the ED2 model is a system of partial differential equations that simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method].
+<!-- TODO Run this description by Mike?
+-->
+The original "Ecosystem Demography" (ED) model [@moorcroft_2001_method] was enhanced by @medvigy_2009_mechanistic (as ED2) with fully-coupled leaf physiology, canopy radiative transfer, and micrometeorology, and each of these components have been further refined over the last decade [culminating in ED2.2; @longo_2019_ed1].
+Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:
+
+- paleoecological studies [@rollinson_2017_emergent];
+- free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration];
+- estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
+- vegetation-climate feedbacks {Swann, etc} <!-- TODO -->
 

### 2019-06-21 analysis/paper/paper.Rmd
[paper] Add brief note about FASET to methods
This is WIP. Need to add much more here.

@@
 
-We performed this study at the University of Michigan Biological Station (Ameriflux site US-UMd, 45.5625$^{\circ}$, -84.6975$^{\circ}$), located in Northern Lower Michigan, USA.
+We performed this study at the University of Michigan Biological Station (UMBS; Ameriflux site US-UMd, 45.5625$^{\circ}$, -84.6975$^{\circ}$), located in Northern Lower Michigan, USA.
 The area surrounding the research station is 87% well-drained upland forest and 13% wetland [@bergen_2007_observing]; the focus of this study is on the former.
@@
 The majority of the forest that is aspen-dominated is undergoing succession to "northern hardwood" (maple, beech, basswood, birch, ash, hemlock), "upland conifer" (red and white pine), or "northern red-oak" ecotypes [@bergen_2007_observing].
+
+Forest stands at UMBS were previously exposed to experimental disturbance as part of the Forest Accelerated Succession Experiment (FASET). <!-- TODO ref -->
+<!-- TODO more on this -->
 

### 2019-06-21 analysis/paper/paper.Rmd
[paper] Add submodel table to paper

@@
 
+These sub-models are summarized in Table X.
+
+| Name                         | Description                                                                          | Color                     |
+|------------------------------|--------------------------------------------------------------------------------------|---------------------------|
+| **Crown model**              |                                                                                      |                           |
+| `closed` (default)           | Cohort crowns take up entire patch area. Competition for light based only on height. | Light                     |
+| `finite`                     | Cohort crown area is proportional to DBH according to PFT-specific allometry.        | Dark                      |
+| **Radiative transfer model** |                                                                                      |                           |
+| `two-stream` (default)       | Two-stream approximation, as in CLM 4.5                                              | Primary (red, blue)       |
+| `multi-scatter`              | Multiple-scatter approximation, following [@zhao_2005_multiple]                      | Secondary (green, orange) |
+| **Trait plasticity**         |                                                                                      |                           |
+| `static` (default)           | SLA and Vcmax are constant                                                           | Cool (blue, green)        |
+| `plastic`                    | SLA increases, and Vcmax decreases, with light level                                 | Warm (red, orange)        |
+
 ## ED2 parameters
@@
 
-| Plant functional type | Species                 |
-|-----------------------|-------------------------|
-| Early hardwood        | _Betula papyrifera_     |
-|                       | _Populus grandidentata_ |
-|                       | _Populus tremuloides_   |
-| Mid hardwood          | _Quercus rubra_         |
-|                       | _Acer rubrum_           |
-|                       | _Acer pensylvaticum_    |
-| Late hardwood         | _Acer saccharum_        |
-|                       | _Fagus grandifolia_     |
-| Pine                  | _Pinus strobus_         |
+| Plant functional type | Species                 | Color  |
+|-----------------------|-------------------------|--------|
+| Early hardwood        | _Betula papyrifera_     | Violet |
+|                       | _Populus grandidentata_ |        |
+|                       | _Populus tremuloides_   |        |
+| Mid hardwood          | _Quercus rubra_         | Blue   |
+|                       | _Acer rubrum_           |        |
+|                       | _Acer pensylvaticum_    |        |
+| Late hardwood         | _Acer saccharum_        | Green  |
+|                       | _Fagus grandifolia_     |        |
+| Pine                  | _Pinus strobus_         | Yellow |
 

### 2019-06-21 analysis/paper/paper.Rmd
[paper] Tweak figure sizes

@@
 
-```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 12)}
+```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 8)}
 bigfacet_cap <- paste(
@@
 
-```{r within-vs-across, fig.cap = within_vs_across_cap, fig.dim = c(6, 6)}
+```{r within-vs-across, fig.cap = within_vs_across_cap, fig.dim = c(9, 4)}
 within_vs_across_cap <- paste(

### 2019-06-21 analysis/drake.R

@@ -257,8 +257,11 @@ plan <- drake_plan(
     scale_fill_manual(
       values = tibble::deframe(both_uncertainty[, c("model", "color")])
     ) +
-    theme(axis.text.x = element_blank(),
-          axis.ticks.x = element_blank())
+    theme(
+      axis.text.x = element_blank(),
+      axis.ticks.x = element_blank(),
+      legend.title = element_blank()
+    )
   #########################################
   # Sensitivity analysis
   #########################################

### 2019-07-01 analysis/paper/paper.Rmd
[paper] Draft first results section

@@
   "by model configuration.",
-  "Red lines are mean (dashed) and min/max (dotted) values",
+  "Individual ensemble members are shown as gray lines.",
+  "The solid line is a LOESS regression (`ggplot2::stat_smooth`)."
+  "Shading represents the middle 80% (10th to 90th percentile) of predictions.",
+  "Black dot with error bars is the mean and min/max value",
   "from Hardiman et al. (2013)."
@@
 `r mdcap(summary_ts_cap)`
+
+Model projections of gross and net primary productivity, aboveground biomass, and leaf area index varied substantially with both parameter value and model structure.
+For every model structural configuration, there were multiple ensemble members that experienced total ecosystem collapse before the end of the simulation.
+However, model runs with finite canopy radius and two-stream radiative transfer had the highest failure rate, with most ensemble members collapsing almost immediately and the remaining few members showing relatively low productivity that declined to nearly zero.
+The remaining model configurations all followed a similar trend of rapid growth in the first simulation decade followed by gradual decline through the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+The ensemble means of these configurations had very similar ecosystem states at peak productivity and at the end of the simulation, as well as the pace of initial rapid growth.
+That being said, ensemble means from model configurations that included finite canopy radius and multiple-scatter radiative transfer sustained peak aboveground biomass and leaf area indices for longer than other configurations, though this was not true of gross or mean primary productivity.
+
+For most model configurations, the ensemble distribution of net primary productivity at the end of the run overlapped observations by Hardiman et al.
+(2013), with an ensemble mean that was slightly lower.
+(The exceptions were model configurations that included the combination of finite canopy radius and two-stream radiative transfer, which consistently underestimated productivity).
+On the other hand, ensemble predictions of leaf area index were consistently lower than observations, with only a subset of individual ensemble members across all configurations even overlapping the lower bounds of the observations.
+In fact, most ensemble members across all configurations failed to reach the observed leaf area index even at their peak.
+A subset of specific parameter combinations from model configurations that included finite canopy radius and multiple-scatter radiative transfer accurately reproduced the mean observed leaf area index.
 

### 2019-07-01 analysis/paper/paper.Rmd
Add results text about lai x pft figure

@@
   if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
+    out
+  } else {
+    invisible(NULL)
+  }
+}
+m <- 0
+htmlfig <- function() {
+  m <<- m + 1
+  out <- paste0("**Figure ", m, "**:")
+  if (grepl("html", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
     out
@@
 param_dist_cap <- paste(
+  htmlfig(),
   "Input parameter distributions from PEcAn trait meta-analysis."
@@
 summary_ts_cap <- paste(
+  htmlfig(),
   "ED2 plot-level ensemble predictions of",
@@
   "Individual ensemble members are shown as gray lines.",
-  "The solid line is a LOESS regression (`ggplot2::stat_smooth`)."
+  "The solid line is a LOESS regression (`ggplot2::stat_smooth`).",
   "Shading represents the middle 80% (10th to 90th percentile) of predictions.",
@@
 
-Model projections of gross and net primary productivity, aboveground biomass, and leaf area index varied substantially with both parameter value and model structure.
+Model projections of gross and net primary productivity, aboveground biomass, and leaf area index varied substantially with both parameter value and model structure (Figure 2).
 For every model structural configuration, there were multiple ensemble members that experienced total ecosystem collapse before the end of the simulation.
@@
 
-Take-away: Parameter uncertainty is as large as, or even larger, than structural uncertainty.
-
 ```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 8)}
 bigfacet_cap <- paste(
+  htmlfig(),
   "ED2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
@@
 
-Take-away: Different parameter combinations produce not just quantitatively different bulk ecosystem properties (e.g.
-NPP, total LAI), but also qualitatively different successional/ecological outcomes.
+Different parameter combinations led to model outputs that had not only different bulk ecosystem properties but also qualitatively different plant communities (Figure 3).
+Most model configurations could produce either a forest that was persistently and completely dominated by early hardwood or pine, or a forest with some coexistence of multiple PFTs.
+The exceptions, again, were model configurations combining finite canopy radius and the two-stream radiative transfer; in these configurations, the only outcomes were immediate ecosystem collapse of all PFTs or survival and mono-dominance of pine.
 

### 2019-07-01 analysis/paper/paper.Rmd
Add Shannon diversity index calculation

@@
 
-```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 7)}
+```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 10)}
 summary_ts_cap <- paste(

### 2019-07-01 analysis/drake.R

@@ -102,15 +102,25 @@ plan <- drake_plan(
       lai = sum(lai_co)
     ) %>%
     ungroup(),
+  # Calculate diversity indices based on LAI
+  diversity = fst(file_in(!!cohort_file)) %>%
+    .[, c("workflow_id", "run_id", "datetime", "pft", "lai_co")] %>%
+    group_by(workflow_id, run_id, year = year(datetime), pft) %>%
+    summarize(lai_co = quantile(lai_co, 0.9)) %>%
+    mutate(lai_p = lai_co / sum(lai_co)) %>%
+    filter(lai_p > 0) %>%
+    summarize(shannon = -sum(lai_p * log(lai_p))),
   jja_means = plot_means %>%
     filter(month(datetime) %in% 6:8) %>%
     group_by(workflow_id, run_id, year = year(datetime)) %>%
     summarize_at(vars(-datetime), mean) %>%
     ungroup() %>%
+    left_join(diversity, by = c("workflow_id", "run_id", "year")) %>%
+    mutate(shannon = if_else(is.na(shannon), 0, shannon)) %>%
     inner_join(current_workflows, by = "workflow_id"),
   jja_long = jja_means %>%
-    gather(variable, value, agb:lai) %>%
-    mutate(variable = factor(variable, c("gpp", "npp", "agb", "lai"))),
+    gather(variable, value, agb:shannon) %>%
+    mutate(variable = factor(variable, c("gpp", "npp", "agb", "lai", "shannon"))),
   jja_summary = jja_long %>%
     group_by(model_split, color, variable, year) %>%
     summarize(lo = quantile(value, 0.1),
@@ -120,7 +130,8 @@ plan <- drake_plan(
       "gpp" = "GPP ~ (kgC ~ year^{-1})",
       "npp" = "NPP ~ (kgC ~ year^{-1})",
       "agb" = "AGB ~ (kgC)",
-      "lai" = "LAI"
+      "lai" = "LAI",
+      "shannon" = "Shannon ~ diversity"
     ), default = label_parsed),
     .default = label_value
   ),
@@ -128,7 +139,7 @@ plan <- drake_plan(
     ~variable, ~low, ~mean, ~hi,
     "lai", 1.8, 4.14, 6.56,
     "npp", 1.68, 3.11, 7.26
-  ) %>% mutate(variable = factor(variable, c("gpp", "npp", "lai", "agb")),
[…]

### 2019-07-01 analysis/scratch/misc.R

@@ -1013,3 +1013,37 @@ known_vars <- tribble(
   ~varname, ~ed_name, ~dimensions,
   "radiation_profile", "FMEAN_RAD_PROFILE_CO", list("radiation", "cohorts")
 )
+
+#########################################
+shannon_index <- function(lai, pft) {
+  total_lai <- sum(lai)
+  p_i <- tapply(lai, pft, function(x) x / total_lai)
+  -sum(p_i * log(p_i), na.rm = TRUE)
+}
+
+simpson_index <- function(lai, pft) {
+  total_lai <- sum(lai)
+  p_i <- tapply(lai, pft, function(x) x / total_lai)
+  sum(p_i ^ 2, na.rm = TRUE)
+}
+
+diversity <- lai_q90 %>%
+  group_by(workflow_id, run_id, year) %>%
+  mutate(total_lai = sum(lai),
+         p_i = lai / total_lai) %>%
+  summarize(
+    shannon = -sum(p_i * log(p_i)),
+    simpson = sum(p_i ^ 2),
+    inv_simpson = 1 / simpson
+  ) %>%
+  ungroup()
+
+jja_means %>%
+  filter(is.na(shannon))
+
+diversity %>%
+  gather(variable, value, shannon, simpson, inv_simpson) %>%
+  ggplot() +
+  aes(x = year, y = value) +
+  geom_line(aes(group = run_id)) +
+  facet_grid(vars(variable), vars(workflow_id))

### 2019-07-02 analysis/paper/paper.Rmd
Drop confidence intervals, color ensembles in summary plot

--- 2019-07-01 analysis/paper/paper.Rmd
@@
 
-```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 7)}
+```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 10)}
 summary_ts_cap <- paste(
   htmlfig(),
-  "ED2 plot-level ensemble predictions of",
+  "ED2 plot-level predictions of",
   "gross (GPP) and net (NPP) primary productivity,",
-  "total leaf area index (LAI), and aboveground biomass (AGB)",
+  "aboveground biomass (AGB),",
+  "total leaf area index (LAI),",
+  "and Shannon diversity index (of plant functional types)",
   "by model configuration.",
-  "Individual ensemble members are shown as gray lines.",
-  "The solid line is a LOESS regression (`ggplot2::stat_smooth`).",
-  "Shading represents the middle 80% (10th to 90th percentile) of predictions.",
+  "The solid line is the ensemble mean.",
   "Black dot with error bars is the mean and min/max value",

### 2019-07-02 analysis/drake.R

@@ -123,7 +123,8 @@ plan <- drake_plan(
     mutate(variable = factor(variable, c("gpp", "npp", "agb", "lai", "shannon"))),
   jja_summary = jja_long %>%
     group_by(model_split, color, variable, year) %>%
-    summarize(lo = quantile(value, 0.1),
+    summarize(avg = mean(value),
+              lo = quantile(value, 0.1),
               hi = quantile(value, 0.9)),
   my_labeller = labeller(
     variable = as_labeller(c(
@@ -143,13 +144,11 @@ plan <- drake_plan(
                year = max(jja_summary$year)),
   summary_ts_plot = ggplot(jja_long) +
     aes(x = year) +
-    geom_line(aes(y = value, group = run_id), color = "grey50", alpha = 0.5) +
-    geom_ribbon(aes(ymin = lo, ymax = hi, fill = color),
-                data = jja_summary, alpha = 0.7) +
-    geom_smooth(aes(y = value), color = "black", size = 1) +
+    geom_line(aes(y = value, group = run_id, color = color), alpha = 0.5) +
+    geom_line(aes(y = avg), color = "black", size = 1, data = jja_summary) +
     geom_pointrange(aes(y = mean, ymin = low, ymax = hi), color = "black",
                     data = hardiman) +
-    scale_fill_identity() +
+    scale_color_identity() +
     facet_grid(vars(variable), vars(model_split), scales = "free_y",
                switch = "y", labeller = my_labeller) +
     theme_cowplot() +

### 2019-07-03 analysis/paper/paper.Rmd
Start working on uncertainty analysis text
Add more results text

@@
 
-Model projections of gross and net primary productivity, aboveground biomass, and leaf area index varied substantially with both parameter value and model structure (Figure 2).
+Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
+The ensemble means of all model configurations followed a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline throughout the rest of the simulation.
+However, the average rate of decline varied by model configuration, particularly the crown model:
+Compared to models with a "closed" canopy (i.e. complete shading; no horizontal competition), models with a "finite" canopy predicted slightly faster decline in net primary productivity.
+This decline was most severe in configurations that combined the finite canopy and two-stream radiative transfer; in these configurations, a disproportionately large number of ensembles experienced complete ecosystem collapse over the course of the simulation.
+
+In general, model configurations with the finite canopy radius model and two-stream radiative transfer were the least productive and diverse.
+
+On average,
+
 For every model structural configuration, there were multiple ensemble members that experienced total ecosystem collapse before the end of the simulation.
@@
 `r mdcap(within_vs_across_cap)`
+
+The importance of structural uncertainty relative to parameter uncertainty depended on the state variable in question.
+Uncertainty in gross and net primary productivity...
 

### 2019-07-04 analysis/paper/paper.Rmd
Update text in first part of results

--- 2019-07-02 analysis/paper/paper.Rmd
@@
 
-Model projections of gross and net primary productivity, aboveground biomass, and leaf area index varied substantially with both parameter value and model structure (Figure 2).
-For every model structural configuration, there were multiple ensemble members that experienced total ecosystem collapse before the end of the simulation.
-However, model runs with finite canopy radius and two-stream radiative transfer had the highest failure rate, with most ensemble members collapsing almost immediately and the remaining few members showing relatively low productivity that declined to nearly zero.
-The remaining model configurations all followed a similar trend of rapid growth in the first simulation decade followed by gradual decline through the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
-The ensemble means of these configurations had very similar ecosystem states at peak productivity and at the end of the simulation, as well as the pace of initial rapid growth.
-That being said, ensemble means from model configurations that included finite canopy radius and multiple-scatter radiative transfer sustained peak aboveground biomass and leaf area indices for longer than other configurations, though this was not true of gross or mean primary productivity.
+Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
+The ensemble means of all model configurations followed a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline throughout the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+However, the average rate of decline varied by model configuration, particularly the crown model:
+Compared to models with a "closed" canopy (i.e.
+complete shading; no horizontal competition), models with a "finite" canopy predicted slightly faster decline in net primary productivity and functional diversity.
+This decline was most severe in configurations that combined the finite canopy and two-stream radiative transfer; in these configurations, a disproportionately large number of ensembles experienced complete ecosystem collapse over the course of the simulation, and the overwhelming fraction of ensemble members predicted near-zero Shannon diversity after 30 years.
+Besides this specific configuration, which consistently under-predicted observed net primary productivity at UMBS, the remaining model configurations had ensemble mean net primary productivity values that were very close to the mean observed value from Hardiman et al.
+(2013).
+Moreover, the ensemble spread of net primary productivity from these configurations was very similar to the uncertainty in the observations, though the distribution was relatively balanced for model predictions but skewed positive for the observations.
 
-For most model configurations, the ensemble distribution of net primary productivity at the end of the run overlapped observations by Hardiman et al.
-(2013), with an ensemble mean that was slightly lower.
-(The exceptions were model configurations that included the combination of finite canopy radius and two-stream radiative transfer, which consistently underestimated productivity).
-On the other hand, ensemble predictions of leaf area index were consistently lower than observations, with only a subset of individual ensemble members across all configurations even overlapping the lower bounds of the observations.
-In fact, most ensemble members across all configurations failed to reach the observed leaf area index even at their peak.
-A subset of specific parameter combinations from model configurations that included finite canopy radius and multiple-scatter radiative transfer accurately reproduced the mean observed leaf area index.
+Patterns in leaf area index were somewhat distinct from patterns in productivity and diversity.
+While model configurations with finite canopy and two-stream radiative transfer again predicted the lowest leaf area index on average, model configurations with finite canopy and multiple-scatter predicted the _highest_ leaf area index on average.
+The latter configurations also had the largest number of ensemble members that fell within the observed distribution of leaf area index; meanwhile, every model configuration predicted an ensemble mean leaf area index at the end of the simulation that was significantly below the observation from Hardiman et al.
+(2013).
 

### 2019-07-04 analysis/paper/paper.Rmd
Note about alternate LAI by PFT analysis

@@
 Different parameter combinations led to model outputs that had not only different bulk ecosystem properties but also qualitatively different plant communities (Figure 3).
-Most model configurations could produce either a forest that was persistently and completely dominated by early hardwood or pine, or a forest with some coexistence of multiple PFTs.
-The exceptions, again, were model configurations combining finite canopy radius and the two-stream radiative transfer; in these configurations, the only outcomes were immediate ecosystem collapse of all PFTs or survival and mono-dominance of pine.
+The most common outcome was immediate and persistent dominance by early hardwood trees, though in some cases, early hardwoods co-existed with late hardwood or pine.
+<!-- TODO Probably need to do a different analysis here.
+Maybe look at all the outputs, identify several different "ecological scenarios", and then analyze the fraction of runs that experienced those probabilities?
+-->
 

### 2019-07-04 analysis/paper/paper.Rmd
Write results for uncertainty barplot

@@
 within_vs_across_cap <- paste(
+  htmlfig(),
   "Comparison of parameteric uncertainty within ensembles (colored bars, colored by model type)",
@@
 
-The importance of structural uncertainty relative to parameter uncertainty depended on the state variable in question.
-Uncertainty in gross and net primary productivity...
-
-Take-away: Overall, structural uncertainty (variance in ensemble means across structures) is comparable or lower than parameter uncertainty (uncertainty within ensembles for a given structure).
-Parameter uncertainty varies significantly across model structures.
-Enabling trait plasticity increases variance in predictions of AGB.
-Enabling the finite canopy radius increases the variance in LAI predictions.
-GPP and NPP have the highest variance by far for the combination of finite canopy radius and two-stream radiative transfer.
+Across all variables, parameter uncertainty (across ensembles within each model configuration) was at least as large, and usually larger, than structural uncertainty (Figure 4).
+The relative importance of parameter uncertainty varied with both model configuration and the variable in question. 
+Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) led to reduced predictive uncertainty for all four variables considered, with the largest uncertainty reduction for gross and net primary productivity.
+Meanwhile, facilitating horizontal competition by enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) substantially reduced uncertainty in aboveground biomass and slightly reduced uncertainty in gross and net primary productivity.
+Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased uncertainty in leaf area index and aboveground biomass, but did not have a consistent effect on uncertainty in predictions of productivity.
 

### 2019-07-04 analysis/paper/paper.Rmd
Add Shannon diversity to all analyses

@@
 Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) led to reduced predictive uncertainty for all four variables considered, with the largest uncertainty reduction for gross and net primary productivity.
-Meanwhile, facilitating horizontal competition by enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) substantially reduced uncertainty in aboveground biomass and slightly reduced uncertainty in gross and net primary productivity.
-Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased uncertainty in leaf area index and aboveground biomass, but did not have a consistent effect on uncertainty in predictions of productivity.
+Meanwhile, facilitating horizontal competition by enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) substantially reduced uncertainty in aboveground biomass and diversity, and slightly reduced uncertainty in gross and net primary productivity.
+Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased uncertainty in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on uncertainty in predictions of productivity.
 

### 2019-07-04 analysis/drake.R

@@ -140,9 +140,11 @@ plan <- drake_plan(
     left_join(diversity, by = c("case", "model_id", "year")) %>%
     mutate(shannon = if_else(is.na(shannon), 0, shannon)) %>%
     inner_join(models, by = "model_id"),
+  use_vars = c("gpp", "npp", "agb", "lai", "shannon"),
+  use_vars_cap = c("GPP", "NPP", "AGB", "LAI", "Shannon"),
   jja_long = jja_means %>%
     gather(variable, value, agb:shannon) %>%
-    mutate(variable = factor(variable, c("gpp", "npp", "agb", "lai", "shannon")),
+    mutate(variable = factor(variable, use_vars),
            param_id = as.numeric(substr(case, 1, 3))),
   jja_summary = jja_long %>%
     group_by(model, color, variable, year) %>%
@@ -164,7 +166,7 @@ plan <- drake_plan(
     ~variable, ~low, ~mean, ~hi,
     "lai", 1.8, 4.14, 6.56,
     "npp", 1.68, 3.11, 7.26
-  ) %>% mutate(variable = factor(variable, c("gpp", "npp", "lai", "agb", "shannon")),
+  ) %>% mutate(variable = factor(variable, use_vars),
                year = max(jja_summary$year)),
   summary_ts_plot = ggplot(jja_long) +
     aes(x = year) +
@@ -259,16 +261,16 @@ plan <- drake_plan(
   time_averages = jja_means %>%
     filter(year > 1910) %>%
     group_by(model, color, case) %>%
-    summarize_at(vars(gpp, npp, agb, lai), mean) %>%
+    summarize_at(use_vars, mean) %>%
     mutate(param_id = as.numeric(substr(case, 1, 3))),
   parameter_uncertainty = time_averages %>%
     group_by(model, color) %>%
-    summarize_at(vars(gpp, npp, agb, lai), var) %>%
+    summarize_at(use_vars, var) %>%
     ungroup() %>%
     arrange(model),
   structure_uncertainty = time_averages %>%
     group_by(model) %>%
-    summarize_at(vars(gpp, npp, agb, lai), mean) %>%
+    summarize_at(use_vars, mean) %>%
     ungroup() %>%
     summarize_at(vars(-model), var) %>%
     mutate(model = "Across-structure",
@@ -279,11 +281,8 @@ plan <- drake_plan(
     mutate(model = fct_inorder(model)),
   within_across_plot = both_uncertainty %>%
[…]

### 2019-07-04 analysis/paper/paper.Rmd
Add more results text

@@
 uncertainty_analysis_cap <- paste(
-  "PEcAn-liked parameter sensitivity and uncertainty analysis, by model type.",
+  htmlfig(),
+  "PEcAn-like parameter sensitivity and uncertainty analysis, by model type.",
   "Elasticity (a) is the normalized sensitivity of the model to a fixed change in the parameter.",
   "pvar (b) is the partial variance, which describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
-  "Input parameter distributions are shown in Figure XXX."
+  "Input parameter distributions are shown in Figure 1."
 )
@@
 
-Take-away: Model parameter uncertainty is sensitive to model configuration.
-For example, turning on the finite canopy radius with the two-stream canopy RTM increases sensitivity across many parameters.
+Model sensitivity to specific parameters ("elasticity") varied depending on the model configuration and the target output variable (Figure 5).
+For predicting C fluxes, biomass, and leaf area, the three most important parameters in terms of model sensitivity were specific leaf area (SLA) for early and late hardwoods, leaf C:N ratio (`c2n_leaf`) for early and mid hardwoods, and V~cmax~ for mid and late hardwoods.
+Both the magnitude and direction of sensitivity to these parameters varied by model configuration.
+For example, in the default ED2 configuration (closed canopy, two-stream radiative transfer, static traits), net primary productivity increased with early hardwood specific leaf area, but when this configuration was altered so SLA and V~cmax~ could vary with light level, net primary productivity decreased with early hardwood specific leaf area.
+Similarly, the model became much more sensitive to early- and mid-hardwood leaf C:N ratio when switching from a closed canopy to a finite canopy configuration.
 
+The parameters that contributed most to overall model predictive uncertainty ("partial variance") differed the parameters to which the model was most sensitive (Figure 5), largely driven by the fact that many of the parameters with the highest sensitivity (e.g.
+SLA, V~cmax~) were also well constrained by data (Figure 1).
+Patterns in partial variance were largely idiosyncratic across model configurations;
+when we considered the top-5 most important parameters for a given output variable and model configuration,
+only five combinations were present in at least half of the model configurations:
+growth respiration, C balance mortality ratio (`mort2`), and stomatal slope (all for early hardwoods).
+
+<!-- TODO Probably need a different analysis here more along the lines of that last sentence. -->
+
+<!--
 ## Details about model structures
@@
 
-<!-- Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation. -->
+Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation.
 
 (TODO)
+-->
 
@@
 
-- Comparisons against other studies
-    - Parameter sensitivity
-        - [@pandit_2018_optimizing] -- Most important parameters for shrubs were SLA, Vcmax, stomatal slope, root:leaf ratio, and root turonver rate (slightly less important: leaf turnover rate, growth respiration, cuticular conductance; much less important: water conductance; basically irrelevant: storage turnover, leaf width)
+- Parameter uncertainty is the dominant driver of model predictive uncertainty.
+- ...but the nature of this parameter uncertainty depends on model configuration.
+
 - Ecological implications of structural uncertainty
@@
     - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
+- Comparisons against other studies
+    - Parameter sensitivity
+        - [@pandit_2018_optimizing] -- Most important parameters for shrubs were SLA, Vcmax, stomatal slope, root:leaf ratio, and root turonver rate (slightly less important: leaf turnover rate, growth respiration, cuticular conductance; much less important: water conductance; basically irrelevant: storage turnover, leaf width)
+        - Compare against @dietze_2014_quantitative and @raczka_2018_what
 - Future directions: Predicting disturbance response

### 2019-07-04 analysis/drake.R

@@ -325,8 +325,10 @@ plan <- drake_plan(
     unnest(raw_sa) %>%
     ungroup(),
   sensitivity_plot_data = sensitivity_results %>%
-    ## mutate(elasticity = if_else(abs(elasticity) > 150, NA_real_, elasticity)) %>%
-    group_by(yvar, xvar) %>%
+    mutate(elasticity = if_else(abs(elasticity) > 75,
+                                75 * sign(elasticity),
+                                elasticity)) %>%
+    group_by(model, yvar) %>%
     mutate(fpvar = pvar / sum(pvar)) %>%
     group_by(xvar) %>%
     mutate(total_pvar = sum(pvar)) %>%
@@ -349,15 +351,30 @@ plan <- drake_plan(
       theme_cowplot() +
       theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
             axis.text.y = element_text(size = rel(0.6)),
-            legend.position = "bottom"),
+            legend.position = "bottom",
+            legend.key.width = unit(3, "lines")),
     transform = map(YYY = c(elasticity, fpvar),
-                    SCALE = c(scale_fill_gradient2(),
-                              scale_fill_continuous(low = "white", high = "blue")))
+                    SCALE = c(scale_fill_gradientn(colors = c("darkred", "darkorange3",
+                                                              "white",
+                                                              "deepskyblue3", "darkblue"),
+                                                   values = scales::rescale(c(
+                                                     -75, -25,
+                                                     0,
+                                                     25, 75))),
+                              scale_fill_continuous(low = "white", high = "darkblue")))
   ),
   sensitivity_plot = target(
     cowplot::plot_grid(sensitivity_plot_piece),
     transform = combine(sensitivity_plot_piece)
-  )
+  ),
+  top_params_fpvar = sensitivity_plot_data %>%
+    group_by(yvar, model) %>%
+    top_n(5, fpvar) %>%
+    arrange(desc(fpvar), .by_group = TRUE) %>%
+    ungroup() %>%
+    select(model, shortname, yvar, xvar, fpvar),
+  top_params_fpvar_count = top_params_fpvar %>%
+    count(yvar, shortname, xvar, sort = TRUE)
 )
 
 if ("--poster" %in% cmdargs) source("analysis/drake_poster.R")

### 2019-07-06 analysis/paper/paper.Rmd
Feedback we can discuss more on Monday

@@
 
-**Temperate forests in the Upper Midwest are important.**
+**Temperate forests in the Upper Midwest are important and changing quickly.**
 {C sequestration}.
@@
 **Therefore, predictions about the future of these forests based on past analogs may not be reliable.**
+BBL: This paragraph obviously needs significant fleshing out, but this should be straightforward; I think we can basically take (with a bit of reworking) material from Chris's 2016 _Ecosphere_ paper: http://dx.doi.org/10.1002/ecs2.1375
 
-**Instead, more reliable predictions can be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
+**More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
 Vegetation models fall broadly along the following spectrum of complexity.
 On one side are relatively simple **"Big leaf" models**---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples?
+BBL: no!
 -->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
 <!-- TODO refs -->
-These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models, though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
+These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models (BBL: I don't think is is true; earliest models like Forest-BGC were standalone), though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
 <!-- TODO refs -->
@@
 Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
-Moreover, these models' **predictions of species composition and ecosystem structure are relevant outside of carbon cycle science**, particularly for forest management and wildlife conservation.
-<!-- TODO refs -->
-However, these **benefits of individual-based models come at the cost of much higher data and computational requirements**, which effectively restrict their application to individual, well-studied sites.
-<!-- TODO refs.
-Also, is this actually true?
--->
-Therefore, a key objective of vegetation modeling research over the last several decades has been the development of approaches that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
+In the middle are models that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive], and have less-intensive data requirements generally.
 In this study, we focus on the Ecosystem Demography model [ED2.2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which constitutes an example of such an approach.
 
-**Vegetation model projections are inherently uncertain, and this uncertainty stems from several sources.**
-_Driver_ uncertainty refers to uncertainty in data about processes that are essential to the system in question but are not represented by the model (e.g. weather and climate for vegetation models).
-_Initial condition_ uncertainty arises from the fact that models have to start somewhere, and the exact conditions at the place and time simulations begin are uncertain.
+BBL: General citations on simple/complex model tradeoffs: [Hawkes (2001)](http://dx.doi.org/10.1016/S0304-3800(00)00267-2), Mladenoff (2004), ...
+Moreover, these models' **predictions of species composition and ecosystem structure are relevant outside of carbon cycle science**, particularly for forest management and wildlife conservation.
+<!-- BBL: maybe move this sentence to discussion?
+TODO refs -->
+
+**Regardless of vegetation model complexity, their projections are inherently uncertain, however.** 
+_Driver_ uncertainty refers to uncertainty in data about processes not represented by the model (e.g. weather and climate for vegetation models).
+_Initial condition_ uncertainty arises from the fact that models have to start somewhere, and the exact conditions at the place and time simulations begin are frequently unknown.
 _Process_ or _structural_ uncertainty arises because vegetation models necessarily only represent a subset of all processes involved in plant biology, and that the process that are included can usually be represented in multiple different ways (i.e.
@@
 Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
-<!-- TODO This whole paragraph is awkward -->
+A key but unknown question in ecosystem modeling is which of the last two categories--structural versus parametric uncertainty--is more important (Fisher et al.
+2015, https://onlinelibrary.wiley.com/doi/full/10.1111/j.1523-1739.2003.00614.x, Ricciuto et al.
+2008).
+(BBL: trying to better link this paragraph into previous and following ones.)
 
 **Several recent studies have looked at parameter uncertainty in ED2.**
-@dietze_2014_quantitative evaluated which ED2 parameters contributed the most to uncertainties in ED2 simulations of productivity across multiple biomes in North America at 1, 5, and 10 year time scales.
-They found that the parameters contributing the most to model uncertainty across all biomes were those related to growth respiration, mortality, stomatal conductance, and water uptake.
+@dietze_2014_quantitative found that the parameters contributing the most to ED2 model uncertainty across all biomes were those related to growth respiration, mortality, stomatal conductance, and water uptake.
 However, they also found that despite some common patterns, parameter uncertainty varied with biome.
@@
 Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large (>20%) impact on predictions of net primary productivity [@viskari_2019_influence].
-The variability in results across these studies demonstrates that parameter contributions to ED2 uncertainty are highly variable, and that additional parameter uncertainty studies are needed.
+(BBL: probably too much detail?)
 
-**While parameter uncertainty ED2 has been previously evaluated, less work has been done on structural uncertainty**.
+**Much less work has been done on structural uncertainty.**.
 Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization (REF - Medlyn?), high-latitude warming (Rogers and Serbin?), {others...}.
@@
 -->
-**However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways.**
-In this study, we evaluate structural uncertainty more precisely/directly by selectively toggling and switching individual processes/model components in ED2 while keeping everything else constant.
+**However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.**
 
-**The focus of this study is on the interaction between parametric and structural uncertainty in the ED2 model.**
+**This study focuses on the interaction between parametric and structural uncertainty in the ED2 model.**
 Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
+(BBL: move to methods)
 Our study is organized around the following guiding questions:
@@
 complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
-We then used a sensitivity and variance decomposition analysis to evaluate the contribution of parameter uncertainty for each model configuration.
+We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results.
 
@@
 
-Forest stands at UMBS were previously exposed to experimental disturbance as part of the Forest Accelerated Succession Experiment (FASET). <!-- TODO ref -->
+Forest stands at UMBS were previously exposed to experimental disturbance as part of the Forest Accelerated Succession Experiment (FASET).
+<!-- TODO ref --> BBL: what's the canonical FoRTE citation?
+Hmm.
+http://dx.doi.org/10.1111/nph.15227
 <!-- TODO more on this -->
@@
 {Brief summary of thermodynamics?}
-{Brief summary of water cycle?}
+{Brief summary of water cycle?} BBL: maybe in one sentence each, but otherwise no, send 'em to a citation.
 
@@
 Three more parameters correspond to the respiration rates of leaves, roots, and "growth maintenance".
+
 Two more parameters control carbon allocation:
 One is the ratio of fine root to leaf biomass (`fineroot2leaf`, or `q`), and another is the ratio of "storage" carbon allocated to reproduction (`r_fract`).
+
 Three parameters control various aspects of adult tree mortality, namely the density-independent mortality rate (`mort3`), and the time scale (`mort1`) and critical carbon balance (`mort2`) of mortality from carbon starvation.
@@
 _Specific leaf area_ (`SLA`) is used to convert leaf biomass to leaf area index, which in turn is used in a variety of calculations related to canopy radiative transfer and micrometeorology.
+
 Two additional parameters describe canopy structure in ways critical to the canopy radiative transfer:
@@
 Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
-The ensemble means of all model configurations followed a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline throughout the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+The ensemble means of all model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline throughout the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
 However, the average rate of decline varied by model configuration, particularly the crown model:
@@
 
-Model sensitivity to specific parameters ("elasticity") varied depending on the model configuration and the target output variable (Figure 5).
+Model sensitivity to specific parameters ("elasticity" _sensu_ ...) varied depending on the model configuration and the target output variable (Figure 5).
 For predicting C fluxes, biomass, and leaf area, the three most important parameters in terms of model sensitivity were specific leaf area (SLA) for early and late hardwoods, leaf C:N ratio (`c2n_leaf`) for early and mid hardwoods, and V~cmax~ for mid and late hardwoods.
@@
         - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
+        - Perhaps note "ecological equifinality" here: there are different ways to arrive at similar outcomes in the model; this is also true in nature? 
 - Implications of uncertainty analysis for future data collection and model development

### 2019-07-09 analysis/paper/paper.Rmd
A few more notes on the multiple scatter model

@@
 - $LAI$ -- Leaf area index
-- TAI -- Total area index (leaf and woody area)
+- $TAI$ -- Total area index (leaf and woody area)
 - $LAI_e$ / $TAI_e$ -- "Effective" area indices (accounting for clumping)
@@
 
+Effective LAI is just true LAI x clumping.
+
 $$ LAI_e = \omega LAI $$
 
+Total area index (TAI) is leaf (LAI) + wood (WAI) area.
+
 $$ TAI_e = LAI_e + WAI $$
+
+Local exposed total area index ($TAI_l$) is the total area index divided by crown area index ($CAI$).
 
@@
 
-Projected area, based on coefficients $\phi_1$ and $\phi_2$.
+Projected area ($a_{proj}$), based on coefficients $\phi_1$ and $\phi_2$.
 
 $$ a_{proj} = \phi_1 + \phi_2 \mu $$
+
+Optical depth ($\lambda$) is projected area over optical depth ($\mu$)
+
+$$ \lambda = \frac{a_{proj}}{\mu} $$
+
+Transmittance for direct radiation of a canopy layer ($\tau_{beam}$) is:
+
+$$ \tau_{beam} = (1 - CAI) + CAI exp\left( -\lambda TAI_l \right) $$
+
+Transmittance for diffuse radiation of a canopy layer ($\tau_{diff}$):
+
+$$ ext_1 = \phi_1 \TAI_l $$
+$$ ext_2 = \phi_2 \TAI_l $$
+$$ \tau_{diff} = (1 - CAI) - CAI exp\left( -ext_1 - ext_2 \right) ext_1^2 e^{exp_1} E_i(-exp_1) + (ext_1 - 1)$$
+
+where $E_i$ is the exponential integral.
+
+Backscatter for
 

### 2019-07-10 analysis/paper/submodel-rtm.Rmd
Add draft of RTM submodel descriptions

---
title: "Submodel descriptions: Radiative transfer"
author: Alexey Shiklomanov
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
stopifnot(
  requireNamespace("purrr", quietly = TRUE),
  requireNamespace("gsl", quietly = TRUE)
)
```

# Common features

## Scattering and backscattering of canopy elements

Following @clm45_note, forward- ($\nu$) and backscattering of leaves or stems ($\nu$) are defined as a function of reflectance ($R$) and transmittance ($T$) of the corresponding materials:

$$ \nu = R + T $$
$$ \omega = \frac{R + T + \frac{1}{4} (R-T)(1 - \chi)^2}{2 (R+T)} $$

Both of these quantities are calculated independently for leaves and wood, and then averaged based on the relative area indices of leaves (LAI, $L$) and wood (WAI, $W$) within a canopy layer.

$$ w_{leaf} = \frac{L}{L + W}$$ 
$$ \nu = \nu_{leaf} w_{leaf} + \nu_{wood} (1 - w_{leaf}) $$
$$ \omega = \omega_{leaf} w_{leaf} + \omega_{wood} (1 - w_{leaf}) $$

## Directional scattering and leaf angle distribution

The directional extinction coefficient ($K(\psi)_i$) can be expressed as:

$$ K(\psi)_i = \frac{G(\psi)}{cos(\psi)} $$

where $G(\psi)$ describes the mean projection per unit leaf area (or "relative projected leaf area") in direction $\psi$.

The leaf angle distribution function used in ED2 is the same as the one used in CLM 4.5 [@clm45_note]:

$$ G(\psi) = \phi_1 + \phi_2 cos(\psi) $$
$$ \phi_1 = 0.5 - 0.633 \chi - 0.33 \chi^2 $$
$$ \phi_2 = 0.877 (1 - 2 \phi_1) $$

where $\chi$ is the _leaf orientation factor_ parameter, defined such that -1 is perfectly vertical leaves, 1 is perfectly horizontal leaves, and 0 is randomly distributed leaf angles.

```{r g-function, echo = FALSE, fig.cap = gfun_cap}
gfun_cap <- "Leaf projection as a function of leaf orientation factor."
phi1_f <- function(orient) {
  0.5 - 0.633 * orient - 0.33 * orient ^ 2
}
phi2_f <- function(orient, phi1 = phi1_f(orient)) {
  0.877 * (1 - 2 * phi1)
}
gfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
  phi1 <- phi1_f(orient)
  phi2 <- phi2_f(phi1 = phi1)
  mu <- cos(theta_rad)
  phi1 + phi2 * mu
}
max_theta <- 89
orient <- c(-0.5, -0.25, 0, 0.25, 0.5)
colors <- colorRampPalette(c("red", "black", "blue"))(length(orient))
plot(c(-90, 90), c(0, 0.9), type = "n",
     xlab = "Zenith angle (degrees)",
     ylab = expression(G(psi)))
legend("bottom", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
invisible(purrr::map2(
    orient, colors,
    ~curve(gfunction(x, .x), -max_theta, max_theta,
           col = .y, add = TRUE)
))
```

```{r kfunction, echo = FALSE, fig.cap = kcap}
kcap <- "K function"
kfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
  gfunction(theta, orient, theta_rad) / cos(theta_rad)
}
plot(c(-70, 70), c(0.2, 1.8), type = "n",
     xlab = "Zenith angle (degrees)",
     ylab = expression(K(psi)))
legend("top", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
invisible(purrr::map2(
  orient, colors,
  ~curve(kfunction(x, .x), -70, 70,
         col = .y, add = TRUE)
))
```

An additional quantity is the average inverse optical depth per unit leaf/wood area ($\bar\mu$):

$$
\bar\mu = \frac{1}{\phi_2}\left( 1 - \frac{\phi_1}{\phi_2} \ln\left( 1 + \frac{\phi_2}{\phi_1}\right) \right)
$$

```{r}
mu_bar <- function(orient) {
  phi1 <- phi1_f(orient)
  phi2 <- phi2_f(phi1 = phi1)
  x <- 1 - phi1 / phi2 * log(1 + phi2 / phi1)
  out <- x / phi2
  out[orient == 0] <- 1
  out
}
curve(mu_bar(x), -0.5, 0.5,
      xlab = "Orient factor",
      ylab = expression(bar(mu)))
```

## Beam backscatter

The beam backscatter (or "upscatter") coefficient for direct radiation, $\beta_0$, is defined as:

$$
\beta_0 = a_s(\psi) \frac{1 + \bar\mu K(\psi)}{\bar\mu K(\psi)}
$$

where $a_s(\psi)$ is the single scattering albedo coefficient, defined as:

$$
a_s(\psi) = \frac{1}{2}
\frac{G(\psi)}{\cos\psi \phi2 + G(\psi)}
\left(
1 - 
\frac{\cos\psi \phi1}{\cos\psi + G(\psi)}
\ln\left(
\frac{\cos\psi \phi1 + \cos\psi \phi2 + G(\psi)}{\cos\psi \phi1}
\right)
\right)
$$

(Note that, for simplicity, $a_s$ here is equivalent to $\frac{a_s}{\omega}$ in @clm45_note equation 3.15, where $\omega$ is the leaf backscatter.)

## Exponential decay of direct radiation

In both the two-stream and multiple scatter schemes, without finite canopy radius, the transmissivity of a layer to direct radiation ($\tau(\psi)_i$) is defined as follows:

$$ \tau(\psi)_i = \exp(-K(\psi)_i L_i) $$

(Note that the inverse optical depth, $\mu_0$, in the two-stream calculation is just $K(\psi)^{-1}$).

## Matrix solution for diffuse fluxes

Both the two-stream and multiple-scatter canopy radiative transfer schemes are set up as the solutions to the following linear matrix equation:

$$ M \times X = Y $$

For $n$ canopy layers, $M$ is a $(2m+2, 2m+2)$ matrix and $X$ and $Y$ $(2m+2)$ vectors.
$X$ is the vector of fluxes (or values that can be converted to fluxes), formatted as follows:

$$
\begin{bmatrix}
F_{up,ground} \\
F_{down,1} \\
F_{up, 1} \\
... \\
F_{down,i} \\
F_{up,i} \\
... \\
F_{down,n} \\
F_{up,n} \\
F_{down,sky}
\end{bmatrix}
$$

$M$ and $Y$ consist of coefficients, which are defined differently in the two models.
Notably, in both cases $M$ is a sparse matrix, composed of zero elements everywhere except the diagonal and its immediate neighbors;
for example, for $n = 3$, $M$ looks like this:

$$
M = \begin{bmatrix}
1 & 0 & 0 & 0 & 0 & 0 \\
m_{2,1} & m_{2,2} & m_{2,3} & 0 & 0 & 0 \\
0 & m_{3,2} & m_{3,3} & m_{4,3} & 0 & 0 \\
0 & 0 & m_{4,3} & m_{4,4} & m_{4,5} & 0 \\
0 & 0 & 0 & m_{5,4} & m_{5,5} & m_{5,6} \\
0 & 0 & 0 & 0 & 0 & 1 \\
\end{bmatrix}
$$

# Multiple scatter

All of this is based on @zhao_2005_multiple.

The $X$ and $Y$ vectors of the matrix problem for diffuse fluxes are defined as follows:

$$
X =
\begin{bmatrix}
SWuO_0 \\
SWd0_1 \\
SWu0_1 \\
... \\
SWd0_i \\
SWu0_i \\
... \\
SWd0_m \\
SWu0_m \\
SWd0_{m + 1}
\end{bmatrix}
$$

$$
Y = 
\begin{bmatrix}
S_0 a_{ground} \\
S_1 r(\psi)_1 \left[ 1 - r_0 r_1 (1 - \alpha_0) (1 - \tau_0) (1 - \alpha_1) (1 - \tau_1) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) \\
S_1 \left[ 1 - r_1 r_2 (1 - \alpha_1) (1 - \tau_1) (1 - \alpha_2) (1 - \tau_2) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) (1 - r(\psi)_1) \\
... \\
S_i r(\psi)_i \left[ i - r_{i-1} r_i (i - \alpha_{i-1}) (i - \tau_{i-1}) (i - \alpha_i) (i - \tau_i) \right] (i - \tau(\psi)_i) (i - \alpha_i) \\
S_i \left[ i - r_i r_{i+1} (i - \alpha_i) (i - \tau_i) (i - \alpha_{i+1}) (i - \tau_{i+1}) \right] (i - \tau(\psi)_i) (i - \alpha_i) (i - r(\psi)_i) \\
... \\
S_m r(\psi)_m \left[ m - r_{m-1} r_m (m - \alpha_{m-1}) (m - \tau_{m-1}) (m - \alpha_m) (m - \tau_m) \right] (m - \tau(\psi)_m) (m - \alpha_m) \\
S_m \left[ m - r_m r_{m+1} (m - \alpha_m) (m - \tau_m) (m - \alpha_{m+1}) (m - \tau_{m+1}) \right] (m - \tau(\psi)_m) (m - \alpha_m) (m - r(\psi)_m) \\
SW_{sky}
\end{bmatrix}
$$

Here, $a_{ground}$ is the albedo of the ground under the canopy and $SW_{sky}$ is the incident shortwave hemispherical flux from the sky;
both are exogenous inputs to the model.
$S_i$ is the direct ("beam") radiation at layer $i$, and is calculated in a loop as follows:

$$ S_i = S_{i + 1} \tau(\psi)_i $$

with $S_{n+1}$ as the incident direct solar flux, an exogenous input.

$M$ is a sparse $(2n + 2)$ by $(2n + 2)$ matrix (where $m$ is the number of vertical canopy layers) that is zero everywhere except on the diagonal and its immediate neighbors.
For example, for $m = 2$:

For $i = 1,2,3,...n$, where $n$ is the number of canopy layers.

$$ m_{1,1} = m_{2n+2,2n+2} = 1 $$
$$ m_{2i,2i-1} = - \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right]$$
$$ m_{2i,2i} = -r_{i-1} \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right] (1 - \alpha_{i-1})(1 - \tau_{i-1}) $$
$$ m_{2i,2i+1} = 1 - r_{i-1} r_i (1 - \alpha_{i-1})(1 - \tau_{i-1})(1 - \alpha_i)(1 - \tau_i) $$
$$ m_{2i+1,2i} = 1 - r_i r_{i+1} (1 - \alpha_i)(1 - \tau_i)(1 - \alpha_{i+1})(1 - \tau_{i+1}) $$
$$ m_{2i,2i} = -r_{i+1} \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right] (1 - \alpha_{i+1})(1 - \tau_{i+1}) $$
$$ m_{2i+1,2i+2} = - \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right]$$

The layer absorption ($\alpha$) and backscatter for direct ($r(\psi)_i$) and diffuse ($r_i$) light are defined as follows:

$$ \alpha_i = 1 - \nu $$
$$ r_i = \omega $$
$$ r(\psi)_i = \beta_0 $$

The _direct_ ("beam") transmittance fraction of canopy layer $i$ at solar zenith angle direction $\psi$ ($\tau(\psi)_i$) is defined as:

$$ \tau(\psi)_i = \exp\left( -K(\psi)_i L_i \right) $$

```{r int-direct, fig.cap = idcap}

idcap <- "Direct transmittance fraction, as a function of LAI"
tau_direct <- function(psi, L, orient) {
  exp(-kfunction(psi, orient) * L)
}
Lmax <- 5
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Direct tranmittance fraction")
legend("topright", c("-0.5", "0", "0.5"),
       col = c("red", "black", "blue"), lty = 1,
       title = "Orient factor")
legend("top", c("0", "30", "60"), lty = 1:3,
       title = "Zenith angle")
curve(tau_direct(0, x, 0), 0, Lmax, add = TRUE)
curve(tau_direct(0, x, -0.5), 0, Lmax, add = TRUE, col = "red")
curve(tau_direct(0, x, 0.5), 0, Lmax, add = TRUE, col = "blue")
curve(tau_direct(30, x, 0), 0, Lmax, add = TRUE, lty = 2)
curve(tau_direct(30, x, -0.5), 0, Lmax, add = TRUE, lty = 2, col = "red")
curve(tau_direct(30, x, 0.5), 0, Lmax, add = TRUE, lty = 2, col = "blue")
curve(tau_direct(60, x, 0), 0, Lmax, add = TRUE, lty = 3)
curve(tau_direct(60, x, -0.5), 0, Lmax, add = TRUE, lty = 3, col = "red")
curve(tau_direct(60, x, 0.5), 0, Lmax, add = TRUE, lty = 3, col = "blue")

```

where $L_i$ is the layer's leaf area index (technically, total area index), and $K(\psi)_i$ is the directional extinction coefficient (see "Directional scattering and leaf angle distribution").

For _diffuse_ radiation,
the interception fraction ($\tau_i$) is defined similarly except that it must be integrated across the hemisphere.
This is simplified by the assumption that incident diffuse radiation is isotropic.

$$
\tau_i = 2 \int_{0}^{\frac{\pi}{2}}
\exp \left( -K(\psi)_i L_i \right) \sin \psi \cos \psi d\psi
$$

For the $K(\psi)$ defined in "Directional scattering and leaf angle distribution", this integral has the following analytical solution:

$$
\tau_i = -\exp(-L_i (\phi_1 + \phi_2))
\left(
\phi_1^2 L_i^2 \exp(\phi_1 L_i) * E_i(-\phi_1 L_i) +
\phi_1 L_i - 1
\right)
$$

where $\phi_1$ and $\phi_2$ are defined as above, and $E_i$ is the exponential integral function.

```{r}

tau_diffuse <- function(L, orient) {
  phi1 <- phi1_f(orient)
  phi2 <- phi2_f(phi1 = phi1)
  ext_diff1 <- phi1 * L
  ext_diff2 <- phi2 * L
  -exp(-ext_diff1 - ext_diff2) *
    (ext_diff1^2 * exp(ext_diff1) * gsl::expint_Ei(-ext_diff1) +
       ext_diff1 - 1)
}
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
purrr::walk2(
  orient, colors,
  ~curve(tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y)
)

```

## Finite canopy radius

Crown area ($C_i$) is computed as a function of DBH.

$$ C_i = a DBH^b $$

where $a$ is 2.490154 and $b$ is 0.8068806 for all of the PFTs used in this study.

```{r}
dbh2ca <- function(dbh) {
  # default parameters for PFTs we use
  b1Ca <- 2.490154
  b2Ca <- 0.8068806
  b1Ca * dbh ^ b2Ca
}
curve(dbh2ca(x), 0, 50,
      xlab = "DBH (cm)",
      ylab = "Crown area index (unitless)")
```

The crown area index ($C_i$) affects the following terms in the following ways:

$$ \tau(\psi)_i = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$

$$
\tau_i = (1 - C_i) - 
C_i \exp(-\frac{L_i}{C_i} (\phi_1 + \phi_2))
\left(
\frac{\phi_1^2 L_i^2}{C_i^2} \exp(\phi_1 \frac{L_i}{C_i}) * E_i(-\phi_1 \frac{L_i}{C_i}) +
\phi_1 \frac{L_i}{C_i} - 1
\right)
$$

# Two-stream

### 2019-07-11 analysis/paper/paper.Rmd
Minor tweaks to RTM notes

@@
 
-$$ ext_1 = \phi_1 \TAI_l $$
-$$ ext_2 = \phi_2 \TAI_l $$
+$$ ext_1 = \phi_1 TAI_l $$
+$$ ext_2 = \phi_2 TAI_l $$
 $$ \tau_{diff} = (1 - CAI) - CAI exp\left( -ext_1 - ext_2 \right) ext_1^2 e^{exp_1} E_i(-exp_1) + (ext_1 - 1)$$
@@
 
-Backscatter for
+Single-scattering albedo ($w_0$):
+
+$$ w_0 = \frac{1}{2} \frac{a_{proj}}{\phi_2 \mu + a_{proj}} \frac{1 - \phi_1 \mu}{\phi_2 \mu a_{proj}}
+  \log \left( 1 + \frac{\phi_2 \mu + a_{proj}}{\phi_1 \mu} \right) $$
+
+Beam backscatter ($\gamma$):
+
+$$ \gamma = w_0 \frac{1 + \bar{\mu} \lambda}{\bar{\mu} \lambda} $$
 

### 2019-07-11 analysis/paper/paper.Rmd
Revise first 2 intro paragraphs

@@
 
-**Temperate forests in the Upper Midwest are important and changing quickly.**
-{C sequestration}.
-{Other ecosystem services}.
+Forests are a defining feature of the US North.
+This region is the most heavily forested region in the US, and these forests are an essential provider of employment and recreation opportunities and other ecosystem services for the region's people, and of habitat for the region's wildlife [@shividenko_2005_forest; @shifley_2012_forests].
+In addition, the forests of the Upper Midwest in particular have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
 However, the fate of these forests is highly uncertain.
-The natural ecological succession of these forests has been significantly altered by the interactions of a significantly different and rapidly changing climate, and novel disturbance regimes under which some natural disturbances (e.g.
-fire) are suppressed while others (e.g.
-harvest; pests and pathogens) may be becoming more frequent.
-<!-- TODO refs -->
-{UMBS ecological background/natural history?} 
-**Therefore, predictions about the future of these forests based on past analogs may not be reliable.**
-BBL: This paragraph obviously needs significant fleshing out, but this should be straightforward; I think we can basically take (with a bit of reworking) material from Chris's 2016 _Ecosphere_ paper: http://dx.doi.org/10.1002/ecs2.1375
+For one, many Upper Midwest forests are undergoing a shift in community composition from early-successional to mid-successional species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
+In addition, these forests face a multitude of direct and indirect pressures from both biotic and abiotic sources [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.
+<!-- TODO More here?
+-->
+Collectively, these processes have culminated in forests that are different from their pre-settlement climax counterparts [@stearns_2002_one; @wolter_2002_recent; @thompson_2013_four], and our ability to make predictions about such non-analog environments based on the past is limited.
 
@@
 Vegetation models fall broadly along the following spectrum of complexity.
-On one side are relatively simple **"Big leaf" models**---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples?
-BBL: no!
--->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
-<!-- TODO refs -->
-These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models (BBL: I don't think is is true; earliest models like Forest-BGC were standalone), though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
+On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
+A common application of these models was simulating the land surface boundary condition for atmospheric general circulation models, though similar models have been (and continue to be) applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
 <!-- TODO refs -->
@@
 -->
-Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
-Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
+Examples of such models are LANDIS [@scheller_2007_design] and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
+Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive]
+
 In the middle are models that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive], and have less-intensive data requirements generally.

### 2019-07-11 analysis/paper/paper.Rmd
Major revisions to introduction

--- 2019-07-11 analysis/paper/paper.Rmd
@@
 
-**Temperate forests in the Upper Midwest are important and changing quickly.**
-{C sequestration}.
-{Other ecosystem services}.
-However, the fate of these forests is highly uncertain.
-The natural ecological succession of these forests has been significantly altered by the interactions of a significantly different and rapidly changing climate, and novel disturbance regimes under which some natural disturbances (e.g.
-fire) are suppressed while others (e.g.
-harvest; pests and pathogens) may be becoming more frequent.
-<!-- TODO refs -->
-{UMBS ecological background/natural history?} 
-**Therefore, predictions about the future of these forests based on past analogs may not be reliable.**
-BBL: This paragraph obviously needs significant fleshing out, but this should be straightforward; I think we can basically take (with a bit of reworking) material from Chris's 2016 _Ecosphere_ paper: http://dx.doi.org/10.1002/ecs2.1375
+**Forests are a defining feature of the US North**.
+This region is the most heavily forested region in the US, and these forests are an essential provider of employment and recreation opportunities and other ecosystem services for the region's people, and of habitat for the region's wildlife [@shividenko_2005_forest; @shifley_2012_forests].
+In addition, the **forests of the Upper Midwest** in particular have been a **net carbon sink** for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
+However, the **fate of these forests is highly uncertain**.
+For one, many Upper Midwest forests are undergoing a **shift in community composition from early-successional to mid-successional** species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
+In addition, these forests face a multitude of **direct and indirect pressures from both biotic and abiotic sources** [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.
+<!-- TODO More here?
+-->
+Collectively, these processes have culminated in forests that are different from their pre-settlement climax counterparts [@stearns_2002_one; @wolter_2002_recent; @thompson_2013_four], and **our ability to make predictions about such non-analog environments based on the past is limited**.
 
 **More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
-Vegetation models fall broadly along the following spectrum of complexity.
-On one side are relatively simple **"Big leaf" models**---such as PNET, SiBCASA, and Biome-BGC <!-- TODO refs; more examples?
-BBL: no!
--->---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
-<!-- TODO refs -->
-These models were designed originally to simulate the land surface boundary condition for atmospheric general circulation models (BBL: I don't think is is true; earliest models like Forest-BGC were standalone), though similar models continue to be applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
+Vegetation models fall broadly along the following spectrum of complexity, which is a direct consequence of trade-offs between **realism and process fidelity** on one hand and **tractability, computational demand, and data requirements** on the other [@hawkes_2000_woody].
+<!-- TODO Might need a better ref here -->
+On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
+A common application of these models was simulating the land surface boundary condition for atmospheric general circulation models, though similar models have been (and continue to be) applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
 <!-- TODO refs -->
@@
 -->
-Examples of such models are LANDIS <!-- TODO ref !--> and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
+Examples of such models are LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
 Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
-In the middle are models that capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive], and have less-intensive data requirements generally.
-In this study, we focus on the Ecosystem Demography model [ED2.2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which constitutes an example of such an approach.
+Between these two extremes are models that use approximations attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
+One specific example of this approach is the **Ecosystem Demography model** [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
 
-BBL: General citations on simple/complex model tradeoffs: [Hawkes (2001)](http://dx.doi.org/10.1016/S0304-3800(00)00267-2), Mladenoff (2004), ...
-Moreover, these models' **predictions of species composition and ecosystem structure are relevant outside of carbon cycle science**, particularly for forest management and wildlife conservation.
-<!-- BBL: maybe move this sentence to discussion?
-TODO refs -->
-
-**Regardless of vegetation model complexity, their projections are inherently uncertain, however.** 
+**Regardless of vegetation model complexity, their projections are inherently uncertain.** 
+This uncertainty comes from many different sources, which can broadly be classified into the following categories:
 _Driver_ uncertainty refers to uncertainty in data about processes not represented by the model (e.g. weather and climate for vegetation models).
@@
 Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
-A key but unknown question in ecosystem modeling is which of the last two categories--structural versus parametric uncertainty--is more important (Fisher et al.
+**The relative importance of these sources of uncertainty in ecological forecasts is a key question because it is directly related to future research priorities**.
+For example, in the atmospheric science community, the insight that meteorological forecasts were most sensitive to initial condition uncertainty (Lorenz REF<!-- TODO others?
+-->) led to a multi-decadal research agenda aimed at constraining the initial state with improved observations, which has directly contributed to a steady and persistent improvement in meteorological forecasts over the last several decades.
+<!-- TODO Holy run-on sentences, Batman!
+--> <!-- TODO REFS!
+-->
+
+**Many ecosystem modeling studies have looked at these categories of uncertainty independently.**
+Perhaps the most work has been done on **parameter uncertainty** in ecosystem models.
+<!-- TODO Consider cutting 50-75% of these next few sentences. Probably way too much detail. -->
+@dietze_2014_quantitative investigated which parameters were most important to ED2 predictions of C sequestration across a diverse range of sites in North America.
+The found that the parameters contributing the most to overall predictive uncertainty were those related to growth respiration, mortality, stomatal conductance, and water uptake, and that parameter uncertainty varied with biome.
+More recently, @raczka_2018_what evaluated parameter uncertainty over 100 year time scale at the Willow Creek Ameriflux site, and found that the most important parameters to model uncertainty were quantum efficiency of photosynthesis, leaf respiration, and soil-plant water transfer.
+Another recent study assessed parameter uncertainty in ED2 as part of its overall goal of parameterizing the model for sagebrush [@pandit_2018_optimizing].
+They found that the most important parameters were specific leaf area, the maximum rate of carbon fixation (V~cmax~), slope of stomatal conductance, the fine-root to leaf carbon ratio, and the fine root turnover rate.
+Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large impact on predictions of net primary productivity [@viskari_2019_influence].
+{**Driver uncertainty** -- a few examples?}
+**Other studies have investigated structural uncertainty.**.
+Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization [@zaehle_2014_evaluation; @medlyn_2015_using; @walker_2015_predicting], and soil C sequestration [@sulman_2018_multiple], among others.
+In particular, @lovenduski_2017_reducing found that model structural uncertainty was the primary source of uncertainty in Earth system model projections of the land C sink.
+This structural uncertainty has been attributed to differences in model representations of key processes, including canopy radiative transfer [@fisher_2017_vegetation], soil biogeochemistry [@wieder_2017_carbon; @sulman_2018_multiple], stomatal conductance [@franks_2018_comparing], and photosynthesis [@rogers_2016_roadmap].
+**However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.**
+**Moreover, comparative analysis of contributions of different types of uncertainty to model projections are rare in the ecosystem modeling literature**.
+<!-- 
+TODO Do we have any examples? Or are these studies entirely non-existent?
+
+Ben's notes:which of the last two categories--structural versus parametric uncertainty--is more important (Fisher et al.
 2015, https://onlinelibrary.wiley.com/doi/full/10.1111/j.1523-1739.2003.00614.x, Ricciuto et al.
 2008).
-(BBL: trying to better link this paragraph into previous and following ones.)
-
-**Several recent studies have looked at parameter uncertainty in ED2.**
-@dietze_2014_quantitative found that the parameters contributing the most to ED2 model uncertainty across all biomes were those related to growth respiration, mortality, stomatal conductance, and water uptake.
-However, they also found that despite some common patterns, parameter uncertainty varied with biome.
-More recently, a similar analysis evaluated parameter uncertainty over 100 year time scale at the Willow Creek Ameriflux site, and found that the most important parameters to model uncertainty were quantum efficiency of photosynthesis, leaf respiration, and soil-plant water transfer [@raczka_2018_what].
-Another recent study assessed parameter uncertainty in ED2 as part of its overall goal of parameterizing the model for sagebrush [@pandit_2018_optimizing].
-They found that the most important parameters were specific leaf area, the maximum rate of carbon fixation (V~cmax~), slope of stomatal conductance, the fine-root to leaf carbon ratio, and the fine root turnover rate.
-Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large (>20%) impact on predictions of net primary productivity [@viskari_2019_influence].
-(BBL: probably too much detail?)
-
-**Much less work has been done on structural uncertainty.**.
-Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization (REF - Medlyn?), high-latitude warming (Rogers and Serbin?), {others...}.
-<!-- TODO -->
-These differences have been attributed to differences in model representations of key processes, including canopy radiative transfer [@fisher_2017_vegetation] and photosynthesis [@rogers_2016_roadmap].
-<!-- TODO More here?
 -->
-**However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.**
 
 **This study focuses on the interaction between parametric and structural uncertainty in the ED2 model.**
-Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
-(BBL: move to methods)
 Our study is organized around the following guiding questions:
@@
 
+<!-- TODO This would probably fit better somewhere else.  -->
+Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
+
 ## Analysis of results
@@
         - Compare ecological resilience/memory between model and reality
+
+Moreover, these models' **predictions of species composition and ecosystem structure are relevant outside of carbon cycle science**, particularly for forest management and wildlife conservation.
+<!-- BBL: maybe move this sentence to discussion?
+TODO refs -->
 

### 2019-07-11 analysis/paper/paper.Rmd
Update references, add GH and OSF links to colophon

@@
 ```
+
+View this project on GitHub in repository [ashiklom/fortebaseline](https://github.com/ashiklom/fortebaseline)
+(This manuscript file is in [`analysis/paper/paper.Rmd`](https://github.com/ashiklom/fortebaseline/blob/master/analysis/paper/paper.Rmd)).
+and in Open Science Framework (OSF) at https://osf.io/dznuf/ .

### 2019-07-11 analysis/paper/paper.Rmd
Fix capitalization and punctuation in colophon

--- 2019-07-11 analysis/paper/paper.Rmd
@@
 ```
+
+View this project on GitHub in repository [ashiklom/fortebaseline](https://github.com/ashiklom/fortebaseline)
+(this manuscript file is in [`analysis/paper/paper.Rmd`](https://github.com/ashiklom/fortebaseline/blob/master/analysis/paper/paper.Rmd))
+and in Open Science Framework (OSF) at https://osf.io/dznuf/ .

### 2019-07-11 analysis/drake.R
Add first draft of NPP vs. LAI pairs plot

@@ -107,6 +107,7 @@ plan <- drake_plan(
   models = distinct(cases, model_id, model, crown, rtm, traits) %>%
     arrange(model) %>%
     mutate(color = RColorBrewer::brewer.pal(n(), "Paired")),
+  model_colors = tibble::deframe(models[, c("model", "color")]),
   plot_means = setDT(fst(!!cohort_file)[, variable_cols])[j = .(
     agb = sum((bleaf + bsapwooda + bstorage) * nplant),
     gpp = sum(fmean_gpp_co * nplant),
@@ -168,6 +169,11 @@ plan <- drake_plan(
     "npp", 1.68, 3.11, 7.26
   ) %>% mutate(variable = factor(variable, use_vars),
                year = max(jja_summary$year)),
+  hardiman_wide = hardiman %>%
+    select(-year) %>%
+    gather(stat, value, low, mean, hi) %>%
+    unite("variable", variable, stat) %>%
+    spread(variable, value),
   summary_ts_plot = ggplot(jja_long) +
     aes(x = year) +
     geom_line(aes(y = value, group = param_id, color = color), alpha = 0.5) +
@@ -374,7 +380,33 @@ plan <- drake_plan(
     ungroup() %>%
     select(model, shortname, yvar, xvar, fpvar),
   top_params_fpvar_count = top_params_fpvar %>%
-    count(yvar, shortname, xvar, sort = TRUE)
+    count(yvar, shortname, xvar, sort = TRUE),
+  #########################################
+  # Pairs plot of time-averaged values
+  #########################################
+  pairs_time_averaged = time_averages %>%
+    ungroup() %>%
+    ggplot() +
+    aes(x = lai, y = npp, color = model) +
+    geom_point() +
+    geom_smooth(method = "lm", se = FALSE) +
+    geom_point(aes(x = lai_mean, y = npp_mean),
+               size = 3,
+               shape = 17,
+               color = "black",
+               inherit.aes = FALSE,
+               data = hardiman_wide) +
+    geom_rect(aes(xmin = lai_low, xmax = lai_hi, ymin = npp_low, ymax = npp_hi),
+              fill = NA,
+              linetype = "dashed",
+              color = "black",
+              inherit.aes = FALSE,
+              data = hardiman_wide) +
+    scale_color_manual(values = model_colors) +
+    coord_cartesian(xlim = c(0, 9), ylim = c(0, 11)) +
[…]

### 2019-07-16 analysis/paper/paper.Rmd
Add first draft of NPP vs. LAI pairs plot
Tighten the first ED2 methods paragraph

@@
 
-Forest stands at UMBS were previously exposed to experimental disturbance as part of the Forest Accelerated Succession Experiment (FASET).
-<!-- TODO ref --> BBL: what's the canonical FoRTE citation?
-Hmm.
-http://dx.doi.org/10.1111/nph.15227
+Forest stands at UMBS were previously exposed to experimental disturbance as part of the Forest Accelerated Succession Experiment (FASET). <!-- TODO ref -->
+{Canonical FoRTE citation [@curtis_2018_forest]}
 <!-- TODO more on this -->
@@
 
-For this study, we used the Ecosystem Demography Model, version 2.2 (ED2).
-Below, we provide only a high-level description of key model features and focus in detail only on submodels and configurations specific to this study.
+For this study, we used the Ecosystem Demography Model, version 2.2 (ED-2.2).
 A full description of the default configuration of the model is provided by @longo_2019_ed1.
+Briefly, ED-2.2 solves the energy, water, and carbon cycles separately for each of multiple "cohorts" of trees of similar composition, size, and age sharing a micro-environment and disturbance history [@moorcroft_2001_method; @medvigy_2009_mechanistic; @longo_2019_ed1].
+Besides the cohort-based approximation of gap dynamics, a distinctive feature of ED-2.2 is that the equations for energy, water, and carbon fluxes are defined in terms of total energy, water, and carbon, which ensures excellent conservation of mass in long-running (multidecadal) simulations [@longo_2019_ed1].
+Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in, among others,
+paleoecological studies [@rollinson_2017_emergent],
+free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration],
+estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation],
+and analyses of regional vegetation-climate feedbacks [@swann_2015_future].
 
-The key feature of the "ecosystem demography" approach underlying the ED2 model is a system of partial differential equations that simulate the emergent mean behavior of "cohorts" of individuals of similar composition, size, and age [@moorcroft_2001_method].
-<!-- TODO Run this description by Mike?
--->
-The original "Ecosystem Demography" (ED) model [@moorcroft_2001_method] was enhanced by @medvigy_2009_mechanistic (as ED2) with fully-coupled leaf physiology, canopy radiative transfer, and micrometeorology, and each of these components have been further refined over the last decade [culminating in ED2.2; @longo_2019_ed1].
-Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in:
-
-- paleoecological studies [@rollinson_2017_emergent];
-- free-air CO~2~ enrichment studies [@dekauwe_2013_forest; @miller_2015_alteration];
-- estimation of potential carbon stocks [@hurtt_2004_beyond_potential_vegetation]
-- vegetation-climate feedbacks {Swann, etc} <!-- TODO -->
-
-ED2 represents both spatial and temporal variability hierarchically.
-The largest spatial unit is the "polygon", within which all elements share the same meteorological forcings.
-Polygons are divided into multiple "sites", each site with its own set of abiotic characteristics such as soil texture and topography.
-Sites are divided into discrete "patches" according to time since last disturbance.
-Finally, each patch is composed of an arbitrary (and dynamic) number of "cohorts", representing the mean behavior of groups of plant individuals of similar size and composition.
-Temporal variability in ED2 is similarly hierarchical.
-Thermodynamics, including energy, water, and eddy (e.g. CO~2~) fluxes, are resolved on the order of seconds.
-Biophysics, including photosynthesis, respiration, and radiative transfer, and the evaluation of energy, water, and CO2 budgets, are resolved on the order of minutes.
-Phenology and plant carbon balance are evaluated daily.
-Cohort dynamics, including structural growth, mortality, and reproduction, are evaluated monthly.
-Finally, patch dynamics (e.g. annual disturbances) are evaluated annually.
-In this study, we consider only a single patch (and, therefore, a single site and polygon) starting from a "near bare ground" condition.
-
-<!-- TODO -->
-{Brief summary of thermodynamics?}
-{Brief summary of water cycle?} BBL: maybe in one sentence each, but otherwise no, send 'em to a citation.
-
-ED2 represents the carbon flux for each cohort as the net sum of contributions from turbulent mixing, leaf photosynthesis and respiration, (fine) root respiration, growth and maintenance respiration, and heterotrophic respiration.
-In addition, each cohort also has a virtual carbon balance that links fast processes like photosynthesis and respiration to slow processes like allocation (to growth, storage, and reproduction) and mortality.
-
-The official ED2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained  https://github.com/ashiklom/ED2/tree/b048950971e91699b78fc566df133caf977fda32.
+The official ED2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained at https://github.com/ashiklom/ED2/tree/b048950971e91699b78fc566df133caf977fda32.
 
@@
 -->
+
+```{r pairs-time, fig.cap = pairs_time_cap, fig.dim = c(6, 6)}
+pairs_time_cap <- paste(
+  htmlfig(),
+  "Time-averaged NPP vs. LAI, with observation (in black)",
+  "and linear regression, by model type."
+)
+readd(pairs_time_averaged)
+```
 

### 2019-07-16 analysis/paper/paper.Rmd
Update methods submodel section

@@
 
-## ED2 submodels
+## Submodels
 
-In this study, we ran a factorial combination of the following ED2 configurations: (1) two-stream vs.
+In this study, we ran a factorial combination of the following ED-2.2 configurations: (1) two-stream vs.
 multiple-scatter canopy radiative transfer models; (2) infinite vs.
@@
 
-Both canopy radiative transfer models in ED2 resolve the full vertical radiation profile within a patch as a function of canopy structure (leaf and wood area indices, crown area, leaf angle distribution) and incident solar radiation.
-The two models differ in the mathematical approximations of the underlying differential radiative transfer equations.
+Both canopy radiative transfer models in ED-2.2 resolve the full vertical radiation profile within a patch as a function of canopy structure (leaf and wood area indices, crown area, leaf angle distribution) and incident solar radiation, following the definitions in CLM 4.5 [@clm45_note].
+Both models also use identical definitions for the optical properties of a single canopy layer as a function of leaf and wood optical properties, leaf angle distribution, canopy clumping, solar zenith angle, and leaf, wood, and crown area indices.
+Both models represent direct (a.k.a.
+"beam") radiation as an exponentially decaying process, and solve for the diffuse (a.k.a.
+"hemispherical", "isotropic") radiation at each layer using a linear matrix equation.
+Where the models differ is in the terms of this matrix equation.
+The two-stream model (default) uses the two-stream approximation originally derived for atmospheric radiative transfer [@meador_1980_two] and later adapted to vegetation canopies [@dickinson_1983_land; @sellers_1985_canopy].
+Meanwhile, the multiple-scatter model was derived from first principles by @zhao_2005_multiple specifically for vegetation canopies, specifically to address known biases and limitations of the two-stream approach [e.g.
+@wang_2003_comparison].
+Both of these models are described in detail in the Supplementary Information. <!-- TODO Exact reference to Appendix -->
 
-- Two-stream model (ED default)
-    - Same general approach as in CLM 4.5 [@clm45_note], but rather than solving for just two layers (sunlit and shaded), solve for each canopy layer.
-    - Set up the two-stream equations for each layer as a linear system and solve the entire set of equations numerically
-    - Similar to approaches used in other models: JULES [@mercado_2007_improving], ORCHIDEE-CAN [@naudts_2015_vertically], CLM(SPA) [@bonan_2014_modeling]
-    - For details, see Supplements S9 and S10 in [@longo_2019_ed1].
-- Multiple scatter model
-    - Based on @zhao_2005_multiple
-
-The crown area submodel in ED2 determines the nature of competition for light between cohorts.
+The crown area submodel in ED-2.2 determines the nature of competition for light between cohorts.
 In the default configuration ("infinite crown area", or "complete shading"), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
 This means that taller cohorts always receive more incoming radiation than shorter cohorts, even when the height difference is small.
-Put another way, this means that there is no horizontal competition, only vertical.
 This has been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous canopies [@fisher_2015_taking].
@@
 | **Radiative transfer model** |                                                                                      |                           |
-| `two-stream` (default)       | Two-stream approximation, as in CLM 4.5                                              | Primary (red, blue)       |
+| `two-stream` (default)       | Two-stream approximation [@sellers_1985_canopy; @clm45_note]                         | Primary (red, blue)       |
 | `multi-scatter`              | Multiple-scatter approximation, following [@zhao_2005_multiple]                      | Secondary (green, orange) |

### 2019-07-16 analysis/paper/paper.Rmd
Minor tweaks to methods

@@
 
-## ED2 parameters
+## ED-2.2 parameters
 
-The full list of parameters in ED2 is large, with over 100 parameters per plant functional type.
+The full list of parameters in ED-2.2 is large, with over 100 parameters per plant functional type.
 A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
@@
 <!-- TODO more on this? -->
-All parameters not described in this section were set to their ED2 PFT-specific defaults.
+All parameters not described in this section were set to their ED-2.2 PFT-specific default values.
 
@@
 
-For each factorial combination of model configurations/submodels (described above), we ran 100 ensemble members from 1901 to 2000.
-Each ensemble member was initialized from ED2's "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
+For each factorial combination of ED-2.2 configurations (described above), we ran 100 ensemble members from 1901 to 2000.
+Each ensemble member was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
 Driving meteorological data was 6-hourly CRU-NCEP <!-- TODO ref.
@@
   htmlfig(),
-  "ED2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
+  "ED-2.2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
 )
@@
 Both the magnitude and direction of sensitivity to these parameters varied by model configuration.
-For example, in the default ED2 configuration (closed canopy, two-stream radiative transfer, static traits), net primary productivity increased with early hardwood specific leaf area, but when this configuration was altered so SLA and V~cmax~ could vary with light level, net primary productivity decreased with early hardwood specific leaf area.
+For example, in the default ED-2.2 configuration (closed canopy, two-stream radiative transfer, static traits), net primary productivity increased with early hardwood specific leaf area, but when this configuration was altered so SLA and V~cmax~ could vary with light level, net primary productivity decreased with early hardwood specific leaf area.
 Similarly, the model became much more sensitive to early- and mid-hardwood leaf C:N ratio when switching from a closed canopy to a finite canopy configuration.

### 2019-07-16 analysis/paper/submodel-rtm.Rmd
Add lots of info to RTM submodel description

@@
 )
+library(fortebaseline)
 ```
@@
 
-Following @clm45_note, forward- ($\nu$) and backscattering of leaves or stems ($\nu$) are defined as a function of reflectance ($R$) and transmittance ($T$) of the corresponding materials:
+Following @clm45_note, forward- ($\nu$) and backscattering of leaves or stems ($\omega$) are defined as a function of reflectance ($R$) and transmittance ($T$) of the corresponding materials:
 
@@
 gfun_cap <- "Leaf projection as a function of leaf orientation factor."
-phi1_f <- function(orient) {
-  0.5 - 0.633 * orient - 0.33 * orient ^ 2
-}
-phi2_f <- function(orient, phi1 = phi1_f(orient)) {
-  0.877 * (1 - 2 * phi1)
-}
-gfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
-  phi1 <- phi1_f(orient)
-  phi2 <- phi2_f(phi1 = phi1)
-  mu <- cos(theta_rad)
-  phi1 + phi2 * mu
-}
 max_theta <- 89
@@
 kcap <- "K function"
-kfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
-  gfunction(theta, orient, theta_rad) / cos(theta_rad)
-}
 plot(c(-70, 70), c(0.2, 1.8), type = "n",
@@
 ```{r}
-mu_bar <- function(orient) {
-  phi1 <- phi1_f(orient)
-  phi2 <- phi2_f(phi1 = phi1)
-  x <- 1 - phi1 / phi2 * log(1 + phi2 / phi1)
-  out <- x / phi2
-  out[orient == 0] <- 1
-  out
-}
 curve(mu_bar(x), -0.5, 0.5,
@@
 a_s(\psi) = \frac{1}{2}
-\frac{G(\psi)}{\cos\psi \phi2 + G(\psi)}
+\frac{G(\psi)}{\cos\psi \phi_2 + G(\psi)}
 \left(
 1 - 
-\frac{\cos\psi \phi1}{\cos\psi + G(\psi)}
+\frac{\cos\psi \phi_1}{\cos\psi + G(\psi)}
 \ln\left(
-\frac{\cos\psi \phi1 + \cos\psi \phi2 + G(\psi)}{\cos\psi \phi1}
+\frac{\cos\psi \phi_1 + \cos\psi \phi_2 + G(\psi)}{\cos\psi \phi_1}
 \right)
@@
 (Note that, for simplicity, $a_s$ here is equivalent to $\frac{a_s}{\omega}$ in @clm45_note equation 3.15, where $\omega$ is the leaf backscatter.)
+
+```{r single-scattering-albedo, fig.cap = ssa_cap}
+ssa_cap <- paste("Single-scattering albedo as a function of ",
+                 "radiation angle and leaf orientation factor.")
+plot(c(-90, 90), c(0, 1.3), type = "n",
+     xlab = expression("Zenith angle" ~ (psi * degree)),
+     ylab = expression("Single-scattering albedo" ~ (a[s])))
+curve(single_scatter_albedo(x, 0), -89, 89, add = TRUE)
+curve(single_scatter_albedo(x, 0.25), -89, 89, col = "blue", add = TRUE)
+
+curve(single_scatter_albedo(x, -0.25), -89, 89, col = "red", add = TRUE)
+legend("topright", c("-0.25", "0", "0.25"), title = "Orient factor",
+       col = c("red", "black", "blue"), lty = 1)
+```
+
+```{r beta0, fig.cap = beta0_cap}
+beta0_cap <- paste(
+  "Backscatter for direct radiation",
+  "as a function of radiation angle",
+  "and leaf orientation factor."
+)
+plot(c(-90, 90), c(0.4, 4), type = "n",
+     xlab = expression("Zenith angle" ~ (psi * degree)),
+     ylab = expression("Direct radiation backscatter" ~ (beta[0])))
+curve(beta0(x, 0), -89, 89, add = TRUE)
+curve(beta0(x, 0.25), -89, 89, col = "blue", add = TRUE)
+curve(beta0(x, -0.25), -89, 89, col = "red", add = TRUE)
+legend("topright", c("-0.25", "0", "0.25"), title = "Orient factor",
+       col = c("red", "black", "blue"), lty = 1)
+```
 
@@
 (Note that the inverse optical depth, $\mu_0$, in the two-stream calculation is just $K(\psi)^{-1}$).
+
+```{r int-direct, fig.cap = idcap}
+idcap <- "Direct transmittance fraction, as a function of LAI"
+Lmax <- 5
+plot(c(0, Lmax), c(0, 1), type = "n",
+     xlab = "Leaf area index",
+     ylab = "Direct tranmittance fraction")
+legend("topright", c("-0.5", "0", "0.5"),
+       col = c("red", "black", "blue"), lty = 1,
+       title = "Orient factor")
+legend("top", c("0", "45"), lty = 1:2,
+       title = "Zenith angle")
+curve(tau_direct(x, 0, 0), 0, Lmax, add = TRUE)
+curve(tau_direct(x, 0, -0.5), 0, Lmax, add = TRUE, col = "red")
+curve(tau_direct(x, 0, 0.5), 0, Lmax, add = TRUE, col = "blue")
+curve(tau_direct(x, 45, 0), 0, Lmax, add = TRUE, lty = 2)
+curve(tau_direct(x, 45, -0.5), 0, Lmax, add = TRUE, lty = 2, col = "red")
+curve(tau_direct(x, 45, 0.5), 0, Lmax, add = TRUE, lty = 3, col = "blue")
+```
 
@@
 
+The _direct_ ("beam") transmittance fraction of canopy layer $i$ at solar zenith angle direction $\psi$ ($\tau(\psi)_i$) is defined above.
+
 The layer absorption ($\alpha$) and backscatter for direct ($r(\psi)_i$) and diffuse ($r_i$) light are defined as follows:
@@
 $$ r(\psi)_i = \beta_0 $$
-
-The _direct_ ("beam") transmittance fraction of canopy layer $i$ at solar zenith angle direction $\psi$ ($\tau(\psi)_i$) is defined as:
-
-$$ \tau(\psi)_i = \exp\left( -K(\psi)_i L_i \right) $$
-
-```{r int-direct, fig.cap = idcap}
-
-idcap <- "Direct transmittance fraction, as a function of LAI"
-tau_direct <- function(psi, L, orient) {
-  exp(-kfunction(psi, orient) * L)
-}
-Lmax <- 5
-plot(c(0, Lmax), c(0, 1), type = "n",
-     xlab = "Leaf area index",
-     ylab = "Direct tranmittance fraction")
-legend("topright", c("-0.5", "0", "0.5"),
-       col = c("red", "black", "blue"), lty = 1,
-       title = "Orient factor")
-legend("top", c("0", "30", "60"), lty = 1:3,
-       title = "Zenith angle")
-curve(tau_direct(0, x, 0), 0, Lmax, add = TRUE)
-curve(tau_direct(0, x, -0.5), 0, Lmax, add = TRUE, col = "red")
-curve(tau_direct(0, x, 0.5), 0, Lmax, add = TRUE, col = "blue")
-curve(tau_direct(30, x, 0), 0, Lmax, add = TRUE, lty = 2)
-curve(tau_direct(30, x, -0.5), 0, Lmax, add = TRUE, lty = 2, col = "red")
-curve(tau_direct(30, x, 0.5), 0, Lmax, add = TRUE, lty = 2, col = "blue")
-curve(tau_direct(60, x, 0), 0, Lmax, add = TRUE, lty = 3)
-curve(tau_direct(60, x, -0.5), 0, Lmax, add = TRUE, lty = 3, col = "red")
-curve(tau_direct(60, x, 0.5), 0, Lmax, add = TRUE, lty = 3, col = "blue")
-
-```
-
-where $L_i$ is the layer's leaf area index (technically, total area index), and $K(\psi)_i$ is the directional extinction coefficient (see "Directional scattering and leaf angle distribution").
 
@@
 
-tau_diffuse <- function(L, orient) {
-  phi1 <- phi1_f(orient)
-  phi2 <- phi2_f(phi1 = phi1)
-  ext_diff1 <- phi1 * L
-  ext_diff2 <- phi2 * L
-  -exp(-ext_diff1 - ext_diff2) *
-    (ext_diff1^2 * exp(ext_diff1) * gsl::expint_Ei(-ext_diff1) +
-       ext_diff1 - 1)
-}
 plot(c(0, Lmax), c(0, 1), type = "n",
@@
   orient, colors,
-  ~curve(tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y)
+  ~curve(ms_tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y)
 )

### 2019-07-16 R/submodel-rtm.R

@@ -1,0 +1,79 @@
+#' Leaf orientation coefficients
+#'
+#' @param orient Leaf orientation factor (-1, 1)
+#' @export
+phi1_f <- function(orient) {
+  0.5 - 0.633 * orient - 0.33 * orient ^ 2
+}
+
+#' @rdname phi1_f
+#' @export
+phi2_f <- function(orient, phi1 = phi1_f(orient)) {
+  0.877 * (1 - 2 * phi1)
+}
+
+#' Relative (G) or absolute (K) projected leaf area in the target direction
+#'
+#' @param theta Projection angle (degrees)
+#' @param orient Leaf orientation factor (-1, 1)
+#' @param theta_rad Projection angle (radians)
+#' @export
+gfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
+  phi1 <- phi1_f(orient)
+  phi2 <- phi2_f(orient)
+  mu <- cos(theta_rad)
+  phi1 + phi2 * mu
+}
+
+#' @rdname gfunction
+#' @export
+kfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
+  gfunction(theta, orient, theta_rad) / cos(theta_rad)
+}
+
+#' Average inverse optical depth
+#'
+#' @inheritParams gfunction
+#' @export
+mu_bar <- function(orient) {
+  phi1 <- phi1_f(orient)
+  phi2 <- phi2_f(orient)
+  x <- 1 - phi1 / phi2 * log(1 + phi2 / phi1)
+  out <- x / phi2
+  out[orient == 0] <- 1
+  out
+}
+
+#' Beam backscatter
+#'
+#' @inheritParams gfunction
+#' @export
+beta0 <- function(theta, orient) {
+  K <- kfunction(theta, orient)
+  mubar <- mu_bar(orient)
+  single_scatter_albedo(theta, orient) * (1 + mubar * K) / (mubar * K)
+}
+
+#' Single-scattering albedo
+#'
+#' @inheritParams gfunction
+#' @export
+single_scatter_albedo <- function(theta, orient, theta_rad = theta * pi / 180) {
+  phi1 <- phi1_f(orient)
+  phi2 <- phi2_f(orient)
+  G <- gfunction(theta, orient)
+  cos_theta <- cos(theta_rad)
+  0.5 * G / (cos(theta_rad) * phi2 + G) *
[…]

### 2019-07-19 analysis/paper/paper.Rmd
First draft of abstract

@@
   - Alexey N. Shiklomanov
+  - Ben Bond-Lamberty
   - Jeff Atkins
-  - Ben Bond-Lamberty
   - Christopher M. Gough
@@
 bibliography: references.bib
-csl: "../templates/ecology.csl" # Insert path for the bib-style
+csl: "../templates/ecology.csl"
 abstract: |
-  Text of abstract
+  Following near-complete deforestation around the end of the 19th century, forests in the US Upper Midwest have been a net carbon sink, driven by their large fraction of fast-growing, early successional species.
+  However, the composition of these forests has recently started to shift toward mid-successional species.
+  Combined with a rapidly changing climate and novel disturbance regime, this means that future biogeochemical cycling and ecological dynamics in these forests is highly uncertain.
+  Demographic vegetation models provide a tool for analyzing this uncertainty and its implications for predictions about the future.
+  In this study, we evaluate the relative importance of two sources of model predictive uncertainty -- "structural" (which processes are represented, and the equations used to represent them) and "parameteric" (plant traits and other parameters used in those equations) -- in simulations of multidecadal carbon cycling and forest succession.
+  Specifically, we seek to answer the following research questions:
+  (1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
+  (2) What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+  To address these questions, used the Ecosystem Demography model v2.2 (ED-2.2) to simulate 100 years of forest dynamics (1900-2000) starting from bare ground at the University of Michigan Biological Station (v2.2).
+  To assess structural uncertainty, we ran ED-2.2 with different representations of key processes important to competition for light, namely crown architecture, canopy radiative transfer, and trait plasticity in shade.
+  To assess parametric uncertainty, we ran ensembles that sample over parameter uncertainties (based on existing trait data and best estimates in the literature).
+  We then compared the magnitude of structural (across representations) and parametric (within each representation) uncertainty, and assess which (if any) of the submodel-parameter combinations could reproduce the observed C fluxes and 100-year succession.
+  Our simulations agreed well with observed net primary productivity after a century of forest growth, but systematically underestimated observed leaf area index.
+  Parameter uncertainty contributed more than structural uncertainty to overall uncertainty in the simulations, but both the overall level of parameteric uncertainty and the relative importance of different parameters varied with model structure.
+  We therefore conclude that:
+  (1) targeted collection of data to constrain vegetation model parameters should be a higher priority than further model development;
+  (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and
+  (3) robust model validation requires multiple different variables, including those related to vegetation productivity, structure, and composition.
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  Dynamic vegetation modeling; carbon cycle; forest succession
 highlights: |

### 2019-07-19 analysis/paper/paper.Rmd
Fix spelling of "parametric"

@@
   Demographic vegetation models provide a tool for analyzing this uncertainty and its implications for predictions about the future.
-  In this study, we evaluate the relative importance of two sources of model predictive uncertainty -- "structural" (which processes are represented, and the equations used to represent them) and "parameteric" (plant traits and other parameters used in those equations) -- in simulations of multidecadal carbon cycling and forest succession.
+  In this study, we evaluate the relative importance of two sources of model predictive uncertainty -- "structural" (which processes are represented, and the equations used to represent them) and "paramteric" (plant traits and other parameters used in those equations) -- in simulations of multidecadal carbon cycling and forest succession.
   Specifically, we seek to answer the following research questions:
   (1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
-  (2) What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+  (2) What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
   To address these questions, used the Ecosystem Demography model v2.2 (ED-2.2) to simulate 100 years of forest dynamics (1900-2000) starting from bare ground at the University of Michigan Biological Station (v2.2).
@@
   Our simulations agreed well with observed net primary productivity after a century of forest growth, but systematically underestimated observed leaf area index.
-  Parameter uncertainty contributed more than structural uncertainty to overall uncertainty in the simulations, but both the overall level of parameteric uncertainty and the relative importance of different parameters varied with model structure.
+  Parameter uncertainty contributed more than structural uncertainty to overall uncertainty in the simulations, but both the overall level of parametric uncertainty and the relative importance of different parameters varied with model structure.
   We therefore conclude that:
@@
 (1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
-(2) What is the cost of considering these processes, in terms of additional parameteric uncertainty?
-(3) What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+(2) What is the cost of considering these processes, in terms of additional parametric uncertainty?
+(3) What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
 To answer these questions, we ran ED2 ensemble simulationes with a factorial combination of submodels related to radiative transfer formulation (two-stream vs.
@@
   htmlfig(),
-  "Comparison of parameteric uncertainty within ensembles (colored bars, colored by model type)",
+  "Comparison of parametric uncertainty within ensembles (colored bars, colored by model type)",
   "and \"structural\" uncertainty (variance in ensemble means; black bar) by output variable.",
@@
 - How much complexity do we need to accurately model ecological dynamics?
-    - Lots of existing work on parameteric uncertainty
+    - Lots of existing work on parametric uncertainty
         - [@dietze_2014_quantitative]
@@
     - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
-    - What is the cost of considering these processes, in terms of additional parameteric uncertainty?
-    - What are the relative contributions of parameteric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+    - What is the cost of considering these processes, in terms of additional parametric uncertainty?
+    - What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
 - Study overview:

### 2019-07-19 analysis/paper/paper.Rmd
Add more parameter descriptions to methods

@@
 
-These sub-models are summarized in Table X.
+These sub-models are summarized in Table 1.
 
@@
 A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
-Instead, we selected a subset of parameters that were identified as important by previous ED2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
+Instead, we selected a subset of parameters guided by (but expanding upon) previous ED2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
 
@@
 The water demand of photosynthesis is controlled by the _stomatal slope_, the sensitivity of stomatal conductance of CO~2~ as a function of CO~2~ concentration and humidity at the leaf surface [@leuning_1995_critical; @ball_1987_model].
-Three more parameters correspond to the respiration rates of leaves, roots, and "growth maintenance".
+Three more parameters correspond to the respiration rates of _leaves_, _roots_, and "_growth maintenance_".
 
@@
 
-Three parameters control various aspects of adult tree mortality, namely the density-independent mortality rate (`mort3`), and the time scale (`mort1`) and critical carbon balance (`mort2`) of mortality from carbon starvation.
-An additional parameter controls seedling mortality.
+Three parameters control various aspects of adult tree mortality.
+The overall adult mortality rate in ED-2.2 ($M$) is the sum of density-independent mortality from aging ($M_I$, plants year^-1^), density-dependent mortality from C starvation ($M_D$), and mortality from cold/frost ($M_F$) (ED-2.2 technically also includes additional term for fire mortality, but we did not include fire in our simulations):
+
+$$
+M = M_I + M_D + M_F
+$$
+
+Density-independent mortality from aging ($M_I$) is a prescribed, PFT-specific parameter (`mort3`).
+Density-dependent mortality from C starvation is calculated as a function of a cohort's C balance limitation:
+
+$$
+M_D = \frac{y_1}{1 + \exp\left[y_2 left( \frac{C_k}{C_k^*} \right)\right]}
+$$
+
+where $C_k$ is the 12 month running mean C balance of the cohort, $C_k^*$ is the running mean ideal C balance if there was no light or water limitation, and $y_1$ (`mort1`; plants year^-1^) and $y_2$ (`mort2`; unitless) are PFT-specific parameters.
+_Seedling mortality_ rate is prescribed as its own PFT-specific parameter.
+
+Several parameters are related to canopy structure and radiative transfer.
 _Specific leaf area_ (`SLA`) is used to convert leaf biomass to leaf area index, which in turn is used in a variety of calculations related to canopy radiative transfer and micrometeorology.
-
-Two additional parameters describe canopy structure in ways critical to the canopy radiative transfer:
 Canopy _clumping factor_ describes how evenly leaf area is distributed in horizontal space (1 being perfectly evenly; 0 being a "black hole" where all leaves are concentrated in a single point); and
@@
 Four parameters control leaf optical properties, namely the fractions of light reflected or transmitted in visible and near-infrared wavelengths (`leaf_(reflect|trans)_(vis|nir)`).
+Details of how these parameters influence canopy radiative transfer are described in the Supplementary Information.
+
 Three other parameters:
@@
 
-By default, ED2 supports 17 different plant functional types, which divide plant species according to photosynthetic pathway (C3 vs.
+By default, ED-2.2 supports 17 different plant functional types, which divide plant species according to photosynthetic pathway (C3 vs.
 C4), growth form (grass vs.

### 2019-07-19 analysis/paper/paper.Rmd
Minor cleanup of paper parameter methods

@@
 
+Table 1: ED-2.2 submodel descriptions
+
 | Name                         | Description                                                                          | Color                     |
@@
 | **Trait plasticity**         |                                                                                      |                           |
-| `static` (default)           | SLA and Vcmax are constant                                                           | Cool (blue, green)        |
-| `plastic`                    | SLA increases, and Vcmax decreases, with light level                                 | Warm (red, orange)        |
+| `static` (default)           | SLA and V~c,max~ are constant                                                        | Cool (blue, green)        |
+| `plastic`                    | SLA increases, and V~c,max~ decreases, with light level                              | Warm (red, orange)        |
 
-## ED-2.2 parameters
+## Parameter descriptions
 
@@
 
+## Plant functional types and parameterization
+
 By default, ED-2.2 supports 17 different plant functional types, which divide plant species according to photosynthetic pathway (C3 vs.
@@
 However, we limited our simulations to the four plant functional types that have any appreciable presence at UMBS: Early, mid, and late temperate deciduous trees and pines.
-The species comprising these plant functional types are shown in Table X.
+The species comprising these plant functional types are shown in Table 2.
 
-Table: Plant functional type-species mappings
+Table 2: Plant functional type-species mappings
 
@@
 
-For each plant functional type, we generated a distribution of parameter values via a trait-meta analysis [@lebauer_2013_facilitating; see also @dietze_2014_quantitative; @raczka_2018_what].
-Prior distributions for this meta-analysis were the same as those used in @dietze_2014_quantitative, @raczka_2018_what, and @viskari_2019_influence.
+For each plant functional type, we generated a distribution of parameter values via the Predictive Ecosystem Analyzer (PEcAn) trait-meta analysis [@lebauer_2013_facilitating; see also @dietze_2014_quantitative; @raczka_2018_what].
+Prior distributions for this meta-analysis were largely adapted from previous ED2 parameter uncertainty studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence], and are shown in detail in Supplementary Information X.
+<!-- TODO Priors supplement; maybe describe entire meta-analysis?
+-->
 Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
-The resulting parameter distributions are shown in Figure X.
+The resulting parameter distributions are shown in Figure 1.
 <!-- TODO more on this? -->

### 2019-07-19 analysis/paper/paper.Rmd
More methods parameter revisions
- Add parameter table
- Tweak text

@@
 $$
-M_D = \frac{y_1}{1 + \exp\left[y_2 left( \frac{C_k}{C_k^*} \right)\right]}
+M_D = \frac{y_1}{1 + \exp\left[y_2 \left( \frac{C_k}{C_k^*} \right)\right]}
 $$
@@
 
-Three other parameters:
+Other parameters:
 
@@
 - Soil-plant water conductance, or "water availability factor" (`water_conductance`)
+- C:N ratio in fine roots (`c2n_fineroot`) and leaves (`c2n_leaf`)
+- Leaf turnover rate (evergreen-only; `leaf_turnover_rate`)
+- Proportion of dispersal that is "global" (`nonlocal_dispersal`) <!-- TODO What is this? Can I drop it from the analysis? -->
+
+The full list of parameters is shown in Table 2.
+
+```{r param-table}
+tbl <- readd(param_table)
+tbl <- subset(tbl, Ran == "yes", select = c(-unit_parsed, -Ran))
+tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
+knitr::kable(tbl, escape = FALSE, caption = paste0(
+  "Table 2: Parameter descriptions. ",
+  "For additional information, see the ",
+  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
+))
+```
 
@@
 However, we limited our simulations to the four plant functional types that have any appreciable presence at UMBS: Early, mid, and late temperate deciduous trees and pines.
-The species comprising these plant functional types are shown in Table 2.
+The species comprising these plant functional types are shown in Table 3.
 
-Table 2: Plant functional type-species mappings
+Table 3: Plant functional type-species mappings
 
@@
 The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
-The entire model execution workflow was conducted using the Predictive Ecosystem Analyzer (PEcAn, www.pecanproject.org).
 
@@
     - Vertical profiles of radiation, traits?
+
+All analyses for this work were performed using R 3.6 (see "Colophon" for full details).
+Code and supporting data for reproducing this analysis are publicly available on the Open Science Framework (OSF) at https://osf.io/dznuf/.
 
@@
 
-```{r param-table}
-tbl <- readd(param_table)
-tbl <- subset(tbl, select = -unit_parsed)
-tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-knitr::kable(tbl, escape = FALSE, caption = paste0(
-  "**Table**: Parameter descriptions. ",
-  "For additional information, see the ",
-  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
-))
-```
-
 ##### pagebreak

### 2019-07-22 analysis/paper/paper.Rmd
Update abstract with current AGU text

@@
 abstract: |
-  Following near-complete deforestation around the end of the 19th century, forests in the US Upper Midwest have been a net carbon sink, driven by their large fraction of fast-growing, early successional species.
-  However, the composition of these forests has recently started to shift toward mid-successional species.
-  Combined with a rapidly changing climate and novel disturbance regime, this means that future biogeochemical cycling and ecological dynamics in these forests is highly uncertain.
-  Demographic vegetation models provide a tool for analyzing this uncertainty and its implications for predictions about the future.
-  In this study, we evaluate the relative importance of two sources of model predictive uncertainty -- "structural" (which processes are represented, and the equations used to represent them) and "paramteric" (plant traits and other parameters used in those equations) -- in simulations of multidecadal carbon cycling and forest succession.
-  Specifically, we seek to answer the following research questions:
-  (1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
-  (2) What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-  To address these questions, used the Ecosystem Demography model v2.2 (ED-2.2) to simulate 100 years of forest dynamics (1900-2000) starting from bare ground at the University of Michigan Biological Station (v2.2).
-  To assess structural uncertainty, we ran ED-2.2 with different representations of key processes important to competition for light, namely crown architecture, canopy radiative transfer, and trait plasticity in shade.
-  To assess parametric uncertainty, we ran ensembles that sample over parameter uncertainties (based on existing trait data and best estimates in the literature).
-  We then compared the magnitude of structural (across representations) and parametric (within each representation) uncertainty, and assess which (if any) of the submodel-parameter combinations could reproduce the observed C fluxes and 100-year succession.
-  Our simulations agreed well with observed net primary productivity after a century of forest growth, but systematically underestimated observed leaf area index.
-  Parameter uncertainty contributed more than structural uncertainty to overall uncertainty in the simulations, but both the overall level of parametric uncertainty and the relative importance of different parameters varied with model structure.
-  We therefore conclude that:
-  (1) targeted collection of data to constrain vegetation model parameters should be a higher priority than further model development;
-  (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and
-  (3) robust model validation requires multiple different variables, including those related to vegetation productivity, structure, and composition.
+  Forests in the US Upper Midwest have been a net C sink for a century, driven by their fast-growing, early successional species. 
+  However, the composition of these regrown forests has started to shift toward mid-successional species. 
+  Combined with changing climate and disturbance, this means that future biogeochemical and ecological dynamics in these forests are highly uncertain. 
+  Demographic vegetation models provide a tool for analyzing this uncertainty and its implications. 
+  Here, we evaluate the relative importance of model "structural" (which processes are represented, and how) and "parameter" (parameters used in those equations) uncertainty, asking: 
+  (1) Which processes are most important to accurately model community succession and C cycling in Upper Midwest forests?
+  (2) What are the relative contributions of parameter vs.
+structural uncertainty?
+We ran ensembles of the Ecosystem Demography model v2.2 (sampling over parameter uncertainty) from 1900 to 2000 with different representations of key processes important to competition for light.
+  We then compared the magnitude of structural (across representations) and parameter (within each representation) uncertainty, and assessed which submodel-parameter combinations reproduced observed C fluxes and succession.
+  Our simulations agreed well with observed net primary productivity but systematically underestimated leaf area index.
+  Overall, parameter uncertainty contributed more than structural uncertainty to total simulation uncertainty, but the amount of parameter uncertainty varied with model structure.
+  We conclude that: (1) targeted data collection to constrain vegetation model parameters should be a higher priority than model development; (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and (3) robust model validation requires scrutiny of multiple different variables, including those related to vegetation productivity, structure, and composition.
 keywords: |

### 2019-07-22 analysis/paper/paper.Rmd
Add observations, modify other analyses, add results text

@@
   "The solid line is the ensemble mean.",
+  "The dashed line is the 80% confidence interval.",
   "Black dot with error bars is the mean and min/max value",
@@
 The ensemble means of all model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline throughout the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
-However, the average rate of decline varied by model configuration, particularly the crown model:
+However, the average rate of decline varied by model configuration.
 Compared to models with a "closed" canopy (i.e.
-complete shading; no horizontal competition), models with a "finite" canopy predicted slightly faster decline in net primary productivity and functional diversity.
+complete shading; no horizontal competition), models with a "finite" canopy predicted faster decline in net primary productivity and functional diversity.
 This decline was most severe in configurations that combined the finite canopy and two-stream radiative transfer; in these configurations, a disproportionately large number of ensembles experienced complete ecosystem collapse over the course of the simulation, and the overwhelming fraction of ensemble members predicted near-zero Shannon diversity after 30 years.
-Besides this specific configuration, which consistently under-predicted observed net primary productivity at UMBS, the remaining model configurations had ensemble mean net primary productivity values that were very close to the mean observed value from Hardiman et al.
-(2013).
-Moreover, the ensemble spread of net primary productivity from these configurations was very similar to the uncertainty in the observations, though the distribution was relatively balanced for model predictions but skewed positive for the observations.
+<!-- TODO Update this sentence -- the NPP is just super low -->
+Besides this specific configuration, which consistently under-predicted observed net primary productivity at UMBS, the remaining model configurations had ensemble mean net primary productivity values that were very close to the values observed at UMBS.
 
 Patterns in leaf area index were somewhat distinct from patterns in productivity and diversity.
-While model configurations with finite canopy and two-stream radiative transfer again predicted the lowest leaf area index on average, model configurations with finite canopy and multiple-scatter predicted the _highest_ leaf area index on average.
-The latter configurations also had the largest number of ensemble members that fell within the observed distribution of leaf area index; meanwhile, every model configuration predicted an ensemble mean leaf area index at the end of the simulation that was significantly below the observation from Hardiman et al.
-(2013).
+While model configurations with finite canopy and two-stream radiative transfer again predicted the lowest leaf area index on average, model configurations with finite canopy and multiple-scatter predicted the _highest_ leaf area index on average (though this was skewed by a relatively small number of extremely high-LAI simulations).
+The latter configurations also had the largest number of ensemble members that fell within the observed distribution of leaf area index; meanwhile, every model configuration predicted an ensemble mean leaf area index at the end of the simulation that was significantly below the UMBS observations.
+
+```{r pairs-time, fig.cap = pairs_time_cap, fig.dim = c(8, 8)}
+pairs_time_cap <- paste(
+  htmlfig(),
+  "Post-1975 growing season averaged NPP vs. LAI, with observation (in black)",
+  "and linear regression, by model type."
+)
+readd(pairs_time_averaged)
+```
+
+In most simulations, the leaf area index was usually too low relative for a given level of net primary productivity, but the relationship between these two variables varied with model configuration (Figure 3).
+The crown model (closed vs.
+finite) had the strongest effect: Compared to the simulations with "closed" crowns, simulations with finite canopy radius generally had lower net primary productivity for the same leaf area index.
+Only a few simulations matched _both_ the observed leaf area index and net primary productivity, and all of them had the finite canopy radius configuration.
 
@@
 -->
-
-```{r pairs-time, fig.cap = pairs_time_cap, fig.dim = c(6, 6)}
-pairs_time_cap <- paste(
-  htmlfig(),
-  "Time-averaged NPP vs. LAI, with observation (in black)",
-  "and linear regression, by model type."
-)
-readd(pairs_time_averaged)
-```
 

### 2019-07-22 analysis/drake.R

@@ -184,31 +184,38 @@ plan <- drake_plan(
   my_labeller = labeller(
     model = function(x) gsub(" ", "\n", x),
     variable = as_labeller(c(
-      "gpp" = "GPP ~ (kgC ~ year^{-1})",
-      "npp" = "NPP ~ (kgC ~ year^{-1})",
+      "gpp" = "GPP ~ (MgC ~ ha^-1 ~ year^-1)",
+      "npp" = "NPP ~ (MgC ~ ha^-1 ~ year^-1)",
       "agb" = "AGB ~ (kgC)",
       "lai" = "LAI",
       "shannon" = "Shannon ~ diversity"
     ), default = label_parsed),
     .default = label_value
   ),
-  hardiman = tribble(
-    ~variable, ~low, ~mean, ~hi,
-    "lai", 1.8, 4.14, 6.56,
-    "npp", 1.68, 3.11, 7.26
+  observations = tribble(
+    ~variable, ~low, ~mean, ~hi, ~source,
+    ## "lai", 1.8, 4.14, 6.56, "Hardiman 2013",
+    ## "npp", 1.68, 3.11, 7.26, "Hardiman 2013",
+    "npp", 1.69 - 1.96 * 0.512, 1.69, 1.69 + 1.96 * 0.512, "UMBS",
+    "lai", 3.97 - 1.96 * 0.423, 3.97, 3.97 + 1.96 * 0.423, "Ameriflux"
   ) %>% mutate(variable = factor(variable, use_vars),
                year = max(jja_summary$year)),
-  hardiman_wide = hardiman %>%
-    select(-year) %>%
+  observations_wide = observations %>%
+    select(-year, -source) %>%
     gather(stat, value, low, mean, hi) %>%
     unite("variable", variable, stat) %>%
     spread(variable, value),
   summary_ts_plot = ggplot(jja_long) +
     aes(x = year) +
-    geom_line(aes(y = value, group = param_id, color = color), alpha = 0.5) +
+    geom_line(aes(y = value, group = param_id, color = color), alpha = 0.25) +
+    geom_line(data = jja_summary, aes(y = hi),
+              color = "black", linetype = "dashed") +
+    geom_line(data = jja_summary, aes(y = lo),
+              color = "black", linetype = "dashed") +
     geom_line(aes(y = avg), color = "black", size = 1, data = jja_summary) +
-    geom_pointrange(aes(y = mean, ymin = low, ymax = hi), color = "black",
[…]

### 2019-07-22 analysis/scratch/umbs-data.R

@@ -40,9 +40,9 @@ umbs_npp <- read_xlsx(umbs_npp_file, range = "A3:C21",
 umbs_npp_summary <- umbs_npp %>%
   summarize(
     mean_npp_wood = mean(npp_wood),
-    sd_npp_wood = sd(sd)
+    sd_npp_wood = mean(sd)
   ) %>%
-  mutate_all()
+  mutate_all(`/`, 1000) # Convert from kgC to MgC (ha-1 yr-1)
 
 # NPP_LAI data
 npp_lai_file <- path(data_dir, "NPP_LAI_All AF Litter Traps .xlsx")

### 2019-07-22 analysis/paper/paper.Rmd
Improve parameter subset LAI TS figure and text

@@
   "Post-1975 growing season averaged NPP vs. LAI, with observation (in black)",
-  "and linear regression, by model type."
+  "and linear regression, by model type.",
+  "Letter labels are selected runs with the same parameters,",
+  "and match the rows in Figure 4. The characteristics of these poitns are",
+  "as follows:",
+  "(A) high NPP and LAI, and relatively high pine LAI;",
+  "(B) high Shannon diversity;",
+  "(C) close to observed NPP and LAI values;",
+  "(D) high early hardwood LAI",
+  "(E) high late hardwood LAI."
 )
@@
 
-In most simulations, the leaf area index was usually too low relative for a given level of net primary productivity, but the relationship between these two variables varied with model configuration (Figure 3).
+In most simulations, the leaf area index was usually too low for a given level of net primary productivity, but the relationship between these two variables varied with model configuration (Figure 3).
 The crown model (closed vs.
@@
   htmlfig(),
-  "ED-2.2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization."
+  "PFT-level predictions of leaf area index (LAI) by model configuration and parameterization.",
+  "Each row of plots is a simulation with the same set of parameters across model configurations.",
+  "Letters correspond to the labels in Figure 3."
 )
@@
 
-Different parameter combinations led to model outputs that had not only different bulk ecosystem properties but also qualitatively different plant communities (Figure 3).
-The most common outcome was immediate and persistent dominance by early hardwood trees, though in some cases, early hardwoods co-existed with late hardwood or pine.
-<!-- TODO Probably need to do a different analysis here.
-Maybe look at all the outputs, identify several different "ecological scenarios", and then analyze the fraction of runs that experienced those probabilities?
--->
+Model structure and parameters interacted in determining the community assembly and dynamics of each simulation (Figure 4).
+Generally, the highest net primary productivity and leaf area index values (averaged across growing seasons from 1975 to 2000) occurred in runs dominated by pine.
+With the exception of the configuration with finite canopy radius and two-stream radiative transfer (under which most simulations had near-zero productivity), parameter values were more important to determining the ecological trajectory than model configuration.
+For example, all model configurations in parameter combination _E_ saw immediate and total dominance by late hardwood followed by eventual die-off.
+However, the magnitude of peak leaf area index and the pace of the subsequent die-off varied with model configuration, with a smaller peak and slower die-off under a finite canopy compared to the default closed canopy.
 

### 2019-07-22 analysis/drake.R

@@ -278,8 +278,9 @@ plan <- drake_plan(
     295, "High diversity", "B",
     160, "Close to observations", "C",
     25, "High early hardwood", "D",
-    50, "High mid hardwood", "E",
-    172, "High late hardwood", "F"
+    # No runs have particularly high mid-hardwood LAI
+    ## 50, "High mid hardwood", "E",
+    172, "High late hardwood", "E"
   ),
   lai_pft_plot = lai_q90 %>%
     inner_join(models, by = "model_id") %>%

### 2019-07-22 analysis/scratch/param-subsets.R

@@ -55,6 +55,9 @@ ta_pft %>%
   ungroup() %>%
   count(param_id, sort = TRUE)
 
+ta_pft %>%
+  filter(param_id == 50, shortname == "Mid")
+
 # Late hardwood -- 172
 ta_pft %>%
   filter(shortname == "Late") %>%

### 2019-07-23 analysis/paper/paper.Rmd
Add better sensitivity plot

@@
   "PEcAn-like parameter sensitivity and uncertainty analysis, by model type.",
-  "Elasticity (a) is the normalized sensitivity of the model to a fixed change in the parameter.",
-  "pvar (b) is the partial variance, which describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
+  "Elasticity (a, b) is the normalized sensitivity of the model to a fixed change in the parameter.",
+  "Partial variance (c, d) describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
   "Input parameter distributions are shown in Figure 1."

### 2019-07-23 R/top-n-sensitivity-plot.R

@@ -1,0 +1,32 @@
+#' Plot the top N sensitivity values
+#'
+#' @param sensitivity_plot_data Formatted sensitivity results
+#' @param y_var Response variable for which to plot sensitivity (character)
+#' @param metric Sensitivity metric to plot, as _symbol_ (e.g. `elasticiity`,
+#'   `pvar`)
+#' @param n Show this many values per panel
+#' @return `ggplot` object of plot
+#' @author Alexey Shiklomanov
+#' @export
+top_n_sensitivity_plot <- function(sensitivity_plot_data, y_var, metric,
+                                   n = 10) {
+  top_x <- sensitivity_plot_data %>%
+    group_by(model, yvar) %>%
+    top_n(n, {{metric}}) %>%
+    ungroup() %>%
+    arrange(yvar, model, {{metric}}) %>%
+    mutate(i = row_number())
+
+  top_x %>%
+    filter(yvar == y_var) %>%
+    ggplot() +
+    aes(x = i, y = {{metric}}, color = shortname) +
+    geom_segment(aes(xend = i, yend = 0)) +
+    geom_point() +
+    facet_wrap(vars(model), scales = "free_y", drop = TRUE) +
+    coord_flip() +
+    scale_x_continuous(
+      breaks = top_x$i,
+      labels = top_x$xvar
+    )
+}

### 2019-07-23 analysis/drake.R

@@ -379,39 +379,18 @@ plan <- drake_plan(
         lvls_revalue(use_vars_cap)
     ),
   sensitivity_plot_piece = target(
-    ggplot(sensitivity_plot_data) +
-      aes(x = shortname, y = trait_alpha, fill = YYY) +
-      geom_tile() +
-      facet_grid(vars(yvar), vars(model)) +
-      SCALE +
-      labs(x = "PFT", y = "Trait") +
-      theme_cowplot() +
-      theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
-            axis.text.y = element_text(size = rel(0.6)),
-            legend.position = "bottom",
-            legend.key.width = unit(3, "lines")),
-    transform = map(YYY = c(elasticity, fpvar),
-                    SCALE = c(scale_fill_gradientn(colors = c("darkred", "darkorange3",
-                                                              "white",
-                                                              "deepskyblue3", "darkblue"),
-                                                   values = scales::rescale(c(
-                                                     -75, -25,
-                                                     0,
-                                                     25, 75))),
-                              scale_fill_continuous(low = "white", high = "darkblue")))
+    top_n_sensitivity_plot(sensitivity_plot_data, .y_var, .metric) +
+      labs(x = .xlab, y = "Trait", color = "PFT") +
+      theme_cowplot(),
+    transform = map(.y_var = rep(c("NPP", "LAI"), 2),
+                    .metric = rep(c("elasticity", "pvar"), each = 2),
+                    .xlab = rep(c("Elasticity", "Partial variance"),
+                                each = 2))
   ),
   sensitivity_plot = target(
-    cowplot::plot_grid(sensitivity_plot_piece),
+    cowplot::plot_grid(sensitivity_plot_piece, nrow = 2),
     transform = combine(sensitivity_plot_piece)
   ),
-  top_params_fpvar = sensitivity_plot_data %>%
-    group_by(yvar, model) %>%
-    top_n(5, fpvar) %>%
-    arrange(desc(fpvar), .by_group = TRUE) %>%
-    ungroup() %>%
-    select(model, shortname, yvar, xvar, fpvar),
-  top_params_fpvar_count = top_params_fpvar %>%
-    count(yvar, shortname, xvar, sort = TRUE),
   #########################################
   # Pairs plot of time-averaged values
   #########################################

### 2019-07-23 analysis/scratch/misc.R

@@ -1403,5 +1403,17 @@ ggplot(ymeans) +
 #########################################
 # Cleaner sensitivity results figure
 #########################################
+top_n_sensitivity_plot(sensitivity_plot_data, "NPP", elasticity) +
+  labs(y = "Partial variance", x = "Trait") +
+  scale_color_manual(values = pfts("color")) +
+  theme_cowplot()
 
-sensitivity_plot_data
+top_n_sensitivity_plot(sensitivity_plot_data, "NPP", elasticity) +
+  labs(y = "Partial variance", x = "Trait") +
+  scale_color_manual(values = pfts("color")) +
+  theme_cowplot()
+
+top_n_sensitivity_plot(sensitivity_plot_data, "NPP", pvar) +
+  labs(y = "Partial variance", x = "Trait") +
+  scale_color_manual(values = pfts("color")) +
+  theme_cowplot()

### 2019-07-23 analysis/paper/paper.Rmd
Update first results paragraph

@@
 Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
-The ensemble means of all model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline throughout the rest of the simulation (with individual ensemble members sometimes undergoing sudden collapse in specific years).
-However, the average rate of decline varied by model configuration.
-Compared to models with a "closed" canopy (i.e.
-complete shading; no horizontal competition), models with a "finite" canopy predicted faster decline in net primary productivity and functional diversity.
-This decline was most severe in configurations that combined the finite canopy and two-stream radiative transfer; in these configurations, a disproportionately large number of ensembles experienced complete ecosystem collapse over the course of the simulation, and the overwhelming fraction of ensemble members predicted near-zero Shannon diversity after 30 years.
-<!-- TODO Update this sentence -- the NPP is just super low -->
-Besides this specific configuration, which consistently under-predicted observed net primary productivity at UMBS, the remaining model configurations had ensemble mean net primary productivity values that were very close to the values observed at UMBS.
-
-Patterns in leaf area index were somewhat distinct from patterns in productivity and diversity.
-While model configurations with finite canopy and two-stream radiative transfer again predicted the lowest leaf area index on average, model configurations with finite canopy and multiple-scatter predicted the _highest_ leaf area index on average (though this was skewed by a relatively small number of extremely high-LAI simulations).
-The latter configurations also had the largest number of ensemble members that fell within the observed distribution of leaf area index; meanwhile, every model configuration predicted an ensemble mean leaf area index at the end of the simulation that was significantly below the UMBS observations.
+One model configuration -- finite canopy radius, two-stream radiative transfer, and plastic traits -- failed to grow at all, regardless of parameter combinations, and a similar configuration with static traits showed much less growth than any other model configurations.
+The remaining configurations all exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+The ensemble mean ecosystem state of most model confugrations had net primary productivity very close to that observed at UMBS, but generally much lower leaf area index.
+However, the ensemble spread around the mean was large -- across all configurations (except the finite canopy - two-stream radiative transfer cases), the spread in net primary productivity was from zero to several times higher than the observed.
+Ensemble variance in leaf area index was lower on average but, in some case (e.g. finite canopy, multiple-scatter radiative transfer, static traits) was over an order of magnitude.
+The predicted Shannon diversity index of plant functional types was, on average, higher in simulations when the canopy architecture was finite rather than closed.
 

### 2019-07-23 analysis/paper/paper.Rmd
Write second methods paragraph

@@
   "Letter labels are selected runs with the same parameters,",
-  "and match the rows in Figure 4. The characteristics of these poitns are",
-  "as follows:",
-  "(A) high NPP and LAI, and relatively high pine LAI;",
-  "(B) high Shannon diversity;",
-  "(C) close to observed NPP and LAI values;",
-  "(D) high early hardwood LAI",
-  "(E) high late hardwood LAI."
+  "and match the rows in Figure 4."
 )
@@
 
-In most simulations, the leaf area index was usually too low for a given level of net primary productivity, but the relationship between these two variables varied with model configuration (Figure 3).
-The crown model (closed vs.
-finite) had the strongest effect: Compared to the simulations with "closed" crowns, simulations with finite canopy radius generally had lower net primary productivity for the same leaf area index.
-Only a few simulations matched _both_ the observed leaf area index and net primary productivity, and all of them had the finite canopy radius configuration.
+In most simulations, the leaf area index was too low for a given level of net primary productivity, but the relationship between these two variables varied with model configuration (Figure 3).
+The crown model (closed vs. finite) had the strongest effect:
+Compared to the simulations with "closed" crowns, simulations with finite canopy radius generally had lower net primary productivity for the same leaf area index.
+Only a few individual simulations matched _both_ the observed leaf area index and net primary productivity, and all of them had the finite canopy radius configuration.
 
@@
   "Each row of plots is a simulation with the same set of parameters across model configurations.",
-  "Letters correspond to the labels in Figure 3."
+  "Letters correspond to the labels in Figure 3.",
+  "Note that these are predictions of total leaf area index summed across all cohorts of the same PFT,",
+  "meaning that higher LAI in one PFT does not necessarily indicate dominance, size, etc."
 )
@@
 
-Model structure and parameters interacted in determining the community assembly and dynamics of each simulation (Figure 4).
-Generally, the highest net primary productivity and leaf area index values (averaged across growing seasons from 1975 to 2000) occurred in runs dominated by pine.
-With the exception of the configuration with finite canopy radius and two-stream radiative transfer (under which most simulations had near-zero productivity), parameter values were more important to determining the ecological trajectory than model configuration.
-For example, all model configurations in parameter combination _E_ saw immediate and total dominance by late hardwood followed by eventual die-off.
-However, the magnitude of peak leaf area index and the pace of the subsequent die-off varied with model configuration, with a smaller peak and slower die-off under a finite canopy compared to the default closed canopy.
+Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure 4).
+Consider simulations _A_ and _C_, the former showing competition of early hardwood with pine, and the latter completely dominated by pine:
+With a closed canopy and static traits, both simulations produced similar overall LAI and NPP values (Figure 3), but moving to a finite canopy dramatically increased the total LAI in _C_ without much affecting _A_, while enabling trait plasticity dramatically increased the NPP of _A_ without much affecting _C_ (Figure 3).
+This example also shows the impacts of different configurations on predicted community dynamics -- namely, in this specific case, enabling trait plasticity helped pine out-perform early hardwood, but only under a closed (rather than finite) canopy.
+Finite canopy radius similarly benefits early hardwood in its competition against late hardwood in simulations _B_ and, to a lesser extent, _D_.
+Simulations _B_ and _D_ also show that the sensitivity of a given parameter combination to model structure is itself dependent on the parameters -- simulation _B_ produced similar ecological dynamics and aggregate state variables regardless of model configuration (except the extreme cases; Figure 3), while simulation _D_ had near-zero NPP and LAI in two cases (closed, two-stream, static and finite, multi-scatter, plastic), comparable NPP and LAI to _B_ with closed canopy and trait plasticity, and higher NPP and LAI than _D_ with a finite canopy, multi-scatter radiative transfer, and static traits (Figure 3).
 
@@
 The relative importance of parameter uncertainty varied with both model configuration and the variable in question.
-Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) led to reduced predictive uncertainty for all four variables considered, with the largest uncertainty reduction for gross and net primary productivity.
-Meanwhile, facilitating horizontal competition by enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) substantially reduced uncertainty in aboveground biomass and diversity, and slightly reduced uncertainty in gross and net primary productivity.
-Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased uncertainty in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on uncertainty in predictions of productivity.
+Facilitating horizontal competition by enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) reduced uncertainty in aboveground biomass, diversity, and primary productivity.
+Using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased uncertainty in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on uncertainty in productivity.
+Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) had effectively no impact on simulation uncertainty (_but see caveat at the beginning of results_.)
 
-```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(15, 15)}
+```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(20, 15)}
 uncertainty_analysis_cap <- paste(

### 2019-07-23 analysis/paper/paper.Rmd
Update within vs across uncertainty paragraph

@@
 
-Across all variables, parameter uncertainty (across ensembles within each model configuration) was at least as large, and usually larger, than structural uncertainty (Figure 4).
+Except for the configurations with persistently low productivity, parameter uncertainty (across ensembles within each model configuration) was much larger than structural uncertainty (variance in ensemble means across configurations; Figure 4).
 The relative importance of parameter uncertainty varied with both model configuration and the variable in question.
-Facilitating horizontal competition by enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) reduced uncertainty in aboveground biomass, diversity, and primary productivity.
-Using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased uncertainty in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on uncertainty in productivity.
-Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) had effectively no impact on simulation uncertainty (_but see caveat at the beginning of results_.)
+Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) slightly reduced variability in primary productivity, slightly increased variability in Shannon diversity, and had mixed effects on aboveground biomass and total leaf area.
+Enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) dramatically decreased variability in primary productivity and Shannon diversity and increased variability in leaf area, and had mixed effects on aboveground biomass.
+reduced variance in aboveground biomass, diversity, and primary productivity.
+Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased variance in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on variance in productivity productivity.
 

### 2019-07-23 analysis/paper/paper.Rmd
Update sensitivity analysis code and text

@@
   htmlfig(),
-  "PEcAn-like parameter sensitivity and uncertainty analysis, by model type.",
-  "Elasticity (a, b) is the normalized sensitivity of the model to a fixed change in the parameter.",
-  "Partial variance (c, d) describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
-  "Input parameter distributions are shown in Figure 1."
+  "PEcAn-like parameter sensitivity and uncertainty analysis, by model type,",
+  "for 1975-2000 averaged growing season net primary productivity (left) and leaf area index (right).",
+  "Elasticity (top) is the normalized sensitivity of the model to a fixed change in the parameter.",
+  "Partial variance (bottom) describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
+  "Each panel shows the top 10 parameter values, in terms of the absolute value of the corresponding quantity.",
+  "Input parameter distributions are shown in Figure 1.",
+  "Note that x-axis scales vary across panels."
 )
@@
 
-Model sensitivity to specific parameters ("elasticity" _sensu_ ...) varied depending on the model configuration and the target output variable (Figure 5).
-For predicting C fluxes, biomass, and leaf area, the three most important parameters in terms of model sensitivity were specific leaf area (SLA) for early and late hardwoods, leaf C:N ratio (`c2n_leaf`) for early and mid hardwoods, and V~cmax~ for mid and late hardwoods.
-Both the magnitude and direction of sensitivity to these parameters varied by model configuration.
-For example, in the default ED-2.2 configuration (closed canopy, two-stream radiative transfer, static traits), net primary productivity increased with early hardwood specific leaf area, but when this configuration was altered so SLA and V~cmax~ could vary with light level, net primary productivity decreased with early hardwood specific leaf area.
-Similarly, the model became much more sensitive to early- and mid-hardwood leaf C:N ratio when switching from a closed canopy to a finite canopy configuration.
+Model sensitivity to specific parameters ["elasticity" _sensu_ @lebauer_2013_facilitating] varied depending on the model configuration and the target output variable (Figure 5).
+The parameter with the single highest elasticity to NPP and LAI across all model configurations was specific leaf area (SLA), though the PFT whose SLA was most important varied with model configuration.
+Other frequently (but not always) important parameters were the leaf C:N ratio (`c2n_leaf`), ratio of fine root to leaf biomass (`fineroot2leaf`), leaf and root respiration rates, and leaf optical properties (`leaf_reflect_vis/nir`).
+<!-- TODO Dig into model configurations. -->
 
-The parameters that contributed most to overall model predictive uncertainty ("partial variance") differed the parameters to which the model was most sensitive (Figure 5), largely driven by the fact that many of the parameters with the highest sensitivity (e.g.
+The parameters that contributed most to overall model predictive uncertainty ("partial variance") differed from parameters to which the model was most sensitive (Figure 5), largely driven by the fact that many of the parameters with the highest sensitivity (e.g.
 SLA, V~cmax~) were also well constrained by data (Figure 1).
-Patterns in partial variance were largely idiosyncratic across model configurations;
-when we considered the top-5 most important parameters for a given output variable and model configuration,
-only five combinations were present in at least half of the model configurations:
-growth respiration, C balance mortality ratio (`mort2`), and stomatal slope (all for early hardwoods).
-
-<!-- TODO Probably need a different analysis here more along the lines of that last sentence. -->
-
-<!--
-## Details about model structures
-
-```{r rad_prof, fig.cap = rad_barplot_cap, fig.dim = c(5, 7)}
-rad_barplot_cap <- paste(
-  "Comparison of radiation profiles under two-stream and multiple scatter",
-  "canopy radiative transfer models."
-)
-```
-
-`r mdcap(rad_barplot_cap)`
-
-Compared to the default two-stream model, the multiple scatter model leads to slightly less beam (direct) radiation in the understory but slightly more diffuse radiation.
-
-(TODO)
--->
+Leaf optical properties of at least one PFT were among the top 10 most important parameters across all configurations, and in multiple cases, were the most important.
+Other often important parameters were plant-soil water conductance (`water_conductance`), parameters related to adult mortality (`mort1,2,3`), and growth respiration.
+Again, patterns in partial variance were largely idiosyncratic across model configurations.
+<!-- TODO Dig into this more -->
+<!-- TODO Check this analysis. Look at pairs plots of trait vs. variable, and my SA spline fits. -->
 

### 2019-07-23 R/top-n-sensitivity-plot.R

@@ -9,12 +9,12 @@
 #' @author Alexey Shiklomanov
 #' @export
 top_n_sensitivity_plot <- function(sensitivity_plot_data, y_var, metric,
-                                   n = 10) {
+                                   n = 10, scales = "free_y") {
   top_x <- sensitivity_plot_data %>%
     group_by(model, yvar) %>%
     top_n(n, abs({{metric}})) %>%
     ungroup() %>%
-    arrange(yvar, model, {{metric}}) %>%
+    arrange(yvar, model, abs({{metric}})) %>%
     mutate(i = row_number())
 
   top_x %>%
@@ -23,7 +23,7 @@ top_n_sensitivity_plot <- function(sensitivity_plot_data, y_var, metric,
     aes(x = i, y = {{metric}}, color = shortname) +
     geom_segment(aes(xend = i, yend = 0)) +
     geom_point() +
-    facet_wrap(vars(model), scales = "free_y", drop = TRUE) +
+    facet_wrap(vars(model), scales = scales, drop = TRUE) +
     coord_flip() +
     scale_x_continuous(
       breaks = top_x$i,

### 2019-07-23 analysis/drake.R

@@ -370,9 +370,9 @@ plan <- drake_plan(
     unnest(raw_sa) %>%
     ungroup(),
   sensitivity_plot_data = sensitivity_results %>%
-    mutate(elasticity = if_else(abs(elasticity) > 75,
-                                75 * sign(elasticity),
-                                elasticity)) %>%
+    ## mutate(elasticity = if_else(abs(elasticity) > 75,
+    ##                             75 * sign(elasticity),
+    ##                             elasticity)) %>%
     group_by(model, yvar) %>%
     mutate(fpvar = pvar / sum(pvar)) %>%
     group_by(xvar) %>%
@@ -387,7 +387,8 @@ plan <- drake_plan(
         lvls_revalue(use_vars_cap)
     ),
   sensitivity_plot_piece = target(
-    top_n_sensitivity_plot(sensitivity_plot_data, .y_var, .metric) +
+    top_n_sensitivity_plot(sensitivity_plot_data, .y_var, .metric,
+                           scales = "free") +
       ggtitle(.y_var) +
       labs(y = .ylab, x = "Trait", color = "PFT") +
       scale_color_manual(values = pfts("color")) +

### 2019-08-23 analysis/paper/paper.Rmd
Add Gough 2019 Ecology reference
Add PDF document output

@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+    pdf_document:
+      latex_engine: xelatex
 bibliography: references.bib
@@
         - Compare ecological resilience/memory between model and reality
+- Other relevant references:
+  - @gough_2019_high -- Canopy structural complexity is the strongest predictor of productivity across NEON sites. Possible future direction?
 

### 2019-09-10 analysis/paper/paper.Rmd
Adjust results text about summary time series fi

@@
 Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
-One model configuration -- finite canopy radius, two-stream radiative transfer, and plastic traits -- failed to grow at all, regardless of parameter combinations, and a similar configuration with static traits showed much less growth than any other model configurations.
-The remaining configurations all exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
-The ensemble mean ecosystem state of most model confugrations had net primary productivity very close to that observed at UMBS, but generally much lower leaf area index.
-However, the ensemble spread around the mean was large -- across all configurations (except the finite canopy - two-stream radiative transfer cases), the spread in net primary productivity was from zero to several times higher than the observed.
-Ensemble variance in leaf area index was lower on average but, in some case (e.g. finite canopy, multiple-scatter radiative transfer, static traits) was over an order of magnitude.
+All model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+Two model configurations -- finite canopy radius and two-stream radiative transfer (regardless of trait plasticity) -- had much lower productivity, biomass, and diversity than the other configurations, which generally had similar ensemble mean values and spread.
+The ensemble means of all model configurations under-predicted observed net primary productivity and leaf area index, both at the peak of initial growth (around 1920) and at the end of the simulations, though peak predicted values were much closer to thte observations.
+However, the ensemble spread around the mean was large:
+across all configurations (except the finite canopy - two-stream radiative transfer cases), the spread in net primary productivity was from zero to nearly twice as high as the observation.
+Ensemble variance in leaf area index was lower on average but, in some cases (especially the combination of finite canopy and multiple-scatter radiative transfer) was over an order of magnitude.
 The predicted Shannon diversity index of plant functional types was, on average, higher in simulations when the canopy architecture was finite rather than closed.

### 2019-09-10 analysis/paper/paper.Rmd
Save figure files as PNG
This makes building the manuscript _much_ faster

@@
 ```{r, setup, echo = FALSE}
-knitr::opts_chunk$set(
+library(knitr)
+library(here)
+library(drake)
+
+fig <- function(...) {
+  filepath <- here("analysis", "figures", ...)
+  knitr::include_graphics(file_in(filepath))
+}
+
+opts_chunk$set(
   collapse = TRUE,
@@
 
-```{r param_dist, fig.cap = param_dist_cap, fig.dim = c(13, 10)}
-readd(param_dist_gg)
+```{r param_dist, fig.cap = param_dist_cap}
+fig("param-dist.png")
 ```
@@
 
-```{r summary-ts-plot, fig.cap = summary_ts_cap, fig.dim = c(10, 10)}
+```{r summary-ts-plot, fig.cap = summary_ts_cap}
 summary_ts_cap <- paste(
@@
   "The dashed line is the 80% confidence interval.",
-  "Black dot with error bars is the mean and min/max value",
-  "from Hardiman et al. (2013)."
+  "Black dot with error bars is the observed mean and min/max value",
+  "(TODO REF)."
 )
-readd(summary_ts_plot)
+fig("summary-ts-plot.png")
 ```
@@
 
-```{r pairs-time, fig.cap = pairs_time_cap, fig.dim = c(8, 8)}
+```{r pairs-time, fig.cap = pairs_time_cap}
 pairs_time_cap <- paste(
@@
 )
-readd(pairs_time_averaged)
+fig("pairs-time-averaged.png")
 ```
 
-In most simulations, the leaf area index was too low for a given level of net primary productivity, but the relationship between these two variables varied with model configuration (Figure 3).
 The crown model (closed vs. finite) had the strongest effect:
@@
 
-```{r bigfacet-plot, fig.cap = bigfacet_cap, fig.dim = c(12, 8)}
+```{r bigfacet-plot, fig.cap = bigfacet_cap}
 bigfacet_cap <- paste(
@@
 )
-readd(lai_pft_plot)
+fig("lai-pft-plot.png")
 ```
@@
 )
-readd(within_across_plot)
+fig("within-across-plot.png")
 ```
@@
 
-```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, fig.dim = c(20, 15)}
+```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap}
 uncertainty_analysis_cap <- paste(
@@
 )
-readd(sensitivity_plot)
+fig("sensitivity-plot.png")
 ```

### 2019-09-10 analysis/drake.R

@@ -10,7 +10,6 @@ if (getRversion() >= "3.6") {
                 mask.ok = c("as.difftime", "date"),
                 exclude = c("intersect", "setdiff", "union", "here",
                             "stamp"))
-  conflictRules("ggplot2", exclude = "ggsave")
   conflictRules("data.table", exclude = c("between", "first", "last",
                                           "transpose", "hour", "isoweek",
                                           "mday", "minute", "month",
@@ -53,6 +52,7 @@ cmdargs <- commandArgs(trailingOnly = TRUE)
 analysis_dir <- dir_create(here("analysis"))
 data_dir <- dir_create(path(analysis_dir, "data"))
 download_dir <- dir_create(path(data_dir, "retrieved"))
+fig_dir <- dir_create(path(analysis_dir, "figures"))
 
 # "Observation time" output files
 cohort_file <- path(download_dir, "all-output-cohort.fst")
@@ -220,6 +220,12 @@ plan <- drake_plan(
           axis.text.x = element_text(angle = 90, vjust = 0.5),
           strip.placement = "outside",
           strip.background = element_blank()),
+  summary_ts_plot_png = ggsave(
+    file_out(!!path(fig_dir, "summary-ts-plot.png")),
+    summary_ts_plot,
+    # TODO: Fix dims so axis titles aren't cut off
+    width = 10, height = 10
+  ),
   #########################################
   # Parameter distributions
   #########################################
@@ -237,7 +243,13 @@ plan <- drake_plan(
     labs(x = "PFT", fill = "PFT") +
     theme_cowplot() +
     theme(axis.title.y = element_blank(),
-          axis.text.x = element_blank()),
+          axis.text.x = element_blank(),
+          legend.position = c(0.7, 0.1)),
+  param_dist_png = ggsave(
+    file_out(!!path(fig_dir, "param-dist.png")),
+    param_dist_gg,
+    width = 15.3, height = 9.9
+  ),
   #########################################
   # LAI by PFT plot
   #########################################
@@ -300,6 +312,11 @@ plan <- drake_plan(
       axis.title.x = element_blank(),
       axis.text.x = element_text(angle = 90, vjust = 0.5)
     ),
+  lai_pft_plot_png = ggsave(
+    file_out(!!path(fig_dir, "lai-pft-plot.png")),
+    lai_pft_plot,
+    width = 11.42, height = 5.67
+  ),
   #########################################
[…]

### 2019-09-10 analysis/paper/paper.Rmd
Tweak Figure 3 results text

@@
   "and linear regression, by model type.",
+  "Each point indicates a single ensemble run with a different parameter combination.",
   "Letter labels are selected runs with the same parameters,",
@@
 
+Model configuration affected the emergent relationship between net primary productivity and leaf area index (Figure 3).
 The crown model (closed vs. finite) had the strongest effect:
 Compared to the simulations with "closed" crowns, simulations with finite canopy radius generally had lower net primary productivity for the same leaf area index.
-Only a few individual simulations matched _both_ the observed leaf area index and net primary productivity, and all of them had the finite canopy radius configuration.
+Although many simulations matched _either_ the observed net primary productivity or the observed leaf area index, only a few individual runs matched _both_.
 

### 2019-09-10 analysis/paper/paper.Rmd
Use bold face for letters in Figure 4 results text

@@
 Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure 4).
-Consider simulations _A_ and _C_, the former showing competition of early hardwood with pine, and the latter completely dominated by pine:
-With a closed canopy and static traits, both simulations produced similar overall LAI and NPP values (Figure 3), but moving to a finite canopy dramatically increased the total LAI in _C_ without much affecting _A_, while enabling trait plasticity dramatically increased the NPP of _A_ without much affecting _C_ (Figure 3).
+Consider simulations **A** and **C**, the former showing competition of early hardwood with pine, and the latter completely dominated by pine:
+With a closed canopy and static traits, both simulations produced similar overall LAI and NPP values (Figure 3), but moving to a finite canopy dramatically increased the total LAI in **C** without much affecting **A**, while enabling trait plasticity dramatically increased the NPP of **A** without much affecting **C** (Figure 3).
 This example also shows the impacts of different configurations on predicted community dynamics -- namely, in this specific case, enabling trait plasticity helped pine out-perform early hardwood, but only under a closed (rather than finite) canopy.
-Finite canopy radius similarly benefits early hardwood in its competition against late hardwood in simulations _B_ and, to a lesser extent, _D_.
-Simulations _B_ and _D_ also show that the sensitivity of a given parameter combination to model structure is itself dependent on the parameters -- simulation _B_ produced similar ecological dynamics and aggregate state variables regardless of model configuration (except the extreme cases; Figure 3), while simulation _D_ had near-zero NPP and LAI in two cases (closed, two-stream, static and finite, multi-scatter, plastic), comparable NPP and LAI to _B_ with closed canopy and trait plasticity, and higher NPP and LAI than _D_ with a finite canopy, multi-scatter radiative transfer, and static traits (Figure 3).
+Finite canopy radius similarly benefits early hardwood in its competition against late hardwood in simulations **B** and, to a lesser extent, **D**.
+Simulations **B** and **D** also show that the sensitivity of a given parameter combination to model structure is itself dependent on the parameters -- simulation **B** produced similar ecological dynamics and aggregate state variables regardless of model configuration (except the extreme cases; Figure 3), while simulation **D** had near-zero NPP and LAI in two cases (closed, two-stream, static and finite, multi-scatter, plastic), comparable NPP and LAI to **B** with closed canopy and trait plasticity, and higher NPP and LAI than **D** with a finite canopy, multi-scatter radiative transfer, and static traits (Figure 3).
 

### 2019-09-10 analysis/paper/paper.Rmd
Remove FASET info from methods
Remove old outline notes from paper

@@
 The majority of the forest that is aspen-dominated is undergoing succession to "northern hardwood" (maple, beech, basswood, birch, ash, hemlock), "upland conifer" (red and white pine), or "northern red-oak" ecotypes [@bergen_2007_observing].
-
-Forest stands at UMBS were previously exposed to experimental disturbance as part of the Forest Accelerated Succession Experiment (FASET). <!-- TODO ref -->
-{Canonical FoRTE citation [@curtis_2018_forest]}
-<!-- TODO more on this -->
 
@@
 
-# Appendix 1: Submodel descriptions
-
-## Radiative transfer: Definitions
-
-- $\omega$ -- Clumping factor
-- $LAI$ -- Leaf area index
-- $TAI$ -- Total area index (leaf and woody area)
-- $LAI_e$ / $TAI_e$ -- "Effective" area indices (accounting for clumping)
-
-## Two-stream RTM
-
-<!-- TODO -->
-
-## Multiple scatter RTM
-
-Effective LAI is just true LAI x clumping.
-
-$$ LAI_e = \omega LAI $$
-
-Total area index (TAI) is leaf (LAI) + wood (WAI) area.
-
-$$ TAI_e = LAI_e + WAI $$
-
-Local exposed total area index ($TAI_l$) is the total area index divided by crown area index ($CAI$).
-
-$$ TAI_{l} = \frac{TAI_e}{CAI} $$
-
-Leaf ($w_l$) and wood ($w_w$) weights:
-
-$$ w_l = \frac{LAI_e}{TAI_e} $$
-$$ w_w = 1 - w_l $$
-
-Projected area ($a_{proj}$), based on coefficients $\phi_1$ and $\phi_2$.
-
-$$ a_{proj} = \phi_1 + \phi_2 \mu $$
-
-Optical depth ($\lambda$) is projected area over optical depth ($\mu$)
-
-$$ \lambda = \frac{a_{proj}}{\mu} $$
-
-Transmittance for direct radiation of a canopy layer ($\tau_{beam}$) is:
-
-$$ \tau_{beam} = (1 - CAI) + CAI exp\left( -\lambda TAI_l \right) $$
-
-Transmittance for diffuse radiation of a canopy layer ($\tau_{diff}$):
-
-$$ ext_1 = \phi_1 TAI_l $$
-$$ ext_2 = \phi_2 TAI_l $$
-$$ \tau_{diff} = (1 - CAI) - CAI exp\left( -ext_1 - ext_2 \right) ext_1^2 e^{exp_1} E_i(-exp_1) + (ext_1 - 1)$$
-
-where $E_i$ is the exponential integral.
-
-Single-scattering albedo ($w_0$):
-
-$$ w_0 = \frac{1}{2} \frac{a_{proj}}{\phi_2 \mu + a_{proj}} \frac{1 - \phi_1 \mu}{\phi_2 \mu a_{proj}}
-  \log \left( 1 + \frac{\phi_2 \mu + a_{proj}}{\phi_1 \mu} \right) $$
-
-Beam backscatter ($\gamma$):
-
-$$ \gamma = w_0 \frac{1 + \bar{\mu} \lambda}{\bar{\mu} \lambda} $$
-
-# Old outline
-
-## Introduction
-
-- Why we need models?
-    - Forests are important, for C sequestration, other ecosystem services
-    - Policy decisions related to forests depend on future projections
-    - We need models to make future projections
-- What kinds of models are out there?
-    - Traditional: "Big leaf" models
-        - Needed to be simple to run on old computers
-        - Don't need as much data
-        - ...but they fail to capture important processes, which matter not only for accurate predictions about C cycle, but also because they are processes we care about (habitat, forestry, tourism, etc.)!
-    - Individual-based models
-        - Much more detailed, but harder to run, and need much more data.
-        - Trade-off between complexity and accuracy.
-- Cohort-based models like ED2 as a middle ground?
-    - Core ED concept [@moorcroft_2001_method]
-    - ED2 developments -- [@medvigy_2009_mechanistic; @medvigy_2011_predicting]
-    - Most recent version -- [@longo_2019_ed1]
-    - Validation/applications of the ED model:
-        - [@rollinson_2017_emergent] -- ED2 paleoclimate
-        - [@miller_2015_alteration] -- ED2 at Duke FACE
-        - [@dekauwe_2013_forest] -- ED2 at Duke and Oak Ridge FACE
-        - [@longo_2019_ed2] -- Model evaluation in Tropics
-- How much complexity do we need to accurately model ecological dynamics?
-    - Lots of existing work on parametric uncertainty
-        - [@dietze_2014_quantitative]
-        - [@raczka_2018_what]
-        - [@pandit_2018_optimizing]
-    - Less on structural uncertainty -- i.e. not just what coefficients do we use, but what _equations_, and more generally, what _processes_ do we even consider?
-- What are the processes that drive succession of ecological communities?
-    - Competition for light [@bazzaz_1979_physiological]
-    - Others? Leaning towards omitting because factorial combination becomes too large.
-        - Competition for water?
-        - Competition for nutrients? (ED's N model isn't very good. Alternatives are being developed, but not in ED mainline yet.)
-- Brief review of community/ecosystem ecology of the Upper Midwest
-    - UMBS is a useful site for testing these theories because of its natural history.
-    - Around 1900, mature forest dominated by white (_Pinus strobus_) and red pine (_Pinus resinosa_), hemlock (_Tsuga canadensis_), and sugar maple (_Acer saccharum_).
-    - But, repeat harvest and burning between 1880 and 1920 [@kilburn_1960_effects] led to replacement by early successional aspen (_Populus grandidentata_ and _Populus tremuloides_), birch (_Betula papyrifera_), and cherry (_Prunus pensylvatica_) [@gates_1930_aspen; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging; @bergen_2007_observing].
-    - Such aspen-dominated secondary forests are sustained by background levels of disturbance, particularly fire [@gates_1930_aspen]
-    - However, due to aggressive fire suppression in the US, these forests are now undergoing succession, primarily to mid/late-successional hardwood (typified by sugar maple and American beech, _Fagus grandifolia_) or upland conifer (typified by red and white pine) [@bergen_2007_observing].
-    - The combination of stand-replacing disturbance ~100 years ago and suppressed disturbance since then makes this an ideal site for testing a vegetation model run from bare ground without having to prescribe disturbance.
-- Guiding questions:
-    - Which of these processes do we need to be able to accurately model what happens ("community succession and C cycle dynamics") to Midwestern forests?
-    - What is the cost of considering these processes, in terms of additional parametric uncertainty?
-    - What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-- Study overview:
-    - ED with factorial combination of submodels related to radiative transfer formulation (two-stream vs.
-multiple scatter), horizontal competition (finite canopy radius vs.
-complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
-    - Ensemble analysis to look at parameter uncertainty
-
-$$ \lambda = \frac{a_{proj}}{\mu} $$
-
-## Methods
-
-### Site description
-
-- Secondary successional forest -- bigtooth aspen (_Populus grandidentata_), northern red oak (_Quercus rubra_), red maple (_Acer rubrum_), paper birch (_Betula papyrifera_), eastern white pine (_Pinus strobus_)
-- Average age -- 95 years (2013)
-- NEP in UMBS tower is 1.58 MG C ha^-1^ yr^-1^ (0.80-1.98) (1999-2006), but with landscape variation (REF?)
-- Heavily logged in late 1800s and early 1900s, disturbed by fire until 1923 (REF)
-- Present day composition typical of upper Great Lakes region
-- Previously, the site of the FASET experiment [@gough_2013_sustained; @gough_2008_multi]
-- UMBS is 87% upland and 13% wetland [@bergen_2007_observing]. We focus on the upland ecosystem, which has the following characteristics (from @bergen_2007_observing):
-    - 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, 1.2% lowland glacial lake
-    - 60.9% aspen-dominated, 16.6% northern hardwood (_Acer saccharum_, _Acer rubrum_, _Fagus grandifolia_, _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_), 13.3% upland conifer (_Pinus strobus_, _Pinus resinosa_)
-    - 1.6% northern red oak (_Quercus rubra_)
-    - 60.9% transitioning from aspen to next succession, of which: 55.8% to northern hardwood, 34.7% to upland conifer, 3.8% to northern red oak, 2.6% to northern red oak-red maple, 3.1% to upland conifer-northern hardwood
-
-### PFT definitions
-
-- **Early successional hardwood**
-    - _Betula papyrifera_ (white/paper birch)
-    - _Populus grandidentata_ (bigtooth aspen)
-    - _Populus tremuloides_ (trembling aspen)
-- **Mid successional hardwood**
-    - _Quercus rubra_ (red oak)
-    - _Acer rubrum_ (red maple)
-    - _Acer pensylvaticum_ (striped maple)
-- **Late successional hardwood**
-    - _Acer saccharum_ (sugar maple)
-    - _Fagus grandifolia_ (American beech)
-- **Pine**
-    - _Pinus strobus_ (white pine)
-- _Omitted candidates_
-    - Early hardwood
-        - _Prunus pensylvatica_ (fire cherry)
-    - Mid hardwood
-        - _Betula alleghaniensis_ (yellow birch)
-        - _Tilia americana_ (basswood)
-        - _Fraxinus americana_ (white ash)
-    - Late hardwood
-        - _Ostyra virginiana_ (hop-hornbeam)
-    - Pine
-        - _Pinus resinosa_ (red pine)
-
 ##### pagebreak

### 2019-09-10 analysis/paper/submodel-rtm.Rmd
Finish MS and 2S RTM code

@@
 ---
-title: "Submodel descriptions: Radiative transfer"
+title: "Description of radiative transfer submodels"
 author: Alexey Shiklomanov
+bibliography: references.bib
 ---
@@
 # Two-stream
+
+# Comparison
+
+First, we generate some common inputs:
+
+```{r, echo = TRUE}
+ipft <- c(2, 1, 1)
+L <- c(1, 1.5, 2)
+p4 <- rrtm::prospect4(1.4, c(40, 60), 0.01, 0.01)
+vis <- p4[(400:700) - 399, , ]
+lr <- colMeans(vis[,,1])
+lt <- colMeans(vis[,,2])
+
+orient <- c(0.2, 0.25)
+
+theta <- 15
+S0_beam <- 0.8
+S0_diff <- 1 - S0_beam
+alb_ground <- 0.1
+
+```
+
+Then we run both models.
+
+```{r}
+ms_out <- multi_scatter(ipft, L, lr, lt, orient,
+                        theta, S0_beam, S0_diff, alb_ground)
+ts_out <- two_stream(ipft, L, lr, lt, orient,
+                     theta, S0_beam, S0_diff, alb_ground)
+
+plot(ms_out$light_level, type = "l")
+lines(ts_out$light_level, col = 2)
+
+plot(ms_out$light_diff_level, type = "l")
+lines(ts_out$light_diff_level, col = 2)
+```
+
+# References

### 2019-09-10 R/submodel-rtm-2s.R

@@ -1,0 +1,219 @@
+#' R implementation of ED2 two-stream radiation model
+#'
+#' @param czen Cosine of angle of incidence (`cosaio`)
+#' @param iota_g Ground (soil + snow) albedo
+#' @param pft PFT identities of each cohort, as integer (ncoh)
+#' @param lai Leaf area index of each cohort (ncoh)
+#' @param wai Wood area index of each cohort (ncoh)
+#' @param cai Crown area of each cohort (ncoh)
+#' @param orient_factor Orient factor (npft)
+#' @param clumping_factor Clumping factor (npft)
+#' @param leaf_reflect Leaf reflectance spectra (nwl * npft)
+#' @param leaf_trans Leaf transmittance spectra (nwl * npft)
+#' @param wood_reflect Wood reflectance spectra (nwl * npft)
+#' @param wood_trans Wood transmittance spectra (nwl * npft)
+#' @param down_sky Normalized diffuse solar spectrum (nwl)
+#' @param down0_sky Normalized direct solar spectrum (nwl)
+#' @param wavelengths Numeric vector of wavelengths to use, in nm
+#'   (nwl). Default is 400:2500.
+#' @return List of outputs
+#' @author Alexey Shiklomanov
+#' @export
+two_stream <- function(ipft, L,
+                       lr, lt, orient,
+                       theta = 15,
+                       S0_beam = 0.8,
+                       S0_diff = 1 - S0_beam,
+                       alb_ground = 0.1) {
+
+  # Cohort variables
+  # Order: bottom -> Top
+  ncoh <- length(ipft)
+  npft <- length(lr)
+  stopifnot(
+    length(L) == ncoh,
+    length(lt) == npft,
+    length(orient) == npft,
+    max(ipft) <= npft
+  )
+
+  # Unpack PFT variables
+  orient <- orient[ipft]
+  lr <- lr[ipft]
+  lt <- lt[ipft]
+
+  # Calculations from sfc_rad
+  leaf_scatter <- lr + lt
[…]

### 2019-09-10 R/submodel-rtm-ms.R

@@ -11,26 +11,42 @@ ms_tau_diffuse <- function(L, orient) {
     (ext_diff1^2 * exp(ext_diff1) * gsl::expint_Ei(-ext_diff1) + ext_diff1 - 1)
 }
 
-if (FALSE) {
+#' Multiple-scatter model
+#'
+#' Cohorts are arranged from bottom (`i = 1`) to top (`i = ncohort`)
+#'
+#' @param ipft Integer indices of PFTs (`icohort`)
+#' @param lr Leaf reflectance (`ipft`)
+#' @param lt Leaf transmittance (`ipft`)
+#' @return List containin cohort-level downward beam radiation, and downward and
+#'   upward diffuse radiation.
+#' @inheritParams tau_direct
+#' @author Alexey Shiklomanov
+#' @export
+multi_scatter <- function(ipft, L,
+                          # PFT-level parameters
+                          lr, lt, orient,
+                          # Site scalars
+                          theta = 15,
+                          S0_beam = 0.8,
+                          S0_diff = 1 - S0_beam,
+                          alb_ground = 0.1) {
 
   # Cohort variables
   # Order: bottom -> Top
-  ipft <- c(2, 1, 1)
-  L <- c(1, 1.5, 2)
   ncoh <- length(ipft)
+  npft <- length(lr)
+  stopifnot(
+    length(L) == ncoh,
+    length(lt) == npft,
+    length(orient) == npft,
+    max(ipft) <= npft
+  )
 
-  # PFT variables
-  p4 <- rrtm::prospect4(1.4, c(40, 60), 0.01, 0.01)
-  vis <- p4[(400:700) - 399, , ]
-  lr <- colMeans(vis[, , 1])[ipft]
-  lt <- colMeans(vis[, , 2])[ipft]
-  orient <- c(0.2, 0.1)[ipft]
-
-  # Site scalars
-  theta <- 15
-  S0_beam <- 0.8
-  S0_diff <- 1 - S0_beam
-  alb_ground <- 0.1
+  # Convert to PFT variables
+  orient <- orient[ipft]
[…]

### 2019-09-10 R/submodel-rtm.R

@@ -77,3 +77,32 @@ tau_direct <- function(L, theta, orient) {
   K <- kfunction(theta, orient)
   exp(-K * L)
 }
+
+#' Integrate light levels from flux estimates
+#'
+#' @param down0 Cohort-level direct downwelling flux
+#' @param down Cohort-level diffuse downwelling flux
+#' @param up Cohort-level diffuse upwelling flux
+#' @return List containing overall light level, and levels of direct and diffuse light
+#' @author Alexey Shiklomanov
+#' @export
+integrate_light <- function(down0, down, up) {
+  # Integrate light levels
+  ncoh <- length(down0) - 1
+  light_level <- numeric(ncoh)
+  light_beam_level <- numeric(ncoh)
+  light_diff_level <- numeric(ncoh)
+
+  for (k in seq_len(ncoh)) {
+    kp1 <- k + 1
+    light_level[k] <- light_level[k] + 0.5 * (down[k] + down[kp1] + down0[k] + down0[kp1])
+    light_beam_level[k] <- light_beam_level[k] + 0.5 * (down0[k] + down0[kp1])
+    light_diff_level[k] <- light_diff_level[k] + 0.5 * (down[k] + down[kp1])
+  }
+
+  list(
+    light_level = light_level,
+    light_beam_level = light_beam_level,
+    light_diff_level = light_diff_level
+  )
+}

### 2019-09-10 inst/ms-julia.jl

@@ -1,0 +1,21 @@
+ipft = [2 1 1]
+L = [1 1.5 2]
+
+ncoh = length(ipft)
+
+lr = [0.045 0.056 0.056]
+lt = [0.025 0.046 0.046]
+orient = [0.2 0.1][ipft]
+
+theta = 15
+S0_beam = 0.8
+S0_diff = 1 - S0_beam
+alb_ground = 0.1
+
+nu = lr + lt
+omega = nu .+ 0.25 .* (lr .- lt) .* (1 .- orient) .^ 2 ./ (2 .* nu)
+
+phi1 = ...
+phi2 = ...
+
+a = [1 - alb_ground ]

### 2019-09-12 analysis/paper/paper.Rmd
Add very janky intro paragraph about structure

@@
 -->
+
+<!-- TODO This paragraph is janky as hell. Needs heavy revision. -->
+Previous work found that an earlier version of ED had difficulty sustaining multiple plant functional types in long simulations [@fisher_2010_assessing].
+One of the primary drivers of successional change in temperate forests is competition for light <!-- TODO REF -- Bazaaz?
+-->, and differences in plants' strategies for light acquisition and utilization are a key driver of niche differentiation in these ecosystems <!-- TODO REF -->.
+To this end, we focused our structural uncertainty analysis on three processes related to competition for light:
+Representation of horizontal competition;
+representation of canopy radiative transfer (especially of diffuse radiation);
+and trait plasticity with light level.
 

### 2019-09-12 analysis/paper/paper.Rmd
Fix typo in introduction

@@
 (3) What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-To answer these questions, we ran ED2 ensemble simulationes with a factorial combination of submodels related to radiative transfer formulation (two-stream vs.
+To answer these questions, we ran ED2 ensemble simulations with a factorial combination of submodels related to radiative transfer formulation (two-stream vs.
 multiple scatter), horizontal competition (finite canopy radius vs.

### 2019-09-12 analysis/paper/paper.Rmd
Add possible TODOs about Fifugre 3
Draft first discussion paragraph

@@
 
+<!-- TODO Use peak instead of post-1975? -->
+<!-- TODO Use a logarithmic, rather than a linear fit? -->
+
 Model configuration affected the emergent relationship between net primary productivity and leaf area index (Figure 3).
@@
 
+The overarching objective of this study was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community assembly.
+
+## ED2.2 vs. reality
+
+Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
+**Ensemble mean trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure 2);
+real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
+This idea that ED2.2's main issue is that it is moving too quickly is supported by the fact that ED2.2 generally **underestimates productivity after 100 years**, but model **predictions at peak productivity are close to reality**.
+One reason ED2.2 is moving too quickly through succession is **missing mechanisms for sustaining productivity through early succession**.
+In particular, recent evidence points to the **important role of low-severity, high-frequency disturbances** in sustaining ecosystem productivity [@gough_2016_disturbance].
+ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the top 10 most important parameters for predictive uncertainty (Figure 6).
+<!-- Maybe move this elsewhere? -->
+<!-- TODO Look at mortality across ED runs. Does it relate at all to results? -->
+However, **our simulations did not include other sources of disturbance**, such as partial harvests, extreme weather, and damage from insects and pathogens, which are common to this region [@amiro_2010_ecosystem] and which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
 - Parameter uncertainty is the dominant driver of model predictive uncertainty.

### 2019-09-12 analysis/paper/chunks.Rmd
Move old discussion notes to chunks.Rmd

@@
 Extent of recovery depends on balance of gross (aboveground?) primary productivity (GPP) and respiration; GPP increase during recovery, while respiration may stay constant [@amiro_2010_ecosystem].
+
+## Old discussion notes
+
+- Parameter uncertainty is the dominant driver of model predictive uncertainty.
+- ...but the nature of this parameter uncertainty depends on model configuration.
+
+- Ecological implications of structural uncertainty
+    - Predictions about ecological succession are very sensitive to the understory light environment.
+        - Need a good representation of canopy structure (finite canopy radius)
+        - How you model canopy radiative transfer matters! Small differences in the model can give significant results.
+    - But, trait plasticity can provide an alternative mechanism for success in the understory.
+        - Important modeling implication for all the emerging research on intra-specific trait variability.
+        - This combination of discrete functional types with plastic traits is closer to reality than purely trait-based ("eco-evolutionary?") models or completely discretized models with static traits on the other.
+        - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
+        - Perhaps note "ecological equifinality" here: there are different ways to arrive at similar outcomes in the model; this is also true in nature? 
+- Implications of uncertainty analysis for future data collection and model development
+    - Looking beyond aggregate variables (e.g. net primary productivity, total LAI) is essential to model evaluation and calibration. It's possible to generate a relatively?
+    - (Expected) Subsurface parameters (e.g. leaf-root ratios) have high sensitivity, but are poorly constrained.
+    - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
+- Comparisons against other studies
+    - Parameter sensitivity
+        - [@pandit_2018_optimizing] -- Most important parameters for shrubs were SLA, Vcmax, stomatal slope, root:leaf ratio, and root turonver rate (slightly less important: leaf turnover rate, growth respiration, cuticular conductance; much less important: water conductance; basically irrelevant: storage turnover, leaf width)
+        - Compare against @dietze_2014_quantitative and @raczka_2018_what
+- Other relevant references:
+  - @gough_2019_high -- Canopy structural complexity is the strongest predictor of productivity across NEON sites. Possible future direction?

### 2019-09-12 analysis/paper/paper.Rmd
Use here::here for git2r::repository
Move old discussion notes to chunks.Rmd

@@
 However, **our simulations did not include other sources of disturbance**, such as partial harvests, extreme weather, and damage from insects and pathogens, which are common to this region [@amiro_2010_ecosystem] and which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
-- Parameter uncertainty is the dominant driver of model predictive uncertainty.
-- ...but the nature of this parameter uncertainty depends on model configuration.
-
-- Ecological implications of structural uncertainty
-    - Predictions about ecological succession are very sensitive to the understory light environment.
-        - Need a good representation of canopy structure (finite canopy radius)
-        - How you model canopy radiative transfer matters! Small differences in the model can give significant results.
-    - But, trait plasticity can provide an alternative mechanism for success in the understory.
-        - Important modeling implication for all the emerging research on intra-specific trait variability.
-        - This combination of discrete functional types with plastic traits is closer to reality than purely trait-based ("eco-evolutionary?") models or completely discretized models with static traits on the other.
-        - Important caveat: Current implementation is based entirely on tropical data, which has more extreme canopy shading than temperate forests (REF), so tropical species may have evolved stronger trait plasticity than temperate species (REF).
-        - Perhaps note "ecological equifinality" here: there are different ways to arrive at similar outcomes in the model; this is also true in nature? 
-- Implications of uncertainty analysis for future data collection and model development
-    - Parameter uncertainty is still a _huge_ source of overall model predictive uncertainty. Different parameters can give _qualitatively_ different ecological outcomes.
-    - Looking beyond aggregate variables (e.g. net primary productivity, total LAI) is essential to model evaluation and calibration. It's possible to generate a relatively?
-    - (Expected) Subsurface parameters (e.g. leaf-root ratios) have high sensitivity, but are poorly constrained.
-    - Trait covariance may alleviate some data limitation (Shiklomanov et al. 2018 [preprint])
-- Comparisons against other studies
-    - Parameter sensitivity
-        - [@pandit_2018_optimizing] -- Most important parameters for shrubs were SLA, Vcmax, stomatal slope, root:leaf ratio, and root turonver rate (slightly less important: leaf turnover rate, growth respiration, cuticular conductance; much less important: water conductance; basically irrelevant: storage turnover, leaf width)
-        - Compare against @dietze_2014_quantitative and @raczka_2018_what
-- Future directions: Predicting disturbance response
-    - FoRTE project -- Forecasting a disturbance.
-        - Compare ecological resilience/memory between model and reality
-- Other relevant references:
-  - @gough_2019_high -- Canopy structural complexity is the strongest predictor of productivity across NEON sites. Possible future direction?
-
-Moreover, these models' **predictions of species composition and ecosystem structure are relevant outside of carbon cycle science**, particularly for forest management and wildlife conservation.
-<!-- BBL: maybe move this sentence to discussion?
-TODO refs -->
 
@@
 # if your Rmd is not in analysis/paper/
-git2r::repository("../..")
+git2r::repository(here::here())
 ```

### 2019-09-12 analysis/paper/paper.Rmd
Add line break in abstract source
Add note about Figure 2 variables
Big-picture outline of 2nd half of discussion

@@
   (1) Which processes are most important to accurately model community succession and C cycling in Upper Midwest forests?
-  (2) What are the relative contributions of parameter vs.
-structural uncertainty?
+  (2) What are the relative contributions of parameter vs. structural uncertainty?
   We ran ensembles of the Ecosystem Demography model v2.2 (sampling over parameter uncertainty) from 1900 to 2000 with different representations of key processes important to competition for light.
@@
 
+<!-- TODO Should one of these be Early Hardwood (or other PFT?) LAI/biomass fraction -->
+<!-- TODO Drop some of these variables? -->
+
 Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
@@
 However, **our simulations did not include other sources of disturbance**, such as partial harvests, extreme weather, and damage from insects and pathogens, which are common to this region [@amiro_2010_ecosystem] and which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
+Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance frequencies and see how they affect ecological dynamics.
+
+- Figure 3:
+  - For the same LAI, closed canopy had higher productivity than finite. Why?
+    - Not zero-sum; finite canopy hurts dominant trees but doesn't improve subdominant species that much <!-- TODO Need support for this! -->
+  - Importance of using multiple observational constraints. Ecological equifinality is a real problem.
+
+## Drivers of uncertainty
+
+**Parameter uncertainty was the dominant driver of uncertainty in our analysis**.
+Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables** -- NPP ranged from 0 to almost 2x the observed, and LAI was up to 3x observed (though more tightly clustered around zero) (Figure 2).
+Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figure 4).
+**Parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.
+
+**Parameter uncertainty is not synonymous with sensitivity**.
+Many of the parameters to which model outputs are most sensitive are already well constrained by data and do not contribute as much to predictive uncertainty (Figure 6).
+Although parameter uncertainty is still large, this means that **targeted data collection works and that we are moving in the right direction**!
+This speaks to the value of looking at not just sensitivity, but partial variance [@lebauer_2013_facilitating; @dietze_2014_quantitative].
+(Compare sensitivity results to these studies, and also to @pandit_2018_optimizing).
+
+Although parameter uncertainty dominated, **structural and parameter uncertainty interacted in important ways**.
+The relative position of different parameter combinations in multivariate space (Figure 3), the total amount of parameter uncertainty (Figure 5), and the relative importance of different parameters to predictive uncertainty (Figure 6) all differed across model structures (Figure 3).
+This is despite the fact that our model structure changes were relatively modest compared to differences between different models.
+Collectively, this means that **sensitivity and uncertainty analyses are not necessarily generalizable across models**.
 

### 2019-09-12 analysis/paper/paper.Rmd
More work on Discussion

@@
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
-This idea that ED2.2's main issue is that it is moving too quickly is supported by the fact that ED2.2 generally **underestimates productivity after 100 years**, but model **predictions at peak productivity are close to reality**.
+This idea that ED2.2's main issue is that it is moving too quickly is supported by the fact that ED2.2 generally underestimates productivity after 100 years, but model **predictions at peak productivity are close to reality**.
 One reason ED2.2 is moving too quickly through succession is **missing mechanisms for sustaining productivity through early succession**.
@@
 **Parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.
+The dominance of parameter uncertainty over structural uncertainty in our results suggests that **targeted data collection should be a higher priority than model development**.
 
@@
 This speaks to the value of looking at not just sensitivity, but partial variance [@lebauer_2013_facilitating; @dietze_2014_quantitative].
-(Compare sensitivity results to these studies, and also to @pandit_2018_optimizing).
+
+So **which parameters are a research priority**?
+Leaf optical properties often do not get a lot of attention, but are important [see also, @viskari_2019_influence].
+Also, soil-plant water conductance and growth respiration -- both of these were important in @dietze_2014_quantitative.
+(Compare also to @raczka_2018_what, @pandit_2018_optimizing).
 

### 2019-09-13 R/submodel-rtm-ms.R
Start thinking about crown area

@@ -1,14 +1,19 @@
 #' Transmissivity to diffuse radiation for the multiple-scatter model
 #'
-#' @inheritParams tau_direct 
+#' @param C Crown area index (CAI) (default = 1)
+#' @inheritParams tau_direct
 #' @export
-ms_tau_diffuse <- function(L, orient) {
+ms_tau_diffuse <- function(L, orient, C = 1) {
+  # Locally-exposed leaf area
+  locetai <- L / C
   phi1 <- phi1_f(orient)
   phi2 <- phi2_f(orient)
-  ext_diff1 <- phi1 * L
-  ext_diff2 <- phi2 * L
-  -exp(-ext_diff1 - ext_diff2) *
-    (ext_diff1^2 * exp(ext_diff1) * gsl::expint_Ei(-ext_diff1) + ext_diff1 - 1)
+  ext_diff1 <- phi1 * locetai
+  ext_diff2 <- phi2 * locetai
+  (1 - C) -
+    C * exp(-ext_diff1 - ext_diff2) *
+    (ext_diff1 ^ 2 * exp(ext_diff1) *
+       gsl::expint_Ei(-ext_diff1) + ext_diff1 - 1)
 }
 
 #' Multiple-scatter model

### 2019-09-13 R/submodel-rtm.R

@@ -70,10 +70,11 @@ single_scatter_albedo <- function(theta, orient, theta_rad = theta * pi / 180) {
 #' Transmissivity to direct radiation
 #'
 #' @param L Leaf (or total) area index
+#' @param C Crown area index (default = 1)
 #' @inheritParams kfunction
 #' @author Alexey Shiklomanov
 #' @export
-tau_direct <- function(L, theta, orient) {
+tau_direct <- function(L, theta, orient, C = 1) {
   K <- kfunction(theta, orient)
   exp(-K * L)
 }

### 2019-09-16 analysis/paper/paper.Rmd
Add acknowledgements TODO
Fix minor typos
Fix drake RMArkdown processing

@@
 
-```{r, setup, echo = FALSE}
+```{r setup, echo = FALSE}
 library(knitr)
@@
 library(drake)
-
-fig <- function(...) {
-  filepath <- here("analysis", "figures", ...)
-  knitr::include_graphics(file_in(filepath))
-}
 
@@
   echo = FALSE,
-  comment = "#>"
+  comment = "#>",
+  out.width = "\\linewidth"
   ## fig.path = "../figures/"
@@
 
-```{r methods_captions}
+```{r param_dist, fig.cap = param_dist_cap}
 param_dist_cap <- paste(
@@
 )
-```
-
-```{r param_dist, fig.cap = param_dist_cap}
-fig("param-dist.png")
+readd(param_dist_knit)
 ```
@@
   "The solid line is the ensemble mean.",
-  "The dashed line is the 80% confidence interval.",
+  "The dashed line is the 80\\% confidence interval.",
   "Black dot with error bars is the observed mean and min/max value",
@@
 )
-fig("summary-ts-plot.png")
+readd(summary_ts_plot_knit)
 ```
@@
 )
-fig("pairs-time-averaged.png")
+readd(pairs_time_averaged_knit)
 ```
@@
 )
-fig("lai-pft-plot.png")
+readd(lai_pft_plot_knit)
 ```
@@
 )
-fig("within-across-plot.png")
+readd(within_across_plot_knit)
 ```
@@
 )
-fig("sensitivity-plot.png")
+readd(sensitivity_plot_knit)
 ```
@@
 
+<!-- TODO Fill in Acknowledgements -->
 This project funded by NSF grant.

### 2019-09-16 analysis/drake.R

@@ -226,6 +226,9 @@ plan <- drake_plan(
     # TODO: Fix dims so axis titles aren't cut off
     width = 10, height = 10
   ),
+  summary_ts_plot_knit = knitr::include_graphics(file_in(!!path(
+    fig_dir, "summary-ts-plot.png"
+  ))),
   #########################################
   # Parameter distributions
   #########################################
@@ -250,6 +253,7 @@ plan <- drake_plan(
     param_dist_gg,
     width = 15.3, height = 9.9
   ),
+  param_dist_knit = knitr::include_graphics(file_in(!!path(fig_dir, "param-dist.png"))),
   #########################################
   # LAI by PFT plot
   #########################################
@@ -317,6 +321,9 @@ plan <- drake_plan(
     lai_pft_plot,
     width = 11.42, height = 5.67
   ),
+  lai_pft_plot_knit = knitr::include_graphics(file_in(!!path(
+    fig_dir, "lai-pft-plot.png"
+  ))),
   #########################################
   # Parameter vs. structure uncertainty
   #########################################
@@ -366,6 +373,9 @@ plan <- drake_plan(
     within_across_plot,
     width = 8.3, height = 5.1
   ),
+  within_across_plot_knit = knitr::include_graphics(file_in(!!path(
+    fig_dir, "within-across-plot.png"
+  ))),
   #########################################
   # Sensitivity analysis
   #########################################
@@ -432,6 +442,9 @@ plan <- drake_plan(
     # TODO: This is huge...
     width = 20, height = 15
   ),
+  sensitivity_plot_knit = knitr::include_graphics(file_in(!!path(
+    fig_dir, "sensitivity-plot.png"
+  ))),
   #########################################
   # Pairs plot of time-averaged values
   #########################################
@@ -477,7 +490,10 @@ plan <- drake_plan(
     pairs_time_averaged,
     height = 5.67,
     width = 8.92
-  )
+  ),
+  pairs_time_averaged_knit = knitr::include_graphics(file_in(!!path(
+    fig_dir, "pairs-time-averaged.png"
+  ))),
 )
 
 if ("--poster" %in% cmdargs) source("analysis/drake_poster.R")
@@ -489,7 +505,7 @@ if ("--paper" %in% cmdargs) {
         knitr_in(!!(path(analysis_dir, "paper", "paper.Rmd"))),
         .format
       ),
-      transform = map(.format = c("html_document", "pdf_document"))),
+      transform = map(.format = c("html_document", "pdf_document")))
     ))
 }
 

### 2019-09-17 analysis/paper/submodel-rtm.Rmd
Start thinking about crown area
Add two-stream submodel code

@@
 
-Following @clm45_note, forward- ($\nu$) and backscattering of leaves or stems ($\omega$) are defined as a function of reflectance ($R$) and transmittance ($T$) of the corresponding materials:
+Following @clm45_note, forward- ($\nu$) and backscattering ($\omega$) of canopy elements (leaves or stems) are defined as a function of those elements' reflectance ($R$) and transmittance ($T$):
 
-$$ \nu = R + T $$
-$$ \omega = \frac{R + T + \frac{1}{4} (R-T)(1 - \chi)^2}{2 (R+T)} $$
+$$ \nu_x = R_x + T_x $$
+$$ \omega_x = \frac{R_x + T_x + \frac{1}{4} (R_x-T_x)(1 - \chi)^2}{2 (R_x+T_x)} $$
 
@@
 \left(
-\phi_1^2 L_i^2 \exp(\phi_1 L_i) * E_i(-\phi_1 L_i) +
+\phi_1^2 L_i^2 \exp(\phi_1 L_i) E_i(-\phi_1 L_i) +
 \phi_1 L_i - 1
@@
 
-## Finite canopy radius
+# Two-stream
+
+The $X$ and $Y$ vectors of the matrix problem for diffuse fluxes are defined as follows:
+
+$$
+X =
+\begin{bmatrix}
+SWuO_0 \\
+SWd0_1 \\
+SWu0_1 \\
+... \\
+SWd0_i \\
+SWu0_i \\
+... \\
+SWd0_m \\
+SWu0_m \\
+SWd0_{m + 1}
+\end{bmatrix}
+$$
+
+$$
+Y = 
+\begin{bmatrix}
+S_0 a_{ground} () \\
+S_1 r(\psi)_1 \left[ 1 - r_0 r_1 (1 - \alpha_0) (1 - \tau_0) (1 - \alpha_1) (1 - \tau_1) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) \\
+S_1 \left[ 1 - r_1 r_2 (1 - \alpha_1) (1 - \tau_1) (1 - \alpha_2) (1 - \tau_2) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) (1 - r(\psi)_1) \\
+... \\
+S_i r(\psi)_i \left[ i - r_{i-1} r_i (i - \alpha_{i-1}) (i - \tau_{i-1}) (i - \alpha_i) (i - \tau_i) \right] (i - \tau(\psi)_i) (i - \alpha_i) \\
+S_i \left[ i - r_i r_{i+1} (i - \alpha_i) (i - \tau_i) (i - \alpha_{i+1}) (i - \tau_{i+1}) \right] (i - \tau(\psi)_i) (i - \alpha_i) (i - r(\psi)_i) \\
+... \\
+S_m r(\psi)_m \left[ m - r_{m-1} r_m (m - \alpha_{m-1}) (m - \tau_{m-1}) (m - \alpha_m) (m - \tau_m) \right] (m - \tau(\psi)_m) (m - \alpha_m) \\
+S_m \left[ m - r_m r_{m+1} (m - \alpha_m) (m - \tau_m) (m - \alpha_{m+1}) (m - \tau_{m+1}) \right] (m - \tau(\psi)_m) (m - \alpha_m) (m - r(\psi)_m) \\
+SW_{sky}
+\end{bmatrix}
+$$
+
+Diffuse transmittance.
+
+```{r}
+
+plot(c(0, Lmax), c(0, 1), type = "n",
+     xlab = "Leaf area index",
+     ylab = "Diffuse transmittance fraction")
+legend("topright", as.character(orient),
+       title = "Orient factor",
+       lty = "solid",
+       col = colors)
+purrr::walk2(
+  orient, colors,
+  ~curve(ts_tau_diffuse(x, .x, lr = 0.056, lt = 0.045, cai = 1), 0, Lmax, add = TRUE, col = .y)
+)
+
+```
+
+Here it is compared with the multi-scatter model.
+
+```{r}
+
+orient_sub <- c(-0.5, 0, 0.5)
+colors_sub <- colors[c(1, 3, 5)]
+plot(c(0, Lmax), c(0, 1), type = "n",
+     xlab = "Leaf area index",
+     ylab = "Diffuse transmittance fraction")
+legend("topright", as.character(orient_sub),
+       title = "Orient factor",
+       lty = "solid",
+       col = colors_sub)
+legend("top", c("multi-scatter", "two-stream"),
+       title = "RTM",
+       lty = c("dashed", "solid"))
+purrr::walk2(
+  orient_sub, colors_sub,
+  ~curve(ms_tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y, lty = "dashed")
+)
+purrr::walk2(
+  orient_sub, colors_sub,
+  ~curve(ts_tau_diffuse(x, .x, lr = 0.056, lt = 0.045, cai = 1), 0, Lmax, add = TRUE, col = .y)
+)
+
+```
+
+# Finite canopy radius
 
@@
 
-The crown area index ($C_i$) affects the following terms in the following ways:
+In both models, the transmittance of direct radiation corrected for crown area index ($\tau(\psi)_i^*$) is calculated as follows:
 
-$$ \tau(\psi)_i = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$
+$$ \tau(\psi)_i^* = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$
+
+```{r}
+tau_psi_star <- function(L, C, theta, orient) {
+  k <- kfunction(theta, orient)
+  1 - C + C * exp(-k * L / C)
+}
+curve(tau_psi_star(1, x, 0, 0), 0, 1,
+      ylim = c(0, 1),
+      xlab = "Crown area index", y = expression(tau(psi) ^ "*"))
+curve(tau_psi_star(2, x, 0, 0), 0, 1, col = 2, add = TRUE)
+curve(tau_psi_star(3, x, 0, 0), 0, 1, col = 3, add = TRUE)
+curve(tau_psi_star(4, x, 0, 0), 0, 1, col = 4, add = TRUE)
+```
+
+The effect of the crown area index on diffuse transmittance varies by model.
+In the **two-stream** model, the crown area index affects the average diffuse optical depth ($\bar{\mu}$) as follows:
 
 $$
-\tau_i = (1 - C_i) - 
-C_i \exp(-\frac{L_i}{C_i} (\phi_1 + \phi_2))
-\left(
-\frac{\phi_1^2 L_i^2}{C_i^2} \exp(\phi_1 \frac{L_i}{C_i}) * E_i(-\phi_1 \frac{L_i}{C_i}) +
-\phi_1 \frac{L_i}{C_i} - 1
-\right)
+\bar{\mu}^* = \frac{-L_i}{\log(1 - C_i + C_i \exp( -L_i / \bar{\mu} ))}
 $$
 
-# Two-stream
+(Note that if $C_i = 1$, this simplifies to $\bar{\mu}^* = \bar{\mu}$).
+
+In the **multiple-scatter** model, if we define the diffuse transmittance expression ($\tau_i$) above as a function of leaf area index, $T_i(L)$,
+then the diffuse transmittance corrected for crown area ($\tau_i^*$) becomes:
+
+$$
+\tau_i^* = (1 - C_i) + C_i T_i\left( \frac{L_i}{C_i} \right)
+$$
+
+Comparing the diffuse radiation response of the two models:
+
+```{r}
+
+cai <- c(0.1, 0.25, 0.5, 1)
+colors_sub <- c("red1", "red2", "red3", "black")
+plot(c(0, Lmax), c(0, 1), type = "n",
+     xlab = "Leaf area index",
+     ylab = "Diffuse transmittance fraction")
+legend("bottomleft", as.character(cai),
+       title = "CAI",
+       lty = "solid",
+       col = colors_sub,
+       inset = c(0, 0.2))
+legend("bottomleft", c("MS", "2S"),
+       title = "RTM",
+       lty = c("dashed", "solid"))
+purrr::walk2(
+  cai, colors_sub,
+  ~curve(ms_tau_diffuse(x, 0, .x), 0, Lmax, add = TRUE, col = .y, lty = "dashed")
+)
+purrr::walk2(
+  cai, colors_sub,
+  ~curve(ts_tau_diffuse(x, 0, lr = 0.056, lt = 0.045, cai = .x), 0, Lmax, add = TRUE, col = .y)
+)
+
+```
 
@@
 alb_ground <- 0.1
-
 ```

### 2019-09-17 R/submodel-rtm-2s.R

@@ -217,3 +217,30 @@ two_stream <- function(ipft, L,
     ll_list
   )
 }
+#' Average inverse optical depth for diffuse radiation: Two-stream
+#'
+#' With correction for finite crown area.
+#'
+#' @inheritParams two_stream
+#' @return
+#' @author Alexey Shiklomanov
+#' @export
+ts_mubar_star <- function(orient_factor, lai, cai = 1) {
+  mubar <- mu_bar(orient_factor)
+  -lai / log(1 - cai + cai * exp(-lai / mubar))
+}
+
+#' Transmissivity of diffuse radiation: Two-stream
+#'
+#' @inheritParams two_stream
+#' @return
+#' @author Alexey Shiklomanov
+#' @export
+ts_tau_diffuse <- function(lai, orient_factor, lr, lt, cai) {
+  mu <- ts_mubar_star(orient_factor, lai, cai)
+  iota <- scatter(lr, lt)
+  beta <- backscatter(lr, lt, orient_factor)
+  epsil <- 1 - 2 * beta
+  lambda <- sqrt((1 - epsil * iota) * (1 - iota)) / mu
+  exp(-lambda * lai)
+}

### 2019-09-17 R/submodel-rtm.R

@@ -107,3 +107,13 @@ integrate_light <- function(down0, down, up) {
     light_diff_level = light_diff_level
   )
 }
+
+scatter <- function(lr, lt) {
+  lr + lt
+}
+
+backscatter <- function(lr, lt, orient) {
+  num <- lr + lt + 0.25 * (lr - lt) * (1 - orient) ^ 2
+  den <- 2 * (lr + lt)
+  num / den
+}

### 2019-09-17 analysis/paper/submodel-rtm.Rmd
Add crown area index comparison plots (draft)

@@
 
-plot(ms_out$light_level, type = "l")
-lines(ts_out$light_level, col = 2)
+plot(ms_out$light_level, type = "l", lty = "dashed",
+     xlab = "Cohort", ylab = "Total light level")
+lines(ts_out$light_level, lty = "solid")
+legend("topleft", c("2S", "MS"), lty = c("solid", "dashed"))
 
-plot(ms_out$light_diff_level, type = "l")
-lines(ts_out$light_diff_level, col = 2)
+plot(ms_out$light_diff_level, type = "l", lty = "dashed",
+     xlab = "Cohort", ylab = "Diffuse light level")
+lines(ts_out$light_diff_level, lty = "solid")
+legend("topleft", c("2S", "MS"), lty = c("solid", "dashed"))
+```
+
+Run both models under different crown area indices.
+
+```{r}
+
+ncoh <- 3
+ipft <- rep(1, ncoh)
+lai <- rep(0.5, ncoh)
+cai <- c(0.1, 0.25, 0.5, 1)
+colors <- c("red1", "red2", "red3", "black")
+ms1 <- lapply(cai, function(.x) multi_scatter(
+  ipft, lai, lr[1], lt[1], orient[1], C = .x,
+  theta, S0_beam, S0_diff, alb_ground
+))
+ts1 <- lapply(cai, function(.x) two_stream(
+  ipft, lai, lr[1], lt[1], orient[1], cai = .x,
+  theta, S0_beam, S0_diff, alb_ground
+))
+## par(mfrow = c(ncoh, 1))
+## for (i in rev(seq_len(ncoh))) {
+##   ms1_ll <- vapply(ms1, function(x) x[["light_level"]][i], numeric(1))
+##   ts1_ll <- vapply(ts1, function(x) x[["light_level"]][i], numeric(1))
+##   matplot(cai, cbind(ts1_ll, ms1_ll), type = "l", col = "black",
+##           ylim = c(0.2, 1))
+##   legend("topright", c("2S", "MS"), lty = 1:2)
+## }
+
+v <- "light_diff_level"
+par(mfrow = c(length(cai), 1))
+for (i in seq_along(cai)) {
+  ms1_ll <- ms1[[i]][[v]]
+  ts1_ll <- ts1[[i]][[v]]
+  matplot(cbind(ts1_ll, ms1_ll), type = "l", col = "black",
+          xlab = "Cohort", ylab = "Total light level",
+          ylim = c(0, 0.4), main = paste("CAI = ", cai[i]))
+  legend("bottomright", c("2S", "MS"), lty = 1:2)
+}
+
 ```

### 2019-09-17 R/submodel-rtm-2s.R

@@ -21,6 +21,7 @@
 #' @export
 two_stream <- function(ipft, L,
                        lr, lt, orient,
+                       cai = rep(1, length(L)),
                        theta = 15,
                        S0_beam = 0.8,
                        S0_diff = 1 - S0_beam,
@@ -65,7 +66,6 @@ two_stream <- function(ipft, L,
   ## leaf_weight <- elai / etai
   ## wood_weight <- 1 - leaf_weight
 
-  cai <- rep(1, length(L))
   czen <- cos(theta * pi / 180)
 
   # Inverse optical depth of direct radiation

### 2019-09-17 R/submodel-rtm-ms.R

@@ -31,6 +31,8 @@ ms_tau_diffuse <- function(L, orient, C = 1) {
 multi_scatter <- function(ipft, L,
                           # PFT-level parameters
                           lr, lt, orient,
+                          # Crown area index
+                          C = 1,
                           # Site scalars
                           theta = 15,
                           S0_beam = 0.8,
@@ -80,14 +82,14 @@ multi_scatter <- function(ipft, L,
   t <- fvector(seq(0, ncoh + 1))
   t[0] <- 0
   t[ncoh + 1] <- 1
-  t[1:ncoh] <- ms_tau_diffuse(L, orient)
+  t[1:ncoh] <- ms_tau_diffuse(L, orient, C)
   ti <- 1 - t   # omt_diff
 
   # Direct transmittance (tau_beam in ED2)
   tpsi <- fvector(seq(0, ncoh + 1))
   tpsi[0] <- 0
   tpsi[ncoh + 1] <- 1
-  tpsi[1:ncoh] <- tau_direct(L, theta, orient)
+  tpsi[1:ncoh] <- tau_direct(L, theta, orient, C)
   tpsii <- 1 - tpsi  # omt_beam
 
   # Direct reflectance (r_beam in ED2)

### 2019-09-18 analysis/paper/paper.Rmd
Suggestions for first part of discussion

@@
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance]. 
+This is a common characteristic of many ecosystem models: they may be relatively insensitive to short-term changes (e.g.
+https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.13626) but have trouble sustaining forest production over long time periods, not capturing forest resilience (https://www.sciencedirect.com/science/article/pii/S0378112716310581).
+This is likely because of species successional changes (https://www.nature.com/articles/s41598-018-23763-y, https://www.mdpi.com/1999-4907/9/8/460) and the role of minor or subtle disturbances (http://dx.doi.org/10.1111/nph.15227).
+Neither of these processes is well captured by current models (https://onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2486.2004.00891.x, maybe B-L 2016?). 
+This is likely to be an increasingly important limitation in a world of rising disturbances and climate-driven species changes.
+
 This idea that ED2.2's main issue is that it is moving too quickly is supported by the fact that ED2.2 generally underestimates productivity after 100 years, but model **predictions at peak productivity are close to reality**.
+(https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.13077)
 One reason ED2.2 is moving too quickly through succession is **missing mechanisms for sustaining productivity through early succession**.
@@
 ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the top 10 most important parameters for predictive uncertainty (Figure 6). 
-<!-- Maybe move this elsewhere? -->
-<!-- TODO Look at mortality across ED runs. Does it relate at all to results? -->
-However, **our simulations did not include other sources of disturbance**, such as partial harvests, extreme weather, and damage from insects and pathogens, which are common to this region [@amiro_2010_ecosystem] and which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
+In addition, **our simulations did not include other sources of disturbance**, such as partial harvests, extreme weather, and damage from insects and pathogens, which are common to this region [@amiro_2010_ecosystem] and which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
+We suggest that how these mortality processes interact with factors such as canopy restructuring and thus radiation transmission (http://dx.doi.org/10.1016/j.foreco.2013.02.031), sustaining the productivity of aging forests over decades (Luyssaert 2008?) is critical to explore in future research.
+
 Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance frequencies and see how they affect ecological dynamics.

### 2019-10-01 analysis/paper/paper.Rmd
Add methods note about MS vs. 2S model
Minor tweaks to methods text

--- 2019-09-16 analysis/paper/paper.Rmd
@@
 @wang_2003_comparison].
+**Compared to the default two-stream model, the multiple-scatter model slightly increases diffuse light levels in the understory.**
 Both of these models are described in detail in the Supplementary Information. <!-- TODO Exact reference to Appendix -->
@@
 This has been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous canopies [@fisher_2015_taking].
-In the alternate configuration ("finite crown area", or "partial shading"), canopies take up only a fraction of the available horizontal area, meaning that multiple cohorts of similar height can receive the same level of light.
+In the alternate configuration ("finite crown area", or "partial shading"), canopies take up only a fraction of the available horizontal area, which generally exposes the understory to more light.
 The horizontal area of crowns is determined by allometric equations from @dietze_2008_changing.

### 2019-10-01 analysis/paper/paper.Rmd
Fix references to minimum_height in text and table

@@
 - Fraction of litter carbon that enters the labile (fast) soil carbon pool (`f_labile`)
-- Minimum plant height for reproduction (`minimum_height`)
+- Minimum plant height for reproduction (`repro_min_h`)
 - Soil-plant water conductance, or "water availability factor" (`water_conductance`)

### 2019-10-02 analysis/paper/paper.Rmd
Add new results outline to paper

@@
 
+## New outline
+
+- Impacts of structure alone
+  - Figure: ED2 runs with default parameters, looking at (a) fluxes, and (b) cohort composition, and (c) cohort light levels
+  - Fluxes are similar, but there is some variability
+  - Composition: Also similar, but a few noteworthy things
+- Default vs. median parameters
+  - Figure: Compare default params, median params, and parameter subsets
+  - Compared to default values, median parameter values are:
+    - Growth resp --- Early, Mid, Late: Higher; Pine: Lower --- Strong penalty on hardwood productivity.
+    - Fine root : leaf ratio --- Early, Late: Lower; Mid, Pine: Higher --- Reduced early and mid hardwood productivity.
+    - Stomatal slope --- Late: Lower; Others: Higher --- Higher productivity for EH, MH, P; lower productivity for LH
+    - Root turnover --- EH, LH: Lower; Pine: Higher --- Higher productivity for EH, MH, and LH
+    - Quantum efficiency -- All: Lower --- Lower productivity
+    - Leaf refl/trans NIR --- All: Higher --- Higher productivity
+    - Mort 3 --- All: Higher --- Lower productivity
+    - Repro min H --- All: Lower --- Lower productivity
+    - Water conductance -- All: Higher --- Higher productivity
+    - SLA --- Early: same; Mid: much higher; Late: higher; Pine: higher --- Propels mid-hardwood into canopy.
+    - Vcmax --- Higher for all
+  - Primary impact is the difference in SLA. High mid-hardwood SLA drives its productivity. (Panel: SLA)
+  - Vcmax and growth respiration roughly balance each other out (because much higher Vcmax across the board, but also higher growth respiration for hardwoods).
+But the combination of SLA, Vcmax, and growth respiration allows pine to quickly beat out early and late hardwood, and eventually to catch up to mid-hardwood.
+  - SLA and Vcmax are both very well-constrained by data, so the ED2 defaults are definitely at fault here. With these median values, we end up with the wrong composition.
+- Structure x parameter uncertainty
+  - Figure: Time series of NPP, LAI, and relative PFT dominance
+  - Parameter uncertainty is much larger than structural uncertainty
+  - Figure: Pairs of NPP, LAI, and PFT fraction, for each structure (first draw all, then pick relevant subsets)
+  - Q: How many runs end up with the right composition (EH-dominated) at peak (~1925), ~1950, and 1999?
+- Details of parameter uncertainty
+  - Figure: Parameter sensitivity and uncertainty analysis
+  - Some parameters are always important. Other parameters matter more for different structures.
+
 ## Summary

### 2019-10-04 analysis/paper/paper.Rmd
Add new text and figs to results

--- 2019-10-01 analysis/paper/paper.Rmd
@@
 
-```{r param_dist, fig.cap = param_dist_cap}
-param_dist_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions from PEcAn trait meta-analysis."
-)
-readd(param_dist_knit)
-```
-
-`r mdcap(param_dist_cap)`
-
 ## Modeling workflow
@@
 
+## Impacts of structure
+
+```{r , fig.cap = structure_default_cap}
+structure_default_cap <- paste(
+  "Annual NPP and LAI from ED2 runs",
+  "with default parameters",
+  "across different configurations."
+)
+readd(structure_compare_default_knit)
+```
+
+Different model structures very closely for the first half of the simulation, at which point they started to diverge.
+Post 1950, the combination of two-stream radiative transfer, finite canopy radius, and plastic traits had the highest sustained productivity, followed by the same configuration except with static traits, with the remaining configurations having very similar results.
+
+```{r , fig.cap = structure_default_light_cap}
+structure_default_light_cap <- paste(
+  "Radiation profiles of cohorts from ED2 runs with",
+  "default parameters, across different configurations."
+  "Each point represents a single cohort of the given PFT,"
+  "position on the y-axis represents the height of that cohort,"
+  "and position on the x-axis represents the relative light level",
+  "received by that cohort."
+)
+readd(structure_default_light_knit)
+```
+
+Differences in productivity across structures are perhaps best understood by looking at differences in resulting radiation profiles at different snapshots in time over the course of the simulation.
+Enabling the finite canopy radius increased the availability of light for understory cohorts throughout the simulation, with the most notable effects up to 1950.
+Under a closed canopy, switching from a two-stream to a multiple-scatter canopy radiative transfer model also enhanced the understory light environment under a closed canopy, but to a lesser extent, and the brightest understory occurred under the combination of the two-stream radiative transfer and finite canopy radius.
+However, none of the configurations were able to sustain any PFTs other than early hardwoods by 1980, and only the unique combination of finite canopy radius and two-stream radiative transfer were able to sustain more than a single dominant cohort (2 early hardwood cohorts with static traits, and 3 with plastic traits).
+
+## Default vs. median parameters
+
+```{r param_dist, fig.cap = param_dist_cap}
+param_dist_cap <- paste(
+  htmlfig(),
+  "Input parameter distributions from PEcAn trait meta-analysis."
+)
+readd(param_dist_knit)
+```
+
+`r mdcap(param_dist_cap)`
+
+Parameter estimates from the trait meta-analysis were often substantially different from the ED2 defaults.
+In many cases, median parameter estimates were different entirely because of the choice of prior distributions (though, in almost all cases, the distributions provided reasonably high likelihood to ED2 default parameter values).
+However, in several important cases, parameter well-constrained by data produced values substantially different from the ED2 defaults.
+For example, ED2 default values for SLA were too low for all PFTs except Early Hardwoods, and default values for Vcmax values were too low for all PFTs.
+
+```{r , fig.cap = default_median_cap}
+default_median_cap <- paste(
+  "Comparing results of ED2 with default parameters and",
+  "posterior median parameters from trait meta-analysis",
+  "across different strucures."
+)
+readd(default_median_fluxes_knit)
+```
+
+```{r}
+default_median_cap <- paste(
+  "Comparing results of ED2 PFT-level LAI",
+  "with default parameters",
+  "posterior median parameters from trait meta-analysis",
+  "across different strucures."
+)
+readd(default_median_lai_knit)
+```
+
+The above differences in median parameter led to ED2 runs that differed from the default runs in several important ways.
+First of all, throughout all simulations and across all model structures, autotrophic respiration was several times higher with median parameter estimates compared to default values.
+This is driven primarily by higher median values for growth and leaf respiration across all PFTs (notably, the ED2 default is a growth respiration of _zero_ for all temperate hardwood PFTs).
+In some cases, this is compensated by an increase in gross primary productivity (driven by higher Vcmax values across all PFTs), leading to similar net primary productivity values to the default parameter runs.
+Second, the variation across model structures is much higher, with every change in model structure exerting a noticeable effect on flux dynamics.
+The most dramatic example is that, with median parameter values, all of the simulations with the finite canopy radius produced much lower productivity and heterotrophic respiration than simulations with a closed canopy.
+Third, even where net primary productivity is similar between default and median runs, the underlying forest composition is very different.
+Namely, under the default parameterization, ED2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
+However, with the median parameterization, ED2 starts out with forests dominated by _mid_-hardwood species that are quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remains the dominant canopy species throughout the rest of the simulation.
+Exploratory analyses (results not shown / supplement) show that the main parameters responsible for this compositional trajectory are SLA (especially the higher SLA of pine and mid-hardwood), Vcmax, and growth respiration.
+While the uncertainty in growth respiration is large, we emphasize that SLA and Vcmax are well constrained by data for all PFTs considered in this analysis.
+
 ## Summary
@@
 `r mdcap(summary_ts_cap)`
+
+<!-- - Structure x parameter uncertainty -->
+<!--   - Figure: Time series of NPP, LAI, and relative PFT dominance -->
+<!--   - Parameter uncertainty is much larger than structural uncertainty -->
+<!--   - Figure: Pairs of NPP, LAI, and PFT fraction, for each structure (first draw all, then pick relevant subsets) -->
+<!--   - Q: How many runs end up with the right composition (EH-dominated) at peak (~1925), ~1950, and 1999? -->
 
@@
 ## Uncertainty analysis
+
+<!-- - Details of parameter uncertainty -->
+<!--   - Figure: Parameter sensitivity and uncertainty analysis -->
+<!--   - Some parameters are always important. Other parameters matter more for different structures. -->
 

### 2019-10-10 analysis/paper/paper.Rmd
Add PFT ternary plot, and other scratch code

@@
   "ED2 plot-level predictions of",
-  "gross (GPP) and net (NPP) primary productivity,",
-  "aboveground biomass (AGB),",
+  "net (NPP) primary productivity and",
   "total leaf area index (LAI),",
-  "and Shannon diversity index (of plant functional types)",
   "by model configuration.",
-  "The solid line is the ensemble mean.",
-  "The dashed line is the 80\\% confidence interval.",
+  "The dashed lines indicate the 10th and 90th quantiles of results.",
   "Black dot with error bars is the observed mean and min/max value",
@@
 
-Model projections of gross and net primary productivity, aboveground biomass, leaf area index, and functional diversity varied with both parameter value and model structure (Figure 2).
+Model projections of net primary productivity (NPP) and leaf area index (LAI) varied with both parameter value and model structure (Figure X).
 All model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
-Two model configurations -- finite canopy radius and two-stream radiative transfer (regardless of trait plasticity) -- had much lower productivity, biomass, and diversity than the other configurations, which generally had similar ensemble mean values and spread.
-The ensemble means of all model configurations under-predicted observed net primary productivity and leaf area index, both at the peak of initial growth (around 1920) and at the end of the simulations, though peak predicted values were much closer to thte observations.
+The ensemble spread within any single configuration---i.e.
+the parameter uncertainty---was similar across all configurations; both NPP and LAI varied 5-10 fold at any given timestep depending on parameter values.
+Within all configurations, most (>90%) ensemble members under-predicted observed present-day NPP and LAI by year 2000, though individual ensemble members from some configurations over-predicted one or both variables.
+Predictions of peak NPP (between 1910 and 1950 for most simulations) were closer to the observed present-day NPP (though also with large ensemble spread), but predictions of LAI during this period were still too low.
+
+```{r pft-jitterplot, fig.cap = pft_jitterplot_cap}
+pft_jitterplot_cap <- paste(
+  htmlfig(),
+  "Fraction of total aboveground biomass (AGB)",
+  "of each plant functional type (PFT)",
+  "predicted by different parameter-configuration combionations",
+  "of ED2."
+)
+readd(pft_jitterplot_knit)
+
+```
+
+```{r pft-ternplot, fig.cap = pft_tern_cap}
+pft_tern_cap <- paste(
+  htmlfig(),
+  "Ternary plot of predicted PFT composition in 1920, 1950, and 1980.",
+  "Ternary coordinates are as follows:",
+  "L is Early Hardwood; T is the sum of Mid and Late Hardwood; R is Pine.",
+  "Triangle fill color indicates the number of points within that bin."
+)
+readd(pft_tern_knit)
+```
+
+Trajectories of species composition also varied widely by model configuration (Figure X, Y).
+Under all configurations, most ensemble members were dominanted by a single PFT, and the sustained presence of more than 2 PFTs was limited to only a handful specific ensemble members across all configurations.
+However, _which_ PFT was dominant varied depending on the parameter combination.
+
 However, the ensemble spread around the mean was large:

### 2019-10-10 analysis/drake/overall-uncertainty.R

@@ -27,24 +27,43 @@ plan <- plan <- bind_plans(plan, drake_plan(
   pft_totals = pft_data %>%
     group_by(case, year) %>%
     summarize_at(vars(-pft, -agb_frac), sum) %>%
-    pivot_longer(-c(case, year), names_to = "variable", values_to = "value"),
+    pivot_longer(-c(case, year), names_to = "variable", values_to = "value")
+))
+
+### Summary plot
+plan <- bind_plans(plan, drake_plan(
   observations = tribble(
     ~variable, ~low, ~mean, ~hi,
     "npp", 6, 6.5, 7,
     "lai", 3.97 - 1.96 * 0.423, 3.97, 3.97 + 1.96 * 0.423
   ) %>%
     expand_grid(model = models$model),
-  summary_ts_plot_gg = bind_rows(scalar_data, pft_totals) %>%
+  ts_both = bind_rows(scalar_data, pft_totals) %>%
     mutate(
       model_id = substr(case, 4, 6),
       variable = stringr::str_remove(variable, "^mmean_") %>%
         stringr::str_remove("_py$")
     ) %>%
-    filter(variable %in% c("npp", "lai")) %>%
+    filter(variable %in% c("npp", "lai")),
+  ts_summary = ts_both %>%
+    group_by(model_id, year, variable) %>%
+    summarize(
+      hi = quantile(value, 0.9),
+      lo = quantile(value, 0.1)
+    ) %>%
+    left_join(models, "model_id"),
+  summary_ts_plot_gg = ts_both %>%
     left_join(models, "model_id") %>%
     ggplot() +
-    aes(x = year, y = value, color = color, group = case) +
-    geom_line(alpha = 0.25) +
+    aes(x = year, color = color) +
+    geom_line(aes(y = value, group = case), alpha = 0.1, size = 0.3) +
+    geom_ribbon(
+      aes(ymin = lo, ymax = hi),
+      data = ts_summary,
+      color = "black",
+      fill = NA,
+      linetype = "dashed"
+    ) +
     geom_pointrange(
       aes(x = 2000, y = mean, ymin = low, ymax = hi),
       data = observations,
@@ -69,8 +88,7 @@ plan <- plan <- bind_plans(plan, drake_plan(
       axis.title = element_blank(),
       axis.text.x = element_text(angle = 90),
[…]

### 2019-10-11 analysis/drake/overall-uncertainty.R
Move PFT plots to their own R script

@@ -100,73 +100,8 @@ plan <- bind_plans(plan, drake_plan(
   ))
 ))
 
-### PFT plot
-plan <- bind_plans(plan, drake_plan(
-  pft_jitterplot_gg = pft_data %>%
-    filter(year %in% c(1910, 1920, 1950, 1980)) %>%
-    mutate(year = factor(year),
-           model_id = substr(case, 4, 6)) %>%
-    left_join(models, "model_id") %>%
-    ggplot() +
-    aes(x = year, y = agb_frac, color = pft) +
-    geom_jitter(size = 0.5, alpha = 0.2, position = position_jitterdodge()) +
-    facet_wrap(vars(model), ncol = 2) +
-    scale_color_manual(values = pfts("color")) +
-    labs(y = "AGB fraction", color = "PFT") +
-    guides(color = guide_legend(override.aes = list(alpha = 1, size = 2))) +
-    theme_cowplot() +
-    theme(
-      axis.title.x = element_blank(),
-      axis.text.x = element_text(angle = 90),
-      strip.background = element_blank(),
-      legend.position = "bottom"
-    ),
-  pft_jitterplot_png = ggsave(
-    file_out("analysis/figures/pft-jitterplot.png"),
-    pft_jitterplot_gg,
-    width = 6.6, height = 6.4
-  ),
-  pft_jitterplot_knit = knitr::include_graphics(file_in(
-    "analysis/figures/pft-jitterplot.png"
-  ))
-))
-
-### PFT ternary plot
-plan <- bind_plans(plan, drake_plan(
-  tern_data = pft_data %>%
-    mutate(model_id = substr(case, 4, 6)) %>%
-    filter(year %in% c(1920, 1950, 1980)) %>%
-    select(case, model_id, year, pft, agb_frac) %>%
-    pivot_wider(names_from = "pft", values_from = "agb_frac") %>%
-    mutate(`Mid/Late` = `Mid hardwood` + `Late hardwood`) %>%
-    select(-`Mid hardwood`, -`Late hardwood`) %>%
-    rename("Early" = "Early hardwood") %>%
-    left_join(models, "model_id"),
-  pft_tern_gg = ggplot(tern_data) +
-    aes(x = Early, y = `Mid/Late`, z = Pine) +
-    ggtern::geom_tri_tern(bins = 10) +
-    facet_grid(
-      vars(year),
-      vars(fct_relabel(model, ~gsub(" ", "\n", .)))
[…]

### 2019-10-11 analysis/drake/pft-composition-plots.R

@@ -1,0 +1,66 @@
+### PFT plot
+plan <- bind_plans(plan, drake_plan(
+  pft_jitterplot_gg = pft_data %>%
+    filter(year %in% c(1910, 1920, 1950, 1980)) %>%
+    mutate(year = factor(year),
+           model_id = substr(case, 4, 6)) %>%
+    left_join(models, "model_id") %>%
+    ggplot() +
+    aes(x = year, y = agb_frac, color = pft) +
+    geom_jitter(size = 0.5, alpha = 0.2, position = position_jitterdodge()) +
+    facet_wrap(vars(model), ncol = 2) +
+    scale_color_manual(values = pfts("color")) +
+    labs(y = "AGB fraction", color = "PFT") +
+    guides(color = guide_legend(override.aes = list(alpha = 1, size = 2))) +
+    theme_cowplot() +
+    theme(
+      axis.title.x = element_blank(),
+      axis.text.x = element_text(angle = 90),
+      strip.background = element_blank(),
+      legend.position = "bottom"
+    ),
+  pft_jitterplot_png = ggsave(
+    file_out("analysis/figures/pft-jitterplot.png"),
+    pft_jitterplot_gg,
+    width = 6.6, height = 6.4
+  ),
+  pft_jitterplot_knit = knitr::include_graphics(file_in(
+    "analysis/figures/pft-jitterplot.png"
+  ))
+))
+
+### PFT ternary plot
+plan <- bind_plans(plan, drake_plan(
+  tern_data = pft_data %>%
+    mutate(model_id = substr(case, 4, 6)) %>%
+    filter(year %in% c(1920, 1950, 1980)) %>%
+    select(case, model_id, year, pft, agb_frac) %>%
+    pivot_wider(names_from = "pft", values_from = "agb_frac") %>%
+    mutate(`Mid/Late` = `Mid hardwood` + `Late hardwood`) %>%
+    select(-`Mid hardwood`, -`Late hardwood`) %>%
+    rename("Early" = "Early hardwood") %>%
+    left_join(models, "model_id"),
+  pft_tern_gg = ggplot(tern_data) +
+    aes(x = Early, y = `Mid/Late`, z = Pine) +
+    ggtern::geom_tri_tern(bins = 10) +
+    facet_grid(
+      vars(year),
+      vars(fct_relabel(model, ~gsub(" ", "\n", .)))
+    ) +
+    scale_fill_viridis_c(na.value = "white") +
[…]

### 2019-10-11 analysis/paper/paper.Rmd
Move PFT plots to their own R script
Add pairs plot figure and text

@@
 Under all configurations, most ensemble members were dominanted by a single PFT, and the sustained presence of more than 2 PFTs was limited to only a handful specific ensemble members across all configurations.
-However, _which_ PFT was dominant varied depending on the parameter combination.
-
-However, the ensemble spread around the mean was large:
-across all configurations (except the finite canopy - two-stream radiative transfer cases), the spread in net primary productivity was from zero to nearly twice as high as the observation.
-Ensemble variance in leaf area index was lower on average but, in some cases (especially the combination of finite canopy and multiple-scatter radiative transfer) was over an order of magnitude.
-The predicted Shannon diversity index of plant functional types was, on average, higher in simulations when the canopy architecture was finite rather than closed.
+However, _which_ PFT was dominant varied depending on the parameter combination, and any of the four PFTs became the dominant one some subset of simulations.
 
@@
   htmlfig(),
-  "Post-1975 growing season averaged NPP vs. LAI, with observation (in black)",
-  "and linear regression, by model type.",
-  "Each point indicates a single ensemble run with a different parameter combination.",
-  "Letter labels are selected runs with the same parameters,",
-  "and match the rows in Figure 4."
+  "ED2 predictions of annual mean NPP and growing-season LAI in 1920, 1950, and 1999",
+  "by model type.",
+  "Each point indicates a single ensemble run with a different set of parameters.",
+  "Solid line is the best-fit generalized additive model (GAM).",
+  "Dashed lines indicate the error bounds of present-day observed NPP and LAI."
 )
-readd(pairs_time_averaged_knit)
+readd(npp_lai_pairs_knit)
 ```
 
-<!-- TODO Use peak instead of post-1975? -->
 <!-- TODO Use a logarithmic, rather than a linear fit? -->
 
-Model configuration affected the emergent relationship between net primary productivity and leaf area index (Figure 3).
-The crown model (closed vs. finite) had the strongest effect:
-Compared to the simulations with "closed" crowns, simulations with finite canopy radius generally had lower net primary productivity for the same leaf area index.
-Although many simulations matched _either_ the observed net primary productivity or the observed leaf area index, only a few individual runs matched _both_.
+The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs.
+Specifically, compared to runs with the default closed canopy, runs with a finite canopy were more likely to have higher LAI values for the same level of NPP, especially later into the simulation.
+Although many simulations matched _either_ the observed net primary productivity or the observed leaf area index, only a small minority of individual runs matched _both_.
 

### 2019-10-11 analysis/drake/overall-uncertainty.R

@@ -102,80 +102,6 @@ plan <- bind_plans(plan, drake_plan(
 
 ### STOP HERE
 stop()
-### Pairs plots
-pft_wide = pft_data %>%
-  mutate(model_id = substr(case, 4, 6)) %>%
-  select(case, pft, year, model_id, agb_frac) %>%
-  setDT() %>%
-  dcast(case + year + model_id ~ pft, value.var = "agb_frac") %>%
-  as_tibble()
-
-both_wide = scalar_data %>%
-  setDT() %>%
-  dcast(case + year ~ variable, value.var = "value") %>%
-  as_tibble() %>%
-  left_join(pft_wide, c("case", "year"))
-
-both_wide %>%
-  filter(year == 1920) %>%
-  select(model_id, starts_with("mmean"), `Early hardwood`:`Pine`) %>%
-  GGally::ggpairs(
-    .,
-    aes(color = model_id),
-    size = 0.5,
-    alpha = 0.2,
-    columns = names(.)[-1]
-  )
-
-tern_data %>%
-  filter(year == 1920, model_id == "FMS") %>%
-  ggplot() +
-  aes(x = Early, y = `Mid/Late`, z = Pine) +
-  ggtern::geom_tri_tern(bins = 10) +
-  scale_fill_viridis_c(na.value = "white") +
-  ggtern::coord_tern()
-
-get_density <- function(x, y, z, ...) {
-  dens <- misc3d::kde3d(x, y, z, ...)
-  ix <- findInterval(x, dens$x)
-  iy <- findInterval(y, dens$y)
-  iz <- findInterval(z, dens$z)
-  ii <- cbind(ix, iy, iz)
-  dens$d[ii]
-}
-
-pft_tern_pl
-
-tern_data2 <- tern_data %>%
-  group_by(model_id) %>%
-  mutate(dens = get_density(`Early hardwood`, `ML`, `Pine`)) %>%
-  ungroup()
-ggplot(tern_data2) +
-  aes(x = `Early hardwood`, y = `ML`, z = `Pine`, color = dens) +
-  geom_point() +
-  facet_wrap(vars(model_id), ncol = 2) +
-  scale_color_viridis_c() +
-  ggtern::coord_tern()
-
-
-pft_data %>%
-  mutate(model_id = substr(case, 4, 6)) %>%
-  left_join(models, "model_id") %>%
-  ggplot() +
-  aes(x = year, y = agb_frac, group = case, color = color) +
[…]

### 2019-10-11 analysis/drake/pairs-plots.R

@@ -1,0 +1,115 @@
+### Pairs plots
+npp_lai_pairs <- function(dat, year) {
+  dat %>%
+    filter(year == !!year) %>%
+    ggplot() +
+    aes(x = mmean_lai_py, y = mmean_npp_py, color = color) +
+    geom_point(size = 0.8) +
+    geom_smooth(method = mgcv::gam, color = "black") +
+    geom_hline(yintercept = c(6, 7), linetype = "dashed") +
+    geom_vline(xintercept = 3.97 + c(-1, 1) * 1.96 * 0.423,
+               linetype = "dashed") +
+    scale_color_identity() +
+    facet_wrap(vars(model), ncol = 2) +
+    labs(x = "Total LAI", y = expression(NPP ~ (MgC ~ ha^-1 ~ year^-1))) +
+    ggtitle(year) +
+    theme_cowplot() +
+    theme(strip.text = element_text(size = 10))
+}
+
+plan <- bind_plans(plan, drake_plan(
+  pft_wide = pft_data %>%
+    mutate(model_id = substr(case, 4, 6)) %>%
+    select(case, pft, year, model_id, agb_frac) %>%
+    setDT() %>%
+    dcast(case + year + model_id ~ pft, value.var = "agb_frac") %>%
+    as_tibble(),
+  lai_wide = pft_totals %>%
+    filter(variable == "mmean_lai_py") %>%
+    select(case, year, mmean_lai_py = value),
+  both_wide = scalar_data %>%
+    setDT() %>%
+    dcast(case + year ~ variable, value.var = "value") %>%
+    as_tibble() %>%
+    left_join(pft_wide, c("case", "year")) %>%
+    left_join(lai_wide, c("case", "year")) %>%
+    mutate(model_id = substr(case, 4, 6)) %>%
+    left_join(models, "model_id"),
+  npp_lai_pairs_yr = target(
+    npp_lai_pairs(both_wide, .year),
+    transform = map(.year = c(1920, 1950, 1999))
+  ),
+  npp_lai_pairs_gg = target(
+    plot_grid(npp_lai_pairs_yr, nrow = 1),
+    transform = combine(npp_lai_pairs_yr)
+  ),
+  npp_lai_pairs_png = ggsave(
+    file_out("analysis/figures/npp-lai-pairs.png"),
+    npp_lai_pairs_gg,
+    width = 12, height = 5.3
[…]

### 2019-10-11 analysis/paper/paper.Rmd
Add PFT LAI figure and update related text

@@
   "Solid line is the best-fit generalized additive model (GAM).",
-  "Dashed lines indicate the error bounds of present-day observed NPP and LAI."
+  "Dashed lines indicate the error bounds of present-day observed NPP and LAI.",
+  "Text labels match the rows in Figure X below."
 )
@@
 
-```{r bigfacet-plot, fig.cap = bigfacet_cap}
-bigfacet_cap <- paste(
+```{r lai-pft-plot, fig.cap = lai_pft_cap}
+lai_pft_cap <- paste(
   htmlfig(),
@@
 
-`r mdcap(bigfacet_cap)`
-
-Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure 4).
-Consider simulations **A** and **C**, the former showing competition of early hardwood with pine, and the latter completely dominated by pine:
-With a closed canopy and static traits, both simulations produced similar overall LAI and NPP values (Figure 3), but moving to a finite canopy dramatically increased the total LAI in **C** without much affecting **A**, while enabling trait plasticity dramatically increased the NPP of **A** without much affecting **C** (Figure 3).
-This example also shows the impacts of different configurations on predicted community dynamics -- namely, in this specific case, enabling trait plasticity helped pine out-perform early hardwood, but only under a closed (rather than finite) canopy.
-Finite canopy radius similarly benefits early hardwood in its competition against late hardwood in simulations **B** and, to a lesser extent, **D**.
-Simulations **B** and **D** also show that the sensitivity of a given parameter combination to model structure is itself dependent on the parameters -- simulation **B** produced similar ecological dynamics and aggregate state variables regardless of model configuration (except the extreme cases; Figure 3), while simulation **D** had near-zero NPP and LAI in two cases (closed, two-stream, static and finite, multi-scatter, plastic), comparable NPP and LAI to **B** with closed canopy and trait plasticity, and higher NPP and LAI than **D** with a finite canopy, multi-scatter radiative transfer, and static traits (Figure 3).
+Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure X).
+Parameter sets **A** and **B** both resulted in similar patterns of immediate and complete dominance by mid- and late-hardwood, respectively, followed by gradual decline to near-zero LAI.
+In both of these simulations, peak productivity was higher and was sustained for longer when the canopy was closed rather than finite, and this difference was larger under a two-stream rather than a multiple-scatter radiative transfer model.
+Trait plasticity had relatively little effect on most of these two simulations, though in parameter set **B**, under the combination of closed canopy, multiple-scatter radiative transfer, and plastic traits, pine overtook the dominant late hardwood PFT starting around 1975.
+Both of these parameter sets also had the lowest LAI of the five selected here, with a peak LAI of 1 in **A** and 2 in **B**.
+Parameter set **C** showed much more variety in compositional dynamics with model structure:
+In all cases, mid-hardwoods began as the dominant species and were eventually overtaken by pine under a closed canopy and plastic traits, or early hardwood in all other cases (though with a closed canopy and static traits, the final LAI was low).
+Parameter sets **D** and **E** had the highest productivity of these five sets, and were characterized by strong dominance of Early Hardwood and Pine, respectively.
 

### 2019-10-11 analysis/drake/pairs-plots.R

@@ -1,20 +1,28 @@
 ### Pairs plots
-npp_lai_pairs <- function(dat, year) {
-  dat %>%
-    filter(year == !!year) %>%
-    ggplot() +
-    aes(x = mmean_lai_py, y = mmean_npp_py, color = color) +
-    geom_point(size = 0.8) +
-    geom_smooth(method = mgcv::gam, color = "black") +
+npp_lai_pairs <- function(dat, year, labs) {
+  dat_sub <- dat %>%
+    filter(!is.na(model), year == !!year)
+  labs2 <- labs %>%
+    inner_join(dat_sub, "param_id") %>%
+    distinct(param_id, model, label, mmean_lai_py, mmean_npp_py)
+
+  ggplot(dat_sub) +
+    aes(x = mmean_lai_py, y = mmean_npp_py) +
+    geom_point(aes(color = color), size = 0.7, alpha = 0.7) +
+    geom_smooth(method = mgcv::gam, color = "gray30") +
     geom_hline(yintercept = c(6, 7), linetype = "dashed") +
     geom_vline(xintercept = 3.97 + c(-1, 1) * 1.96 * 0.423,
                linetype = "dashed") +
+    geom_point(color = "black", data = labs2) +
+    geom_label_repel(aes(label = label), data = labs2,
+                     min.segment.length = 0) +
     scale_color_identity() +
     facet_wrap(vars(model), ncol = 2) +
     labs(x = "Total LAI", y = expression(NPP ~ (MgC ~ ha^-1 ~ year^-1))) +
     ggtitle(year) +
     theme_cowplot() +
     theme(strip.text = element_text(size = 10))
+
 }
 
 plan <- bind_plans(plan, drake_plan(
@@ -33,10 +41,14 @@ plan <- bind_plans(plan, drake_plan(
     as_tibble() %>%
     left_join(pft_wide, c("case", "year")) %>%
     left_join(lai_wide, c("case", "year")) %>%
-    mutate(model_id = substr(case, 4, 6)) %>%
-    left_join(models, "model_id"),
+    left_join(
+      cases %>%
+        select(case, model_id, param_id),
+      c("case", "model_id")
+    ) %>%
+    left_join(models, c("model_id")),
   npp_lai_pairs_yr = target(
-    npp_lai_pairs(both_wide, .year),
+    npp_lai_pairs(both_wide, .year, use_params),
[…]

### 2019-10-11 analysis/drake/specific-parameters.R

@@ -1,0 +1,68 @@
+plan <- bind_plans(plan, drake_plan(
+  use_params = tribble(
+    ~param_id, ~label, ~why,
+    399, "A", "MH-dominated in 1999, high NPP",
+    39,  "B", "LH-dominated in 1999, high NPP",
+    362, "C", "High competition in 1999, high NPP",
+    284, "D", "EH-dominated in 1999, high NPP",
+    114, "E", "Pine-dominated in 1999, high NPP"
+  ),
+  lai_pft_plot_gg = pft_data %>%
+    mutate(param_id = as.numeric(substr(case, 0, 3))) %>%
+    inner_join(use_params, "param_id") %>%
+    inner_join(cases, c("case", "param_id")) %>%
+    ggplot() +
+    aes(x = year, y = mmean_lai_py, color = pft) +
+    geom_line() +
+    facet_grid(
+      vars(label),
+      vars(fct_relabel(model, ~gsub(" ", "\n", .x))),
+      scales = "free_y"
+    ) +
+    labs(y = "Leaf area index", color = "PFT") +
+    scale_color_manual(values = pfts("color")) +
+    theme_cowplot() +
+    theme(
+      axis.title.x = element_blank(),
+      axis.text.x = element_text(angle = 90),
+      legend.position = "bottom"
+    ),
+  lai_pft_plot_png = ggsave(
+    file_in("analysis/figures/lai-pft.png"),
+    lai_pft_plot_gg,
+    width = 9.7, height = 6.1
+  ),
+  lai_pft_plot_knit = knitr::include_graphics(file_in(
+    "analysis/figures/lai-pft.png"
+  ))
+))
+
+### STOP HERE
+stop()
+
+dev.size()
+
+all_params <- both_wide %>%
+  filter(year == 1999) %>%
+  mutate(param_id = as.numeric(substr(case, 0, 3))) %>%
+  count(param_id) %>%
+  filter(n == 8)
+
+both_wide %>%
+  mutate(param_id = as.numeric(substr(case, 0, 3))) %>%
+  semi_join(all_params, "param_id") %>%
+  rename(EH = `Early hardwood`, MH = `Mid hardwood`, LH = `Late hardwood`) %>%
+  rename_all(~gsub("mmean_", "", .x)) %>%
+  rename_all(~gsub("_py", "", .x)) %>%
+  mutate(
[…]

### 2019-10-14 analysis/paper/paper.Rmd
Compile paper separately from drake
Makes it easier to work on it on different machines.

@@
 
-```{r setup, echo = FALSE}
+```{r setup, include = FALSE}
 library(knitr)
 library(here)
-library(drake)
+
+fig <- function(...) {
+  knitr::include_graphics(here::here("analysis", "figures", ...))
+}
 
@@
 ```{r param-table}
-tbl <- readd(param_table)
+tbl <- readr::read_csv(here(
+  "analysis", "data", "derived-data", "parameter-table.csv"
+))
 tbl <- subset(tbl, Ran == "yes", select = c(-unit_parsed, -Ran))
 tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-knitr::kable(tbl, escape = FALSE, caption = paste0(
+kable(tbl, escape = FALSE, caption = paste0(
   "Table 2: Parameter descriptions. ",
@@
 Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
-The resulting parameter distributions are shown in Figure 1.
-<!-- TODO more on this? -->
 All parameters not described in this section were set to their ED-2.2 PFT-specific default values.
@@
 
-## Impacts of structure
+## Impacts of structure alone
 
@@
 structure_default_cap <- paste(
+  htmlfig(),
   "Annual NPP and LAI from ED2 runs",
@@
 )
-readd(structure_compare_default_knit)
+fig("structure-compare-default.png")
 ```
@@
 structure_default_light_cap <- paste(
+  htmlfig(),
   "Radiation profiles of cohorts from ED2 runs with",
-  "default parameters, across different configurations."
-  "Each point represents a single cohort of the given PFT,"
-  "position on the y-axis represents the height of that cohort,"
+  "default parameters, across different configurations.",
+  "Each point represents a single cohort of the given PFT,",
+  "position on the y-axis represents the height of that cohort,",
   "and position on the x-axis represents the relative light level",
@@
 )
-readd(structure_default_light_knit)
+fig("default-light-levels.png")
 ```
@@
 )
-readd(param_dist_knit)
+fig("param-dist.png")
 ```
-
-`r mdcap(param_dist_cap)`
 
@@
 
-```{r , fig.cap = default_median_cap}
-default_median_cap <- paste(
+```{r default-median-flux, fig.cap = default_median_flux_cap}
+default_median_flux_cap <- paste(
+  htmlfig(),
   "Comparing results of ED2 with default parameters and",
@@
 )
-readd(default_median_fluxes_knit)
+fig("default-median-fluxes.png")
 ```
 
-```{r}
-default_median_cap <- paste(
+```{r default-median-lai , fig.cap = default_median_lai_cap}
+default_median_lai_cap <- paste(
+  htmlfig(),
   "Comparing results of ED2 PFT-level LAI",
@@
 )
-readd(default_median_lai_knit)
+fig("default-median-lai.png")
 ```
@@
 
-## Summary
+## Structure vs. parameter uncertainty
 
@@
 )
-readd(summary_ts_plot_knit)
+fig("summary-ts-plot.png")
 ```
 
-`r mdcap(summary_ts_cap)`
-
 <!-- - Structure x parameter uncertainty -->
-<!--   - Figure: Time series of NPP, LAI, and relative PFT dominance -->
 <!--   - Parameter uncertainty is much larger than structural uncertainty -->
@@
 The ensemble spread within any single configuration---i.e.
-the parameter uncertainty---was similar across all configurations; both NPP and LAI varied 5-10 fold at any given timestep depending on parameter values.
+the parameter uncertainty---was similar and large across all configurations; both NPP and LAI varied 5-10 fold at any given timestep depending on parameter values.
 Within all configurations, most (>90%) ensemble members under-predicted observed present-day NPP and LAI by year 2000, though individual ensemble members from some configurations over-predicted one or both variables.
@@
 )
-readd(pft_jitterplot_knit)
-
+fig("pft-jitterplot.png")
 ```
@@
 )
-readd(pft_tern_knit)
+fig("pft-tern.png")
 ```
@@
 )
-readd(npp_lai_pairs_knit)
+fig("npp-lai-pairs.png")
 ```
@@
 )
-readd(lai_pft_plot_knit)
+fig("lai-pft.png")
 ```
@@
 
-```{r within-vs-across, fig.cap = within_vs_across_cap, fig.dim = c(9, 4)}
+```{r within-vs-across, fig.cap = within_vs_across_cap, fig.dim = c(9, 4), eval = FALSE}
 within_vs_across_cap <- paste(
@@
 
-`r mdcap(within_vs_across_cap)`
-
 Except for the configurations with persistently low productivity, parameter uncertainty (across ensembles within each model configuration) was much larger than structural uncertainty (variance in ensemble means across configurations; Figure 4).
@@
 
-```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap}
+```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, eval = FALSE}
 uncertainty_analysis_cap <- paste(
   htmlfig(),
-  "PEcAn-like parameter sensitivity and uncertainty analysis, by model type,",
+  "Parameter sensitivity and uncertainty analysis, by model type,",
   "for 1975-2000 averaged growing season net primary productivity (left) and leaf area index (right).",
@@
 ```
-
-`r mdcap(uncertainty_analysis_cap)`
 

### 2019-10-14 analysis/drake.R

@@ -64,19 +64,6 @@ walk(src_files, possibly(source, NULL))
 
 if ("--poster" %in% cmdargs) source("analysis/drake_poster.R")
 if ("--postgrad" %in% cmdargs) source("analysis/drake_postgrad.R")
-if ("--paper" %in% cmdargs) {
-  plan <- bind_plans(plan, drake_plan(
-  param_table = file_in(!!(path(data_dir, "derived-data",
-                                "parameter-table.csv"))) %>%
-    read_csv(),
-    paper = target(
-      rmarkdown::render(
-        knitr_in(!!(path(analysis_dir, "paper", "paper.Rmd"))),
-        .format
-      ),
-      transform = map(.format = c("html_document", "pdf_document")))
-    ))
-}
 
 # Parallelism configuration. Not sure which of these is better...
 future::plan(future.callr::callr)

### 2019-10-14 analysis/paper/paper.Rmd
Draft partial variance text

@@
 
+<!--
 Except for the configurations with persistently low productivity, parameter uncertainty (across ensembles within each model configuration) was much larger than structural uncertainty (variance in ensemble means across configurations; Figure 4).
@@
 Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased variance in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on variance in productivity productivity.
+-->
 
-```{r uncertainty-analysis, fig.cap = uncertainty_analysis_cap, eval = FALSE}
-uncertainty_analysis_cap <- paste(
+```{r partial-variance-npp, fig.cap = partial_variance_npp_cap}
+partial_variance_npp_cap <- paste(
   htmlfig(),
-  "Parameter sensitivity and uncertainty analysis, by model type,",
-  "for 1975-2000 averaged growing season net primary productivity (left) and leaf area index (right).",
-  "Elasticity (top) is the normalized sensitivity of the model to a fixed change in the parameter.",
-  "Partial variance (bottom) describes the overall contribution of the parameter to model predictive uncertainty based on the combination of parameter uncertainty and model sensitivity.",
-  "Each panel shows the top 10 parameter values, in terms of the absolute value of the corresponding quantity.",
+  "Contributions of individual parameters to predicted uncertainty in NPP,",
+  "by model type.",
+  "Panels on the left are for the \"peak\" NPP period (averaged over 1925-1950),",
+  "and on the right are for the end of the simulation (averaged over 1975-1999)",
+  "Each panel shows the top 8 parameter values in terms of their variance contribution.",
   "Input parameter distributions are shown in Figure 1.",
@@
 )
-readd(sensitivity_plot_knit)
+fig("partial-variance-npp.png")
 ```
 
-Model sensitivity to specific parameters ["elasticity" _sensu_ @lebauer_2013_facilitating] varied depending on the model configuration and the target output variable (Figure 5).
-The parameter with the single highest elasticity to NPP and LAI across all model configurations was specific leaf area (SLA), though the PFT whose SLA was most important varied with model configuration.
-Other frequently (but not always) important parameters were the leaf C:N ratio (`c2n_leaf`), ratio of fine root to leaf biomass (`fineroot2leaf`), leaf and root respiration rates, and leaf optical properties (`leaf_reflect_vis/nir`).
+```{r partial-variance-lai, fig.cap = partial_variance_lai_cap}
+partial_variance_lai_cap <- paste(
+  htmlfig(),
+  "Same as above, but for total leaf area index (LAI)."
+)
+fig("partial-variance-lai.png")
+```
+
+<!-- Model sensitivity to specific parameters ["elasticity" _sensu_ @lebauer_2013_facilitating] varied depending on the model configuration and the target output variable (Figure 5). -->
+<!-- The parameter with the single highest elasticity to NPP and LAI across all model configurations was specific leaf area (SLA), though the PFT whose SLA was most important varied with model configuration.
+-->
+<!-- Other frequently (but not always) important parameters were the leaf C:N ratio (`c2n_leaf`), ratio of fine root to leaf biomass (`fineroot2leaf`), leaf and root respiration rates, and leaf optical properties (`leaf_reflect_vis/nir`).
+-->
 <!-- TODO Dig into model configurations. -->
 
-The parameters that contributed most to overall model predictive uncertainty ("partial variance") differed from parameters to which the model was most sensitive (Figure 5), largely driven by the fact that many of the parameters with the highest sensitivity (e.g.
+The parameters that contributed most to overall model predictive uncertainty ("partial variance") varied with model structure, time period considered, and output variable (NPP or LAI).
+In general, uncertainties in pine growth respiration factor and early hardwood plant-soil water conductance contributed the most to parameter uncertainty in NPP during the "peak" period (1925-1950).
+{TODO: Fill in the rest of this text...}
+
+<!-- This is old text... -->
+<!-- ...differed from parameters to which the model was most sensitive (Figure 5), largely driven by the fact that many of the parameters with the highest sensitivity (e.g.
 SLA, V~cmax~) were also well constrained by data (Figure 1).
-Leaf optical properties of at least one PFT were among the top 10 most important parameters across all configurations, and in multiple cases, were the most important.
-Other often important parameters were plant-soil water conductance (`water_conductance`), parameters related to adult mortality (`mort1,2,3`), and growth respiration.
-Again, patterns in partial variance were largely idiosyncratic across model configurations.
+-->
+<!-- Leaf optical properties of at least one PFT were among the top 10 most important parameters across all configurations, and in multiple cases, were the most important. -->
+<!-- Other often important parameters were plant-soil water conductance (`water_conductance`), parameters related to adult mortality (`mort1,2,3`), and growth respiration. -->
+<!-- Again, patterns in partial variance were largely idiosyncratic across model configurations. -->
 <!-- TODO Dig into this more -->
 <!-- TODO Check this analysis. Look at pairs plots of trait vs. variable, and my SA spline fits. -->
+
+{TODO: Basic analysis of contributions from parameter interactions, a la Raczka et al. 2018? I expect these will be important.}
 

### 2019-10-14 analysis/drake/sensitivity-analysis.R

@@ -111,9 +111,93 @@ stop()
 
 saveRDS(sensitivity_raw_results, "analysis/data/retrieved/sensitivity-raw-results.rds")
 
+library(tidyverse)
+library(here)
+library(fortebaseline)
+sensitivity_raw_results <- readRDS(here(
+  "analysis",
+  "data",
+  "retrieved",
+  "sensitivity-raw-results.rds"
+))
+
 sensitivity_proc = sensitivity_raw_results %>%
   unnest(result) %>%
-  unnest(result)
+  unnest(result) %>%
+  ungroup()
+
+sensitivity_sub <- sensitivity_proc %>%
+  filter(
+    grepl("mmean_", yvar),
+    is.finite(sensitivity),
+    is.finite(elasticity)
+  ) %>%
+  group_by(model_id, yvar) %>%
+  mutate(fpvar = pvar / sum(pvar)) %>%
+  group_by(xvar) %>%
+  mutate(total_pvar = sum(pvar)) %>%
+  ungroup() %>%
+  separate(xvar, c("shortname", "xvar"), extra = "merge") %>%
+  mutate(
+    shortname = factor(shortname, pfts("shortname")),
+    model = factor(model_id, c("CTS", "CTP", "CMS", "CMP",
+                               "FTS", "FTP", "FMS", "FMP"))
+  )
+
+sensplot <- function(dat, sgroup, metric, yvar) {
+  metric <- rlang::enquo(metric)
+  dat %>%
+    filter(sgroup == !!sgroup) %>%
+    top_n_sensitivity_plot(yvar, !!metric, 8, "free") +
+    facet_wrap(vars(model), scales = "free", drop = TRUE, ncol = 2,
+               dir = "v") +
+    scale_color_manual(values = pfts("color")) +
+    labs(color = "PFT") +
+    cowplot::theme_cowplot() +
+    theme(
+      axis.title.y = element_blank(),
+      axis.text.x = element_text(size = 10),
+      axis.text.y = element_text(size = 12)
+    )
+}
+
+pn1 <- sensplot(sensitivity_sub, "1920-1950", fpvar, "mmean_npp_py") +
+  ggtitle("1920-1950") +
+  labs(y = "Partial variance") +
+  guides(color = FALSE)
+pn2 <- sensplot(sensitivity_sub, "1975-1999", fpvar, "mmean_npp_py") +
[…]

### 2019-10-24 analysis/paper/paper.Rmd
A few minor revisions from meeting with Ben

@@
 <!-- TODO This paragraph is janky as hell. Needs heavy revision. -->
+(Chris/Jeff: Help)
 Previous work found that an earlier version of ED had difficulty sustaining multiple plant functional types in long simulations [@fisher_2010_assessing].
-One of the primary drivers of successional change in temperate forests is competition for light <!-- TODO REF -- Bazaaz?
+One of the primary drivers of successional change in temperate forests is competition for light (REF Bazzaz 19XX) <!-- TODO REF -- Bazaaz?
 -->, and differences in plants' strategies for light acquisition and utilization are a key driver of niche differentiation in these ecosystems <!-- TODO REF -->.
+Plants compete for light in three different ways:
+Growing quickly (early successional, pioneer species),
+structural complexity / morphological change / filling in gaps in existing canopies,
+and adapting traits to maximize light (and other resource) capture and utilization.
 To this end, we focused our structural uncertainty analysis on three processes related to competition for light:
@@
 complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
-We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results.
+We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results in the context of ongoing climate change.
 
@@
 
-Different model structures very closely for the first half of the simulation, at which point they started to diverge.
+Different model structures agree for the first half of the simulation, at which point they started to diverge.
 Post 1950, the combination of two-stream radiative transfer, finite canopy radius, and plastic traits had the highest sustained productivity, followed by the same configuration except with static traits, with the remaining configurations having very similar results.
+Interestingly, the effects of changing individual model configurations are not additive --- for example, enabling trait plasticity had a different effect depending on whether the two-stream or multiple-scatter canopy RTM was used.
 
@@
 
-```{r partial-variance-lai, fig.cap = partial_variance_lai_cap}
-partial_variance_lai_cap <- paste(
-  htmlfig(),
-  "Same as above, but for total leaf area index (LAI)."
-)
-fig("partial-variance-lai.png")
-```
-
 <!-- Model sensitivity to specific parameters ["elasticity" _sensu_ @lebauer_2013_facilitating] varied depending on the model configuration and the target output variable (Figure 5). -->
@@
 
-The parameters that contributed most to overall model predictive uncertainty ("partial variance") varied with model structure, time period considered, and output variable (NPP or LAI).
+The parameters that contributed most to overall model predictive uncertainty ("partial variance") varied with model structure and time period considered.
 In general, uncertainties in pine growth respiration factor and early hardwood plant-soil water conductance contributed the most to parameter uncertainty in NPP during the "peak" period (1925-1950).
 {TODO: Fill in the rest of this text...}
+
+Parameter sensitivity for total LAI was similar to that of NPP (results not shown).
 

### 2019-10-24 analysis/paper/paper.Rmd
Few more minor tweaks from Ben meeting

@@
 Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
-**Ensemble mean trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure 2);
+**Ensemble mean trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure TODO);
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
@@
 **Parameter uncertainty was the dominant driver of uncertainty in our analysis**.
-Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables** -- NPP ranged from 0 to almost 2x the observed, and LAI was up to 3x observed (though more tightly clustered around zero) (Figure 2).
-Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figure 4).
+Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables** -- NPP ranged from 0 to almost 2x the observed, and LAI was up to 3x observed (though more tightly clustered around zero) (Figure TODO).
+Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figure TODO).
 **Parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.
-The dominance of parameter uncertainty over structural uncertainty in our results suggests that **targeted data collection should be a higher priority than model development**.
 
-**Parameter uncertainty is not synonymous with sensitivity**.
-Many of the parameters to which model outputs are most sensitive are already well constrained by data and do not contribute as much to predictive uncertainty (Figure 6).
-Although parameter uncertainty is still large, this means that **targeted data collection works and that we are moving in the right direction**!
-This speaks to the value of looking at not just sensitivity, but partial variance [@lebauer_2013_facilitating; @dietze_2014_quantitative].
+The dominance of parameter uncertainty over structural uncertainty in our results suggests that **constraining parameter uncertainty should be a higher priority than model development**.
+Parameter uncertainty can be constrained in two different ways:
+Using observational data if available;
+and, for parameters "that cannot be directly observed", using "quantitative/algorithmic" model calibration techniques (e.g. Fer, other papers).
+For example, leaf optical properties (i.e. reflectance, transmittance) are both measurable and highly influential, and therefore should be a priority for new data collection.
+In contrast, "growth respiration factor" is an empirical coefficient and not measurable, but highly influential (c.f. Dietze, Raczka), and therefore should be constrained through calibration.
 
+(
 So **which parameters are a research priority**?
@@
 (Compare also to @raczka_2018_what, @pandit_2018_optimizing).
+)
+
+**Parameter uncertainty is not synonymous with sensitivity**.
+Many of the parameters to which model outputs are most sensitive are already well constrained by data and do not contribute as much to predictive uncertainty (Figure TODO).
+SLA and Vcmax fall into this category.
+Although parameter uncertainty is still large, this means that **targeted data collection works and that we are moving in the right direction**!
+This speaks to the value of looking at not just sensitivity, but partial variance [@lebauer_2013_facilitating; @dietze_2014_quantitative].
 
 Although parameter uncertainty dominated, **structural and parameter uncertainty interacted in important ways**.
-The relative position of different parameter combinations in multivariate space (Figure 3), the total amount of parameter uncertainty (Figure 5), and the relative importance of different parameters to predictive uncertainty (Figure 6) all differed across model structures (Figure 3).
-This is despite the fact that our model structure changes were relatively modest compared to differences between different models.
+Individual model structure changes had relatively modest impacts.
+
+* Interactive effects of different structures -- not additive
+* Interactions between parameter and structural uncertainty -- 
+  - For example, the relative importance of different parameters to predictive uncertainty (Figure TODO) differed across model structures (Figure 3).
+
 Collectively, this means that **sensitivity and uncertainty analyses are not necessarily generalizable across models**.

### 2019-10-25 analysis/paper/paper.Rmd
Add more results to abstract

@@
   We then compared the magnitude of structural (across representations) and parameter (within each representation) uncertainty, and assessed which submodel-parameter combinations reproduced observed C fluxes and succession.
-  Our simulations agreed well with observed net primary productivity but systematically underestimated leaf area index.
-  Overall, parameter uncertainty contributed more than structural uncertainty to total simulation uncertainty, but the amount of parameter uncertainty varied with model structure.
+  On average, our simulations underestimated both net primary productivity (NPP) and leaf area index (LAI) after 100 simulation years, though NPP estimates at peak productivity (20-50 years into the simulations) were closer to observations.
+  However, parameter uncertainty was large; at any given timestep, both NPP and LAI ranged from ~0 to >200% of their mean value.
+  Although many simulations matched either the observed NPP or observed LAI at some point during the simulation, only a small minority of individual runs matched both.
+  Structural uncertainty was less important than parameter uncertainty, but model configuration did influence the ecological dynamics within a given parameter set, and affected which parameters were most influential.
   We conclude that: (1) targeted data collection to constrain vegetation model parameters should be a higher priority than model development; (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and (3) robust model validation requires scrutiny of multiple different variables, including those related to vegetation productivity, structure, and composition.

### 2019-10-25 analysis/paper/paper.Rmd
Incorporate Ben's comments about introduction

@@
 However, the **fate of these forests is highly uncertain**.
-For one, many Upper Midwest forests are undergoing a **shift in community composition from early-successional to mid-successional** species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
+Many Upper Midwest forests are undergoing a **shift in community composition from early-successional to mid-successional** species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
 In addition, these forests face a multitude of **direct and indirect pressures from both biotic and abiotic sources** [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.
@@
 <!-- TODO Might need a better ref here -->
-On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site.
-A common application of these models was simulating the land surface boundary condition for atmospheric general circulation models, though similar models have been (and continue to be) applied successfully to simulate biogeochemical processes in many ecosystem science contexts.
-<!-- TODO refs -->
+On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site <!-- TODO REFS -->
 On the other end of the spectrum are **"individual-based" models** (a.k.a.
@@
 Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
-Between these two extremes are models that use approximations attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
+Between these two extremes are models that use approximations to attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
 One specific example of this approach is the **Ecosystem Demography model** [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
+<!-- BBL: Perhaps mention ED2's influence in ESMs and LSMs -->
 
@@
 **The relative importance of these sources of uncertainty in ecological forecasts is a key question because it is directly related to future research priorities**.
-For example, in the atmospheric science community, the insight that meteorological forecasts were most sensitive to initial condition uncertainty (Lorenz REF<!-- TODO others?
--->) led to a multi-decadal research agenda aimed at constraining the initial state with improved observations, which has directly contributed to a steady and persistent improvement in meteorological forecasts over the last several decades.
-<!-- TODO Holy run-on sentences, Batman!
---> <!-- TODO REFS!
--->
 
 **Many ecosystem modeling studies have looked at these categories of uncertainty independently.**
-Perhaps the most work has been done on **parameter uncertainty** in ecosystem models.
 <!-- TODO Consider cutting 50-75% of these next few sentences. Probably way too much detail. -->
@@
 More recently, @raczka_2018_what evaluated parameter uncertainty over 100 year time scale at the Willow Creek Ameriflux site, and found that the most important parameters to model uncertainty were quantum efficiency of photosynthesis, leaf respiration, and soil-plant water transfer.
-Another recent study assessed parameter uncertainty in ED2 as part of its overall goal of parameterizing the model for sagebrush [@pandit_2018_optimizing].
-They found that the most important parameters were specific leaf area, the maximum rate of carbon fixation (V~cmax~), slope of stomatal conductance, the fine-root to leaf carbon ratio, and the fine root turnover rate.
 Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large impact on predictions of net primary productivity [@viskari_2019_influence].
 {**Driver uncertainty** -- a few examples?}
+
 **Other studies have investigated structural uncertainty.**.
@@
 
-Ben's notes:which of the last two categories--structural versus parametric uncertainty--is more important (Fisher et al.
+Ben's notes:which of the last two categories--structural versus parameter uncertainty--is more important (Fisher et al.
 2015, https://onlinelibrary.wiley.com/doi/full/10.1111/j.1523-1739.2003.00614.x, Ricciuto et al.
@@
 
-**This study focuses on the interaction between parametric and structural uncertainty in the ED2 model.**
+**This study focuses on the interaction between parameter and structural uncertainty in the ED2 model.**
 Our study is organized around the following guiding questions:
 (1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
-(2) What is the cost of considering these processes, in terms of additional parametric uncertainty?
-(3) What are the relative contributions of parametric uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-To answer these questions, we ran ED2 ensemble simulations with a factorial combination of submodels related to radiative transfer formulation (two-stream vs.
-multiple scatter), horizontal competition (finite canopy radius vs.
-complete shading), and trait plasticity (whether or not SLA and Vcmax vary with light level)
+(2) What is the cost of considering these processes, in terms of additional parameter uncertainty?
+(3) What are the relative contributions of parameter uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
+We ran ED2 ensemble simulations with a factorial combination of submodels related to radiative transfer formulation, horizontal competition, and trait plasticity.
 We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results in the context of ongoing climate change.

### 2019-10-25 analysis/paper/paper.Rmd
Minor revisions to methods
Per Ben's comments

@@
 Besides the cohort-based approximation of gap dynamics, a distinctive feature of ED-2.2 is that the equations for energy, water, and carbon fluxes are defined in terms of total energy, water, and carbon, which ensures excellent conservation of mass in long-running (multidecadal) simulations [@longo_2019_ed1].
-Various versions of the ED model have been validated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in, among others,
+Various versions of the ED model have been evaluated in boreal [@medvigy_2011_predicting], temperate [@medvigy_2009_mechanistic; @dietze_2014_quantitative; @raczka_2018_what], and tropical biomes [@moorcroft_2001_method; @longo_2019_ed2], and have been applied in, among others,
 paleoecological studies [@rollinson_2017_emergent],
@@
 The two-stream model (default) uses the two-stream approximation originally derived for atmospheric radiative transfer [@meador_1980_two] and later adapted to vegetation canopies [@dickinson_1983_land; @sellers_1985_canopy].
-Meanwhile, the multiple-scatter model was derived from first principles by @zhao_2005_multiple specifically for vegetation canopies, specifically to address known biases and limitations of the two-stream approach [e.g.
+Meanwhile, the multiple-scatter model was derived from first principles by @zhao_2005_multiple specifically for vegetation canopies to address known biases and limitations of the two-stream approach [e.g.
 @wang_2003_comparison].
@@
 The third submodel we evaluated was trait plasticity.
-In the default configuration, all cohorts of a given plant functional type will have the same parameters, regardless of environmental conditions.
+In the default configuration, all cohorts of a given plant functional type have the same parameters, regardless of environmental conditions.
 This ignores the globally-documented intraspecific trait variability as a function of light level [@niinemets_2010_review; @keenan_2016_global].

### 2019-10-25 analysis/paper/paper.Rmd
Minor wording revisions to results

@@
 Enabling the finite canopy radius increased the availability of light for understory cohorts throughout the simulation, with the most notable effects up to 1950.
-Under a closed canopy, switching from a two-stream to a multiple-scatter canopy radiative transfer model also enhanced the understory light environment under a closed canopy, but to a lesser extent, and the brightest understory occurred under the combination of the two-stream radiative transfer and finite canopy radius.
+Under a closed canopy, switching from a two-stream to a multiple-scatter canopy radiative transfer model also enhanced the understory light environment, but to a lesser extent, and the brightest understory occurred under the combination of the two-stream radiative transfer and finite canopy radius.
 However, none of the configurations were able to sustain any PFTs other than early hardwoods by 1980, and only the unique combination of finite canopy radius and two-stream radiative transfer were able to sustain more than a single dominant cohort (2 early hardwood cohorts with static traits, and 3 with plastic traits).
@@
 In many cases, median parameter estimates were different entirely because of the choice of prior distributions (though, in almost all cases, the distributions provided reasonably high likelihood to ED2 default parameter values).
-However, in several important cases, parameter well-constrained by data produced values substantially different from the ED2 defaults.
+However, in several important cases, parameters well-constrained by data produced values substantially different from the ED2 defaults.
 For example, ED2 default values for SLA were too low for all PFTs except Early Hardwoods, and default values for Vcmax values were too low for all PFTs.
@@
 The above differences in median parameter led to ED2 runs that differed from the default runs in several important ways.
-First of all, throughout all simulations and across all model structures, autotrophic respiration was several times higher with median parameter estimates compared to default values.
+First, throughout all simulations and across all model structures, autotrophic respiration was several times higher with median parameter estimates compared to default values.
 This is driven primarily by higher median values for growth and leaf respiration across all PFTs (notably, the ED2 default is a growth respiration of _zero_ for all temperate hardwood PFTs).
@@
 Third, even where net primary productivity is similar between default and median runs, the underlying forest composition is very different.
-Namely, under the default parameterization, ED2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
+Under the default parameterization, ED2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
 However, with the median parameterization, ED2 starts out with forests dominated by _mid_-hardwood species that are quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remains the dominant canopy species throughout the rest of the simulation.
-Exploratory analyses (results not shown / supplement) show that the main parameters responsible for this compositional trajectory are SLA (especially the higher SLA of pine and mid-hardwood), Vcmax, and growth respiration.
+Exploratory analyses (results not shown) show that the main parameters responsible for this compositional trajectory are SLA (especially the higher SLA of pine and mid-hardwood), Vcmax, and growth respiration.
 While the uncertainty in growth respiration is large, we emphasize that SLA and Vcmax are well constrained by data for all PFTs considered in this analysis.
@@
 
-Model projections of net primary productivity (NPP) and leaf area index (LAI) varied with both parameter value and model structure (Figure X).
+Net primary productivity (NPP) and leaf area index (LAI) varied with both parameter value and model structure (Figure X).
 All model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
@@
   htmlfig(),
-  "Comparison of parametric uncertainty within ensembles (colored bars, colored by model type)",
+  "Comparison of parameter uncertainty within ensembles (colored bars, colored by model type)",
   "and \"structural\" uncertainty (variance in ensemble means; black bar) by output variable.",

### 2019-10-25 analysis/paper/paper.Rmd
Remove ternary plot
Too confusing, not well implemented.

@@
 
-```{r pft-ternplot, fig.cap = pft_tern_cap}
-pft_tern_cap <- paste(
-  htmlfig(),
-  "Ternary plot of predicted PFT composition in 1920, 1950, and 1980.",
-  "Ternary coordinates are as follows:",
-  "L is Early Hardwood; T is the sum of Mid and Late Hardwood; R is Pine.",
-  "Triangle fill color indicates the number of points within that bin."
-)
-fig("pft-tern.png")
-```
-
 Trajectories of species composition also varied widely by model configuration (Figure X, Y).

### 2019-10-25 analysis/paper/paper.Rmd
Remove note about parameter interactions
Add intro Pandit paper text as discussion comment

@@
 
-{TODO: Basic analysis of contributions from parameter interactions, a la Raczka et al. 2018? I expect these will be important.}
-
 # Discussion
@@
 )
+
+<!-- From Introduction: -->
+<!-- Another recent study assessed parameter uncertainty in ED2 as part of its overall goal of parameterizing the model for sagebrush [@pandit_2018_optimizing]. -->
+<!-- They found that the most important parameters were specific leaf area, the maximum rate of carbon fixation (V~cmax~), slope of stomatal conductance, the fine-root to leaf carbon ratio, and the fine root turnover rate.
+-->
 

### 2019-10-25 analysis/paper/paper.Rmd
Update modeling workflow methods text

@@
 
-For each factorial combination of ED-2.2 configurations (described above), we ran 100 ensemble members from 1901 to 2000.
-Each ensemble member was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
+For each factorial combination of model configurations (described above), we ran ED-2.2 once with its default parameter values, once at the posterior median estimates from the meta-analysis, and 500 times with parameters sampled randomly from the meta-analysis posterior distributions.
+For the latter, the same 500 parameter samples were used for all model structures to allow us to independently evaluate the impacts of model structure at specific parameter values.
+Each ensemble member ran from January 1, 1902 through December 31, 1999 and was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
 Driving meteorological data was 6-hourly CRU-NCEP <!-- TODO ref.
 More here?
---> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric].
+--> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
 Soil texture was set to 92% sand, 7% silt, and 1% clay, per site-level observations in [@gough_2010_wood].
 The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
-
-<!-- TODO This would probably fit better somewhere else.  -->
-Driver uncertainty is outside the scope of this study, and initial condition uncertainty is minimized by our experimental design, which is conditioned on a specific initial condition (near bare ground).
+By using the same initial conditions and drivers, we do not consider initial condition and driver uncertainty, which are outside the scope of this study.
 

### 2019-10-25 analysis/paper/paper.Rmd
Fix LaTeX table formatting

@@
 library(here)
+library(magrittr)
+library(kableExtra)
 
@@
 | Name | Description | Color |
-|------------------------------|--------------------------------------------------------------------------------------|---------------------------|
-| **Crown model**              |                                                                                      |                           |
-| `closed` (default)           | Cohort crowns take up entire patch area. Competition for light based only on height. | Light                     |
-| `finite`                     | Cohort crown area is proportional to DBH according to PFT-specific allometry.        | Dark                      |
-| **Radiative transfer model** |                                                                                      |                           |
-| `two-stream` (default)       | Two-stream approximation [@sellers_1985_canopy; @clm45_note]                         | Primary (red, blue)       |
-| `multi-scatter`              | Multiple-scatter approximation, following [@zhao_2005_multiple]                      | Secondary (green, orange) |
-| **Trait plasticity**         |                                                                                      |                           |
-| `static` (default)           | SLA and V~c,max~ are constant                                                        | Cool (blue, green)        |
-| `plastic`                    | SLA increases, and V~c,max~ decreases, with light level                              | Warm (red, orange)        |
+|------|-------------|-------|
+| **Crown model** |||
+| **C**losed (default) | Cohort crowns take up entire patch area. Competition for light based only on height. | Light |
+| **F**inite | Cohort crown area is proportional to DBH according to PFT-specific allometry. | Dark |
+| **Radiative transfer model** |||
+| **T**wo-stream (default) | Two-stream approximation [@sellers_1985_canopy; @clm45_note] | Primary (red, blue) |
+| **M**ulti-scatter | Multiple-scatter approximation, following [@zhao_2005_multiple] | Secondary (green, orange) |
+| **Trait plasticity** |||
+| **S**tatic (default) | SLA and V~c,max~ are constant | Cool (blue, green) |
+| **P**lastic | SLA increases, and V~c,max~ decreases, with light level | Warm (red, orange) |
 
@@
 A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
-Instead, we selected a subset of parameters guided by (but expanding upon) previous ED2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
+Instead, we selected a subset of parameters guided by (but expanding upon) previous ED-2.2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
 
@@
 tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-kable(tbl, escape = FALSE, caption = paste0(
+kable(tbl, escape = is_latex_output(), caption = paste0(
   "Table 2: Parameter descriptions. ",
@@
   "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
-))
+), booktabs = TRUE) %>%
+  kable_styling(latex_options = "scale_down")
 ```

### 2019-10-25 analysis/paper/paper.Rmd
Draft analysis of results text; rearrange results

@@
 
+First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values (Figure 1).
+To understand the physical and ecological mechanisms underlying these differences, we also compared the cohort-level composition, height, and canopy radiation profiles in these simulations in four stages of the simulation (Figure 2).
+
+Second, we compared simulations with ED-2.2 default parameters against simulations using median parameter estimates from a trait meta-analysis (Figure 3).
+Here, we looked predictions of annual mean C fluxes (gross and net primary productivity, and autotrophic and heterotrophic respiration; Figure 4) and of summer LAI by PFT (Figure 5).
+
+Third, we considered the full range of parameter uncertainty by looking at predictions across all parameter ensemble members.
+We begin with predictions annual mean NPP and summer mean total LAI across all parameters and structures and compare these results to contemporary (averaged over years TODO) observations of NPP and LAI at UMBS.
+We look at these variables independently as time series (Figure 6) and as joint predictions in three years (Figure 7).
+We then consider impacts on composition by looking at predicted PFT fractions of total aboveground biomass (AGB) at four different years in the simulation (Figure 8) and at time series of LAI by PFT for five parameter combinations spanning the range of predicted ecological outcomes (Figure 9).
+
+Finally, we assessed which parameters contributed the most to parameter uncertainty for each model structure.
+We did this by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what.
+Briefly, this analysis contains uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance").
+For more details, see @lebauer_2013_facilitating.
+
+First, we assessed how ED-2.2 predictions of plot-level ecosystem properties and composition varied across model structures and parameter values.
+For plot-level ecosystem properties, we examined mean annual gross (GPP) and net primary productivity (NPP), and autotrophic (RA) and heterotrophic respiration (RH), as well as total leaf area index (LAI) averaged June-August.
+We compared predicted NPP and LAI values against measurements of NPP and LAI at UMBS averaged over years XXXX-YYYY. <!-- TODO -->
+Although these observations occurred later than the years of the simulation, we found that the distributions of ED-2.2 predictions of NPP and LAI were relatively stable after 1975 and therefore constituted a reasonable approximation of what ED-2.2 would have predicted out to the present day.
+For ease of interpretation, we only show the top eight parameters in terms of partial variance for every model configuration.
+
+For composition, we looked at annual means of summer aboveground biomass and leaf area index
+
 - Overall summary:
@@
 
+```{r pairs-time, fig.cap = pairs_time_cap}
+pairs_time_cap <- paste(
+  htmlfig(),
+  "ED2 predictions of annual mean NPP and growing-season LAI in 1920, 1950, and 1999",
+  "by model type.",
+  "Each point indicates a single ensemble run with a different set of parameters.",
+  "Solid line is the best-fit generalized additive model (GAM).",
+  "Dashed lines indicate the error bounds of present-day observed NPP and LAI.",
+  "Text labels match the rows in Figure X below."
+)
+fig("npp-lai-pairs.png")
+```
+
+The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs.
+Specifically, compared to runs with the default closed canopy, runs with a finite canopy were more likely to have higher LAI values for the same level of NPP, especially later into the simulation.
+Although many simulations matched _either_ the observed net primary productivity or the observed leaf area index, only a small minority of individual runs matched _both_.
+
 ```{r pft-jitterplot, fig.cap = pft_jitterplot_cap}
@@
 However, _which_ PFT was dominant varied depending on the parameter combination, and any of the four PFTs became the dominant one some subset of simulations.
-
-```{r pairs-time, fig.cap = pairs_time_cap}
-pairs_time_cap <- paste(
-  htmlfig(),
-  "ED2 predictions of annual mean NPP and growing-season LAI in 1920, 1950, and 1999",
-  "by model type.",
-  "Each point indicates a single ensemble run with a different set of parameters.",
-  "Solid line is the best-fit generalized additive model (GAM).",
-  "Dashed lines indicate the error bounds of present-day observed NPP and LAI.",
-  "Text labels match the rows in Figure X below."
-)
-fig("npp-lai-pairs.png")
-```
-
-<!-- TODO Use a logarithmic, rather than a linear fit? -->
-
-The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs.
-Specifically, compared to runs with the default closed canopy, runs with a finite canopy were more likely to have higher LAI values for the same level of NPP, especially later into the simulation.
-Although many simulations matched _either_ the observed net primary productivity or the observed leaf area index, only a small minority of individual runs matched _both_.
 

### 2019-10-25 analysis/paper/paper.Rmd
Remove old analysis of results text/outline

@@
 
-First, we assessed how ED-2.2 predictions of plot-level ecosystem properties and composition varied across model structures and parameter values.
-For plot-level ecosystem properties, we examined mean annual gross (GPP) and net primary productivity (NPP), and autotrophic (RA) and heterotrophic respiration (RH), as well as total leaf area index (LAI) averaged June-August.
-We compared predicted NPP and LAI values against measurements of NPP and LAI at UMBS averaged over years XXXX-YYYY. <!-- TODO -->
-Although these observations occurred later than the years of the simulation, we found that the distributions of ED-2.2 predictions of NPP and LAI were relatively stable after 1975 and therefore constituted a reasonable approximation of what ED-2.2 would have predicted out to the present day.
-For ease of interpretation, we only show the top eight parameters in terms of partial variance for every model configuration.
-
-For composition, we looked at annual means of summer aboveground biomass and leaf area index
-
-- Overall summary:
-    - Overall parameter uncertainty -- Time-series plots of total annual GPP and NPP, and maximum LAI and AGB, averaged across ensemble members
-    - Impact of parameter uncertainty on ecological trajectories -- plot of max annual LAI by PFT
-    - Both compared against observations of recently observed LAI and NPP at UMBS (e.g. FASET, FoRTE)
-- Uncertainty analysis:
-    - Parameter vs. structural uncertainty -- Variance in mean annual GPP from 19XX to 19YY across ensembles within model structure, and compare against 
-    - Breakdown of parameter uncertainty -- PEcAn sensitivity and uncertainty analysis [@lebauer_2013_facilitating; @dietze_2014_quantitative]
-- Details of structural uncertainty:
-    - Vertical profiles of radiation, traits?
-
 All analyses for this work were performed using R 3.6 (see "Colophon" for full details).

### 2019-10-25 analysis/paper/paper.Rmd
Add Figure numbers as comments
...to help keep track in text

@@
 
-```{r , fig.cap = structure_default_cap}
+```{r structure-compare-default, fig.cap = structure_default_cap}
+### Figure 1
 structure_default_cap <- paste(
@@
 
-```{r , fig.cap = structure_default_light_cap}
+```{r structure-compare-light, fig.cap = structure_default_light_cap}
+### Figure 2
 structure_default_light_cap <- paste(
@@
 ```{r param_dist, fig.cap = param_dist_cap}
+### Figure 3
 param_dist_cap <- paste(
@@
 ```{r default-median-flux, fig.cap = default_median_flux_cap}
+### Figure 4
 default_median_flux_cap <- paste(
@@
 ```{r default-median-lai , fig.cap = default_median_lai_cap}
+### Figure 5
 default_median_lai_cap <- paste(
@@
 ```{r summary-ts-plot, fig.cap = summary_ts_cap}
+### Figure 6
 summary_ts_cap <- paste(
@@
 ```{r pairs-time, fig.cap = pairs_time_cap}
+### Figure 7
 pairs_time_cap <- paste(
@@
 ```{r pft-jitterplot, fig.cap = pft_jitterplot_cap}
+### Figure 8
 pft_jitterplot_cap <- paste(
@@
 ```{r lai-pft-plot, fig.cap = lai_pft_cap}
+### Figure 9
 lai_pft_cap <- paste(
@@
 ```{r partial-variance-npp, fig.cap = partial_variance_npp_cap}
+### Figure 10
 partial_variance_npp_cap <- paste(

### 2019-10-25 analysis/paper/paper.Rmd
Remove some old results/discussion text

@@
 
-<!-- - Details of parameter uncertainty -->
-<!--   - Figure: Parameter sensitivity and uncertainty analysis -->
-<!--   - Some parameters are always important. Other parameters matter more for different structures. -->
-
-```{r within-vs-across, fig.cap = within_vs_across_cap, fig.dim = c(9, 4), eval = FALSE}
-within_vs_across_cap <- paste(
-  htmlfig(),
-  "Comparison of parameter uncertainty within ensembles (colored bars, colored by model type)",
-  "and \"structural\" uncertainty (variance in ensemble means; black bar) by output variable.",
-  "Output variables are expressed as growing season averages for all years after 1910."
-)
-readd(within_across_plot_knit)
-```
-
 <!--
@@
 ```
-
-<!-- Model sensitivity to specific parameters ["elasticity" _sensu_ @lebauer_2013_facilitating] varied depending on the model configuration and the target output variable (Figure 5). -->
-<!-- The parameter with the single highest elasticity to NPP and LAI across all model configurations was specific leaf area (SLA), though the PFT whose SLA was most important varied with model configuration.
--->
-<!-- Other frequently (but not always) important parameters were the leaf C:N ratio (`c2n_leaf`), ratio of fine root to leaf biomass (`fineroot2leaf`), leaf and root respiration rates, and leaf optical properties (`leaf_reflect_vis/nir`).
--->
-<!-- TODO Dig into model configurations. -->
 

### 2019-10-25 analysis/paper/paper.Rmd
Replace several "radiative transfer"s with RTM

@@
 In this study, we ran a factorial combination of the following ED-2.2 configurations: (1) two-stream vs.
-multiple-scatter canopy radiative transfer models; (2) infinite vs.
+multiple-scatter canopy radiative transfer models (RTMs); (2) infinite vs.
 finite crown area (a.k.a.
@@
 
-Both canopy radiative transfer models in ED-2.2 resolve the full vertical radiation profile within a patch as a function of canopy structure (leaf and wood area indices, crown area, leaf angle distribution) and incident solar radiation, following the definitions in CLM 4.5 [@clm45_note].
+Both canopy RTMs in ED-2.2 resolve the full vertical radiation profile within a patch as a function of canopy structure (leaf and wood area indices, crown area, leaf angle distribution) and incident solar radiation, following the definitions in CLM 4.5 [@clm45_note].
 Both models also use identical definitions for the optical properties of a single canopy layer as a function of leaf and wood optical properties, leaf angle distribution, canopy clumping, solar zenith angle, and leaf, wood, and crown area indices.
@@
 | **F**inite | Cohort crown area is proportional to DBH according to PFT-specific allometry. | Dark |
-| **Radiative transfer model** |||
+| **Radiative transfer model (RTM)** |||
 | **T**wo-stream (default) | Two-stream approximation [@sellers_1985_canopy; @clm45_note] | Primary (red, blue) |
@@
 Parameter sets **A** and **B** both resulted in similar patterns of immediate and complete dominance by mid- and late-hardwood, respectively, followed by gradual decline to near-zero LAI.
-In both of these simulations, peak productivity was higher and was sustained for longer when the canopy was closed rather than finite, and this difference was larger under a two-stream rather than a multiple-scatter radiative transfer model.
-Trait plasticity had relatively little effect on most of these two simulations, though in parameter set **B**, under the combination of closed canopy, multiple-scatter radiative transfer, and plastic traits, pine overtook the dominant late hardwood PFT starting around 1975.
+In both of these simulations, peak productivity was higher and was sustained for longer when the canopy was closed rather than finite, and this difference was larger under a two-stream rather than a multiple-scatter RTM.
+Trait plasticity had relatively little effect on most of these two simulations, though in parameter set **B**, under the combination of closed canopy, multiple-scatter RTM, and plastic traits, pine overtook the dominant late hardwood PFT starting around 1975.
 Both of these parameter sets also had the lowest LAI of the five selected here, with a peak LAI of 1 in **A** and 2 in **B**.

### 2019-10-25 analysis/paper/paper.Rmd
Clean up first results section

@@
 
-Different model structures agree for the first half of the simulation, at which point they started to diverge.
-Post 1950, the combination of two-stream radiative transfer, finite canopy radius, and plastic traits had the highest sustained productivity, followed by the same configuration except with static traits, with the remaining configurations having very similar results.
-Interestingly, the effects of changing individual model configurations are not additive --- for example, enabling trait plasticity had a different effect depending on whether the two-stream or multiple-scatter canopy RTM was used.
+Different model structures agreed for the first half of the simulation, at which point they started to diverge (Figure 1).
+Post 1950, the combination of two-stream RTM, finite canopy radius, and plastic traits had the highest productivity, followed by the same configuration with static traits, with the remaining configurations having very similar results.
+Interestingly, the effects of changing individual model configurations were not additive --- for example, enabling trait plasticity had a much stronger positive effect when paired with the two-stream canopy RTM than with the multiple-scatter canopy RTM.
 
@@
 
-Differences in productivity across structures are perhaps best understood by looking at differences in resulting radiation profiles at different snapshots in time over the course of the simulation.
-Enabling the finite canopy radius increased the availability of light for understory cohorts throughout the simulation, with the most notable effects up to 1950.
-Under a closed canopy, switching from a two-stream to a multiple-scatter canopy radiative transfer model also enhanced the understory light environment, but to a lesser extent, and the brightest understory occurred under the combination of the two-stream radiative transfer and finite canopy radius.
-However, none of the configurations were able to sustain any PFTs other than early hardwoods by 1980, and only the unique combination of finite canopy radius and two-stream radiative transfer were able to sustain more than a single dominant cohort (2 early hardwood cohorts with static traits, and 3 with plastic traits).
+Differences in productivity across model structures were related to differences in composition and vertical canopy radiation profiles over the course of the simulation (Figure 2).
+Enabling the finite canopy radius increased the availability of light for understory cohorts throughout the simulation, particularly in the first half of the simulation.
+Again, however, the effects of model structures were not strictly additive:
+Switching from a two-stream to a multiple-scatter canopy RTM enhanced the understory light environment (albeit slightly) under a closed canopy, but the brightest understory occurred under the combination of the two-stream RTM and finite canopy.
+Despite these differences, by 1980 none of the configurations was able to sustain any PFTs other than early hardwoods, and only configurations with finite canopy and two-stream RTM were able to sustain more than a single dominant cohort (two early hardwood cohorts with static traits, and three with plastic traits).
 

### 2019-10-25 analysis/paper/paper.Rmd
Tighten first par. of 2nd results section

@@
 
-Parameter estimates from the trait meta-analysis were often substantially different from the ED2 defaults.
-In many cases, median parameter estimates were different entirely because of the choice of prior distributions (though, in almost all cases, the distributions provided reasonably high likelihood to ED2 default parameter values).
+Parameter estimates from the trait meta-analysis were often substantially different from the ED-2.2 defaults (Figure 3).
+In many cases, median parameter estimates were different entirely because of the choice of prior distributions.
 However, in several important cases, parameters well-constrained by data produced values substantially different from the ED2 defaults.
-For example, ED2 default values for SLA were too low for all PFTs except Early Hardwoods, and default values for Vcmax values were too low for all PFTs.
+For example, ED-2.2 default values for V~cmax~ were too low for all PFTs, and default values for SLA were too low for all PFTs except Early Hardwoods.
 

### 2019-10-25 analysis/paper/paper.Rmd
Revise median parameter results text

@@
 
+These differences in parameter values resulted in ED-2.2 runs that differed from the default runs in several important ways (Figures 4, 5).
+First, in every model configuration (except finite canopy, two-stream RTM, and static traits), gross primary productivity (GPP) reached a higher peak, and did so earlier under median parameters (Figure 4).
+This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure 3).
+However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure 4), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure 3).
+The resulting net primary productivity (NPP) had similar peak values under default and median parameters, though the temporal patterns were quite different in most model configurations (Figure 4).
+
+Second, the variation across model structures was much higher, with every change in model structure exerting a noticeable effect on flux variability (Figure 4).
+The most dramatic example is that, with median parameter values, simulations with the finite canopy radius produced much lower net productivity and heterotrophic respiration than simulations with a closed canopy in the second half of the simulation.
+Yet again, interactions between configurations were significant.
+For example, under a closed canopy configuration, enabling trait plasticity caused interannual variability in NPP to decrease with a two-stream RTM and increase with a multiple-scatter RTM.
+
 ```{r default-median-lai , fig.cap = default_median_lai_cap}
@@
 
-The above differences in median parameter led to ED2 runs that differed from the default runs in several important ways.
-First, throughout all simulations and across all model structures, autotrophic respiration was several times higher with median parameter estimates compared to default values.
-This is driven primarily by higher median values for growth and leaf respiration across all PFTs (notably, the ED2 default is a growth respiration of _zero_ for all temperate hardwood PFTs).
-In some cases, this is compensated by an increase in gross primary productivity (driven by higher Vcmax values across all PFTs), leading to similar net primary productivity values to the default parameter runs.
-Second, the variation across model structures is much higher, with every change in model structure exerting a noticeable effect on flux dynamics.
-The most dramatic example is that, with median parameter values, all of the simulations with the finite canopy radius produced much lower productivity and heterotrophic respiration than simulations with a closed canopy.
-Third, even where net primary productivity is similar between default and median runs, the underlying forest composition is very different.
-Under the default parameterization, ED2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
-However, with the median parameterization, ED2 starts out with forests dominated by _mid_-hardwood species that are quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remains the dominant canopy species throughout the rest of the simulation.
-Exploratory analyses (results not shown) show that the main parameters responsible for this compositional trajectory are SLA (especially the higher SLA of pine and mid-hardwood), Vcmax, and growth respiration.
-While the uncertainty in growth respiration is large, we emphasize that SLA and Vcmax are well constrained by data for all PFTs considered in this analysis.
+Third, even where NPP was similar between default and median runs, the underlying forest composition was very different.
+With the default parameterization, ED-2.2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
+However, with the median parameterization, ED-2.2 starts out with forests dominated by _mid_-hardwood species that are quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remains the dominant canopy species throughout the rest of the simulation.
+Exploratory analyses (results not shown) show that the main parameters responsible for this compositional trajectory are SLA (especially the higher SLA of pine and mid-hardwood), V~cmax~, and growth respiration.
+While the uncertainty in growth respiration is large, we emphasize that SLA and V~cmax~ are well constrained by data for all PFTs considered in this analysis.
 

### 2019-10-25 analysis/paper/paper.Rmd
More cleaning of results text

@@
 
-Third, even where NPP was similar between default and median runs, the underlying forest composition was very different.
+Third, even where NPP was similar between default and median runs, the underlying forest composition was very different (Figure 5).
 With the default parameterization, ED-2.2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
@@
 
-<!-- - Structure x parameter uncertainty -->
-<!--   - Parameter uncertainty is much larger than structural uncertainty -->
-<!--   - Figure: Pairs of NPP, LAI, and PFT fraction, for each structure (first draw all, then pick relevant subsets) -->
-<!--   - Q: How many runs end up with the right composition (EH-dominated) at peak (~1925), ~1950, and 1999? -->
-
-<!-- TODO Should one of these be Early Hardwood (or other PFT?) LAI/biomass fraction -->
-<!-- TODO Drop some of these variables? -->
-
-Net primary productivity (NPP) and leaf area index (LAI) varied with both parameter value and model structure (Figure X).
-All model configurations exhibited a similar trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+Net primary productivity (NPP) and total leaf area index (LAI) varied with both parameter value and model structure (Figure 6).
+All model configurations exhibited a similar mean trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
 The ensemble spread within any single configuration---i.e.
-the parameter uncertainty---was similar and large across all configurations; both NPP and LAI varied 5-10 fold at any given timestep depending on parameter values.
+the parameter uncertainty---was similar and large across all configurations; both NPP and LAI varied from close to zero to more than twice the mean value at any given timestep.
 Within all configurations, most (>90%) ensemble members under-predicted observed present-day NPP and LAI by year 2000, though individual ensemble members from some configurations over-predicted one or both variables.
@@
 
-The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs.
-Specifically, compared to runs with the default closed canopy, runs with a finite canopy were more likely to have higher LAI values for the same level of NPP, especially later into the simulation.
-Although many simulations matched _either_ the observed net primary productivity or the observed leaf area index, only a small minority of individual runs matched _both_.
+The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs (Figure 7).
+Compared to runs with a closed canopy, runs with a finite canopy usually had higher LAI for the same NPP, especially later into the simulation.
+Although many simulations matched _either_ the observed NPP or the observed LAI, only a small minority of individual runs matched _both_.
 
@@
   "of each plant functional type (PFT)",
-  "predicted by different parameter-configuration combionations",
-  "of ED2."
+  "predicted by different parameter-configuration combinations",
+  "of ED2.",
+  "Each point represents the AGB fraction of a single PFT",
+  "in a single ensemble member at the given point in time."
 )
@@
 
-Trajectories of species composition also varied widely by model configuration (Figure X, Y).
-Under all configurations, most ensemble members were dominanted by a single PFT, and the sustained presence of more than 2 PFTs was limited to only a handful specific ensemble members across all configurations.
-However, _which_ PFT was dominant varied depending on the parameter combination, and any of the four PFTs became the dominant one some subset of simulations.
+The large parameter uncertainty in NPP and LAI also extended to species composition (Figure 8).
+Under all configurations, at any moment in time, most ensemble members were dominated by a single PFT, and the sustained coexistence of more than two PFTs was limited to only a handful specific ensemble members across all configurations.
+However, _which_ PFT was dominant depended strongly on parameter values, and any of the four PFTs became the dominant one in some subset of simulations.
 
@@
   "Each row of plots is a simulation with the same set of parameters across model configurations.",
-  "Letters correspond to the labels in Figure 3.",
+  "Letters correspond to the labels in Figure 7.",
   "Note that these are predictions of total leaf area index summed across all cohorts of the same PFT,",
@@
 
-Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure X).
+Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure 9).
 Parameter sets **A** and **B** both resulted in similar patterns of immediate and complete dominance by mid- and late-hardwood, respectively, followed by gradual decline to near-zero LAI.
-In both of these simulations, peak productivity was higher and was sustained for longer when the canopy was closed rather than finite, and this difference was larger under a two-stream rather than a multiple-scatter RTM.
+In both of these simulations, peak productivity was higher and more sustained when the canopy was closed rather than finite, and this difference was larger under a two-stream rather than a multiple-scatter RTM.
 Trait plasticity had relatively little effect on most of these two simulations, though in parameter set **B**, under the combination of closed canopy, multiple-scatter RTM, and plastic traits, pine overtook the dominant late hardwood PFT starting around 1975.
@@
 Parameter set **C** showed much more variety in compositional dynamics with model structure:
-In all cases, mid-hardwoods began as the dominant species and were eventually overtaken by pine under a closed canopy and plastic traits, or early hardwood in all other cases (though with a closed canopy and static traits, the final LAI was low).
+In all cases, mid-hardwoods began as the dominant species and were eventually overtaken by either pine (closed canopy and plastic traits) or early hardwood (all other cases; though with a closed canopy and static traits, the final LAI was low).
 Parameter sets **D** and **E** had the highest productivity of these five sets, and were characterized by strong dominance of Early Hardwood and Pine, respectively.
-
-## Uncertainty analysis
-
-<!--
-Except for the configurations with persistently low productivity, parameter uncertainty (across ensembles within each model configuration) was much larger than structural uncertainty (variance in ensemble means across configurations; Figure 4).
-The relative importance of parameter uncertainty varied with both model configuration and the variable in question.
-Allowing SLA and V~cmax~ to vary with light level (warm colors, as opposed to static traits -- cool colors) slightly reduced variability in primary productivity, slightly increased variability in Shannon diversity, and had mixed effects on aboveground biomass and total leaf area.
-Enabling the finite canopy radius (dark colors; compared to closed canopy -- light colors) dramatically decreased variability in primary productivity and Shannon diversity and increased variability in leaf area, and had mixed effects on aboveground biomass.
-reduced variance in aboveground biomass, diversity, and primary productivity.
-Finally, using a multiple-scatter (secondary colors) rather than a two-stream (primary colors) radiative transfer scheme increased variance in leaf area index, aboveground biomass, and diversity, but did not have a consistent effect on variance in productivity productivity.
--->
 
@@
   "Each panel shows the top 8 parameter values in terms of their variance contribution.",
-  "Input parameter distributions are shown in Figure 1.",
+  "Input parameter distributions are shown in Figure 3.",
   "Note that x-axis scales vary across panels."
@@
 
-The parameters that contributed most to overall model predictive uncertainty ("partial variance") varied with model structure and time period considered.
-In general, uncertainties in pine growth respiration factor and early hardwood plant-soil water conductance contributed the most to parameter uncertainty in NPP during the "peak" period (1925-1950).
-{TODO: Fill in the rest of this text...}
-
-Parameter sensitivity for total LAI was similar to that of NPP (results not shown).
-
-<!-- This is old text... -->
-<!-- ...differed from parameters to which the model was most sensitive (Figure 5), largely driven by the fact that many of the parameters with the highest sensitivity (e.g.
-SLA, V~cmax~) were also well constrained by data (Figure 1).
--->
-<!-- Leaf optical properties of at least one PFT were among the top 10 most important parameters across all configurations, and in multiple cases, were the most important. -->
-<!-- Other often important parameters were plant-soil water conductance (`water_conductance`), parameters related to adult mortality (`mort1,2,3`), and growth respiration. -->
-<!-- Again, patterns in partial variance were largely idiosyncratic across model configurations. -->
-<!-- TODO Dig into this more -->
-<!-- TODO Check this analysis. Look at pairs plots of trait vs. variable, and my SA spline fits. -->
+The parameters that contributed most to predictive uncertainty in NPP during the "peak" period (1920-1950) ("partial variance") varied with model structure (Figure 10).
+The "water availability factor" (a.k.a.
+plant-soil water conductance) for early hardwoods was the most or second-most important parameter for all but one model configuration (the ED-2.2 default---closed canopy, two-stream RTM, static traits---where it was not even in the top 8), and was particularly important in simulations with a finite canopy.
+The growth respiration factor, particularly of pine and, to a lesser extent, early hardwood, was also among the most influential parameters across most configurations.
+Which parameters were most important, and how important they were, varied with model configuration.
+For instance, the C balance ratio of mortality (`mort2`) of early hardwoods was among the top four most important parameters when the canopy was closed but mattered less when the canopy was finite.
+Similarly, parameters related to leaf optical properties had relatively higher importance in simulations with the two-stream model than the multiple-scatter model.
 

### 2019-10-25 analysis/paper/paper.Rmd
Flesh out uncertainty drivers in discussion

@@
 
-**Parameter uncertainty was the dominant driver of uncertainty in our analysis**.
-Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables** -- NPP ranged from 0 to almost 2x the observed, and LAI was up to 3x observed (though more tightly clustered around zero) (Figure TODO).
-Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figure TODO).
-**Parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.
+Our second question was: What is the relative contribution of structural and parameter uncertainty to C cycling and ecological dynamics at UMBS?
+We found that **parameter uncertainty was by far the dominant driver of uncertainty in our simulations**.
+Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables**, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures 6, 7).
+Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figures 8. 9).
+Therefore, **parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others (REFS).
 
 The dominance of parameter uncertainty over structural uncertainty in our results suggests that **constraining parameter uncertainty should be a higher priority than model development**.
-Parameter uncertainty can be constrained in two different ways:
-Using observational data if available;
-and, for parameters "that cannot be directly observed", using "quantitative/algorithmic" model calibration techniques (e.g. Fer, other papers).
-For example, leaf optical properties (i.e. reflectance, transmittance) are both measurable and highly influential, and therefore should be a priority for new data collection.
-In contrast, "growth respiration factor" is an empirical coefficient and not measurable, but highly influential (c.f. Dietze, Raczka), and therefore should be constrained through calibration.
+Parameter uncertainty can be reduced in two different ways:
+Directly constraining parameters using observational data (for parameters that can be observed directly, such as plant functional traits) [@lebauer_2013_facilitating],
+or indirectly constraining parameters through algorithmic model calibration techniques [@fer_2018_linking].
+For example, leaf optical properties (i.e.
+reflectance, transmittance) are both measurable and highly influential [see also @viskari_2019_influence], and therefore should be a priority for new data collection.
+In contrast, "growth respiration factor" is an empirical coefficient and not measurable, but highly influential [see also @dietze_2014_quantitative; @raczka_2018_what], and therefore should be constrained through calibration.
 
-(
-So **which parameters are a research priority**?
-Leaf optical properties often do not get a lot of attention, but are important [see also, @viskari_2019_influence].
-Also, soil-plant water conductance and growth respiration -- both of these were important in @dietze_2014_quantitative.
-(Compare also to @raczka_2018_what, @pandit_2018_optimizing).
-)
+**Importantly, efforts to reduce model uncertainty are most effective when they target parameters that are poorly constrained and to which the model is most sensitive**.
+For parameters that are already well constrained by data, the marginal value of additional measurements may be limited.
+For example, ED-2.2 is highly sensitive to SLA [@dietze_2014_quantitative; @pandit_2018_optimizing], but thanks to the large number of SLA observations, we were able to estimate it very precisely (Figure 3) and, as a result, it was not a significant source of uncertainty in our analysis (Figure 10).
+At the same time, some parameters have relatively modest impacts on model outputs---for instance, nonlocal dispersal and background mortality rate in our study (Figure 3, 10)---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
+Ultimately, this points to the value of model-data feedbacks [@dietze_2013_improving] in advancing ecological research and setting future research agendas.
 
-<!-- From Introduction: -->
-<!-- Another recent study assessed parameter uncertainty in ED2 as part of its overall goal of parameterizing the model for sagebrush [@pandit_2018_optimizing]. -->
-<!-- They found that the most important parameters were specific leaf area, the maximum rate of carbon fixation (V~cmax~), slope of stomatal conductance, the fine-root to leaf carbon ratio, and the fine root turnover rate.
--->
-
-**Parameter uncertainty is not synonymous with sensitivity**.
-Many of the parameters to which model outputs are most sensitive are already well constrained by data and do not contribute as much to predictive uncertainty (Figure TODO).
-SLA and Vcmax fall into this category.
-Although parameter uncertainty is still large, this means that **targeted data collection works and that we are moving in the right direction**!
-This speaks to the value of looking at not just sensitivity, but partial variance [@lebauer_2013_facilitating; @dietze_2014_quantitative].
-
-Although parameter uncertainty dominated, **structural and parameter uncertainty interacted in important ways**.
-Individual model structure changes had relatively modest impacts.
-
-* Interactive effects of different structures -- not additive
-* Interactions between parameter and structural uncertainty -- 
-  - For example, the relative importance of different parameters to predictive uncertainty (Figure TODO) differed across model structures (Figure 3).
-
-Collectively, this means that **sensitivity and uncertainty analyses are not necessarily generalizable across models**.
+Although parameter uncertainty was larger in magnitude than structural uncertainty, model structure played an important role in shaping the nature of that uncertainty.
+The consequences of individual model structure changes were *usually* modest, but with several dramatic exceptions (Figure 4).
+Moreover, we found bi-directional interactions between structural and parameter uncertainty:
+Different parameter values (especially ones far removed from ED-2.2 defaults) influenced ED-2.2 sensitivity to structural changes (Figures 4, 5),
+and changes to model structure in turn changed the relative importance of specific parameters to overall uncertainty (Figure 10).
+Collectively, this means that **sensitivity and uncertainty analyses are not necessarily generalizable across models**, and highlights the importance of model intercomparison studies.
+It also reinforces our argument that changing model structure---particularly if such changes increase model complexity---should be pursued with extreme caution.
 
@@
 
-<!-- TODO -->
+{TODO}
 

### 2019-10-28 analysis/paper/paper.Rmd
Use bookdown::pdf_document2 and references

@@
 ---
-title: "FoRTE baseline"
+title: |
+  Structure and parameter uncertainty in centennial-scale projections
+  of community succession in an Upper Midwest temperate forest
 author:
@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
-    pdf_document:
+    bookdown::pdf_document2:
       latex_engine: xelatex
@@
   out <- paste0("**Figure ", m, "**:")
-  if (grepl("html", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
+  if (knitr::is_html_output()) {
     out
@@
 
-These sub-models are summarized in Table 1.
+These sub-models are summarized in Table \@ref(tab:submodels).
 
-Table 1: ED-2.2 submodel descriptions
+Table: (\#tab:submodels) ED-2.2 submodel descriptions
 
@@
 
+```{r param-table}
+tbl <- readr::read_csv(here(
+  "analysis", "data", "derived-data", "parameter-table.csv"
+))
+tbl <- subset(tbl, Ran == "yes", select = c(-unit_parsed, -Ran))
+tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
+kable(tbl, escape = is_latex_output(), caption = paste0(
+  "Parameter descriptions. ",
+  "For additional information, see the ",
+  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
+), booktabs = TRUE) %>%
+  kable_styling(latex_options = "scale_down")
+```
+
 The full list of parameters in ED-2.2 is large, with over 100 parameters per plant functional type.
 A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
-Instead, we selected a subset of parameters guided by (but expanding upon) previous ED-2.2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
+Instead, we selected a subset of parameters (Table \@ref(tab:param-table)) guided by (but expanding upon) previous ED-2.2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
 
@@
 
-The full list of parameters is shown in Table 2.
-
-```{r param-table}
-tbl <- readr::read_csv(here(
-  "analysis", "data", "derived-data", "parameter-table.csv"
-))
-tbl <- subset(tbl, Ran == "yes", select = c(-unit_parsed, -Ran))
-tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-kable(tbl, escape = is_latex_output(), caption = paste0(
-  "Table 2: Parameter descriptions. ",
-  "For additional information, see the ",
-  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
-), booktabs = TRUE) %>%
-  kable_styling(latex_options = "scale_down")
-```
-
 ## Plant functional types and parameterization
@@
 
-Table 3: Plant functional type-species mappings
+Table: (\#tab:pft-species) Plant functional type-species mappings
 
@@
 
-First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values (Figure 1).
-To understand the physical and ecological mechanisms underlying these differences, we also compared the cohort-level composition, height, and canopy radiation profiles in these simulations in four stages of the simulation (Figure 2).
+First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values (Figure \@ref(fig:structure-compare-default)).
+To understand the physical and ecological mechanisms underlying these differences, we also compared the cohort-level composition, height, and canopy radiation profiles in these simulations in four stages of the simulation (Figure \@ref(fig:structure-compare-light)).
 
-Second, we compared simulations with ED-2.2 default parameters against simulations using median parameter estimates from a trait meta-analysis (Figure 3).
-Here, we looked predictions of annual mean C fluxes (gross and net primary productivity, and autotrophic and heterotrophic respiration; Figure 4) and of summer LAI by PFT (Figure 5).
+Second, we compared simulations with ED-2.2 default parameters against simulations using median parameter estimates from a trait meta-analysis (Figure \@ref(fig:param-dist)).
+Here, we looked predictions of annual mean C fluxes (gross and net primary productivity, and autotrophic and heterotrophic respiration; Figure \@ref(fig:default-median-flux)) and of summer LAI by PFT (Figure \@ref(fig:default-median-lai)).
 
@@
 We begin with predictions annual mean NPP and summer mean total LAI across all parameters and structures and compare these results to contemporary (averaged over years TODO) observations of NPP and LAI at UMBS.
-We look at these variables independently as time series (Figure 6) and as joint predictions in three years (Figure 7).
-We then consider impacts on composition by looking at predicted PFT fractions of total aboveground biomass (AGB) at four different years in the simulation (Figure 8) and at time series of LAI by PFT for five parameter combinations spanning the range of predicted ecological outcomes (Figure 9).
+We look at these variables independently as time series (Figure \@ref(fig:summary-ts-plot)) and as joint predictions in three years (Figure \@ref(fig:pairs-time)).
+We then consider impacts on composition by looking at predicted PFT fractions of total aboveground biomass (AGB) at four different years in the simulation (Figure \@ref(fig:pft-jitterplot)) and at time series of LAI by PFT for five parameter combinations spanning the range of predicted ecological outcomes (Figure \@ref(fig:lai-pft-plot)).
 
 Finally, we assessed which parameters contributed the most to parameter uncertainty for each model structure.
-We did this by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what.
+We did this by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what (Figure \@ref(fig:partial-variance-npp)).
 Briefly, this analysis contains uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance").
@@
 
-```{r param_dist, fig.cap = param_dist_cap}
+```{r param-dist, fig.cap = param_dist_cap}
 ### Figure 3
@@
 
-```{r default-median-lai , fig.cap = default_median_lai_cap}
+```{r default-median-lai, fig.cap = default_median_lai_cap}
 ### Figure 5

### 2019-10-28 analysis/paper/paper.Rmd
Fix remaining figure references

@@
 
-Different model structures agreed for the first half of the simulation, at which point they started to diverge (Figure 1).
+Different model structures agreed for the first half of the simulation, at which point they started to diverge (Figure \@ref(fig:structure-compare-default)).
 Post 1950, the combination of two-stream RTM, finite canopy radius, and plastic traits had the highest productivity, followed by the same configuration with static traits, with the remaining configurations having very similar results.
@@
 
-Differences in productivity across model structures were related to differences in composition and vertical canopy radiation profiles over the course of the simulation (Figure 2).
+Differences in productivity across model structures were related to differences in composition and vertical canopy radiation profiles over the course of the simulation (Figure \@ref(fig:structure-compare-light)).
 Enabling the finite canopy radius increased the availability of light for understory cohorts throughout the simulation, particularly in the first half of the simulation.
@@
 
-Parameter estimates from the trait meta-analysis were often substantially different from the ED-2.2 defaults (Figure 3).
+Parameter estimates from the trait meta-analysis were often substantially different from the ED-2.2 defaults (Figure \@ref(fig:param-dist)).
 In many cases, median parameter estimates were different entirely because of the choice of prior distributions.
@@
 
-These differences in parameter values resulted in ED-2.2 runs that differed from the default runs in several important ways (Figures 4, 5).
-First, in every model configuration (except finite canopy, two-stream RTM, and static traits), gross primary productivity (GPP) reached a higher peak, and did so earlier under median parameters (Figure 4).
-This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure 3).
-However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure 4), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure 3).
-The resulting net primary productivity (NPP) had similar peak values under default and median parameters, though the temporal patterns were quite different in most model configurations (Figure 4).
+These differences in parameter values resulted in ED-2.2 runs that differed from the default runs in several important ways (Figures \@ref(fig:default-median-flux), \@ref(fig:default-median-lai)).
+First, in every model configuration (except finite canopy, two-stream RTM, and static traits), gross primary productivity (GPP) reached a higher peak, and did so earlier under median parameters (Figure \@ref(fig:default-median-flux)).
+This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure \@ref(fig:param-dist)).
+However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure \@ref(fig:default-median-flux)), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure \@ref(fig:param-dist)).
+The resulting net primary productivity (NPP) had similar peak values under default and median parameters, though the temporal patterns were quite different in most model configurations (Figure \@ref(fig:default-median-flux)).
 
-Second, the variation across model structures was much higher, with every change in model structure exerting a noticeable effect on flux variability (Figure 4).
+Second, the variation across model structures was much higher, with every change in model structure exerting a noticeable effect on flux variability (Figure \@ref(fig:default-median-flux)).
 The most dramatic example is that, with median parameter values, simulations with the finite canopy radius produced much lower net productivity and heterotrophic respiration than simulations with a closed canopy in the second half of the simulation.
@@
 
-Third, even where NPP was similar between default and median runs, the underlying forest composition was very different (Figure 5).
+Third, even where NPP was similar between default and median runs, the underlying forest composition was very different (Figure \@ref(fig:default-median-lai)).
 With the default parameterization, ED-2.2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
@@
 
-Net primary productivity (NPP) and total leaf area index (LAI) varied with both parameter value and model structure (Figure 6).
+Net primary productivity (NPP) and total leaf area index (LAI) varied with both parameter value and model structure (Figure \@ref(fig:summary-ts-plot)).
 All model configurations exhibited a similar mean trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
@@
 
-The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs (Figure 7).
+The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs (Figure \@ref(fig:pairs-time)).
 Compared to runs with a closed canopy, runs with a finite canopy usually had higher LAI for the same NPP, especially later into the simulation.
@@
 
-The large parameter uncertainty in NPP and LAI also extended to species composition (Figure 8).
+The large parameter uncertainty in NPP and LAI also extended to species composition (Figure \@ref(fig:pft-jitterplot)).
 Under all configurations, at any moment in time, most ensemble members were dominated by a single PFT, and the sustained coexistence of more than two PFTs was limited to only a handful specific ensemble members across all configurations.
@@
 
-Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure 9).
+Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure \@ref(fig:lai-pft-plot)).
 Parameter sets **A** and **B** both resulted in similar patterns of immediate and complete dominance by mid- and late-hardwood, respectively, followed by gradual decline to near-zero LAI.
@@
 
-The parameters that contributed most to predictive uncertainty in NPP during the "peak" period (1920-1950) ("partial variance") varied with model structure (Figure 10).
+The parameters that contributed most to predictive uncertainty in NPP during the "peak" period (1920-1950) ("partial variance") varied with model structure (Figure \@ref(fig:partial-variance-npp)).
 The "water availability factor" (a.k.a.
@@
 Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
-**Ensemble mean trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure TODO);
+**Ensemble mean trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure \@ref(fig:summary-ts-plot));
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
@@
 In particular, recent evidence points to the **important role of low-severity, high-frequency disturbances** in sustaining ecosystem productivity [@gough_2016_disturbance].
-ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the top 10 most important parameters for predictive uncertainty (Figure 6).
+ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the top 10 most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
 <!-- Maybe move this elsewhere? -->
@@
 
-- Figure 3:
+- Figure \@ref(fig:pairs-time):
   - For the same LAI, closed canopy had higher productivity than finite. Why?
@@
 We found that **parameter uncertainty was by far the dominant driver of uncertainty in our simulations**.
-Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables**, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures 6, 7).
-Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figures 8. 9).
+Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables**, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures \@ref(fig:summary-ts-plot), \@ref(fig:pairs-time)).
+Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figures \@ref(fig:pft-jitterplot). \@ref(fig:lai-pft-plot)).
 Therefore, **parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others (REFS).
@@
 For example, leaf optical properties (i.e.
-reflectance, transmittance) are both measurable and highly influential [see also @viskari_2019_influence], and therefore should be a priority for new data collection.
-In contrast, "growth respiration factor" is an empirical coefficient and not measurable, but highly influential [see also @dietze_2014_quantitative; @raczka_2018_what], and therefore should be constrained through calibration.
+reflectance, transmittance) are both measurable and highly influential [Figure \@ref(fig:partial-variance-npp); see also @viskari_2019_influence], and therefore should be a priority for new data collection.
+In contrast, "growth respiration factor" is an empirical coefficient and not measurable, but highly influential [Figure \@ref(fig:partial-variance-npp); see also @dietze_2014_quantitative; @raczka_2018_what], and therefore should be constrained through calibration.
 
@@
 For parameters that are already well constrained by data, the marginal value of additional measurements may be limited.
-For example, ED-2.2 is highly sensitive to SLA [@dietze_2014_quantitative; @pandit_2018_optimizing], but thanks to the large number of SLA observations, we were able to estimate it very precisely (Figure 3) and, as a result, it was not a significant source of uncertainty in our analysis (Figure 10).
-At the same time, some parameters have relatively modest impacts on model outputs---for instance, nonlocal dispersal and background mortality rate in our study (Figure 3, 10)---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
+For example, ED-2.2 is highly sensitive to SLA [@dietze_2014_quantitative; @pandit_2018_optimizing], but thanks to the large number of SLA observations, we were able to estimate it very precisely (Figure \@ref(fig:param-dist)) and, as a result, it was not a significant source of uncertainty in our analysis (Figure \@ref(fig:partial-variance-npp)).
+At the same time, some poorly constrained parameters have relatively modest impacts on model outputs---for instance, nonlocal dispersal and background mortality rate in our study (Figure \@ref(fig:param-dist), \@ref(fig:partial-variance-npp))---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
 Ultimately, this points to the value of model-data feedbacks [@dietze_2013_improving] in advancing ecological research and setting future research agendas.
@@
 Although parameter uncertainty was larger in magnitude than structural uncertainty, model structure played an important role in shaping the nature of that uncertainty.
-The consequences of individual model structure changes were *usually* modest, but with several dramatic exceptions (Figure 4).
+The consequences of individual model structure changes were *usually* modest, but with several dramatic exceptions (Figure \@ref(fig:default-median-flux)).
 Moreover, we found bi-directional interactions between structural and parameter uncertainty:
-Different parameter values (especially ones far removed from ED-2.2 defaults) influenced ED-2.2 sensitivity to structural changes (Figures 4, 5),
-and changes to model structure in turn changed the relative importance of specific parameters to overall uncertainty (Figure 10).
+Different parameter values (especially ones far removed from ED-2.2 defaults) influenced ED-2.2 sensitivity to structural changes (Figures \@ref(fig:default-median-flux), \@ref(fig:default-median-lai)),
+and changes to model structure in turn changed the relative importance of specific parameters to overall uncertainty (Figure \@ref(fig:partial-variance-npp)).
 Collectively, this means that **sensitivity and uncertainty analyses are not necessarily generalizable across models**, and highlights the importance of model intercomparison studies.

### 2019-10-28 analysis/paper/paper.Rmd
Add default runs to summary TS plot (disabled)

@@
 
-## ED2.2 vs. reality
+## ED-2.2 vs. reality
 
 Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
-**Ensemble mean trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure \@ref(fig:summary-ts-plot));
+**Ensemble average trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure \@ref(fig:summary-ts-plot));
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].

### 2019-10-28 analysis/drake/overall-uncertainty.R

@@ -57,6 +57,12 @@ plan <- bind_plans(plan, drake_plan(
     ggplot() +
     aes(x = year, color = color) +
     geom_line(aes(y = value, group = case), alpha = 0.1, size = 0.3) +
+    ## # ED2 default results.
+    ## geom_line(
+    ##   aes(y = value),
+    ##   data = rename(structure_default_data, variable = name),
+    ##   linetype = "solid", color = "black"
+    ## ) +
     geom_ribbon(
       aes(ymin = lo, ymax = hi),
       data = ts_summary,

### 2019-10-28 analysis/drake/structure-results.R

@@ -55,14 +55,14 @@ plan <- bind_plans(plan, drake_plan(
 
 ### Structure NPP and LAI figure
 plan <- bind_plans(plan, drake_plan(
-  structure_compare_default_gg = default_annual %>%
+  structure_default_data = default_annual %>%
     select(runtype, model_id = casename, year,
            npp = mmean_npp_py, lai) %>%
     left_join(models, "model_id") %>%
     # Convert kgC m-2 yr-1 to MgC ha-1 yr-1
     mutate(npp = npp * 10) %>%
-    pivot_longer(c(npp, lai)) %>%
-    ggplot() +
+    pivot_longer(c(npp, lai)),
+  structure_compare_default_gg = ggplot(structure_default_data) +
     aes(x = year, y = value, color = model) +
     geom_line() +
     guides(color = guide_legend(title = "Model",

### 2019-10-28 analysis/paper/paper.Rmd
Fix NPP/LAI pairs figure. Update figure captions

@@
   htmlfig(),
-  "ED2 predictions of annual mean NPP and growing-season LAI in 1920, 1950, and 1999",
+  "ED-2.2 predictions of annual mean NPP and growing-season LAI in 1925 and 1999",
   "by model type.",
@@
   htmlfig(),
-  "Contributions of individual parameters to predicted uncertainty in NPP,",
+  "Contributions of individual parameters to predicted uncertainty in",
+  "average net primary productivity (NPP) over the period 1920-1950",
   "by model type.",
-  "Panels on the left are for the \"peak\" NPP period (averaged over 1925-1950),",
-  "and on the right are for the end of the simulation (averaged over 1975-1999)",
-  "Each panel shows the top 8 parameter values in terms of their variance contribution.",
+  "Each panel shows the top 8 parameter values by their fractional contribution",
+  'to parameter uncertainty ("partial variance").',
   "Input parameter distributions are shown in Figure 3.",
@@
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
-This idea that ED2.2's main issue is that it is moving too quickly is supported by the fact that ED2.2 generally underestimates productivity after 100 years, but model **predictions at peak productivity are close to reality**.
-One reason ED2.2 is moving too quickly through succession is **missing mechanisms for sustaining productivity through early succession**.
-In particular, recent evidence points to the **important role of low-severity, high-frequency disturbances** in sustaining ecosystem productivity [@gough_2016_disturbance].
-ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the top 10 most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
+Although ED-2.2 underestimates productivity after 100 years, model **predictions at peak productivity are close to reality**.
+One possible reason for the excessive pace of succession in ED-2.2 is **missing mechanisms for sustaining productivity through early succession**.
+In particular, recent evidence points to the **importance of low-severity, high-frequency disturbances** in sustaining ecosystem productivity [@gough_2016_disturbance].
+ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
 <!-- Maybe move this elsewhere? -->
 <!-- TODO Look at mortality across ED runs. Does it relate at all to results? -->
-However, **our simulations did not include other sources of disturbance**, such as partial harvests, extreme weather, and damage from insects and pathogens, which are common to this region [@amiro_2010_ecosystem] and which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
+However, although our simulations used real meteorology (which included occasional extreme weather events), **we did not consider other sources of disturbance** common to this region [such as harvests, insects and pathogens, and fire; @amiro_2010_ecosystem], which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
 Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance frequencies and see how they affect ecological dynamics.
+
+Under all but a small minority of parameter combinations, **ED-2.2 was unable to create---let alone sustain---a forest with any appreciable functional diversity**.
+The large majority of runs tended towards
 

### 2019-10-28 analysis/drake/pairs-plots.R

@@ -14,8 +14,8 @@ npp_lai_pairs <- function(dat, year, labs) {
     geom_vline(xintercept = 3.97 + c(-1, 1) * 1.96 * 0.423,
                linetype = "dashed") +
     geom_point(color = "black", data = labs2) +
-    geom_label_repel(aes(label = label), data = labs2,
-                     min.segment.length = 0) +
+    geom_text_repel(aes(label = label), data = labs2,
+                    min.segment.length = 0) +
     scale_color_identity() +
     facet_wrap(vars(model), ncol = 2) +
     labs(x = "Total LAI", y = expression(NPP ~ (MgC ~ ha^-1 ~ year^-1))) +
@@ -49,7 +49,7 @@ plan <- bind_plans(plan, drake_plan(
     left_join(models, c("model_id")),
   npp_lai_pairs_yr = target(
     npp_lai_pairs(both_wide, .year, use_params),
-    transform = map(.year = c(1920, 1950, 1999))
+    transform = map(.year = c(1925, 1999))
   ),
   npp_lai_pairs_gg = target(
     plot_grid(npp_lai_pairs_yr, nrow = 1),

### 2019-10-28 analysis/paper/paper.Rmd
Add more discussion text. Other draft code.

@@
 We begin with predictions annual mean NPP and summer mean total LAI across all parameters and structures and compare these results to contemporary (averaged over years TODO) observations of NPP and LAI at UMBS.
-We look at these variables independently as time series (Figure \@ref(fig:summary-ts-plot)) and as joint predictions in three years (Figure \@ref(fig:pairs-time)).
+We look at these variables independently as time series (Figure \@ref(fig:summary-ts-plot)) and as joint predictions in two periods (Figure \@ref(fig:pairs-time)).
 We then consider impacts on composition by looking at predicted PFT fractions of total aboveground biomass (AGB) at four different years in the simulation (Figure \@ref(fig:pft-jitterplot)) and at time series of LAI by PFT for five parameter combinations spanning the range of predicted ecological outcomes (Figure \@ref(fig:lai-pft-plot)).
@@
   htmlfig(),
-  "ED-2.2 predictions of annual mean NPP and growing-season LAI in 1925 and 1999",
+  "ED-2.2 predictions of annual mean NPP and growing-season LAI",
+  "averaged over 1925-1950 (left) and 1975-1999 (right)",
   "by model type.",
@@
   "Dashed lines indicate the error bounds of present-day observed NPP and LAI.",
-  "Text labels match the rows in Figure X below."
+  "Text labels match the rows in Figure 9 below."
 )
@@
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
-Although ED-2.2 underestimates productivity after 100 years, model **predictions at peak productivity are close to reality**.
 One possible reason for the excessive pace of succession in ED-2.2 is **missing mechanisms for sustaining productivity through early succession**.
@@
 Under all but a small minority of parameter combinations, **ED-2.2 was unable to create---let alone sustain---a forest with any appreciable functional diversity**.
-The large majority of runs tended towards
+The large majority of runs produced dominance by a single PFT, though _which_ PFT dominated depended on parameter combination.
+This highlights the difficulty of representing plant functional diversity in vegetation models, which has been true in other models as well [@fisher_2015_taking].
+To the extent that ED-2.2 is "correct" in how it represents processes, this result also indicates that the plant functional trait space of realized niches is small---i.e.
+multiple different trees can co-exist only under very specific combinations of plant traits and environmental conditions (REFS).
 
-- Figure \@ref(fig:pairs-time):
-  - For the same LAI, closed canopy had higher productivity than finite. Why?
-    - Not zero-sum; finite canopy hurts dominant trees but doesn't improve subdominant species that much <!-- TODO Need support for this! -->
-  - Importance of using multiple observational constraints. Ecological equifinality is a real problem.
+Most ED-2.2 runs under-predicted both NPP and LAI, both in the peak period (1920-1950) and towards the end of the simulation (1975-1999) (Figures \@ref(fig:summary-ts-plot), \@ref(fig:pairs-time)).
+However, of our five selected representative runs, the two that most closely matched present day composition at UMBS (early hardwood-dominated; **C** and **D** in Figure \@ref(fig:lai-pft-plot)) also closely matched the observed NPP and LAI (Figure \@ref(fig:pairs-time)).
+This is encouraging because it suggests that it is possible to get both aggregate ecosystem C fluxes and PFT composition right.
+Moreover, given that many important ED-2.2 parameters were well-constrained in our analysis (\@ref(fig:param-dist)), this is possible without using physically unrealistic parameter values.
+We therefore suggest that for model calibration and validation, comparisons against multiple output variables---including variables related to composition---can yield maximum constraint and are not unreasonably strict.
 

### 2019-10-28 analysis/drake/pairs-plots.R

@@ -1,7 +1,10 @@
 ### Pairs plots
-npp_lai_pairs <- function(dat, year, labs) {
+npp_lai_pairs <- function(dat, start_year, end_year, labs) {
   dat_sub <- dat %>%
-    filter(!is.na(model), year == !!year)
+    filter(!is.na(model), year >= !!start_year, year <= !!end_year) %>%
+    group_by(param_id, model, color) %>%
+    summarize_at(vars(mmean_lai_py, mmean_npp_py), mean) %>%
+    ungroup()
   labs2 <- labs %>%
     inner_join(dat_sub, "param_id") %>%
     distinct(param_id, model, label, mmean_lai_py, mmean_npp_py)
@@ -19,10 +22,9 @@ npp_lai_pairs <- function(dat, year, labs) {
     scale_color_identity() +
     facet_wrap(vars(model), ncol = 2) +
     labs(x = "Total LAI", y = expression(NPP ~ (MgC ~ ha^-1 ~ year^-1))) +
-    ggtitle(year) +
+    ggtitle(sprintf("%d - %d", start_year, end_year)) +
     theme_cowplot() +
     theme(strip.text = element_text(size = 10))
-
 }
 
 plan <- bind_plans(plan, drake_plan(
@@ -48,8 +50,9 @@ plan <- bind_plans(plan, drake_plan(
     ) %>%
     left_join(models, c("model_id")),
   npp_lai_pairs_yr = target(
-    npp_lai_pairs(both_wide, .year, use_params),
-    transform = map(.year = c(1925, 1999))
+    npp_lai_pairs(both_wide, .ayear, .zyear, use_params),
+    transform = map(.ayear = c(1920, 1975),
+                    .zyear = c(1950, 1999))
   ),
   npp_lai_pairs_gg = target(
     plot_grid(npp_lai_pairs_yr, nrow = 1),

### 2019-10-28 analysis/drake/specific-parameters.R

@@ -7,6 +7,14 @@ plan <- bind_plans(plan, drake_plan(
     284, "D", "EH-dominated in 1999, high NPP",
     114, "E", "Pine-dominated in 1999, high NPP"
   ),
+  use_param_values = params %>%
+    inner_join(use_params, "param_id") %>%
+    inner_join(pfts(), c("name" = "bety_name")) %>%
+    select(-why, -name, -shortname, -num, -color) %>%
+    ## filter(label %in% c("C", "D")) %>%
+    pivot_longer(-c(param_id, label, pft),
+                 names_to = "trait",
+                 values_to = "value"),
   lai_pft_plot_gg = pft_data %>%
     mutate(param_id = as.numeric(substr(case, 0, 3))) %>%
     inner_join(use_params, "param_id") %>%

### 2019-10-28 analysis/drake/trait-distribution.R

@@ -31,6 +31,8 @@ plan <- plan <- bind_plans(plan, drake_plan(
                data = ed2_default_params, size = 2) +
     geom_point(aes(y = Median, color = "median", shape = "median"),
                data = trait_distribution, size = 2) +
+    # Parameter values
+    ## geom_text(aes(y = value, label = label), data = use_param_values) +
     facet_wrap(vars(trait), scales = "free_y") +
     scale_fill_manual(values = pfts("color")) +
     scale_color_manual(values = c("default" = "red1", "median" = "blue1")) +

### 2019-10-28 analysis/paper/paper.Rmd
Minor cleanup of paper before sending to coauthors

@@
   Although many simulations matched either the observed NPP or observed LAI at some point during the simulation, only a small minority of individual runs matched both.
+  The few runs that matched observed NPP and LAI also matched the present-day PFT mixture (early hardwood-dominated) at UMBS.
   Structural uncertainty was less important than parameter uncertainty, but model configuration did influence the ecological dynamics within a given parameter set, and affected which parameters were most influential.
@@
 highlights: |
-  These are the highlights.
+  TODO
 ---
@@
 **More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
-Vegetation models fall broadly along the following spectrum of complexity, which is a direct consequence of trade-offs between **realism and process fidelity** on one hand and **tractability, computational demand, and data requirements** on the other [@hawkes_2000_woody].
+Vegetation models fall along a spectrum of complexity, which is a direct consequence of trade-offs between **realism and process fidelity** on one hand and **tractability, computational demand, and data requirements** on the other [@hawkes_2000_woody].
 <!-- TODO Might need a better ref here -->
-On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the (weighted) average of all the vegetation at that site <!-- TODO REFS -->
-On the other end of the spectrum are **"individual-based" models** (a.k.a.
-"gap models"), which explicitly simulate multiple indivduals competing for resources at a single site [@shugart_2015_computer].
-<!-- TODO other refs?
--->
-Examples of such models are LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME. <!-- TODO ref --> <!-- TODO Other examples? -->
-Because these models can explicitly represent inter-specific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
-Between these two extremes are models that use approximations to attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individual plants without the need for explicitly simulating each individual [@purves_2008_predictive].
-One specific example of this approach is the **Ecosystem Demography model** [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
+On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the weighted average of all the vegetation at that site <!-- TODO REFS -->
+On the other are **"individual-based" or "gap" models**---such as LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME [@shuman_2015_forest]---which explicitly simulate individuals competing for resources at a single site [@shugart_2015_computer].
+Because these models explicitly represent interspecific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
+Between these two extremes are models that use approximations to attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individuals without explicitly simulating each individual [@purves_2008_predictive].
+One example of this approach is the **Ecosystem Demography model** [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
 <!-- BBL: Perhaps mention ED2's influence in ESMs and LSMs -->
@@
 **Regardless of vegetation model complexity, their projections are inherently uncertain.** 
-This uncertainty comes from many different sources, which can broadly be classified into the following categories:
-_Driver_ uncertainty refers to uncertainty in data about processes not represented by the model (e.g. weather and climate for vegetation models).
-_Initial condition_ uncertainty arises from the fact that models have to start somewhere, and the exact conditions at the place and time simulations begin are frequently unknown.
-_Process_ or _structural_ uncertainty arises because vegetation models necessarily only represent a subset of all processes involved in plant biology, and that the process that are included can usually be represented in multiple different ways (i.e.
+This uncertainty comes from many different sources, which can broadly be classified as follows:
+_Driver_ uncertainty is uncertainty in data about processes not represented by the model (e.g. weather and climate for vegetation models).
+_Initial condition_ uncertainty arises from the fact that the exact conditions at the place and time simulations begin are frequently unknown.
+_Process_ or _structural_ uncertainty arises because models necessarily represent only a subset of all processes involved in plant biology, and that those processes that are included can usually be represented in multiple different ways (i.e.
 which processes are included, and how they are represented).
 Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
-**The relative importance of these sources of uncertainty in ecological forecasts is a key question because it is directly related to future research priorities**.
+**The relative importance of these sources of uncertainty in ecological forecasts is important as a driver of future research priorities**.
 
@@
 <!-- TODO This paragraph is janky as hell. Needs heavy revision. -->
-(Chris/Jeff: Help)
+{Chris/Jeff: Help!}
 Previous work found that an earlier version of ED had difficulty sustaining multiple plant functional types in long simulations [@fisher_2010_assessing].
@@
 
-**This study focuses on the interaction between parameter and structural uncertainty in the ED2 model.**
+**This study focuses on the interaction between parameter and structural uncertainty.**
 Our study is organized around the following guiding questions:
-(1) Which processes related to light utilization are most important to accurately modeling community succession and C cycle dynamics in temperate forests of the Upper Midwest?
-(2) What is the cost of considering these processes, in terms of additional parameter uncertainty?
-(3) What are the relative contributions of parameter uncertainty (data limitation) vs. structural uncertainty (theoretical limitation?)?
-We ran ED2 ensemble simulations with a factorial combination of submodels related to radiative transfer formulation, horizontal competition, and trait plasticity.
+(1) Which processes related to light utilization are most important to accurately model community succession and C cycling in Upper Midwest forests?
+(2) What are the relative contributions of parameter uncertainty (data limitation) vs. structural uncertainty (theoretical limitation)?
+We ran ensemble simulations of the Ecosystem Demography (ED-2.2) model with a factorial combination of submodels related to radiative transfer formulation, horizontal competition, and trait plasticity.
 We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results in the context of ongoing climate change.
@@
 This composition is a legacy of the site's disturbance history:
-The site was intensively logged in the late 1800s and early 1900s, and experienced regularly recurring fires until the mid-1920s, at which point a regime of active fire suppression started that has persisted to the present day.
-<!-- TODO ref -->
-As a result, the average stand age in 2013 was 95 years. <!-- TODO ref -->
+The site was intensively logged in the late 1800s and early 1900s, and experienced regularly recurring fires until the mid-1920s, at which point a regime of active fire suppression started that has persisted to the present day (REF).
+As a result, the average stand age in 2013 was 95 years (REF).
 The majority of the forest that is aspen-dominated is undergoing succession to "northern hardwood" (maple, beech, basswood, birch, ash, hemlock), "upland conifer" (red and white pine), or "northern red-oak" ecotypes [@bergen_2007_observing].
@@
 In this study, we ran a factorial combination of the following ED-2.2 configurations: (1) two-stream vs.
-multiple-scatter canopy radiative transfer models (RTMs); (2) infinite vs.
+multiple-scatter canopy radiative transfer models (RTMs); (2) closed canopy vs.
 finite crown area (a.k.a.
@@
 The crown area submodel in ED-2.2 determines the nature of competition for light between cohorts.
-In the default configuration ("infinite crown area", or "complete shading"), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
+In the default configuration ("closed canopy"), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
 This means that taller cohorts always receive more incoming radiation than shorter cohorts, even when the height difference is small.
-This has been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous canopies [@fisher_2015_taking].
-In the alternate configuration ("finite crown area", or "partial shading"), canopies take up only a fraction of the available horizontal area, which generally exposes the understory to more light.
+Similar approaches have been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous forest patches [@fisher_2015_taking].
+In the alternate configuration ("finite crown area"), canopies take up only a fraction of the available horizontal area, which generally exposes the understory to more light.
 The horizontal area of crowns is determined by allometric equations from @dietze_2008_changing.
@@
 
-Other parameters:
+Other parameters (TODO, briefly):
 
@@
 Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
-All parameters not described in this section were set to their ED-2.2 PFT-specific default values.
+All parameters not described in this section were set to their ED-2.2 PFT-specific default values for all simulations.
 
@@
 Each ensemble member ran from January 1, 1902 through December 31, 1999 and was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
-Driving meteorological data was 6-hourly CRU-NCEP <!-- TODO ref.
-More here?
---> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
-Soil texture was set to 92% sand, 7% silt, and 1% clay, per site-level observations in [@gough_2010_wood].
+Driving meteorological data was 6-hourly CRU-NCEP (REF) combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
+Soil texture was set to 92% sand, 7% silt, and 1% clay, based on site-level observations [@gough_2010_wood].
 The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
@@
 
-The overarching objective of this study was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community assembly.
-
 ## ED-2.2 vs. reality
 
+The overarching objective of this study was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community assembly.
 Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
@@
 <!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-##### pagebreak
+<!-- ##### pagebreak -->
+
+\newpage
 
@@
 
-##### pagebreak
+<!-- ##### pagebreak -->
+
+\newpage
 

### 2019-11-04 analysis/paper/paper.Rmd
Add DOCX export

@@
       fig_caption: yes
-      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
     bookdown::pdf_document2:
       latex_engine: xelatex
+    bookdown::word_document2:
+      reference_docx: "../templates/taylor-francis-article.docx" # Insert path for the DOCX file
+always_allow_html: true
 bibliography: references.bib
@@
 tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-kable(tbl, escape = is_latex_output(), caption = paste0(
-  "Parameter descriptions. ",
-  "For additional information, see the ",
+cap <- paste(
+  "Parameter descriptions.",
+  "For additional information, see the",
   "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
-), booktabs = TRUE) %>%
+)
+if (is_latex_output() || is_html_output()) {
+  kable(tbl, escape = is_latex_output(), caption = cap, booktabs = TRUE) %>%
     kable_styling(latex_options = "scale_down")
+} else {
+  # Probably DOCX output
+  cap <- paste("Table: (\\#tab:param-table)", cap)
+  flextable::flextable(tbl) %>%
+    flextable::autofit() %>%
+    flextable::fit_to_width(7.5) %>%
+    flextable::set_caption(cap)
+}
 ```

### 2019-11-04 analysis/templates/template.Rmd
Add DOCX export

---
title: |
  My title
author:
- Author 1
- Author 2
date: '`r Sys.Date()`'
output:
  bookdown::word_document2
abstract: |
  My abstract.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Heading 1

## Heading 2

### Heading 3

#### Heading 4

##### Heading 5 - this is for the page break with MS Word

In the MS Word document, the following modifications must be made in the 'Modify Style' menu:

+ Set the font color to ‘white’ (rather than ‘Automatic’).
+ Select the smallest font size (8 rather than 11).
+ Select ‘Page break before’ in the ‘Line and Page Breaks’ tab.
+ Set the line spacing to ‘Exactly’ and ‘1 pt’ in the ‘Indents and Spacing’ tab.

After these tweaks, the ‘Heading 5’ style will no longer format a heading of level 5. Instead it will insert a very small and white (and, thus, invisible) line followed by a page break.

Now knit this doc, and edit the styles in the resulting docx file. Then rename as 'template' and use this in the yml:

```
output:
  word_document:
    reference_docx: templates/template.docx
```

Don't forget line numbers for JAS! Headings are bold, underline, nothing. And Arial Narrow 10pt for tables.

See here for more information: http://rmarkdown.rstudio.com/articles_docx.html

```{r , fig.cap = "This is a figure caption"}
plot(Sepal.Length ~ Sepal.Width, data = iris)
```

```{r}
tab <- flextable::flextable(head(iris))
tab <- flextable::autofit(tab)
tab <- flextable::set_caption(tab, "Table: (\\#tab:table1) This is a table.")
tab
```

### 2019-11-11 analysis/paper/paper.Rmd
Address some comments in paper

@@
   Structure and parameter uncertainty in centennial-scale projections
-  of community succession in an Upper Midwest temperate forest
+  of community-carbon cycling in a temperate forest
 author:
@@
   (2) What are the relative contributions of parameter vs. structural uncertainty?
-  We ran ensembles of the Ecosystem Demography model v2.2 (sampling over parameter uncertainty) from 1900 to 2000 with different representations of key processes important to competition for light.
+  We ran ensembles of the Ecosystem Demography model v2.2 (sampling over parameter uncertainty) at the University of Michigan Biological Station (UMBS) from 1900 to 2000 with different representations of key processes important to competition for light.
   We then compared the magnitude of structural (across representations) and parameter (within each representation) uncertainty, and assessed which submodel-parameter combinations reproduced observed C fluxes and succession.
@@
   Although many simulations matched either the observed NPP or observed LAI at some point during the simulation, only a small minority of individual runs matched both.
-  The few runs that matched observed NPP and LAI also matched the present-day PFT mixture (early hardwood-dominated) at UMBS.
+  The few runs that matched observed NPP and LAI also matched the present-day composition (early hardwood-dominated).
   Structural uncertainty was less important than parameter uncertainty, but model configuration did influence the ecological dynamics within a given parameter set, and affected which parameters were most influential.
-  We conclude that: (1) targeted data collection to constrain vegetation model parameters should be a higher priority than model development; (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and (3) robust model validation requires scrutiny of multiple different variables, including those related to vegetation productivity, structure, and composition.
+  We conclude that: (1) targeted data collection to constrain vegetation model parameters should be a higher priority than model development; (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and (3) robust model validation requires scrutiny of multiple different variables, including those related to vegetation productivity and composition.
 keywords: |
@@
 @dietze_2014_quantitative investigated which parameters were most important to ED2 predictions of C sequestration across a diverse range of sites in North America.
-The found that the parameters contributing the most to overall predictive uncertainty were those related to growth respiration, mortality, stomatal conductance, and water uptake, and that parameter uncertainty varied with biome.
+They found that the parameters contributing the most to overall predictive uncertainty were those related to growth respiration, mortality, stomatal conductance, and water uptake, and that parameter uncertainty varied with biome.
 More recently, @raczka_2018_what evaluated parameter uncertainty over 100 year time scale at the Willow Creek Ameriflux site, and found that the most important parameters to model uncertainty were quantum efficiency of photosynthesis, leaf respiration, and soil-plant water transfer.
@@
 This composition is a legacy of the site's disturbance history:
-The site was intensively logged in the late 1800s and early 1900s, and experienced regularly recurring fires until the mid-1920s, at which point a regime of active fire suppression started that has persisted to the present day (REF).
-As a result, the average stand age in 2013 was 95 years (REF).
+The site was intensively logged and burned in the late 1800s and early 1900s, and experienced regularly recurring fires until the mid-1920s, at which point a regime of active fire suppression started that has persisted to the present day [@kilburn_1960_effects; @barnes_1981_michigan; @friedman_2005_regional_legacies_logging].
+As a result, the average stand age in 2013 was 95 years [@gough_2013_sustained].
 The majority of the forest that is aspen-dominated is undergoing succession to "northern hardwood" (maple, beech, basswood, birch, ash, hemlock), "upland conifer" (red and white pine), or "northern red-oak" ecotypes [@bergen_2007_observing].
@@
 
-For this study, we used the Ecosystem Demography Model, version 2.2 (ED-2.2).
+For this study, we used the Ecosystem Demography Model, version 2.2 (ED-2.2) to simulate forest C cycling and plant functional type (PFT) composition.
 A full description of the default configuration of the model is provided by @longo_2019_ed1.
@@
 <!-- TODO Fill in Acknowledgements -->
-This project funded by NSF grant.
-Cyberinfrastructure provided by Pacific Northwest National Laboratory (PNNL).
+This project was supported by NSF Award 1655095.
+Cyberinfrastructure capabilities were provided by the Pacific Northwest National Laboratory (PNNL).
 Data from University of Michigan Biological Station (UMBS).

### 2019-11-11 analysis/paper/paper.Rmd
Revise methods-submodels section

@@
 
-In this study, we ran a factorial combination of the following ED-2.2 configurations: (1) two-stream vs.
+For our analysis of model structure, we chose to target processes of known ecological importance.
+Specifically, we ran ED-2.2 with a factorial combination of the following configurations: (1) two-stream vs.
 multiple-scatter canopy radiative transfer models (RTMs); (2) closed canopy vs.
@@
 partial shading); and (3) static vs.
-light-plastic traits.
+plastic (varying with light level) traits.
 
@@
 "hemispherical", "isotropic") radiation at each layer using a linear matrix equation.
-Where the models differ is in the terms of this matrix equation.
+The models differ in the terms of this matrix equation.
 The two-stream model (default) uses the two-stream approximation originally derived for atmospheric radiative transfer [@meador_1980_two] and later adapted to vegetation canopies [@dickinson_1983_land; @sellers_1985_canopy].
@@
 In the default configuration ("closed canopy"), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
-This means that taller cohorts always receive more incoming radiation than shorter cohorts, even when the height difference is small.
+This means that taller cohorts always receive more incoming radiation---including 100% of the direct radiation---than shorter cohorts, even when the height difference is small.
 Similar approaches have been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous forest patches [@fisher_2015_taking].
-In the alternate configuration ("finite crown area"), canopies take up only a fraction of the available horizontal area, which generally exposes the understory to more light.
+In the alternate configuration ("finite crown area"), canopies take up only a fraction of the available horizontal area, which allows more light (including some direct radiation) to reach subdominant cohorts.
 The horizontal area of crowns is determined by allometric equations from @dietze_2008_changing.
@@
 This ignores the globally-documented intraspecific trait variability as a function of light level [@niinemets_2010_review; @keenan_2016_global].
-<!-- TODO Also, classical ecological succession theory bazzaz_1979_physiological?
--->
 In the alternate configuration, as light level decreases (trees become more shaded), specific leaf area increases and V~cmax~ decreases, following empirical relationships from the tropics [@lloyd_2010_optimisation].

### 2019-11-11 analysis/paper/paper.Rmd
Major trim of parameter methods section

@@
 Instead, we selected a subset of parameters (Table \@ref(tab:param-table)) guided by (but expanding upon) previous ED-2.2 sensitivity studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
-
-Three parameters are related to leaf-level physiology:
-Following the enzyme-kinetic model of @farquhar_1980_biochemical, the rate of photosynthesis is the minimum of light-limited and enzyme-limited reactions.
-The former are controlled by the _quantum efficiency_ parameter---maximum efficiency with which absorbed photosynthetically active radiation is converted to CO~2~.
-The latter are controlled by _V~cmax~_, the maximum rate of carbon fixation by Rubisco.
-The water demand of photosynthesis is controlled by the _stomatal slope_, the sensitivity of stomatal conductance of CO~2~ as a function of CO~2~ concentration and humidity at the leaf surface [@leuning_1995_critical; @ball_1987_model].
-Three more parameters correspond to the respiration rates of _leaves_, _roots_, and "_growth maintenance_".
-
-Two more parameters control carbon allocation:
-One is the ratio of fine root to leaf biomass (`fineroot2leaf`, or `q`), and another is the ratio of "storage" carbon allocated to reproduction (`r_fract`).
-
-Three parameters control various aspects of adult tree mortality.
-The overall adult mortality rate in ED-2.2 ($M$) is the sum of density-independent mortality from aging ($M_I$, plants year^-1^), density-dependent mortality from C starvation ($M_D$), and mortality from cold/frost ($M_F$) (ED-2.2 technically also includes additional term for fire mortality, but we did not include fire in our simulations):
-
-$$
-M = M_I + M_D + M_F
-$$
-
-Density-independent mortality from aging ($M_I$) is a prescribed, PFT-specific parameter (`mort3`).
-Density-dependent mortality from C starvation is calculated as a function of a cohort's C balance limitation:
-
-$$
-M_D = \frac{y_1}{1 + \exp\left[y_2 \left( \frac{C_k}{C_k^*} \right)\right]}
-$$
-
-where $C_k$ is the 12 month running mean C balance of the cohort, $C_k^*$ is the running mean ideal C balance if there was no light or water limitation, and $y_1$ (`mort1`; plants year^-1^) and $y_2$ (`mort2`; unitless) are PFT-specific parameters.
-_Seedling mortality_ rate is prescribed as its own PFT-specific parameter.
-
-Several parameters are related to canopy structure and radiative transfer.
-_Specific leaf area_ (`SLA`) is used to convert leaf biomass to leaf area index, which in turn is used in a variety of calculations related to canopy radiative transfer and micrometeorology.
-Canopy _clumping factor_ describes how evenly leaf area is distributed in horizontal space (1 being perfectly evenly; 0 being a "black hole" where all leaves are concentrated in a single point); and
-leaf _orientation factor_ describes the average distribution of leaf angles (-1 being perfectly vertical, 1 being perfectly horizontal, and 0 being random).
-Four parameters control leaf optical properties, namely the fractions of light reflected or transmitted in visible and near-infrared wavelengths (`leaf_(reflect|trans)_(vis|nir)`).
-Details of how these parameters influence canopy radiative transfer are described in the Supplementary Information.
-
-Other parameters (TODO, briefly):
-
-- Fraction of litter carbon that enters the labile (fast) soil carbon pool (`f_labile`)
-- Minimum plant height for reproduction (`repro_min_h`)
-- Soil-plant water conductance, or "water availability factor" (`water_conductance`)
-- C:N ratio in fine roots (`c2n_fineroot`) and leaves (`c2n_leaf`)
-- Leaf turnover rate (evergreen-only; `leaf_turnover_rate`)
-- Proportion of dispersal that is "global" (`nonlocal_dispersal`) <!-- TODO What is this? Can I drop it from the analysis? -->
+Collectively, these parameters cover a range of physiological and ecological processes relevant to C cycling and plant community composition, including 
+leaf physiology (light and dark reactions of photosynthesis; respiration; evapotranspiration),
+carbon allocation,
+plant organ stoichiometry,
+canopy architecture and radiative transfer,
+dispersal,
+reproduction,
+and mortality.
+A full description of these parameters and their associated processes in ED-2.2 is provided in @longo_2019_ed1.
 

### 2019-11-11 analysis/paper/parameters.Rmd
Major trim of parameter methods section

Three parameters are related to leaf-level physiology:
Following the enzyme-kinetic model of @farquhar_1980_biochemical, the rate of photosynthesis is the minimum of light-limited and enzyme-limited reactions.
The former are controlled by the _quantum efficiency_ parameter---maximum efficiency with which absorbed photosynthetically active radiation is converted to CO~2~.
The latter are controlled by _V~cmax~_, the maximum rate of carbon fixation by Rubisco.
The water demand of photosynthesis is controlled by the _stomatal slope_, the sensitivity of stomatal conductance of CO~2~ as a function of CO~2~ concentration and humidity at the leaf surface [@leuning_1995_critical; @ball_1987_model].
Three more parameters correspond to the respiration rates of _leaves_, _roots_, and "_growth maintenance_".

Two more parameters control carbon allocation:
One is the ratio of fine root to leaf biomass (`fineroot2leaf`, or `q`), and another is the ratio of "storage" carbon allocated to reproduction (`r_fract`).

Three parameters control various aspects of adult tree mortality.
The overall adult mortality rate in ED-2.2 ($M$) is the sum of density-independent mortality from aging ($M_I$, plants year^-1^), density-dependent mortality from C starvation ($M_D$), and mortality from cold/frost ($M_F$) (ED-2.2 technically also includes additional term for fire mortality, but we did not include fire in our simulations):

$$
M = M_I + M_D + M_F
$$

Density-independent mortality from aging ($M_I$) is a prescribed, PFT-specific parameter (`mort3`).
Density-dependent mortality from C starvation is calculated as a function of a cohort's C balance limitation:

$$
M_D = \frac{y_1}{1 + \exp\left[y_2 \left( \frac{C_k}{C_k^*} \right)\right]}
$$

where $C_k$ is the 12 month running mean C balance of the cohort, $C_k^*$ is the running mean ideal C balance if there was no light or water limitation, and $y_1$ (`mort1`; plants year^-1^) and $y_2$ (`mort2`; unitless) are PFT-specific parameters.
_Seedling mortality_ rate is prescribed as its own PFT-specific parameter.

Several parameters are related to canopy structure and radiative transfer.
_Specific leaf area_ (`SLA`) is used to convert leaf biomass to leaf area index, which in turn is used in a variety of calculations related to canopy radiative transfer and micrometeorology.
Canopy _clumping factor_ describes how evenly leaf area is distributed in horizontal space (1 being perfectly evenly; 0 being a "black hole" where all leaves are concentrated in a single point); and
leaf _orientation factor_ describes the average distribution of leaf angles (-1 being perfectly vertical, 1 being perfectly horizontal, and 0 being random).
Four parameters control leaf optical properties, namely the fractions of light reflected or transmitted in visible and near-infrared wavelengths (`leaf_(reflect|trans)_(vis|nir)`).
Details of how these parameters influence canopy radiative transfer are described in the Supplementary Information.

Other parameters (TODO, briefly):

- Fraction of litter carbon that enters the labile (fast) soil carbon pool (`f_labile`)
- Minimum plant height for reproduction (`repro_min_h`)
- Soil-plant water conductance, or "water availability factor" (`water_conductance`)
- C:N ratio in fine roots (`c2n_fineroot`) and leaves (`c2n_leaf`)
- Leaf turnover rate (evergreen-only; `leaf_turnover_rate`)
- Proportion of dispersal that is "global" (`nonlocal_dispersal`) <!-- TODO What is this? Can I drop it from the analysis? -->

### 2019-11-11 analysis/paper/paper.Rmd
Move Figure 2 to Supplement (S1)

@@
 First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values (Figure \@ref(fig:structure-compare-default)).
-To understand the physical and ecological mechanisms underlying these differences, we also compared the cohort-level composition, height, and canopy radiation profiles in these simulations in four stages of the simulation (Figure \@ref(fig:structure-compare-light)).
+To understand the physical and ecological mechanisms underlying these differences, we also compared the cohort-level composition, height, and canopy radiation profiles in these simulations in four stages of the simulation (Figure S1).
 
@@
 We then consider impacts on composition by looking at predicted PFT fractions of total aboveground biomass (AGB) at four different years in the simulation (Figure \@ref(fig:pft-jitterplot)) and at time series of LAI by PFT for five parameter combinations spanning the range of predicted ecological outcomes (Figure \@ref(fig:lai-pft-plot)).
-
 Finally, we assessed which parameters contributed the most to parameter uncertainty for each model structure.
@@
 
-Different model structures agreed for the first half of the simulation, at which point they started to diverge (Figure \@ref(fig:structure-compare-default)).
+Different model structures agreed for the first half of the century-long simulation, at which point they started to diverge (Figure \@ref(fig:structure-compare-default)).
 Post 1950, the combination of two-stream RTM, finite canopy radius, and plastic traits had the highest productivity, followed by the same configuration with static traits, with the remaining configurations having very similar results.
@@
 
-```{r structure-compare-light, fig.cap = structure_default_light_cap}
-### Figure 2
-structure_default_light_cap <- paste(
-  htmlfig(),
-  "Radiation profiles of cohorts from ED2 runs with",
-  "default parameters, across different configurations.",
-  "Each point represents a single cohort of the given PFT,",
-  "position on the y-axis represents the height of that cohort,",
-  "and position on the x-axis represents the relative light level",
-  "received by that cohort."
-)
-fig("default-light-levels.png")
-```
-
-Differences in productivity across model structures were related to differences in composition and vertical canopy radiation profiles over the course of the simulation (Figure \@ref(fig:structure-compare-light)).
+Differences in productivity across model structures were related to differences in composition and vertical canopy radiation profiles over the course of the simulation (Figure S1).
 Enabling the finite canopy radius increased the availability of light for understory cohorts throughout the simulation, particularly in the first half of the simulation.

### 2019-11-11 analysis/paper/paper.Rmd
Move param-dist figure to supplement (S2)

@@
 
-Second, we compared simulations with ED-2.2 default parameters against simulations using median parameter estimates from a trait meta-analysis (Figure \@ref(fig:param-dist)).
+Second, we compared simulations with ED-2.2 default parameters against simulations using median parameter estimates from a trait meta-analysis (Figure S2).
 Here, we looked predictions of annual mean C fluxes (gross and net primary productivity, and autotrophic and heterotrophic respiration; Figure \@ref(fig:default-median-flux)) and of summer LAI by PFT (Figure \@ref(fig:default-median-lai)).
@@
 
-```{r param-dist, fig.cap = param_dist_cap}
-### Figure 3
-param_dist_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions from PEcAn trait meta-analysis."
-)
-fig("param-dist.png")
-```
-
-Parameter estimates from the trait meta-analysis were often substantially different from the ED-2.2 defaults (Figure \@ref(fig:param-dist)).
+Parameter estimates from the trait meta-analysis were often substantially different from the ED-2.2 defaults (Figure S2).
 In many cases, median parameter estimates were different entirely because of the choice of prior distributions.
@@
 First, in every model configuration (except finite canopy, two-stream RTM, and static traits), gross primary productivity (GPP) reached a higher peak, and did so earlier under median parameters (Figure \@ref(fig:default-median-flux)).
-This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure \@ref(fig:param-dist)).
-However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure \@ref(fig:default-median-flux)), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure \@ref(fig:param-dist)).
+This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure S2).
+However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure \@ref(fig:default-median-flux)), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure S2).
 The resulting net primary productivity (NPP) had similar peak values under default and median parameters, though the temporal patterns were quite different in most model configurations (Figure \@ref(fig:default-median-flux)).
@@
 This is encouraging because it suggests that it is possible to get both aggregate ecosystem C fluxes and PFT composition right.
-Moreover, given that many important ED-2.2 parameters were well-constrained in our analysis (\@ref(fig:param-dist)), this is possible without using physically unrealistic parameter values.
+Moreover, given that many important ED-2.2 parameters were well-constrained in our analysis (Figure S2), this is possible without using physically unrealistic parameter values.
 We therefore suggest that for model calibration and validation, comparisons against multiple output variables---including variables related to composition---can yield maximum constraint and are not unreasonably strict.
@@
 For parameters that are already well constrained by data, the marginal value of additional measurements may be limited.
-For example, ED-2.2 is highly sensitive to SLA [@dietze_2014_quantitative; @pandit_2018_optimizing], but thanks to the large number of SLA observations, we were able to estimate it very precisely (Figure \@ref(fig:param-dist)) and, as a result, it was not a significant source of uncertainty in our analysis (Figure \@ref(fig:partial-variance-npp)).
-At the same time, some poorly constrained parameters have relatively modest impacts on model outputs---for instance, nonlocal dispersal and background mortality rate in our study (Figure \@ref(fig:param-dist), \@ref(fig:partial-variance-npp))---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
+For example, ED-2.2 is highly sensitive to SLA [@dietze_2014_quantitative; @pandit_2018_optimizing], but thanks to the large number of SLA observations, we were able to estimate it very precisely (Figure S2) and, as a result, it was not a significant source of uncertainty in our analysis (Figure \@ref(fig:partial-variance-npp)).
+At the same time, some poorly constrained parameters have relatively modest impacts on model outputs---for instance, nonlocal dispersal and background mortality rate in our study (Figure S2, \@ref(fig:partial-variance-npp))---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
 Ultimately, this points to the value of model-data feedbacks [@dietze_2013_improving] in advancing ecological research and setting future research agendas.

### 2019-11-11 analysis/paper/paper.Rmd
Minor revisions and comments in results and disc.

@@
 The large parameter uncertainty in NPP and LAI also extended to species composition (Figure \@ref(fig:pft-jitterplot)).
-Under all configurations, at any moment in time, most ensemble members were dominated by a single PFT, and the sustained coexistence of more than two PFTs was limited to only a handful specific ensemble members across all configurations.
+Under all configurations, at any moment in time, most ensemble members were dominated by a single PFT, and the sustained coexistence of more than two PFTs was limited to a small set of specific ensemble members across all configurations.
 However, _which_ PFT was dominant depended strongly on parameter values, and any of the four PFTs became the dominant one in some subset of simulations.
@@
 Parameter set **C** showed much more variety in compositional dynamics with model structure:
-In all cases, mid-hardwoods began as the dominant species and were eventually overtaken by either pine (closed canopy and plastic traits) or early hardwood (all other cases; though with a closed canopy and static traits, the final LAI was low).
+In all cases, mid-hardwoods began as the dominant PFT and were eventually overtaken by either pine (closed canopy and plastic traits) or early hardwood (all other cases; though with a closed canopy and static traits, the final LAI was low).
 Parameter sets **D** and **E** had the highest productivity of these five sets, and were characterized by strong dominance of Early Hardwood and Pine, respectively.
@@
 
-The overarching objective of this study was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community assembly.
+The overarching objective of this study was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community succession.
 Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?

### 2019-11-11 analysis/paper/paper.Rmd
Add (and address some) comments from Jeff

@@
 Previous work found that an earlier version of ED had difficulty sustaining multiple plant functional types in long simulations [@fisher_2010_assessing].
-One of the primary drivers of successional change in temperate forests is competition for light (REF Bazzaz 19XX) <!-- TODO REF -- Bazaaz?
+One of the primary drivers of successional change in temperate forests is competition for light [@canham_1990_light]  <!-- TODO REF -- Bazaaz?
 -->, and differences in plants' strategies for light acquisition and utilization are a key driver of niche differentiation in these ecosystems <!-- TODO REF -->.
@@
 We did this by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what (Figure \@ref(fig:partial-variance-npp)).
-Briefly, this analysis contains uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance").
+Briefly, this analysis uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance").
 For more details, see @lebauer_2013_facilitating.
@@
 Different model structures agreed for the first half of the century-long simulation, at which point they started to diverge (Figure \@ref(fig:structure-compare-default)).
-Post 1950, the combination of two-stream RTM, finite canopy radius, and plastic traits had the highest productivity, followed by the same configuration with static traits, with the remaining configurations having very similar results.
+Post 1950, the combination of two-stream RTM, finite canopy radius, and plastic traits had the highest productivity, followed by the same configuration with static traits, with the remaining configurations having similar results.
 Interestingly, the effects of changing individual model configurations were not additive --- for example, enabling trait plasticity had a much stronger positive effect when paired with the two-stream canopy RTM than with the multiple-scatter canopy RTM.
@@
 
-Third, even where NPP was similar between default and median runs, the underlying forest composition was very different (Figure \@ref(fig:default-median-lai)).
+Third, even where NPP was similar between default and median runs, the underlying forest composition was different (Figure \@ref(fig:default-median-lai)).
 With the default parameterization, ED-2.2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
@@
 However, although our simulations used real meteorology (which included occasional extreme weather events), **we did not consider other sources of disturbance** common to this region [such as harvests, insects and pathogens, and fire; @amiro_2010_ecosystem], which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
-Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance frequencies and see how they affect ecological dynamics.
+Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance regimes and see how they affect ecological dynamics.
 

### 2019-11-12 analysis/paper/paper.Rmd
Formatting tweaks for IR release

@@
 
-**Forests are a defining feature of the US North**.
+Forests are a defining feature of the US North.
 This region is the most heavily forested region in the US, and these forests are an essential provider of employment and recreation opportunities and other ecosystem services for the region's people, and of habitat for the region's wildlife [@shividenko_2005_forest; @shifley_2012_forests].
-In addition, the **forests of the Upper Midwest** in particular have been a **net carbon sink** for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
-However, the **fate of these forests is highly uncertain**.
-Many Upper Midwest forests are undergoing a **shift in community composition from early-successional to mid-successional** species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
-In addition, these forests face a multitude of **direct and indirect pressures from both biotic and abiotic sources** [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.
+In addition, the forests of the Upper Midwest in particular have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
+However, the fate of these forests is highly uncertain.
+Many Upper Midwest forests are undergoing a shift in community composition from early-successional to mid-successional species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
+In addition, these forests face a multitude of direct and indirect pressures from both biotic and abiotic sources [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.
 <!-- TODO More here?
 -->
-Collectively, these processes have culminated in forests that are different from their pre-settlement climax counterparts [@stearns_2002_one; @wolter_2002_recent; @thompson_2013_four], and **our ability to make predictions about such non-analog environments based on the past is limited**.
+Collectively, these processes have culminated in forests that are different from their pre-settlement climax counterparts [@stearns_2002_one; @wolter_2002_recent; @thompson_2013_four], and our ability to make predictions about such non-analog environments based on the past is limited.
 
-**More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.**
-Vegetation models fall along a spectrum of complexity, which is a direct consequence of trade-offs between **realism and process fidelity** on one hand and **tractability, computational demand, and data requirements** on the other [@hawkes_2000_woody].
+More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.
+Vegetation models fall along a spectrum of complexity, which is a direct consequence of trade-offs between realism and process fidelity on one hand and tractability, computational demand, and data requirements on the other [@hawkes_2000_woody].
 <!-- TODO Might need a better ref here -->
-On one side are relatively simple **"big leaf" models**---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the weighted average of all the vegetation at that site <!-- TODO REFS -->
-On the other are **"individual-based" or "gap" models**---such as LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME [@shuman_2015_forest]---which explicitly simulate individuals competing for resources at a single site [@shugart_2015_computer].
+On one side are relatively simple "big leaf" models---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the weighted average of all the vegetation at that site <!-- TODO REFS -->
+On the other are "individual-based" or "gap" models---such as LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME [@shuman_2015_forest]---which explicitly simulate individuals competing for resources at a single site [@shugart_2015_computer].
 Because these models explicitly represent interspecific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
 Between these two extremes are models that use approximations to attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individuals without explicitly simulating each individual [@purves_2008_predictive].
-One example of this approach is the **Ecosystem Demography model** [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
+One example of this approach is the Ecosystem Demography model [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
 <!-- BBL: Perhaps mention ED2's influence in ESMs and LSMs -->
 
-**Regardless of vegetation model complexity, their projections are inherently uncertain.**
+Regardless of vegetation model complexity, their projections are inherently uncertain.
 This uncertainty comes from many different sources, which can broadly be classified as follows:
@@
 Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
-**The relative importance of these sources of uncertainty in ecological forecasts is important as a driver of future research priorities**.
+The relative importance of these sources of uncertainty in ecological forecasts is important as a driver of future research priorities.
 
-**Many ecosystem modeling studies have looked at these categories of uncertainty independently.**
+Many ecosystem modeling studies have looked at these categories of uncertainty independently.
 <!-- TODO Consider cutting 50-75% of these next few sentences. Probably way too much detail. -->
@@
 Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large impact on predictions of net primary productivity [@viskari_2019_influence].
-{**Driver uncertainty** -- a few examples?}
+<!-- {Driver uncertainty -- a few examples?} -->
 
-**Other studies have investigated structural uncertainty.**.
+Other studies have investigated structural uncertainty..
 Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization [@zaehle_2014_evaluation; @medlyn_2015_using; @walker_2015_predicting], and soil C sequestration [@sulman_2018_multiple], among others.
@@
 This structural uncertainty has been attributed to differences in model representations of key processes, including canopy radiative transfer [@fisher_2017_vegetation], soil biogeochemistry [@wieder_2017_carbon; @sulman_2018_multiple], stomatal conductance [@franks_2018_comparing], and photosynthesis [@rogers_2016_roadmap].
-**However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.**
-**Moreover, comparative analysis of contributions of different types of uncertainty to model projections are rare in the ecosystem modeling literature**.
+However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.
+Moreover, comparative analysis of contributions of different types of uncertainty to model projections are rare in the ecosystem modeling literature.
 <!--
@@
 
-**This study focuses on the interaction between parameter and structural uncertainty.**
+This study focuses on the interaction between parameter and structural uncertainty.
 Our study is organized around the following guiding questions:
@@
 @wang_2003_comparison].
-**Compared to the default two-stream model, the multiple-scatter model slightly increases diffuse light levels in the understory.**
+Compared to the default two-stream model, the multiple-scatter model slightly increases diffuse light levels in the understory.
 Both of these models are described in detail in the Supplementary Information. <!-- TODO Exact reference to Appendix -->
@@
 Each ensemble member ran from January 1, 1902 through December 31, 1999 and was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
-Driving meteorological data was 6-hourly CRU-NCEP (REF) combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
+Driving meteorological data was 6-hourly CRU-NCEP <!-- REF --> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
 Soil texture was set to 92% sand, 7% silt, and 1% clay, based on site-level observations [@gough_2010_wood].
@@
   "The dashed lines indicate the 10th and 90th quantiles of results.",
-  "Black dot with error bars is the observed mean and min/max value",
-   "(TODO REF)."
+  "Black dot with error bars is the observed mean and min/max value."
+  # TODO REF
 )
@@
 Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
-**Ensemble average trajectory matches traditional understanding of succession**---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but **happens too quickly** (Figure \@ref(fig:summary-ts-plot));
+Ensemble average trajectory matches traditional understanding of succession---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but happens too quickly (Figure \@ref(fig:summary-ts-plot));
 real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
-One possible reason for the excessive pace of succession in ED-2.2 is **missing mechanisms for sustaining productivity through early succession**.
-In particular, recent evidence points to the **importance of low-severity, high-frequency disturbances** in sustaining ecosystem productivity [@gough_2016_disturbance].
+One possible reason for the excessive pace of succession in ED-2.2 is missing mechanisms for sustaining productivity through early succession.
+In particular, recent evidence points to the importance of low-severity, high-frequency disturbances in sustaining ecosystem productivity [@gough_2016_disturbance].
 ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
@@
 <!-- TODO Look at mortality across ED runs. Does it relate at all to results? -->
-However, although our simulations used real meteorology (which included occasional extreme weather events), **we did not consider other sources of disturbance** common to this region [such as harvests, insects and pathogens, and fire; @amiro_2010_ecosystem], which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
+However, although our simulations used real meteorology (which included occasional extreme weather events), we did not consider other sources of disturbance common to this region [such as harvests, insects and pathogens, and fire; @amiro_2010_ecosystem], which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
 Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance regimes and see how they affect ecological dynamics.
 
-Under all but a small minority of parameter combinations, **ED-2.2 was unable to create---let alone sustain---a forest with any appreciable functional diversity**.
+Under all but a small minority of parameter combinations, ED-2.2 was unable to create---let alone sustain---a forest with any appreciable functional diversity.
 The large majority of runs produced dominance by a single PFT, though _which_ PFT dominated depended on parameter combination.
@@
 To the extent that ED-2.2 is "correct" in how it represents processes, this result also indicates that the plant functional trait space of realized niches is small---i.e.
-multiple different trees can co-exist only under very specific combinations of plant traits and environmental conditions (REFS).
+multiple different trees can co-exist only under very specific combinations of plant traits and environmental conditions.
+<!-- TODO REF -->
 
@@
 Our second question was: What is the relative contribution of structural and parameter uncertainty to C cycling and ecological dynamics at UMBS?
-We found that **parameter uncertainty was by far the dominant driver of uncertainty in our simulations**.
-Different parameter combinations led to **huge ranges in ensemble predictions of ensemble variables**, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures \@ref(fig:summary-ts-plot), \@ref(fig:pairs-time)).
-Moreover, different parameter combinations produced **qualitatively different ecological narratives** (Figures \@ref(fig:pft-jitterplot). \@ref(fig:lai-pft-plot)).
-Therefore, **parameter uncertainty has implications beyond C cycling**---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others (REFS).
+We found that parameter uncertainty was by far the dominant driver of uncertainty in our simulations.
+Different parameter combinations led to huge ranges in ensemble predictions of ensemble variables, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures \@ref(fig:summary-ts-plot), \@ref(fig:pairs-time)).
+Moreover, different parameter combinations produced qualitatively different ecological narratives (Figures \@ref(fig:pft-jitterplot). \@ref(fig:lai-pft-plot)).
+Therefore, parameter uncertainty has implications beyond C cycling---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.
+<!-- TODO REFS -->
 
-The dominance of parameter uncertainty over structural uncertainty in our results suggests that **constraining parameter uncertainty should be a higher priority than model development**.
+The dominance of parameter uncertainty over structural uncertainty in our results suggests that constraining parameter uncertainty should be a higher priority than model development.
 Parameter uncertainty can be reduced in two different ways:
@@
 
-**Importantly, efforts to reduce model uncertainty are most effective when they target parameters that are poorly constrained and to which the model is most sensitive**.
+Importantly, efforts to reduce model uncertainty are most effective when they target parameters that are poorly constrained and to which the model is most sensitive.
 For parameters that are already well constrained by data, the marginal value of additional measurements may be limited.
@@
 and changes to model structure in turn changed the relative importance of specific parameters to overall uncertainty (Figure \@ref(fig:partial-variance-npp)).
-Collectively, this means that **sensitivity and uncertainty analyses are not necessarily generalizable across models**, and highlights the importance of model intercomparison studies.
+Collectively, this means that sensitivity and uncertainty analyses are not necessarily generalizable across models, and highlights the importance of model intercomparison studies.
 It also reinforces our argument that changing model structure---particularly if such changes increase model complexity---should be pursued with extreme caution.
-
-# Conclusions
-
-{TODO}
 
@@
 Cyberinfrastructure capabilities were provided by the Pacific Northwest National Laboratory (PNNL).
-Data from University of Michigan Biological Station (UMBS).
-Data from TRY (TODO: Specific TRY statement).
+Additional data were provided by the University of Michigan Biological Station (UMBS) and the TRY project.
 

### 2019-11-14 analysis/paper/paper.Rmd
Revise introduction paragraph per C and J input

--- 2019-11-11 analysis/paper/paper.Rmd
@@
 
-<!-- TODO This paragraph is janky as hell. Needs heavy revision. -->
-{Chris/Jeff: Help!}
-Previous work found that an earlier version of ED had difficulty sustaining multiple plant functional types in long simulations [@fisher_2010_assessing].
-One of the primary drivers of successional change in temperate forests is competition for light [@canham_1990_light]  <!-- TODO REF -- Bazaaz?
--->, and differences in plants' strategies for light acquisition and utilization are a key driver of niche differentiation in these ecosystems <!-- TODO REF -->.
-Plants compete for light in three different ways:
-Growing quickly (early successional, pioneer species),
-structural complexity / morphological change / filling in gaps in existing canopies,
-and adapting traits to maximize light (and other resource) capture and utilization.
-To this end, we focused our structural uncertainty analysis on three processes related to competition for light:
+Resource---particularly light---availability and distribution drives demographic variation in plant C fixation and community structural changes through succession.
+Observations show that competition for light drives niche differentiation [@silvertown_2004_plant] and consequently successional change [@canham_1990_light; @bazzaz_1993_successional], particularly in closed canopies.
+Prior ED versions were limited in their capacity to sustain multiple plant functional types during stand development due in part to unrealistically aggressive shading of the understory and a lack of consideration for tree crown area, such that the marginal height advantage of a vegetation cohort resulted in a disproportionally large growth advantage [@fisher_2010_assessing].
+Alternative approaches that accommodate horizontal competition for light [e.g.
+perfect plasticity assumption, @strigul_2008_scaling; @weng_2015_scaling] reduce the relative impact of height on light availability and growth, and may therefore promote competition [@fisher_2017_vegetation].
+In addition, trait plasticity may permit individuals to acclimate and species to adapt to changing light environments in ways that improve light use efficiency [@niinemets_1998_analysis; @lloyd_2010_optimisation; @keenan_2016_global].
+To this end, we focused our structural uncertainty analysis on three processes related to competition for- and utilization efficiency of light:
 Representation of horizontal competition;

### 2019-11-15 analysis/paper/paper.Rmd
Rough draft of new intro paragraph
Got the idea down -- will wordsmith later.

@@
 
-Forests are a defining feature of the US North.
-This region is the most heavily forested region in the US, and these forests are an essential provider of employment and recreation opportunities and other ecosystem services for the region's people, and of habitat for the region's wildlife [@shividenko_2005_forest; @shifley_2012_forests].
-In addition, the forests of the Upper Midwest in particular have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
-However, the fate of these forests is highly uncertain.
-Many Upper Midwest forests are undergoing a shift in community composition from early-successional to mid-successional species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
+Human land use patterns over the last 500 years, and particularly over the last century, have led to decline in primary forests across most of the Earth's surface [@hurtt_2011_harmonization].
+However, in many places, subsequent changes in land use (e.g. abandonment of agriculture and pasture) have caused much of this area to regrow as secondary forest [@hurtt_2011_harmonization].
+Understanding the drivers behind this regrowth, and the ability to predict its future course, are therefore essential to shaping global policy around climate, biodiversity, and natural resource management.
+
+The temperate forests of the Upper Midwest are a model system for studying secondary forest ecology and biogeochemistry.
+These have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
+More recently, many Upper Midwest forests are undergoing a shift in community composition from early-successional to mid-successional species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
 In addition, these forests face a multitude of direct and indirect pressures from both biotic and abiotic sources [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.

### 2019-11-18 analysis/paper/paper.Rmd
Major revisions to introduction
Significantly shorten uncertainty section. Minor cleanup in a few places.

@@
 The temperate forests of the Upper Midwest are a model system for studying secondary forest ecology and biogeochemistry.
-These have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
+These forests have been a net carbon sink for most of the past century, as fast-growing early successional species established and thrived following near-complete deforestation around the end of the 19th Century [@birdsey_2006_forest; @williams_2012_carbon].
 More recently, many Upper Midwest forests are undergoing a shift in community composition from early-successional to mid-successional species, and the ecological and biogeochemical consequences of this transition are not well-established [@gough_2016_disturbance].
-In addition, these forests face a multitude of direct and indirect pressures from both biotic and abiotic sources [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us], non-native insects and pathogens [@lovett_2016_nonnative], and more.
-<!-- TODO More here?
--->
+In addition, these forests face a multitude of direct and indirect pressures from both biotic and abiotic sources [@shifley_2016_future], including longer and more severe droughts [@gustafson_2016_implications; @swanston_2017_vulnerability; @lienard_2016_us] and non-native insects and pathogens [@lovett_2016_nonnative].
 Collectively, these processes have culminated in forests that are different from their pre-settlement climax counterparts [@stearns_2002_one; @wolter_2002_recent; @thompson_2013_four], and our ability to make predictions about such non-analog environments based on the past is limited.
@@
 More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.
-Vegetation models fall along a spectrum of complexity, which is a direct consequence of trade-offs between realism and process fidelity on one hand and tractability, computational demand, and data requirements on the other [@hawkes_2000_woody].
+Vegetation models fall along a spectrum of complexity, with trade-offs between realism and process fidelity on one hand and computational demand and data requirements on the other [@hawkes_2000_woody].
 <!-- TODO Might need a better ref here -->
-On one side are relatively simple "big leaf" models---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location consists essentially of a single large "plant" whose characteristics are the weighted average of all the vegetation at that site <!-- TODO REFS -->
-On the other are "individual-based" or "gap" models---such as LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME [@shuman_2015_forest]---which explicitly simulate individuals competing for resources at a single site [@shugart_2015_computer].
-Because these models explicitly represent interspecific differences in plant productivity, resource allocation, and stress tolerance (among others), as well as the competition that emerges out of these differences, they may be better able to represent changes in ecosystem-scale processes, especially in no-analog conditions [@purves_2008_predictive].
-Between these two extremes are models that use approximations to attempt to capture the emergent biogeochemical and ecological outcomes of interactions between individuals without explicitly simulating each individual [@purves_2008_predictive].
+On one side are "big leaf" models---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location is represented by a single average "plant". <!-- TODO REFS -->
+On the other are "individual-based" (or "gap") models---such as LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME [@shuman_2015_forest]---that explicitly simulate competition between multiple individuals [@shugart_2015_computer].
+In the middle are models that approximate the emergent outcomes of competitive interactions among different individuals without explicitly simulating each individual [@purves_2008_predictive].
 One example of this approach is the Ecosystem Demography model [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
-<!-- BBL: Perhaps mention ED2's influence in ESMs and LSMs -->
 
 Regardless of vegetation model complexity, their projections are inherently uncertain.
-This uncertainty comes from many different sources, which can broadly be classified as follows:
-_Driver_ uncertainty is uncertainty in data about processes not represented by the model (e.g. weather and climate for vegetation models).
-_Initial condition_ uncertainty arises from the fact that the exact conditions at the place and time simulations begin are frequently unknown.
-_Process_ or _structural_ uncertainty arises because models necessarily represent only a subset of all processes involved in plant biology, and that those processes that are included can usually be represented in multiple different ways (i.e.
-which processes are included, and how they are represented).
-Finally, _parameter_ uncertainty arises because of natural variability and imperfect calibration of the above process representations.
-The relative importance of these sources of uncertainty in ecological forecasts is important as a driver of future research priorities.
-
-Many ecosystem modeling studies have looked at these categories of uncertainty independently.
-<!-- TODO Consider cutting 50-75% of these next few sentences. Probably way too much detail. -->
-@dietze_2014_quantitative investigated which parameters were most important to ED2 predictions of C sequestration across a diverse range of sites in North America.
-They found that the parameters contributing the most to overall predictive uncertainty were those related to growth respiration, mortality, stomatal conductance, and water uptake, and that parameter uncertainty varied with biome.
-More recently, @raczka_2018_what evaluated parameter uncertainty over 100 year time scale at the Willow Creek Ameriflux site, and found that the most important parameters to model uncertainty were quantum efficiency of photosynthesis, leaf respiration, and soil-plant water transfer.
-Another study looked specifically at contributions to uncertainty from parameters related to canopy radiative transfer, and found that parameters related to both canopy structure and leaf optical properties had a large impact on predictions of net primary productivity [@viskari_2019_influence].
-<!-- {Driver uncertainty -- a few examples?} -->
-
-Other studies have investigated structural uncertainty..
-Numerous vegetation model intercomparison studies have demonstrated that different models produce significantly different projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization [@zaehle_2014_evaluation; @medlyn_2015_using; @walker_2015_predicting], and soil C sequestration [@sulman_2018_multiple], among others.
-In particular, @lovenduski_2017_reducing found that model structural uncertainty was the primary source of uncertainty in Earth system model projections of the land C sink.
+This uncertainty comes from many different sources, of which we consider two in this study:
+_Structural_ uncertainty refers to which process are included and how they are represented.
+Model intercomparison studies have shown the importance of different model structures to projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization [@zaehle_2014_evaluation; @medlyn_2015_using; @walker_2015_predicting], and soil C sequestration [@sulman_2018_multiple], and
+@lovenduski_2017_reducing found that model structural uncertainty was the primary source of uncertainty in Earth system model projections of the land C sink.
 This structural uncertainty has been attributed to differences in model representations of key processes, including canopy radiative transfer [@fisher_2017_vegetation], soil biogeochemistry [@wieder_2017_carbon; @sulman_2018_multiple], stomatal conductance [@franks_2018_comparing], and photosynthesis [@rogers_2016_roadmap].
 However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.
-Moreover, comparative analysis of contributions of different types of uncertainty to model projections are rare in the ecosystem modeling literature.
+Meanwhile, _parameter_ uncertainty arises because of natural variability and imperfect calibration in process representations.
+Previous analyses of the Ecosystem Demography model in particular have highlighted the importance of parameter uncertainty (and reduction thereof) to predictions of primary productivity and composition [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
+While many modeling studies have looked at structural and parameter uncertainty independently, comparative analysis of contributions of both uncertainty sources are rare in the ecosystem modeling literature.
+
 <!--
-TODO Do we have any examples? Or are these studies entirely non-existent?
-
 Ben's notes:which of the last two categories--structural versus parameter uncertainty--is more important (Fisher et al.

### 2019-11-18 analysis/paper/paper.Rmd
Clean up and shorten methods text

@@
 
-Resource---particularly light---availability and distribution drives demographic variation in plant C fixation and community structural changes through succession.
+Light availability and distribution drives demographic variation in plant C fixation and community structural changes through succession.
 Observations show that competition for light drives niche differentiation [@silvertown_2004_plant] and consequently successional change [@canham_1990_light; @bazzaz_1993_successional], particularly in closed canopies.
@@
 Compared to the default two-stream model, the multiple-scatter model slightly increases diffuse light levels in the understory.
-Both of these models are described in detail in the Supplementary Information. <!-- TODO Exact reference to Appendix -->
 
@@
 
-By default, ED-2.2 supports 17 different plant functional types, which divide plant species according to photosynthetic pathway (C3 vs.
-C4), growth form (grass vs.
-tree), leaf phenology habit (deciduous vs.
+By default, ED-2.2 supports 17 different plant functional types, which divide plant species according to growth form (grass vs.
+tree), photosynthetic pathway (C3 vs.
+C4), leaf habit (deciduous vs.
 evergreen), biome (e.g.
@@
 However, we limited our simulations to the four plant functional types that have any appreciable presence at UMBS: Early, mid, and late temperate deciduous trees and pines.
-The species comprising these plant functional types are shown in Table 3.
+The species comprising these plant functional types are shown in Table \@ref(tab:pft-species).
 
@@
 Each ensemble member ran from January 1, 1902 through December 31, 1999 and was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
-Driving meteorological data was 6-hourly CRU-NCEP <!-- REF --> combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
+Driving meteorological data were 6-hourly CRU-NCEP [@cruncep_pecan] combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
 Soil texture was set to 92% sand, 7% silt, and 1% clay, based on site-level observations [@gough_2010_wood].
 The initial soil moisture profile was set to the average soil moisture profile reported in the UMBS Ameriflux ancillary data (https://ameriflux.lbl.gov/sites/siteinfo/US-UMd).
-By using the same initial conditions and drivers, we do not consider initial condition and driver uncertainty, which are outside the scope of this study.
 
@@
 
-First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values (Figure \@ref(fig:structure-compare-default)).
-To understand the physical and ecological mechanisms underlying these differences, we also compared the cohort-level composition, height, and canopy radiation profiles in these simulations in four stages of the simulation (Figure S1).
-
-Second, we compared simulations with ED-2.2 default parameters against simulations using median parameter estimates from a trait meta-analysis (Figure S2).
-Here, we looked predictions of annual mean C fluxes (gross and net primary productivity, and autotrophic and heterotrophic respiration; Figure \@ref(fig:default-median-flux)) and of summer LAI by PFT (Figure \@ref(fig:default-median-lai)).
-
-Third, we considered the full range of parameter uncertainty by looking at predictions across all parameter ensemble members.
-We begin with predictions annual mean NPP and summer mean total LAI across all parameters and structures and compare these results to contemporary (averaged over years TODO) observations of NPP and LAI at UMBS.
-We look at these variables independently as time series (Figure \@ref(fig:summary-ts-plot)) and as joint predictions in two periods (Figure \@ref(fig:pairs-time)).
-We then consider impacts on composition by looking at predicted PFT fractions of total aboveground biomass (AGB) at four different years in the simulation (Figure \@ref(fig:pft-jitterplot)) and at time series of LAI by PFT for five parameter combinations spanning the range of predicted ecological outcomes (Figure \@ref(fig:lai-pft-plot)).
-Finally, we assessed which parameters contributed the most to parameter uncertainty for each model structure.
-We did this by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what (Figure \@ref(fig:partial-variance-npp)).
-Briefly, this analysis uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance").
-For more details, see @lebauer_2013_facilitating.
+First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values.
+Second, we then compared these default parameter simulations to simulations using median parameter estimates from a trait meta-analysis.
+Third, we considered the full range of parameter uncertainty by looking at predictions of NPP, LAI, and composition across all parameter ensemble members.
+Finally, we assessed which parameters contributed the most to parameter uncertainty for each model structure by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what.
+Briefly, this analysis uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance"; see @lebauer_2013_facilitating for more details).
 

### 2019-11-18 analysis/paper/paper.Rmd
No TOC in PDF output. Update refs.

@@
       latex_engine: xelatex
+      toc: no
     bookdown::word_document2:

### 2019-11-18 analysis/paper/paper.Rmd
Revise Figure 2 -- Remove RH, fix labels & caption

@@
   htmlfig(),
-  "Comparing results of ED2 with default parameters and",
-  "posterior median parameters from trait meta-analysis",
+  "ED-2.2 predictions of mean annual",
+  "gross (GPP) and net primary productivity (NPP)",
+  "and autotrophic respiration (RA) from simulations using",
+  "ED-2.2 defualt parameter values (red) and",
+  "posterior median parameters from trait meta-analysis (blue)",
   "across different strucures."

### 2019-11-18 analysis/drake/structure-results.R

@@ -154,21 +154,35 @@ plan <- plan <- bind_plans(plan, drake_plan(
     pivot_longer(-c(casename:datetime)) %>%
     annual_mean(),
   default_median_fluxes_gg = both_long_s %>%
-    filter(grepl("mmean_(gpp|npp|plresp|rh)_py", name)) %>%
+    filter(grepl("mmean_(gpp|npp|plresp)_py", name)) %>%
     mutate(
       name = stringr::str_remove(name, "mmean_") %>%
         stringr::str_remove("_py") %>%
-        factor(c("gpp", "plresp", "npp", "rh"), c("GPP", "RA", "NPP", "RH")),
+        factor(c("gpp", "plresp", "npp"), c("GPP", "RA", "NPP")),
       # Unit conversion
       value = value * 10
     ) %>%
+    left_join(models, c("casename" = "model_id")) %>%
     ggplot() +
     aes(x = year, y = value, color = runtype) +
     geom_line() +
-    facet_grid(vars(name), vars(casename), scales = "free_y") +
+    facet_grid(
+      vars(name), vars(model),
+      scales = "free_y",
+      labeller = labeller(
+        .default = label_value,
+        model = function(labels) gsub(" ", "\n", labels)
+      ),
+      switch = "y"
+    ) +
     cowplot::theme_cowplot() +
-    theme(axis.text.x = element_text(angle = 90),
-          axis.title = element_blank()),
+    labs(y = expression("Flux" ~ ("MgC" ~ "ha"^-1 ~ "year"^-1)),
+         color = "Parameters") +
+    theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
+          axis.title.x = element_blank(),
+          legend.position = "bottom",
+          strip.background = element_blank(),
+          strip.placement = "outside"),
   default_median_fluxes_png = ggsave(
     file_out("analysis/figures/default-median-fluxes.png"),
     default_median_fluxes_gg,

### 2019-11-18 analysis/paper/paper.Rmd
Polish default vs. median LAI plot & cap

--- 2019-11-18 analysis/paper/paper.Rmd
@@
   htmlfig(),
-  "Comparing results of ED2 with default parameters and",
-  "posterior median parameters from trait meta-analysis",
+  "ED-2.2 predictions of mean annual",
+  "gross (GPP) and net primary productivity (NPP)",
+  "and autotrophic respiration (RA) from simulations using",
+  "ED-2.2 default parameter values (red) and",
+  "posterior median parameters from trait meta-analysis (blue)",
   "across different strucures."
@@
   htmlfig(),
-  "Comparing results of ED2 PFT-level LAI",
-  "with default parameters",
-  "posterior median parameters from trait meta-analysis",
+  "ED-2.2 predictions of mean growing season leaf area index (LAI)",
+  "by plant functional type (PFT), using",
+  "ED-2.2 default parameter values (top) and",
+  "posterior median parameters from trait meta-analysis (bottom)",
   "across different strucures."

### 2019-11-18 analysis/drake/structure-results.R

@@ -193,7 +193,7 @@ plan <- plan <- bind_plans(plan, drake_plan(
   ))
 ))
 
-plan <- plan <- bind_plans(plan, drake_plan(
+plan <- bind_plans(plan, drake_plan(
   both_long_p = both_results %>%
     select(casename, runtype, pft_py) %>%
     unnest(pft_py) %>%
@@ -204,16 +204,26 @@ plan <- plan <- bind_plans(plan, drake_plan(
     annual_mean(),
   default_median_lai_gg = both_long_p %>%
     filter(name == "mmean_lai_py") %>%
+    left_join(models, c("casename" = "model_id")) %>%
     ggplot() +
     aes(x = year, y = value, color = pft, group = pft) +
     geom_line() +
-    facet_grid(vars(runtype), vars(casename)) +
+    facet_grid(
+      vars(runtype), vars(model),
+      labeller = labeller(
+        .default = label_value,
+        model = function(labels) gsub(" ", "\n", labels)
+      )
+    ) +
     scale_color_manual(values = pfts("color")) +
-    labs(y = "LAI") +
+    labs(y = "LAI", color = "Plant functional type") +
+    guides(color = guide_legend(override.aes = list(size = 2))) +
     cowplot::theme_cowplot() +
     theme(
       axis.title.x = element_blank(),
-      axis.text.x = element_text(angle = 90)
+      axis.text.x = element_text(angle = 90, vjust = 0.5),
+      legend.position = "bottom",
+      strip.background = element_blank()
     ),
   default_median_lai_png = ggsave(
     file_out("analysis/figures/default-median-lai.png"),

### 2019-11-18 analysis/paper/paper.Rmd
Add production efficiency to summary TS plot

@@
   "ED2 plot-level predictions of",
-  "net (NPP) primary productivity and",
+  "net (NPP) primary productivity,",
   "total leaf area index (LAI),",
+  "and production efficiency (NPP / LAI)",
   "by model configuration.",
-  "The dashed lines indicate the 10th and 90th quantiles of results.",
   "Black dot with error bars is the observed mean and min/max value."

### 2019-11-18 analysis/drake/overall-uncertainty.R

@@ -32,10 +32,12 @@ plan <- plan <- bind_plans(plan, drake_plan(
 
 ### Summary plot
 plan <- bind_plans(plan, drake_plan(
-  observations = tribble(
+  observations =
+    tribble(
     ~variable, ~low, ~mean, ~hi,
     "npp", 6, 6.5, 7,
-    "lai", 3.97 - 1.96 * 0.423, 3.97, 3.97 + 1.96 * 0.423
+    "lai", 3.97 - 1.96 * 0.423, 3.97, 3.97 + 1.96 * 0.423,
+    "prod_eff", 6 / (3.97 - 1.96 * 0.423), 6.5 / 3.97, 7 / (3.97 + 1.96 * 0.423)
   ) %>%
     expand_grid(model = models$model),
   ts_both = bind_rows(scalar_data, pft_totals) %>%
@@ -45,31 +47,23 @@ plan <- bind_plans(plan, drake_plan(
         stringr::str_remove("_py$")
     ) %>%
     filter(variable %in% c("npp", "lai")),
-  ts_summary = ts_both %>%
+  ts_both2 = ts_both %>%
+    pivot_wider(names_from = "variable", values_from = "value") %>%
+    mutate(prod_eff = npp / lai) %>%
+    pivot_longer(c(npp, lai, prod_eff), names_to = "variable", values_to = "value") %>%
+    filter(!(variable == "prod_eff" & (abs(value) > 40 | value < -5))),
+  ts_summary = ts_both2 %>%
     group_by(model_id, year, variable) %>%
     summarize(
-      hi = quantile(value, 0.9),
-      lo = quantile(value, 0.1)
+      hi = quantile(value, 0.9, na.rm = TRUE),
+      lo = quantile(value, 0.1, na.rm = TRUE)
     ) %>%
     left_join(models, "model_id"),
-  summary_ts_plot_gg = ts_both %>%
+  summary_ts_plot_gg = ts_both2 %>%
     left_join(models, "model_id") %>%
     ggplot() +
     aes(x = year, color = color) +
     geom_line(aes(y = value, group = case), alpha = 0.1, size = 0.3) +
-    ## # ED2 default results.
-    ## geom_line(
-    ##   aes(y = value),
-    ##   data = rename(structure_default_data, variable = name),
-    ##   linetype = "solid", color = "black"
-    ## ) +
-    geom_ribbon(
-      aes(ymin = lo, ymax = hi),
-      data = ts_summary,
-      color = "black",
[…]

### 2019-11-19 analysis/paper/paper.Rmd
Add diversity to summary TS plot
- Add PFT AGB fracs. plot to supplement
- Revise text

@@
   "ED2 plot-level predictions of",
-  "net (NPP) primary productivity,",
+  "net primary productivity (NPP),",
   "total leaf area index (LAI),",
-  "and production efficiency (NPP / LAI)",
+  "production efficiency (NPP / LAI)",
+  "and effective number of PFTs (N~PFT,eff~)",
   "by model configuration.",
+  "Effective number of PFTs was calculated as the inverse",
+  "of the Simpson diversity index with proportions expressed as",
+  "fraction of aboveground biomass.",
   "Black dot with error bars is the observed mean and min/max value."
@@
 
-Net primary productivity (NPP) and total leaf area index (LAI) varied with both parameter value and model structure (Figure \@ref(fig:summary-ts-plot)).
+Net primary productivity (NPP), total leaf area index (LAI), their ratio, and PFT diversity varied with both parameter value and model structure (Figure \@ref(fig:summary-ts-plot)).
 All model configurations exhibited a similar mean trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
@@
 Predictions of peak NPP (between 1910 and 1950 for most simulations) were closer to the observed present-day NPP (though also with large ensemble spread), but predictions of LAI during this period were still too low.
-
-```{r pairs-time, fig.cap = pairs_time_cap}
-### Figure 7
-pairs_time_cap <- paste(
-  htmlfig(),
-  "ED-2.2 predictions of annual mean NPP and growing-season LAI",
-  "averaged over 1925-1950 (left) and 1975-1999 (right)",
-  "by model type.",
-  "Each point indicates a single ensemble run with a different set of parameters.",
-  "Solid line is the best-fit generalized additive model (GAM).",
-  "Dashed lines indicate the error bounds of present-day observed NPP and LAI.",
-  "Text labels match the rows in Figure 9 below."
-)
-fig("npp-lai-pairs.png")
-```
-
-The emergent relationship between NPP and LAI was similar across model configurations early in the runs, but became more sensitive to model configuration later into the runs (Figure \@ref(fig:pairs-time)).
-Compared to runs with a closed canopy, runs with a finite canopy usually had higher LAI for the same NPP, especially later into the simulation.
+The closed canopy configuration usually had a higher NPP for the same LAI than the finite canopy configuration, especially later into the simulation.
 Although many simulations matched _either_ the observed NPP or the observed LAI, only a small minority of individual runs matched _both_.
 
-```{r pft-jitterplot, fig.cap = pft_jitterplot_cap}
-### Figure 8
-pft_jitterplot_cap <- paste(
-  htmlfig(),
-  "Fraction of total aboveground biomass (AGB)",
-  "of each plant functional type (PFT)",
-  "predicted by different parameter-configuration combinations",
-  "of ED2.",
-  "Each point represents the AGB fraction of a single PFT",
-  "in a single ensemble member at the given point in time."
-)
-fig("pft-jitterplot.png")
-```
-
-The large parameter uncertainty in NPP and LAI also extended to species composition (Figure \@ref(fig:pft-jitterplot)).
+The large parameter uncertainty in NPP and LAI also extended to species composition (Figure \@ref(fig:summary-ts-plot)).
 Under all configurations, at any moment in time, most ensemble members were dominated by a single PFT, and the sustained coexistence of more than two PFTs was limited to a small set of specific ensemble members across all configurations.
-However, _which_ PFT was dominant depended strongly on parameter values, and any of the four PFTs became the dominant one in some subset of simulations.
+However, _which_ PFT was dominant depended strongly on parameter values, and any of the four PFTs became the dominant one in some subset of simulations (Figure S3).
 

### 2019-11-19 analysis/drake/overall-uncertainty.R

@@ -51,7 +51,8 @@ plan <- bind_plans(plan, drake_plan(
     pivot_wider(names_from = "variable", values_from = "value") %>%
     mutate(prod_eff = npp / lai) %>%
     pivot_longer(c(npp, lai, prod_eff), names_to = "variable", values_to = "value") %>%
-    filter(!(variable == "prod_eff" & (abs(value) > 40 | value < -5))),
+    filter(!(variable == "prod_eff" & (abs(value) > 40 | value < -5))) %>%
+    bind_rows(pft_n_effective %>% mutate(model_id = substr(case, 4, 6))),
   ts_summary = ts_both2 %>%
     group_by(model_id, year, variable) %>%
     summarize(
@@ -59,14 +60,32 @@ plan <- bind_plans(plan, drake_plan(
       lo = quantile(value, 0.1, na.rm = TRUE)
     ) %>%
     left_join(models, "model_id"),
+  obs_plot = observations %>%
+    mutate(
+      variable = factor(variable, c("npp", "lai", "prod_eff", "d2"), c(
+        "atop(NPP, (MgC ~ ha^{-1} ~ year^{-1}))",
+        "LAI",
+        "'NPP / LAI'",
+        "N['PFT,eff']"
+      ))
+    ),
   summary_ts_plot_gg = ts_both2 %>%
     left_join(models, "model_id") %>%
+    filter(variable %in% c("npp", "lai", "prod_eff", "d2")) %>%
+    mutate(
+      variable = factor(variable, c("npp", "lai", "prod_eff", "d2"), c(
+        "atop(NPP, (MgC ~ ha^{-1} ~ year^{-1}))",
+        "LAI",
+        "'NPP / LAI'",
+        "N['PFT,eff']"
+      ))
+    ) %>%
     ggplot() +
     aes(x = year, color = color) +
     geom_line(aes(y = value, group = case), alpha = 0.1, size = 0.3) +
     geom_pointrange(
       aes(x = 2000, y = mean, ymin = low, ymax = hi),
-      data = observations,
+      data = obs_plot,
       color = "black",
       inherit.aes = FALSE
     ) +
@@ -75,19 +94,13 @@ plan <- bind_plans(plan, drake_plan(
       vars(fct_relabel(model, ~gsub(" ", "\n", .))),
       scales = "free_y",
       switch = "y",
-      labeller = labeller(
-        variable = as_labeller(c(
[…]

### 2019-11-19 analysis/drake/pft-composition-plots.R

@@ -63,4 +63,50 @@ plan <- bind_plans(plan, drake_plan(
   ))
 ))
 
+### Effective number of PFTs
+plan <- bind_plans(plan, drake_plan(
+  pft_n_effective = pft_data %>%
+    group_by(case, year) %>%
+    summarize(
+      simpson = sum(agb_frac ^ 2),
+      d2 = 1 / simpson
+    ) %>%
+    ungroup() %>%
+    pivot_longer(
+      c(simpson, d2),
+      names_to = "variable",
+      values_to = "value"
+    )
+))
+
+### PFT AGB fraction time series
+plan <- bind_plans(plan, drake_plan(
+  pft_ts_plot = pft_data %>%
+    mutate(model_id = substr(case, 4, 6)) %>%
+    left_join(models, "model_id") %>%
+    ggplot() +
+    aes(x = year, y = agb_frac, group = case, color = color) +
+    geom_line(alpha = 0.1, size = 0.3) +
+    facet_grid(
+      vars(pft), vars(model),
+      labeller = labeller(
+        model = function(labels) gsub(" ", "\n", labels),
+        pft = label_value
+      )
+    ) +
+    ylab("AGB fraction") +
+    scale_color_identity() +
+    theme_cowplot() +
+    theme(
+      axis.text.x = element_text(angle = 90, vjust = 0.5),
+      axis.title.x = element_blank(),
+      strip.background = element_blank()
+    ),
+  pft_ts_plot_png = ggsave(
+    file_out("analysis/figures/pft-ts-plot.png"),
+    pft_ts_plot,
+    width = 10, height = 6
+  )
+))
+
 stop()

### 2019-11-19 analysis/paper/paper.Rmd
Add parameter combinations to summary TS plot

@@
   "fraction of aboveground biomass.",
+  "Each black line represents the same set of parameters across all configurations,",
+  "and corresponds to the letters in Figure 5.",
   "Black dot with error bars is the observed mean and min/max value."

### 2019-11-19 analysis/drake/overall-uncertainty.R

@@ -69,7 +69,7 @@ plan <- bind_plans(plan, drake_plan(
         "N['PFT,eff']"
       ))
     ),
-  summary_ts_plot_gg = ts_both2 %>%
+  summary_ts_data = ts_both2 %>%
     left_join(models, "model_id") %>%
     filter(variable %in% c("npp", "lai", "prod_eff", "d2")) %>%
     mutate(
@@ -79,10 +79,15 @@ plan <- bind_plans(plan, drake_plan(
         "'NPP / LAI'",
         "N['PFT,eff']"
       ))
-    ) %>%
-    ggplot() +
+    ),
+  summary_ts_plot_gg = ggplot(summary_ts_data) +
     aes(x = year, color = color) +
     geom_line(aes(y = value, group = case), alpha = 0.1, size = 0.3) +
+    geom_line(
+      aes(y = value, group = case, linetype = label),
+      color = "black",
+      data = ts_params
+    ) +
     geom_pointrange(
       aes(x = 2000, y = mean, ymin = low, ymax = hi),
       data = obs_plot,
@@ -97,6 +102,14 @@ plan <- bind_plans(plan, drake_plan(
       labeller = label_parsed
     ) +
     scale_color_identity() +
+    scale_linetype_manual(values = c(
+      A = "solid",
+      B = "longdash",
+      C = "dotted",
+      D = "twodash",
+      E = "dotdash"
+    )) +
+    labs(linetype = "Param. set") +
     theme_cowplot() +
     theme(
       axis.title = element_blank(),
@@ -114,8 +127,16 @@ plan <- bind_plans(plan, drake_plan(
   ))
 ))
 
+plan <- bind_plans(plan, drake_plan(
+  ts_params = summary_ts_data %>%
+    mutate(param_id = as.numeric(substr(case, 0, 3))) %>%
+    inner_join(use_params, "param_id") %>%
+    mutate(label = fct_rev(label))
+))
+
 ### STOP HERE
 stop()
+
 ### Download files
 cohort_osf <- "..."
 pft_osf <- "..."

### 2019-11-19 analysis/paper/paper.Rmd
Tweak captions
Make sure ED-2.2 is everywhere.

--- 2019-11-19 analysis/paper/paper.Rmd
@@
   htmlfig(),
-  "ED-2.2 predictions of mean growing season leaf area index (LAI)",
+  "ED-2.2 plot-level predictions of",
+  "mean growing season leaf area index (LAI)",
   "by plant functional type (PFT), using",
-  "ED-2.2 default parameter values (top) and",
+  "default parameter values (top) and",
   "posterior median parameters from trait meta-analysis (bottom)",
@@
   htmlfig(),
-  "ED2 plot-level predictions of",
+  "ED-2.2 plot-level predictions of",
   "net primary productivity (NPP),",
@@
   htmlfig(),
-  "PFT-level predictions of leaf area index (LAI) by model configuration and parameterization.",
+  "ED-2.2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization.",
   "Each row of plots is a simulation with the same set of parameters across model configurations.",
@@
   "Contributions of individual parameters to predicted uncertainty in",
-  "average net primary productivity (NPP) over the period 1920-1950",
+  "ED-2.2 predictions of average net primary productivity (NPP)",
+  "over the period 1920-1950",
   "by model type.",
@@
 
-Most ED-2.2 runs under-predicted both NPP and LAI, both in the peak period (1920-1950) and towards the end of the simulation (1975-1999) (Figures \@ref(fig:summary-ts-plot), \@ref(fig:pairs-time)).
-However, of our five selected representative runs, the two that most closely matched present day composition at UMBS (early hardwood-dominated; **C** and **D** in Figure \@ref(fig:lai-pft-plot)) also closely matched the observed NPP and LAI (Figure \@ref(fig:pairs-time)).
+Most ED-2.2 runs under-predicted both NPP and LAI, both in the peak period (1920-1950) and towards the end of the simulation (1975-1999) (Figures \@ref(fig:summary-ts-plot)).
+However, of our five selected representative runs, the two that most closely matched present day composition at UMBS (early hardwood-dominated; **C** and **D** in Figure \@ref(fig:lai-pft-plot)) also closely matched the observed NPP and LAI (Figure \@ref(fig:summary-ts-plot)).
 This is encouraging because it suggests that it is possible to get both aggregate ecosystem C fluxes and PFT composition right.
@@
 We found that parameter uncertainty was by far the dominant driver of uncertainty in our simulations.
-Different parameter combinations led to huge ranges in ensemble predictions of ensemble variables, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures \@ref(fig:summary-ts-plot), \@ref(fig:pairs-time)).
-Moreover, different parameter combinations produced qualitatively different ecological narratives (Figures \@ref(fig:pft-jitterplot). \@ref(fig:lai-pft-plot)).
+Different parameter combinations led to huge ranges in ensemble predictions of ensemble variables, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures \@ref(fig:summary-ts-plot)).
+Moreover, different parameter combinations produced qualitatively different ecological narratives (Figures \@ref(fig:lai-pft-plot)).
 Therefore, parameter uncertainty has implications beyond C cycling---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.

### 2019-11-19 analysis/paper/paper.Rmd
Change model aesthetics, esp. in summary plot
Use color for labels, gray shading for CI, and don't show individual
ensemble members.

@@
   htmlfig(),
-  "ED-2.2 plot-level predictions of",
+  "ED-2.2 ensemble plot-level predictions of",
   "net primary productivity (NPP),",
@@
   "by model configuration.",
+  "Shading indicates the middle 50\\% (dark), middle 95\\% (medium),",
+  "and entire range (light) of ensemble simulations",
+  "(NPP/LAI results are clipped to (-5, 40) for readability,",
+  "as a few isolated ensemble members produced extreme values at specific",
+  "time points).",
   "Effective number of PFTs was calculated as the inverse",
@@
   "fraction of aboveground biomass.",
-  "Each black line represents the same set of parameters across all configurations,",
+  "Each colored line represents the same set of parameters across all configurations,",
   "and corresponds to the letters in Figure 5.",

### 2019-11-19 analysis/drake/overall-uncertainty.R

@@ -56,66 +56,55 @@ plan <- bind_plans(plan, drake_plan(
   ts_summary = ts_both2 %>%
     group_by(model_id, year, variable) %>%
     summarize(
-      hi = quantile(value, 0.9, na.rm = TRUE),
-      lo = quantile(value, 0.1, na.rm = TRUE)
+      max = max(value, na.rm = TRUE),
+      hi2 = quantile(value, 0.95, na.rm = TRUE),
+      hi1 = quantile(value, 0.75, na.rm = TRUE),
+      lo1 = quantile(value, 0.25, na.rm = TRUE),
+      lo2 = quantile(value, 0.05, na.rm = TRUE),
+      min = min(value, na.rm = TRUE)
     ) %>%
-    left_join(models, "model_id"),
-  obs_plot = observations %>%
-    mutate(
-      variable = factor(variable, c("npp", "lai", "prod_eff", "d2"), c(
-        "atop(NPP, (MgC ~ ha^{-1} ~ year^{-1}))",
-        "LAI",
-        "'NPP / LAI'",
-        "N['PFT,eff']"
-      ))
-    ),
+    left_join(models, "model_id") %>%
+    filter(variable %in% c("npp", "lai", "prod_eff", "d2")),
   summary_ts_data = ts_both2 %>%
     left_join(models, "model_id") %>%
-    filter(variable %in% c("npp", "lai", "prod_eff", "d2")) %>%
-    mutate(
-      variable = factor(variable, c("npp", "lai", "prod_eff", "d2"), c(
-        "atop(NPP, (MgC ~ ha^{-1} ~ year^{-1}))",
-        "LAI",
-        "'NPP / LAI'",
-        "N['PFT,eff']"
-      ))
-    ),
-  summary_ts_plot_gg = ggplot(summary_ts_data) +
-    aes(x = year, color = color) +
-    geom_line(aes(y = value, group = case), alpha = 0.1, size = 0.3) +
+    filter(variable %in% c("npp", "lai", "prod_eff", "d2")),
+  summary_ts_plot_gg = ggplot(ts_summary) +
+    aes(x = year) +
+    geom_ribbon(aes(ymin = min, ymax = max), fill = "gray90") +
+    geom_ribbon(aes(ymin = lo2, ymax = hi2), fill = "gray70") +
+    geom_ribbon(aes(ymin = lo1, ymax = hi1), fill = "gray40") +
     geom_line(
-      aes(y = value, group = case, linetype = label),
[…]

### 2019-11-19 analysis/drake/structure-results.R

@@ -63,28 +63,31 @@ plan <- bind_plans(plan, drake_plan(
     mutate(npp = npp * 10) %>%
     pivot_longer(c(npp, lai)),
   structure_compare_default_gg = ggplot(structure_default_data) +
-    aes(x = year, y = value, color = model) +
+    aes(x = year, y = value, linetype = crown, color = traits, group = model) +
     geom_line() +
-    guides(color = guide_legend(title = "Model",
-                                override.aes = list(size = 2))) +
+    labs(linetype = "Crown model", color = "Trait plasticity") +
     facet_grid(
-      vars(name), vars(rtm),
+      vars(name = factor(name, c("npp", "lai"))), vars(rtm),
       scales = "free_y",
       switch = "y",
-      labeller = labeller(name = as_labeller(c(
-        "npp" = "NPP ~ (MgC ~ ha^-1 ~ yr^-1)",
-        "lai" = "LAI"
-      ), default = label_parsed))
+      labeller = labeller(
+        name = as_labeller(c(
+          "npp" = "atop(NPP, (MgC ~ ha^-1 ~ yr^-1))",
+          "lai" = "LAI"
+        ), default = label_parsed),
+        rtm = function(labels) paste0("Canopy RTM: ", labels)
+      )
     ) +
-    scale_color_manual(values = model_colors) +
+    scale_color_manual(values = c("#F5793A", "#0F2080")) +
     cowplot::theme_cowplot() +
     theme(axis.title = element_blank(),
+          axis.text.x = element_text(angle = 90, vjust = 0.5),
           strip.background = element_blank(),
           strip.placement = "outside"),
   structure_compare_default_png = ggsave(
     file_out("analysis/figures/structure-compare-default.png"),
     structure_compare_default_gg,
-    width = 10.1, height = 7
+    width = 6.5, height = 4.2
   ),
   structure_compare_default_knit = knitr::include_graphics(file_in(
     "analysis/figures/structure-compare-default.png"

### 2019-11-19 analysis/paper/paper.Rmd
Revise introduction. Semifinal version.

@@
 
+<!-- TODO First two intro paragraphs are weak -->
 Human land use patterns over the last 500 years, and particularly over the last century, have led to decline in primary forests across most of the Earth's surface [@hurtt_2011_harmonization].
@@
 More reliable predictions can in theory be obtained by using dynamic vegetation models that explicitly represent processes involved in forest growth and mortality.
-Vegetation models fall along a spectrum of complexity, with trade-offs between realism and process fidelity on one hand and computational demand and data requirements on the other [@hawkes_2000_woody].
-<!-- TODO Might need a better ref here -->
-On one side are "big leaf" models---such as PNET, SiBCASA, and Biome-BGC---in which the vegetation at a particular location is represented by a single average "plant". <!-- TODO REFS -->
+Vegetation models fall along a spectrum of complexity, with trade-offs between realism and process fidelity on one hand and computational demand and data requirements on the other.
+On one side are "big leaf" models---such as PNET [@aber_1992_pnet], SiBCASA [@schaefer_2008_sibcasa], and Biome-BGC [@running_1993_generalization]---in which the vegetation at a particular location is represented by a single average "plant".
 On the other are "individual-based" (or "gap") models---such as LANDIS [@mladenoff_2004_landis; @scheller_2007_design] and UVAFME [@shuman_2015_forest]---that explicitly simulate competition between multiple individuals [@shugart_2015_computer].
 In the middle are models that approximate the emergent outcomes of competitive interactions among different individuals without explicitly simulating each individual [@purves_2008_predictive].
-One example of this approach is the Ecosystem Demography model [ED2; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
+One example of this approach is the Ecosystem Demography model [ED; @moorcroft_2001_method; @medvigy_2011_predicting; @longo_2019_ed1], which is the focus of this study.
 
@@
 _Structural_ uncertainty refers to which process are included and how they are represented.
-Model intercomparison studies have shown the importance of different model structures to projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties], response to CO~2~ fertilization [@zaehle_2014_evaluation; @medlyn_2015_using; @walker_2015_predicting], and soil C sequestration [@sulman_2018_multiple], and
-@lovenduski_2017_reducing found that model structural uncertainty was the primary source of uncertainty in Earth system model projections of the land C sink.
+Model intercomparison studies have shown the importance of different model structures to projections of overall land carbon sequestration [@friedlingstein_2006_climate; @friedlingstein_2014_uncertainties; @lovenduski_2017_reducing], response to CO~2~ fertilization [@zaehle_2014_evaluation; @medlyn_2015_using; @walker_2015_predicting], and soil C sequestration [@sulman_2018_multiple].
 This structural uncertainty has been attributed to differences in model representations of key processes, including canopy radiative transfer [@fisher_2017_vegetation], soil biogeochemistry [@wieder_2017_carbon; @sulman_2018_multiple], stomatal conductance [@franks_2018_comparing], and photosynthesis [@rogers_2016_roadmap].
-However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because different models are different from each other in too many different ways, meaning that structural effects are confounded with other uncertainty sources.
-Meanwhile, _parameter_ uncertainty arises because of natural variability and imperfect calibration in process representations.
-Previous analyses of the Ecosystem Demography model in particular have highlighted the importance of parameter uncertainty (and reduction thereof) to predictions of primary productivity and composition [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
+However, the extent to which specific processes contribute to model uncertainty is difficult to evaluate from model intercomparisons because models differ from each other in too many ways, meaning that structural effects are confounded with other uncertainty sources.
+At the same time, _parameter_ uncertainty arises because of natural variability and imperfect calibration in process representations.
+Previous analyses of the ED model in particular have highlighted the importance of parameter uncertainty (and reduction thereof) to predictions of primary productivity and composition [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence].
 While many modeling studies have looked at structural and parameter uncertainty independently, comparative analysis of contributions of both uncertainty sources are rare in the ecosystem modeling literature.
 
-<!--
-Ben's notes:which of the last two categories--structural versus parameter uncertainty--is more important (Fisher et al.
-2015, https://onlinelibrary.wiley.com/doi/full/10.1111/j.1523-1739.2003.00614.x, Ricciuto et al.
-2008).
--->
-
+<!-- TODO This paragraph still needs work  -->
 Light availability and distribution drives demographic variation in plant C fixation and community structural changes through succession.
@@
 (2) What are the relative contributions of parameter uncertainty (data limitation) vs. structural uncertainty (theoretical limitation)?
-We ran ensemble simulations of the Ecosystem Demography (ED-2.2) model with a factorial combination of submodels related to radiative transfer formulation, horizontal competition, and trait plasticity.
-We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results in the context of ongoing climate change.
+We ran ensemble simulations of the Ecosystem Demography (ED-2.2) model with a factorial combination of submodels related to canopy radiative transfer formulation, horizontal competition, and trait plasticity.
+We then used sensitivity and variance decomposition analyses to evaluate the contribution of parameter uncertainty for each model configuration, exploring the ecological implications of the results in the context of ongoing global change.
 

### 2019-11-19 analysis/paper/paper.Rmd
Revise methods. Semifinal version.

@@
 The landscape geography of the UMBS upland forest is 20.4% moraine, 37.8% high outwash plain, 31.3% low outwash plain, 5.7% lake-margin terrace, 3.6% ice-contact, and 1.2% lowland glacial lake [@bergen_2007_observing].
-Most of the UMBS upland forest canopy is dominated by temperate deciduous early-successional species, most importantly _Populus grandidentata_ (bigtooth aspen) and, to a lesser extent, _Betula papyrifera_ (paper birch), with _Acer saccharum_ (sugar maple), _Acer rubrum_ (red maple), _Fagus grandifolia_ (American beech), _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_ (eastern hemlock), _Quercus rubra_ (northern red oak), _Pinus strobus_ (white pine), and _Pinus resinosa_ (red pine) existing in various fractions in the understory (and, in patches, in the canopy).
+The UMBS upland forest canopy is dominated by temperate deciduous early-successional species, most importantly _Populus grandidentata_ (bigtooth aspen) and, to a lesser extent, _Betula papyrifera_ (paper birch), with _Acer saccharum_ (sugar maple), _Acer rubrum_ (red maple), _Fagus grandifolia_ (American beech), _Tilia americana_ (basswood), _Betula alleghaniensis_ (yellow birch), _Fraxinus americana_ (white ash), _Tsuga canadensis_ (eastern hemlock), _Quercus rubra_ (northern red oak), _Pinus strobus_ (white pine), and _Pinus resinosa_ (red pine) existing in various fractions in the understory (and, in patches, in the canopy).
 This composition is a legacy of the site's disturbance history:
@@
 
-The official ED2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained at https://github.com/ashiklom/ED2/tree/b048950971e91699b78fc566df133caf977fda32.
+The official ED-2.2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained at https://github.com/ashiklom/ED2/tree/b048950971e91699b78fc566df133caf977fda32.
 
@@
 plastic (varying with light level) traits.
+These sub-models are summarized in Table \@ref(tab:submodels).
 
@@
 The models differ in the terms of this matrix equation.
-The two-stream model (default) uses the two-stream approximation originally derived for atmospheric radiative transfer [@meador_1980_two] and later adapted to vegetation canopies [@dickinson_1983_land; @sellers_1985_canopy].
-Meanwhile, the multiple-scatter model was derived from first principles by @zhao_2005_multiple specifically for vegetation canopies to address known biases and limitations of the two-stream approach [e.g.
+The default model (_two-stream_) uses a special multi-canopy solution [derived in @longo_2019_ed1] of the two-stream approximation for vegetation canopies [@dickinson_1983_land; @sellers_1985_canopy].
+Meanwhile, the _multiple-scatter_ model was derived from first principles by @zhao_2005_multiple specifically for vegetation canopies to address known biases and limitations of the two-stream approach [e.g.
 @wang_2003_comparison].
@@
 The crown area submodel in ED-2.2 determines the nature of competition for light between cohorts.
-In the default configuration ("closed canopy"), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
+In the default configuration (_closed_), the leaf area of a cohort is distributed across the entire horizontal area of a patch.
 This means that taller cohorts always receive more incoming radiation---including 100% of the direct radiation---than shorter cohorts, even when the height difference is small.
 Similar approaches have been shown to excessively suppress competition from sub-dominant individuals and result in unrealistically homogeneous forest patches [@fisher_2015_taking].
-In the alternate configuration ("finite crown area"), canopies take up only a fraction of the available horizontal area, which allows more light (including some direct radiation) to reach subdominant cohorts.
+In the alternate configuration (_finite_), canopies take up only a fraction of the available horizontal area, which allows more light (including some direct radiation) to reach subdominant cohorts.
 The horizontal area of crowns is determined by allometric equations from @dietze_2008_changing.
@@
 The third submodel we evaluated was trait plasticity.
-In the default configuration, all cohorts of a given plant functional type have the same parameters, regardless of environmental conditions.
-This ignores the globally-documented intraspecific trait variability as a function of light level [@niinemets_2010_review; @keenan_2016_global].
-In the alternate configuration, as light level decreases (trees become more shaded), specific leaf area increases and V~cmax~ decreases, following empirical relationships from the tropics [@lloyd_2010_optimisation].
-
-These sub-models are summarized in Table \@ref(tab:submodels).
+In the default configuration (_static_), all cohorts of a given PFT have the same parameters, regardless of environmental conditions.
+This ignores globally-documented intraspecific trait variability with light level and relative canopy position [@niinemets_2010_review; @keenan_2016_global].
+In the alternate configuration (_plastic_), cohorts that receive less light increase their specific leaf area and decrease their V~cmax~, following empirical relationships from the tropics [@lloyd_2010_optimisation].
 
@@
 
-The full list of parameters in ED-2.2 is large, with over 100 parameters per plant functional type.
+The full list of parameters in ED-2.2 is large, with over 100 parameters per PFT.
 A full sensitivity and uncertainty analysis across all of these parameters is outside the scope of this study.
@@
 
-By default, ED-2.2 supports 17 different plant functional types, which divide plant species according to growth form (grass vs.
-tree), photosynthetic pathway (C3 vs.
-C4), leaf habit (deciduous vs.
-evergreen), biome (e.g.
-temperate vs.
-tropical), and successional status (e.g.
-early, mid, late).
-However, we limited our simulations to the four plant functional types that have any appreciable presence at UMBS: Early, mid, and late temperate deciduous trees and pines.
-The species comprising these plant functional types are shown in Table \@ref(tab:pft-species).
+By default, ED-2.2 supports 17 different PFTs, which divide plant species according to growth form, photosynthetic pathway, leaf habit, biome, and successional status.
+However, we limited our simulations to the four PFTs that have any appreciable presence at UMBS: Early-, mid-, and late-successional temperate deciduous ("hardwood") trees and pines.
+The species comprising these PFTs are shown in Table \@ref(tab:pft-species).
 
@@
 For each plant functional type, we generated a distribution of parameter values via the Predictive Ecosystem Analyzer (PEcAn) trait-meta analysis [@lebauer_2013_facilitating; see also @dietze_2014_quantitative; @raczka_2018_what].
-Prior distributions for this meta-analysis were largely adapted from previous ED2 parameter uncertainty studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence], and are shown in detail in Supplementary Information X.
-<!-- TODO Priors supplement; maybe describe entire meta-analysis?
--->
-Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from Shiklomanov (2019 {dissertation, chapter 3}).
+Prior distributions for this meta-analysis were largely adapted from previous ED parameter uncertainty studies [@dietze_2014_quantitative; @raczka_2018_what; @viskari_2019_influence], and are shown in detail in Supplementary Table S1.
+Species trait data for this meta-analysis came from existing records in the BETY database [www.betydb.org, @lebauer_2017_betydb], as well as from publicly available records in the TRY database [www.try-db.org, @kattge_2011_try] and, specifically for leaf optical properties, from @shiklomanov_dissertation.
 All parameters not described in this section were set to their ED-2.2 PFT-specific default values for all simulations.
@@
 For each factorial combination of model configurations (described above), we ran ED-2.2 once with its default parameter values, once at the posterior median estimates from the meta-analysis, and 500 times with parameters sampled randomly from the meta-analysis posterior distributions.
-For the latter, the same 500 parameter samples were used for all model structures to allow us to independently evaluate the impacts of model structure at specific parameter values.
-Each ensemble member ran from January 1, 1902 through December 31, 1999 and was initialized from a "near-bare ground" condition: An equal number of seedlings of each plant functional type (see previous section) at the minimum resolvable size.
+For the latter, the same parameter samples were used for all model structures to allow us to independently evaluate the impacts of model structure at specific parameter values.
+We ran each ensemble member from January 1, 1902 through December 31, 1999 and initialized them from a "near-bare ground" condition: An equal number of seedlings of each included PFT at the minimum resolvable size.
 Driving meteorological data were 6-hourly CRU-NCEP [@cruncep_pecan] combined with an annual atmospheric CO~2~ record from Law-Dome ice core [@etheridge_1998_historical] and Mauna Loa observatory [@thoning_1989_atmospheric] linearly interpolated to 3-hourly resolution.
@@
 
-First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) and summer (June, July, August) total leaf area index (LAI) across model structures while holding ED-2.2 parameters at their default values.
-Second, we then compared these default parameter simulations to simulations using median parameter estimates from a trait meta-analysis.
+First, we evaluated the impact of model structure independent of any changes to parameters by comparing predictions of average annual net primary productivity (NPP) summer (June, July, August) total leaf area index (LAI), the NPP:LAI ratio, and the effective number of PFTs across model structures while holding ED-2.2 parameters at their default values.
+We calculated the effective number of PFTs ($N_{\text{PFT,eff}}$) as the inverse of the Simpson diversity index:
+
+$$
+N_{\text{PFT,eff}} = \left( \sum_i {{f_{\text{AGB,i}}}^2} \right) ^ {-1}
+$$
+
+where $f_{\text{AGB,i}}$ is the fraction of aboveground biomass for PFT $i$ at a particular time.
+
+Second, we compared these default parameter simulations to simulations using median parameter estimates from a trait meta-analysis.
 Third, we considered the full range of parameter uncertainty by looking at predictions of NPP, LAI, and composition across all parameter ensemble members.
 Finally, we assessed which parameters contributed the most to parameter uncertainty for each model structure by performing a sensitivity and variance decomposition analysis following @lebauer_2013_facilitating, @dietze_2014_quantitative, and @raczka_2018_what.
-Briefly, this analysis uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ("partial variance"; see @lebauer_2013_facilitating for more details).
+Briefly, this analysis uses a spline to estimate the relationship between individual parameters and output variables and then uses these splines to predict the variance in the output variable attributable to individual parameters ["partial variance"; see @lebauer_2013_facilitating for more details].
 
-All analyses for this work were performed using R 3.6 (see "Colophon" for full details).
+All analyses for this work were performed using R 3.6 [@r_361].
 Code and supporting data for reproducing this analysis are publicly available on the Open Science Framework (OSF) at https://osf.io/dznuf/.

### 2019-11-19 analysis/paper/paper.Rmd
Revise first results section

@@
 Post 1950, the combination of two-stream RTM, finite canopy radius, and plastic traits had the highest productivity, followed by the same configuration with static traits, with the remaining configurations having similar results.
-Interestingly, the effects of changing individual model configurations were not additive --- for example, enabling trait plasticity had a much stronger positive effect when paired with the two-stream canopy RTM than with the multiple-scatter canopy RTM.
+The effects of changing individual model configurations were not additive---for example, enabling trait plasticity had a much stronger positive effect when paired with the two-stream canopy RTM than with the multiple-scatter canopy RTM.
 
@@
 In many cases, median parameter estimates were different entirely because of the choice of prior distributions.
-However, in several important cases, parameters well-constrained by data produced values substantially different from the ED2 defaults.
-For example, ED-2.2 default values for V~cmax~ were too low for all PFTs, and default values for SLA were too low for all PFTs except Early Hardwoods.
+However, in several important cases, parameters well-constrained by data produced values substantially different from the model defaults.
+For example, default values for V~cmax~ were too low for all PFTs, and default values for SLA were too low for all PFTs except Early Hardwoods.
 
@@
   "and autotrophic respiration (RA) from simulations using",
-  "ED-2.2 default parameter values (red) and",
+  "default parameter values (red) and",
   "posterior median parameters from trait meta-analysis (blue)",
-  "across different strucures."
+  "across different strucures.",
+  "Median and default parameter values are shown in Figure S2."
 )
@@
 ```
-
-These differences in parameter values resulted in ED-2.2 runs that differed from the default runs in several important ways (Figures \@ref(fig:default-median-flux), \@ref(fig:default-median-lai)).
-First, in every model configuration (except finite canopy, two-stream RTM, and static traits), gross primary productivity (GPP) reached a higher peak, and did so earlier under median parameters (Figure \@ref(fig:default-median-flux)).
-This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure S2).
-However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure \@ref(fig:default-median-flux)), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure S2).
-The resulting net primary productivity (NPP) had similar peak values under default and median parameters, though the temporal patterns were quite different in most model configurations (Figure \@ref(fig:default-median-flux)).
-
-Second, the variation across model structures was much higher, with every change in model structure exerting a noticeable effect on flux variability (Figure \@ref(fig:default-median-flux)).
-The most dramatic example is that, with median parameter values, simulations with the finite canopy radius produced much lower net productivity and heterotrophic respiration than simulations with a closed canopy in the second half of the simulation.
-Yet again, interactions between configurations were significant.
-For example, under a closed canopy configuration, enabling trait plasticity caused interannual variability in NPP to decrease with a two-stream RTM and increase with a multiple-scatter RTM.
 
@@
 
+These differences in parameter values resulted in ED-2.2 runs that differed from the default runs in several important ways (Figures \@ref(fig:default-median-flux), \@ref(fig:default-median-lai)).
+First, in every model configuration (except finite canopy, two-stream RTM, and static traits), gross primary productivity (GPP) reached a higher peak, and did so earlier, under median parameters (Figure \@ref(fig:default-median-flux)).
+This was driven primarily by higher V~cmax~ and SLA values for most PFTs (Figure S2).
+However, this increase in GPP was offset by a large increase in autotrophic respiration (RA; Figure \@ref(fig:default-median-flux)), driven primarily by higher growth and leaf respiration rates across all PFTs (notably, the ED-2.2 default is _zero growth respiration_ for all temperate hardwood PFTs; Figure S2).
+The resulting net primary productivity (NPP) had similar peak values under default and median parameters, though the temporal patterns were quite different in most model configurations (Figure \@ref(fig:default-median-flux)).
+
+Second, the variation across model structures was much higher, with every change in model structure exerting a noticeable effect on flux variability (Figure \@ref(fig:default-median-flux)).
+The most dramatic example is that, with median parameter values, simulations with the finite canopy radius produced much lower net productivity than simulations with a closed canopy in the second half of the simulation.
+Again, interactions between configurations were significant; for example, under a closed canopy configuration, enabling trait plasticity caused interannual variability in NPP to decrease with a two-stream RTM and increase with a multiple-scatter RTM.
+
 Third, even where NPP was similar between default and median runs, the underlying forest composition was different (Figure \@ref(fig:default-median-lai)).
-With the default parameterization, ED-2.2 produces forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually die out.
-However, with the median parameterization, ED-2.2 starts out with forests dominated by _mid_-hardwood species that are quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remains the dominant canopy species throughout the rest of the simulation.
-Exploratory analyses (results not shown) show that the main parameters responsible for this compositional trajectory are SLA (especially the higher SLA of pine and mid-hardwood), V~cmax~, and growth respiration.
+With the default parameterization, ED-2.2 produced forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually died out.
+However, with the median parameterization, ED-2.2 started out with forests dominated by _mid_-hardwood species that were quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remained the dominant canopy species throughout the rest of the simulation.
+Exploratory analyses (results not shown) showed that the main parameters responsible for this compositional trajectory were SLA (especially the higher SLA of pine and mid-hardwood), V~cmax~, and growth respiration.
 While the uncertainty in growth respiration is large, we emphasize that SLA and V~cmax~ are well constrained by data for all PFTs considered in this analysis.

### 2019-11-19 analysis/paper/paper.Rmd
Revise second half of results. Semifinal version.

@@
 However, in several important cases, parameters well-constrained by data produced values substantially different from the model defaults.
-For example, default values for V~cmax~ were too low for all PFTs, and default values for SLA were too low for all PFTs except Early Hardwoods.
+For example, default values for V~cmax~ were too low for all PFTs, and default values for SLA were too low for all PFTs except early hardwoods.
 
@@
 Third, even where NPP was similar between default and median runs, the underlying forest composition was different (Figure \@ref(fig:default-median-lai)).
-With the default parameterization, ED-2.2 produced forests dominated by Early Hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually died out.
+With the default parameterization, ED-2.2 produced forests dominated by early hardwoods, with an understory of small mid- and even smaller late-hardwood trees that eventually died out.
 However, with the median parameterization, ED-2.2 started out with forests dominated by _mid_-hardwood species that were quickly (pre-1910 to 1930, depending on configuration) replaced with pine, which remained the dominant canopy species throughout the rest of the simulation.
@@
 Net primary productivity (NPP), total leaf area index (LAI), their ratio, and PFT diversity varied with both parameter value and model structure (Figure \@ref(fig:summary-ts-plot)).
-All model configurations exhibited a similar mean trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975 (with individual ensemble members sometimes undergoing sudden collapse in specific years).
+All model configurations exhibited a similar mean trajectory of rapid growth in the first 15-20 years followed by a gradual decline to a lower, stable value around 1975.
 The ensemble spread within any single configuration---i.e.
-the parameter uncertainty---was similar and large across all configurations; both NPP and LAI varied from close to zero to more than twice the mean value at any given timestep.
-Within all configurations, most (>90%) ensemble members under-predicted observed present-day NPP and LAI by year 2000, though individual ensemble members from some configurations over-predicted one or both variables.
-Predictions of peak NPP (between 1910 and 1950 for most simulations) were closer to the observed present-day NPP (though also with large ensemble spread), but predictions of LAI during this period were still too low.
-The closed canopy configuration usually had a higher NPP for the same LAI than the finite canopy configuration, especially later into the simulation.
-Although many simulations matched _either_ the observed NPP or the observed LAI, only a small minority of individual runs matched _both_.
+parameter uncertainty---was similar and large across all configurations; both NPP and LAI varied from close to zero to more than twice the mean value at any given timestep.
+Within all configurations, most (~95%) ensemble members under-predicted observed present-day NPP and LAI by year 2000, though a small number of ensemble members from some configurations over-predicted one or both variables.
+Peak NPP (between 1910 and 1950 for most simulations) was closer to the observed present-day NPP (though also with large ensemble spread), but LAI during this period was still too low.
+The NPP to LAI ratio ("production efficiency") was higher in the closed canopy configuration than the finite canopy configuration, especially later into the simulation.
+Although many simulations matched _either_ the observed NPP or the observed LAI, only a small minority of individual runs matched _both_ simultaneously (Figure S3).
 
@@
 Under all configurations, at any moment in time, most ensemble members were dominated by a single PFT, and the sustained coexistence of more than two PFTs was limited to a small set of specific ensemble members across all configurations.
-However, _which_ PFT was dominant depended strongly on parameter values, and any of the four PFTs became the dominant one in some subset of simulations (Figure S3).
+However, _which_ PFT was dominant depended strongly on parameter values, and any of the four PFTs became the dominant one in some subset of simulations (Figure S4).
 
@@
 In all cases, mid-hardwoods began as the dominant PFT and were eventually overtaken by either pine (closed canopy and plastic traits) or early hardwood (all other cases; though with a closed canopy and static traits, the final LAI was low).
-Parameter sets **D** and **E** had the highest productivity of these five sets, and were characterized by strong dominance of Early Hardwood and Pine, respectively.
+Parameter sets **D** and **E** had the highest productivity of these five sets, and were characterized by strong dominance of early hardwood and Pine, respectively.
 
@@
 The parameters that contributed most to predictive uncertainty in NPP during the "peak" period (1920-1950) ("partial variance") varied with model structure (Figure \@ref(fig:partial-variance-npp)).
-The "water availability factor" (a.k.a.
-plant-soil water conductance) for early hardwoods was the most or second-most important parameter for all but one model configuration (the ED-2.2 default---closed canopy, two-stream RTM, static traits---where it was not even in the top 8), and was particularly important in simulations with a finite canopy.
-The growth respiration factor, particularly of pine and, to a lesser extent, early hardwood, was also among the most influential parameters across most configurations.
+The plant-soil water conductance (a.k.a.
+"water availability factor") for early hardwoods was the most or second-most important parameter for all but one model configuration (the ED-2.2 default---closed canopy, two-stream RTM, static traits---where it was not even in the top 8), and was particularly important in simulations with a finite canopy.
+The growth respiration factor, particularly of Pine and, to a lesser extent, early hardwood, was also among the most influential parameters across most configurations.
 Which parameters were most important, and how important they were, varied with model configuration.
-For instance, the C balance ratio of mortality (`mort2`) of early hardwoods was among the top four most important parameters when the canopy was closed but mattered less when the canopy was finite.
+For instance, the C balance ratio of mortality (`mort2`) of early Hardwoods was among the top four most important parameters when the canopy was closed but mattered less when the canopy was finite.
 Similarly, parameters related to leaf optical properties had relatively higher importance in simulations with the two-stream model than the multiple-scatter model.

### 2019-11-19 analysis/paper/paper.Rmd
Add CG text to discussion

@@
 
-The overarching objective of this study was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community succession.
-Our first question was: To what extent can the model, under any configuration, reproduce the observed reality?
-Ensemble average trajectory matches traditional understanding of succession---rapid growth to peak productivity, followed by gradual decline [@odum_1969_strategy_ecosystem_development]---but happens too quickly (Figure \@ref(fig:summary-ts-plot));
-real ecosystems sustain peak productivity for longer [@gough_2016_disturbance].
+Our overarching objective was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community succession.
+Our first question was: 
+Which model structural and parameter configurations reproduce observed NPP and community compositional changes?
+We found that model-observation fidelity was variable, with modelled community succession generally departing from observations and modelled NPP more comparable to observed values and trends.
+
+Modelled community succession, which was variable across model structural and parameter configurations, poorly predicted the observed transition of dominance from early to mixed successional hardwoods during the first century of regrowth.
+<!-- TODO Highlight the large variability! -->
+The community successional dynamics observed at our study site are representative of temperate forests globally, with short-lived early hardwoods initially occupying a majority of the canopy and over time gradually declining as longer-lived mid- and late-hardwood species emerge as dominant [@wolter_2002_recent; @gough_2007_legacy; @sommerfeld_2018_???].
+At our site, _Populus_, a genus of early hardwoods, is abundant in the canopy during the first 50 to 75 years of succession and, upon peak maturity, declines precipitously at a rate of 40% per decade [@gough_2010_wood].
+Filling this canopy void over time are mid-successional _Quercus_ and _Acer_, and later successional _Pinus_ [@scheuermann_2018_???].
+However, with the exception of a select few parameter combinations (e.g.
+C and D in Figure \@ref(fig:lai-pft-plot)), modelled community succession exhibited only modest changes in functional diversity over time.
+<!-- TODO Fix from here on -->
+Moreover, a single PFT emerged as dominant in a majority of runs, though _which_ PFT dominated depended on parameter combination.
+This mismatch between observations and simulations reinforces the difficulty of accurately representing plant functional diversity in vegetation models [@fisher_2015_taking].
+Specific to our analysis, our results indicate that ED-2.2 perceives limited niche overlap among PFTs, resulting in an unrealistically low level of PFT coexistence. <!-- TODO REFS -->
+
+Unlike community succession, most model configurations produced similarly realistic C flux successional trajectories.
+In particular, the use of median parameter values resulted in century-long C flux dynamics that mirror the trajectory of site-level observations [@gough_2007_legacy] and, more broadly, those from other temperate forests [@law_2001_carbon; @law_2003_changes; @irvine_2004_age; @tang_2009_soil; @kashian_2013_postfire].
+In our study, modelled and observed C fluxes generally increased rapidly in the first quarter century before gradually declining, a pattern that is consistent with ecological theory [@odum_1969_strategy_ecosystem_development] and global observations [@pregitzer_2004_carbon; @luyssaert_2008_old; @gough_2016_disturbance].
+At odd with site-level observations, however, is the extent of simulated declines in production over time.
+Rather than decreasing, long-term meteorological flux tower observations indicate net ecosystem production increased two-fold over a period of 15 years in the century-old UMBS forest, possibly because the decline of senscent early hardwoods released mid and late hardwoods poised for rapid growth [@gough_2016_disturbance; @curtis_2018_forest].
 One possible reason for the excessive pace of succession in ED-2.2 is missing mechanisms for sustaining productivity through early succession.
 In particular, recent evidence points to the importance of low-severity, high-frequency disturbances in sustaining ecosystem productivity [@gough_2016_disturbance].
-ED2.2 has a mortality scheme, and indeed, parameters related to mortality were often among the most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
-<!-- Maybe move this elsewhere? -->
-<!-- TODO Look at mortality across ED runs. Does it relate at all to results? -->
+ED-2.2 has a mortality scheme, and indeed, parameters related to mortality were occasionally among the most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
 However, although our simulations used real meteorology (which included occasional extreme weather events), we did not consider other sources of disturbance common to this region [such as harvests, insects and pathogens, and fire; @amiro_2010_ecosystem], which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
-Future work will: (1) Calibrate ED2.2 disturbance response to experimental observations (FoRTE), and (2) run long-term simulations with prescribed disturbance regimes and see how they affect ecological dynamics.
+Future work should evaluate the impacts of prescribed disturbance regimes on multidecadal predictions of ecological dynamics and biogeochemistry.
 

### 2019-11-19 analysis/paper/paper.Rmd
Significant revisions to first discussion section

@@
   "production efficiency (NPP / LAI)",
-  "and effective number of PFTs (N~PFT,eff~)",
+  "and effective number of PFTs ($N_{\\text{PFT,eff}}$)",
   "by model configuration.",
@@
 
-Our overarching objective was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem productivity and community succession.
+Our overarching objective was to investigate how model structure and parameters interact in multidecadal simulations of forest ecosystem production and community succession.
 Our first question was:
@@
 
-Modelled community succession, which was variable across model structural and parameter configurations, poorly predicted the observed transition of dominance from early to mixed successional hardwoods during the first century of regrowth.
-<!-- TODO Highlight the large variability! -->
-The community successional dynamics observed at our study site are representative of temperate forests globally, with short-lived early hardwoods initially occupying a majority of the canopy and over time gradually declining as longer-lived mid- and late-hardwood species emerge as dominant [@wolter_2002_recent; @gough_2007_legacy; @sommerfeld_2018_???].
+The community successional dynamics observed at our study site are representative of temperate forests globally, with short-lived early hardwoods initially occupying a majority of the canopy and over time gradually declining as longer-lived mid- and late-hardwood species emerge as dominant [@wolter_2002_recent; @gough_2007_legacy; @sommerfeld_2018_patterns].
 At our site, _Populus_, a genus of early hardwoods, is abundant in the canopy during the first 50 to 75 years of succession and, upon peak maturity, declines precipitously at a rate of 40% per decade [@gough_2010_wood].
-Filling this canopy void over time are mid-successional _Quercus_ and _Acer_, and later successional _Pinus_ [@scheuermann_2018_???].
-However, with the exception of a select few parameter combinations (e.g.
-C and D in Figure \@ref(fig:lai-pft-plot)), modelled community succession exhibited only modest changes in functional diversity over time.
-<!-- TODO Fix from here on -->
-Moreover, a single PFT emerged as dominant in a majority of runs, though _which_ PFT dominated depended on parameter combination.
-This mismatch between observations and simulations reinforces the difficulty of accurately representing plant functional diversity in vegetation models [@fisher_2015_taking].
-Specific to our analysis, our results indicate that ED-2.2 perceives limited niche overlap among PFTs, resulting in an unrealistically low level of PFT coexistence. <!-- TODO REFS -->
+Filling this canopy void over time are mid-successional _Quercus_ and _Acer_, and later successional _Pinus_ [@scheuermann_2018_effects].
+However, modelled community composition, while highly variable, overwhelmingly favored dominance by a single PFT (Figure \@ref(fig:summary-ts-plot)), with only a few parameter combinations exhibiting more realistic ecological dynamics (Figure \@ref(fig:lai-pft-plot)).
+This tendency towards mono-dominance reinforces the difficulty of accurately representing plant functional diversity in vegetation models [@fisher_2015_taking].
+Specific to our analysis, our results indicate that environmental filtering and competitive exclusion are potent forces in ED-2.2, and that realistic levels of coexistence and competition can only be achieved under precise combinations of the multivariate parameter space.
 
-Unlike community succession, most model configurations produced similarly realistic C flux successional trajectories.
-In particular, the use of median parameter values resulted in century-long C flux dynamics that mirror the trajectory of site-level observations [@gough_2007_legacy] and, more broadly, those from other temperate forests [@law_2001_carbon; @law_2003_changes; @irvine_2004_age; @tang_2009_soil; @kashian_2013_postfire].
-In our study, modelled and observed C fluxes generally increased rapidly in the first quarter century before gradually declining, a pattern that is consistent with ecological theory [@odum_1969_strategy_ecosystem_development] and global observations [@pregitzer_2004_carbon; @luyssaert_2008_old; @gough_2016_disturbance].
-At odd with site-level observations, however, is the extent of simulated declines in production over time.
+Unlike community succession, most model configurations produced more realistic C flux successional trajectories.
+In particular, the use of median parameter values resulted in century-long C flux dynamics (Figure \@ref(fig:default-median-flux)) that mirror the trajectory of site-level observations [@gough_2007_legacy] and, more broadly, those from other temperate forests [@law_2001_carbon; @law_2003_changes; @irvine_2004_age; @tang_2009_soil; @kashian_2013_postfire].
+In our study, modelled and observed C fluxes generally increased rapidly in the first quarter century before gradually declining (Figures \@ref(fig:default-median-flux) and \@ref(fig:summary-ts-plot)), a pattern that is consistent with ecological theory [@odum_1969_strategy_ecosystem_development] and global observations [@pregitzer_2004_carbon; @luyssaert_2008_old; @gough_2016_disturbance].
+At odds with site-level observations, however, is the extent of simulated declines in production over time.
 Rather than decreasing, long-term meteorological flux tower observations indicate net ecosystem production increased two-fold over a period of 15 years in the century-old UMBS forest, possibly because the decline of senscent early hardwoods released mid and late hardwoods poised for rapid growth [@gough_2016_disturbance; @curtis_2018_forest].
+This mismatch may be driven by the aforementioned tendency of ED-2.2 towards monodominance; in the absence of an understory, there is nothing to compensate declining production among dominant trees.
+
+When considering the coupling of composition and C fluxes, we found that the successional arc of production was similar among model configurations but improved when simulations more closely represented observed forest composition.
+Most ED-2.2 runs under-predicted NPP and LAI during the peak period (1920-1950) and towards the end of the simulation (1975-1999) (Figure \@ref(fig:summary-ts-plot)).
+However, of five representative runs, the two that most closely matched present-day composition at UMBS (early hardwood dominated; **C** and **D** in Figure \@ref(fig:lai-pft-plot)) also exhibited the greatest fidelity with observed NPP and LAI (Figure \@ref(fig:summary-ts-plot) and S3).
+These results are encouraging, demonstrating an ecologically realistic dependency of C fluxes on canopy structure and composition [@pan_2011_large; @he_2012_relationships; @reich_2012_key].
+Moreover, we show that such realistic outcomes are possible without tuning important and directly observable parameters like SLA and V~cmax~ to unrealistic values (Figure S2).
+We therefore argue that, although challenging, it is not unreasonable to expect demographic vegetation models to accurately simulate both community composition and ecosystem biogeochemistry, and in fact, accuracy in the latter may depend on the former.
+
+OLD DISTURBANCE TEXT:
 One possible reason for the excessive pace of succession in ED-2.2 is missing mechanisms for sustaining productivity through early succession.
@@
 Future work should evaluate the impacts of prescribed disturbance regimes on multidecadal predictions of ecological dynamics and biogeochemistry.
-
-Under all but a small minority of parameter combinations, ED-2.2 was unable to create---let alone sustain---a forest with any appreciable functional diversity.
-The large majority of runs produced dominance by a single PFT, though _which_ PFT dominated depended on parameter combination.
-This highlights the difficulty of representing plant functional diversity in vegetation models, which has been true in other models as well [@fisher_2015_taking].
-To the extent that ED-2.2 is "correct" in how it represents processes, this result also indicates that the plant functional trait space of realized niches is small---i.e.
-multiple different trees can co-exist only under very specific combinations of plant traits and environmental conditions.
-<!-- TODO REF -->
-
-Most ED-2.2 runs under-predicted both NPP and LAI, both in the peak period (1920-1950) and towards the end of the simulation (1975-1999) (Figures \@ref(fig:summary-ts-plot)).
-However, of our five selected representative runs, the two that most closely matched present day composition at UMBS (early hardwood-dominated; **C** and **D** in Figure \@ref(fig:lai-pft-plot)) also closely matched the observed NPP and LAI (Figure \@ref(fig:summary-ts-plot)).
-This is encouraging because it suggests that it is possible to get both aggregate ecosystem C fluxes and PFT composition right.
-Moreover, given that many important ED-2.2 parameters were well-constrained in our analysis (Figure S2), this is possible without using physically unrealistic parameter values.
-We therefore suggest that for model calibration and validation, comparisons against multiple output variables---including variables related to composition---can yield maximum constraint and are not unreasonably strict.
 

### 2019-11-19 analysis/paper/paper.Rmd
Drop old disturbance text

@@
 
-OLD DISTURBANCE TEXT:
-One possible reason for the excessive pace of succession in ED-2.2 is missing mechanisms for sustaining productivity through early succession.
-In particular, recent evidence points to the importance of low-severity, high-frequency disturbances in sustaining ecosystem productivity [@gough_2016_disturbance].
-ED-2.2 has a mortality scheme, and indeed, parameters related to mortality were occasionally among the most important parameters for predictive uncertainty (Figure \@ref(fig:partial-variance-npp)).
-However, although our simulations used real meteorology (which included occasional extreme weather events), we did not consider other sources of disturbance common to this region [such as harvests, insects and pathogens, and fire; @amiro_2010_ecosystem], which may contribute to sustaining productivity and dominance of early-successional forests [@gough_2016_disturbance].
-Future work should evaluate the impacts of prescribed disturbance regimes on multidecadal predictions of ecological dynamics and biogeochemistry.
-
 ## Drivers of uncertainty

### 2019-11-20 analysis/paper/paper.Rmd
Spell acknowledgments correctly
Minor revisions to second half of discussion

@@
 We found that parameter uncertainty was by far the dominant driver of uncertainty in our simulations.
-Different parameter combinations led to huge ranges in ensemble predictions of ensemble variables, with NPP and LAI ranging from zero to more than twice the ensemble mean estimates (Figures \@ref(fig:summary-ts-plot)).
-Moreover, different parameter combinations produced qualitatively different ecological narratives (Figures \@ref(fig:lai-pft-plot)).
-Therefore, parameter uncertainty has implications beyond C cycling---differences in species composition and relative abundance have implications for forest management and wildlife conservation, among others.
-<!-- TODO REFS -->
+Different parameter combinations led to not only huge ranges in ensemble predictions of stand-level NPP and LAI, but also a multitude of qualitatively different ecological narratives (Figures \@ref(fig:summary-ts-plot), \@ref(fig:lai-pft-plot)).
+Parameter uncertainty in demographic models therefore has broad implications for C cycle science, community ecology, and forest management.
 
@@
 For example, ED-2.2 is highly sensitive to SLA [@dietze_2014_quantitative; @pandit_2018_optimizing], but thanks to the large number of SLA observations, we were able to estimate it very precisely (Figure S2) and, as a result, it was not a significant source of uncertainty in our analysis (Figure \@ref(fig:partial-variance-npp)).
-At the same time, some poorly constrained parameters have relatively modest impacts on model outputs---for instance, nonlocal dispersal and background mortality rate in our study (Figure S2, \@ref(fig:partial-variance-npp))---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
+At the same time, some poorly constrained parameters have relatively modest impacts on model outputs---for instance, rates of nonlocal dispersal and background mortality in our study (Figure S2, \@ref(fig:partial-variance-npp))---and therefore, reducing their uncertainties is not an effective way to improve model predictions.
 Ultimately, this points to the value of model-data feedbacks [@dietze_2013_improving] in advancing ecological research and setting future research agendas.
@@
 and changes to model structure in turn changed the relative importance of specific parameters to overall uncertainty (Figure \@ref(fig:partial-variance-npp)).
-Collectively, this means that sensitivity and uncertainty analyses are not necessarily generalizable across models, and highlights the importance of model intercomparison studies.
+Collectively, this means that sensitivity and uncertainty analyses may not necessarily be generalizable across models, and highlights the importance of model intercomparison studies.
 It also reinforces our argument that changing model structure---particularly if such changes increase model complexity---should be pursued with extreme caution.
 
-# Acknowledgements
+# Acknowledgments
 

### 2019-11-20 analysis/paper/paper.Rmd
Major revisions to abstract

@@
 abstract: |
-  Forests in the US Upper Midwest have been a net C sink for a century, driven by their fast-growing, early successional species.
-  However, the composition of these regrown forests has started to shift toward mid-successional species.
-  Combined with changing climate and disturbance, this means that future biogeochemical and ecological dynamics in these forests are highly uncertain.
-  Demographic vegetation models provide a tool for analyzing this uncertainty and its implications.
-  Here, we evaluate the relative importance of model "structural" (which processes are represented, and how) and "parameter" (parameters used in those equations) uncertainty, asking:
-  (1) Which processes are most important to accurately model community succession and C cycling in Upper Midwest forests?
-  (2) What are the relative contributions of parameter vs. structural uncertainty?
-  We ran ensembles of the Ecosystem Demography model v2.2 (sampling over parameter uncertainty) at the University of Michigan Biological Station (UMBS) from 1900 to 2000 with different representations of key processes important to competition for light.
-  We then compared the magnitude of structural (across representations) and parameter (within each representation) uncertainty, and assessed which submodel-parameter combinations reproduced observed C fluxes and succession.
-  On average, our simulations underestimated both net primary productivity (NPP) and leaf area index (LAI) after 100 simulation years, though NPP estimates at peak productivity (20-50 years into the simulations) were closer to observations.
-  However, parameter uncertainty was large; at any given timestep, both NPP and LAI ranged from ~0 to >200% of their mean value.
-  Although many simulations matched either the observed NPP or observed LAI at some point during the simulation, only a small minority of individual runs matched both.
-  The few runs that matched observed NPP and LAI also matched the present-day composition (early hardwood-dominated).
-  Structural uncertainty was less important than parameter uncertainty, but model configuration did influence the ecological dynamics within a given parameter set, and affected which parameters were most influential.
-  We conclude that: (1) targeted data collection to constrain vegetation model parameters should be a higher priority than model development; (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and (3) robust model validation requires scrutiny of multiple different variables, including those related to vegetation productivity and composition.
+  Regrowth of secondary forests following changes in human land-use patterns is a widespread phenomenon globally.
+  The land-use and disturbance history of the US Upper Midwest makes the temperate forests of this region a model system for studying the ecology and biogeochemistry of secondary forests.
+  Demographic vegetation models encapsulate our best understanding of forest function, and whether or not these models can accurately reproduce observed multidecadal patterns of forest succession is an important indication of our ability to predict the responses of terrestrial ecosystems to future global change.
+  In this study, we test the accuracy and precision with which a demographic vegetation model---ED-2.2---can simulate C cycling and community composition in 100 years of temperate forest regrowth following stand-replacing disturbance, asking:
+  (1) Which processes and parameters are most important to accurately model community composition and C cycling in Upper Midwest temperate forests?
+  (2) What is the relative importance of model structure (which processes are represented and how) versus model parameter values to these predictions?
+  Previous work has shown that demographic model predictions of ecological dynamics are highly sensitive to how these models represent competition for light, echoing a large body of ecological research on the importance of this process to forest succession.
+  Therefore, we ran ensembles of the Ecosystem Demography model v2.2 (sampling over parameter uncertainty) at the University of Michigan Biological Station (UMBS) from 1900 to 2000 with different representations of key processes important to competition for light.
+  We then compared the magnitude of structural (across representations) and parameter (within each representation) uncertainty, and assessed which submodel-parameter combinations best reproduced observed C fluxes and succession.
+  On average, our simulations underestimated both net primary production (NPP) and leaf area index (LAI) after 100 simulation years, though NPP estimates at peak production (20-50 years into the simulations) were closer to observations.
+  In addition, most simulations failed to capture observed competitive interactions, predicting persistent dominance by a single plant functional type (PFT).
+  However, parameter uncertainty was large; at any given timestep, both NPP and LAI ranged from ~0 to >200% of their mean value, and depending on parameter value, stands could end up dominated by any one of the four PFTs we considered.
+  Encouragingly, representative runs that best matched present-day composition (early hardwood-dominated) also exhibited the greatest fidelity with present-day observed NPP and LAI, demonstrating an ecologically realistic dependency of C fluxes on canopy structure and composition.
+  Structural uncertainty was less important than parameter uncertainty, but model structure did influence the ecological dynamics within a given parameter set, and affected which parameters were most important to predictive uncertainty.
+  We conclude that:
+  (1) targeted data collection to constrain vegetation model parameters should be a higher priority than model development;
+  (2) parameter sensitivity analyses are not necessarily generalizable across model structures; and
+  (3) it is challenging but not unreasonable to expect demographic vegetation models to accurately simulate both community composition and ecosystem biogeochemistry, and in fact, accuracy in the latter may depend on the former.
 keywords: |

### 2019-11-20 analysis/paper/paper.Rmd
Move figures to end. Revise DOCX template.

@@
 
+out_format <- knitr::opts_knit$get("rmarkdown.pandoc.to")
+
+pgbreak <- function() {
+  if (grepl("docx", out_format)) {
+    return("##### pagebreak")
+  } else if (is_latex_output()) {
+    return("\\newline")
+  }
+}
+
 opts_chunk$set(
@@
   out <- sprintf("**Figure %d**: %s", n, text)
-  if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
+  if (grepl("gfm", out_format, ignore.case = TRUE)) {
     out
@@
 
-Table: (\#tab:submodels) ED-2.2 submodel descriptions
-
-| Name | Description | Color |
-|------|-------------|-------|
-| **Crown model** |||
-| **C**losed (default) | Cohort crowns take up entire patch area. Competition for light based only on height. | Light |
-| **F**inite | Cohort crown area is proportional to DBH according to PFT-specific allometry. | Dark |
-| **Radiative transfer model (RTM)** |||
-| **T**wo-stream (default) | Two-stream approximation [@sellers_1985_canopy; @clm45_note] | Primary (red, blue) |
-| **M**ulti-scatter | Multiple-scatter approximation, following [@zhao_2005_multiple] | Secondary (green, orange) |
-| **Trait plasticity** |||
-| **S**tatic (default) | SLA and V~c,max~ are constant | Cool (blue, green) |
-| **P**lastic | SLA increases, and V~c,max~ decreases, with light level | Warm (red, orange) |
-
 ## Parameter descriptions
-
-```{r param-table}
-tbl <- readr::read_csv(here(
-  "analysis", "data", "derived-data", "parameter-table.csv"
-))
-tbl <- subset(tbl, Ran == "yes", select = c(-unit_parsed, -Ran))
-tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
-cap <- paste(
-  "Parameter descriptions.",
-  "For additional information, see the",
-  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
-)
-if (is_latex_output() || is_html_output()) {
-  kable(tbl, escape = is_latex_output(), caption = cap, booktabs = TRUE) %>%
-    kable_styling(latex_options = "scale_down")
-} else {
-  # Probably DOCX output
-  cap <- paste("Table: (\\#tab:param-table)", cap)
-  flextable::flextable(tbl) %>%
-    flextable::autofit() %>%
-    flextable::fit_to_width(7.5) %>%
-    flextable::set_caption(cap)
-}
-```
 
@@
 The species comprising these PFTs are shown in Table \@ref(tab:pft-species).
-
-Table: (\#tab:pft-species) Plant functional type-species mappings
-
-| Plant functional type | Species                 | Color  |
-|-----------------------|-------------------------|--------|
-| Early hardwood        | _Betula papyrifera_     | Violet |
-|                       | _Populus grandidentata_ |        |
-|                       | _Populus tremuloides_   |        |
-| Mid hardwood          | _Quercus rubra_         | Blue   |
-|                       | _Acer rubrum_           |        |
-|                       | _Acer pensylvaticum_    |        |
-| Late hardwood         | _Acer saccharum_        | Green  |
-|                       | _Fagus grandifolia_     |        |
-| Pine                  | _Pinus strobus_         | Yellow |
 
@@
 
-```{r structure-compare-default, fig.cap = structure_default_cap}
-### Figure 1
-structure_default_cap <- paste(
-  htmlfig(),
-  "Annual NPP and LAI from ED2 runs",
-  "with default parameters",
-  "across different configurations."
-)
-fig("structure-compare-default.png")
-```
-
 Different model structures agreed for the first half of the century-long simulation, at which point they started to diverge (Figure \@ref(fig:structure-compare-default)).
@@
 For example, default values for V~cmax~ were too low for all PFTs, and default values for SLA were too low for all PFTs except early hardwoods.
-
-```{r default-median-flux, fig.cap = default_median_flux_cap}
-### Figure 4
-default_median_flux_cap <- paste(
-  htmlfig(),
-  "ED-2.2 predictions of mean annual",
-  "gross (GPP) and net primary productivity (NPP)",
-  "and autotrophic respiration (RA) from simulations using",
-  "default parameter values (red) and",
-  "posterior median parameters from trait meta-analysis (blue)",
-  "across different strucures.",
-  "Median and default parameter values are shown in Figure S2."
-)
-fig("default-median-fluxes.png")
-```
-
-```{r default-median-lai, fig.cap = default_median_lai_cap}
-### Figure 5
-default_median_lai_cap <- paste(
-  htmlfig(),
-  "ED-2.2 plot-level predictions of",
-  "mean growing season leaf area index (LAI)",
-  "by plant functional type (PFT), using",
-  "default parameter values (top) and",
-  "posterior median parameters from trait meta-analysis (bottom)",
-  "across different strucures."
-)
-fig("default-median-lai.png")
-```
 
@@
 
-```{r summary-ts-plot, fig.cap = summary_ts_cap}
-### Figure 6
-summary_ts_cap <- paste(
-  htmlfig(),
-  "ED-2.2 ensemble plot-level predictions of",
-  "net primary productivity (NPP),",
-  "total leaf area index (LAI),",
-  "production efficiency (NPP / LAI)",
-  "and effective number of PFTs ($N_{\\text{PFT,eff}}$)",
-  "by model configuration.",
-  "Shading indicates the middle 50\\% (dark), middle 95\\% (medium),",
-  "and entire range (light) of ensemble simulations",
-  "(NPP/LAI results are clipped to (-5, 40) for readability,",
-  "as a few isolated ensemble members produced extreme values at specific",
-  "time points).",
-  "Effective number of PFTs was calculated as the inverse",
-  "of the Simpson diversity index with proportions expressed as",
-  "fraction of aboveground biomass.",
-  "Each colored line represents the same set of parameters across all configurations,",
-  "and corresponds to the letters in Figure 5.",
-  "Black dot with error bars is the observed mean and min/max value."
-  # TODO REF
-)
-fig("summary-ts-plot.png")
-```
-
 Net primary productivity (NPP), total leaf area index (LAI), their ratio, and PFT diversity varied with both parameter value and model structure (Figure \@ref(fig:summary-ts-plot)).
@@
 
-```{r lai-pft-plot, fig.cap = lai_pft_cap}
-### Figure 9
-lai_pft_cap <- paste(
-  htmlfig(),
-  "ED-2.2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization.",
-  "Each row of plots is a simulation with the same set of parameters across model configurations.",
-  "Letters correspond to the labels in Figure 7.",
-  "Note that these are predictions of total leaf area index summed across all cohorts of the same PFT,",
-  "meaning that higher LAI in one PFT does not necessarily indicate dominance, size, etc."
-)
-fig("lai-pft.png")
-```
-
 Model structure and parameters interacted to determine the community assembly and dynamics of each simulation, with consequences for overall ecosystem productivity that varied by model structure (Figure \@ref(fig:lai-pft-plot)).
@@
 Parameter sets **D** and **E** had the highest productivity of these five sets, and were characterized by strong dominance of early hardwood and Pine, respectively.
-
-```{r partial-variance-npp, fig.cap = partial_variance_npp_cap}
-### Figure 10
-partial_variance_npp_cap <- paste(
-  htmlfig(),
-  "Contributions of individual parameters to predicted uncertainty in",
-  "ED-2.2 predictions of average net primary productivity (NPP)",
-  "over the period 1920-1950",
-  "by model type.",
-  "Each panel shows the top 8 parameter values by their fractional contribution",
-  'to parameter uncertainty ("partial variance").',
-  "Input parameter distributions are shown in Figure 3.",
-  "Note that x-axis scales vary across panels."
-)
-fig("partial-variance-npp.png")
-```
 
@@
 
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-<!-- ##### pagebreak -->
-
-\newpage
+`r pgbreak()`
 
@@
 
-<!-- ##### pagebreak -->
+`r pgbreak()`
 
-\newpage
+# Tables
+
+Table: (\#tab:submodels) ED-2.2 submodel descriptions
+
+| Name | Description | Color |
+|------|-------------|-------|
+| **Crown model** |||
+| **C**losed (default) | Cohort crowns take up entire patch area. Competition for light based only on height. | Light |
+| **F**inite | Cohort crown area is proportional to DBH according to PFT-specific allometry. | Dark |
+| **Radiative transfer model (RTM)** |||
+| **T**wo-stream (default) | Two-stream approximation [@sellers_1985_canopy; @clm45_note] | Primary (red, blue) |
+| **M**ulti-scatter | Multiple-scatter approximation, following [@zhao_2005_multiple] | Secondary (green, orange) |
+| **Trait plasticity** |||
+| **S**tatic (default) | SLA and V~c,max~ are constant | Cool (blue, green) |
+| **P**lastic | SLA increases, and V~c,max~ decreases, with light level | Warm (red, orange) |
+
+`r pgbreak()`
+
+```{r param-table}
+tbl <- readr::read_csv(here(
+  "analysis", "data", "derived-data", "parameter-table.csv"
+))
+tbl <- subset(tbl, Ran == "yes", select = c(-unit_parsed, -Ran))
+tbl[["ED Name"]] <- sprintf("`%s`", tbl[["ED Name"]])
+cap <- paste(
+  "Parameter descriptions.",
+  "For additional information, see the",
+  "[ED2 model wiki](https://github.com/EDmodel/ED2/wiki/PFT-parameters)."
+)
+if (is_latex_output() || is_html_output()) {
+  kable(tbl, escape = is_latex_output(), caption = cap, booktabs = TRUE) %>%
+    kable_styling(latex_options = "scale_down")
+} else {
+  # Probably DOCX output
+  cap <- paste("Table: (\\#tab:param-table)", cap)
+  flextable::flextable(tbl) %>%
+    flextable::autofit() %>%
+    flextable::fit_to_width(7.5) %>%
+    flextable::set_caption(cap)
+}
+```
+
+`r pgbreak()`
+
+Table: (\#tab:pft-species) Plant functional type-species mappings
+
+| Plant functional type | Species                 | Color  |
+|-----------------------|-------------------------|--------|
+| Early hardwood        | _Betula papyrifera_     | Violet |
+|                       | _Populus grandidentata_ |        |
+|                       | _Populus tremuloides_   |        |
+| Mid hardwood          | _Quercus rubra_         | Blue   |
+|                       | _Acer rubrum_           |        |
+|                       | _Acer pensylvaticum_    |        |
+| Late hardwood         | _Acer saccharum_        | Green  |
+|                       | _Fagus grandifolia_     |        |
+| Pine                  | _Pinus strobus_         | Yellow |
+
+`r pgbreak()`
+
+# Figures
+
+```{r structure-compare-default, fig.cap = structure_default_cap}
+### Figure 1
+structure_default_cap <- paste(
+  htmlfig(),
+  "Annual NPP and LAI from ED2 runs",
+  "with default parameters",
+  "across different configurations."
+)
+fig("structure-compare-default.png")
+```
+
+`r pgbreak()`
+
+```{r default-median-flux, fig.cap = default_median_flux_cap}
+### Figure 4
+default_median_flux_cap <- paste(
+  htmlfig(),
+  "ED-2.2 predictions of mean annual",
+  "gross (GPP) and net primary productivity (NPP)",
+  "and autotrophic respiration (RA) from simulations using",
+  "default parameter values (red) and",
+  "posterior median parameters from trait meta-analysis (blue)",
+  "across different strucures.",
+  "Median and default parameter values are shown in Figure S2."
+)
+fig("default-median-fluxes.png")
+```
+
+`r pgbreak()`
+
+```{r default-median-lai, fig.cap = default_median_lai_cap}
+### Figure 5
+default_median_lai_cap <- paste(
+  htmlfig(),
+  "ED-2.2 plot-level predictions of",
+  "mean growing season leaf area index (LAI)",
+  "by plant functional type (PFT), using",
+  "default parameter values (top) and",
+  "posterior median parameters from trait meta-analysis (bottom)",
+  "across different strucures."
+)
+fig("default-median-lai.png")
+```
+
+`r pgbreak()`
+
+```{r summary-ts-plot, fig.cap = summary_ts_cap}
+### Figure 6
+summary_ts_cap <- paste(
+  htmlfig(),
+  "ED-2.2 ensemble plot-level predictions of",
+  "net primary productivity (NPP),",
+  "total leaf area index (LAI),",
+  "production efficiency (NPP / LAI)",
+  "and effective number of PFTs ($N_{\\text{PFT,eff}}$)",
+  "by model configuration.",
+  "Shading indicates the middle 50\\% (dark), middle 95\\% (medium),",
+  "and entire range (light) of ensemble simulations",
+  "(NPP/LAI results are clipped to (-5, 40) for readability,",
+  "as a few isolated ensemble members produced extreme values at specific",
+  "time points).",
+  "Effective number of PFTs was calculated as the inverse",
+  "of the Simpson diversity index with proportions expressed as",
+  "fraction of aboveground biomass.",
+  "Each colored line represents the same set of parameters across all configurations,",
+  "and corresponds to the letters in Figure 5.",
+  "Black dot with error bars is the observed mean and min/max value."
+  # TODO REF
+)
+fig("summary-ts-plot.png")
+```
+
+`r pgbreak()`
+
+```{r lai-pft-plot, fig.cap = lai_pft_cap}
+### Figure 9
+lai_pft_cap <- paste(
+  htmlfig(),
+  "ED-2.2 PFT-level predictions of leaf area index (LAI) by model configuration and parameterization.",
+  "Each row of plots is a simulation with the same set of parameters across model configurations.",
+  "Letters correspond to the labels in Figure 7.",
+  "Note that these are predictions of total leaf area index summed across all cohorts of the same PFT,",
+  "meaning that higher LAI in one PFT does not necessarily indicate dominance, size, etc."
+)
+fig("lai-pft.png")
+```
+
+`r pgbreak()`
+
+```{r partial-variance-npp, fig.cap = partial_variance_npp_cap}
+### Figure 10
+partial_variance_npp_cap <- paste(
+  htmlfig(),
+  "Contributions of individual parameters to predicted uncertainty in",
+  "ED-2.2 predictions of average net primary productivity (NPP)",
+  "over the period 1920-1950",
+  "by model type.",
+  "Each panel shows the top 8 parameter values by their fractional contribution",
+  'to parameter uncertainty ("partial variance").',
+  "Input parameter distributions are shown in Figure 3.",
+  "Note that x-axis scales vary across panels."
+)
+fig("partial-variance-npp.png")
+```
+
+`r pgbreak()`
 

### 2019-11-20 analysis/drake/sensitivity-analysis.R
Minor formatting and figure fixes

@@ -141,9 +141,8 @@ plan <- bind_plans(plan, drake_plan(
     ungroup() %>%
     separate(xvar, c("shortname", "xvar"), extra = "merge") %>%
     mutate(
-      shortname = factor(shortname, pfts("shortname")),
-      model = factor(model_id, c("CTS", "CTP", "CMS", "CMP",
-                                 "FTS", "FTP", "FMS", "FMP")),
+      shortname = factor(shortname, pfts("shortname"), pfts("pft")),
+      model = factor(model_id, levels(models$model_id), levels(models$model)),
       xvar = factor(xvar, ed2_param_table[["ED Name"]],
                     ed2_param_table[["Display name"]])
     ),

### 2019-11-20 analysis/paper/paper.Rmd
Minor formatting and figure fixes
More formatting tweaks for DOCX export

@@
 author:
-  - Alexey N. Shiklomanov
-  - Ben Bond-Lamberty
-  - Jeff Atkins
-  - Christopher M. Gough
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+  - Alexey N. Shiklomanov$^{a *}$, Ben Bond-Lamberty$^a$, Jeffrey Atkins$^b$, Christopher M. Gough$^b$
+  - $^a$ *Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, USA*; $^b$ *Department of Biology, Virginia Commonwealth University, Richmond, VA, USA*
+  - "* Corresponding author: alexey.shiklomanov@pnnl.gov, (301) 314-6713; 5825 University Research Ct, Suite 3500, College Park, MD 20740"
 output:
@@
 abstract: |
+  **Abstract**:
   Regrowth of secondary forests following changes in human land-use patterns is a widespread phenomenon globally.
@@
   } else if (is_latex_output()) {
-    return("\\newline")
+    return("\\newpage")
   }
@@
 
-The official ED-2.2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained at https://github.com/ashiklom/ED2/tree/b048950971e91699b78fc566df133caf977fda32.
+The official ED-2.2 source code is available at https://github.com/EDmodel/ED2, and the exact version used in this study (which includes minor revisions to accommodate the requirements of this study) can be obtained at https://github.com/ashiklom/ED2/tree/b048950.
 
@@
 `r pgbreak()`
-
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r}
-# what commit is this file at? You may need to change the path value
-# if your Rmd is not in analysis/paper/
-git2r::repository(here::here())
-```
-
-View this project on GitHub in repository [ashiklom/fortebaseline](https://github.com/ashiklom/fortebaseline)
-(this manuscript file is in [`analysis/paper/paper.Rmd`](https://github.com/ashiklom/fortebaseline/blob/master/analysis/paper/paper.Rmd))
-and in Open Science Framework (OSF) at https://osf.io/dznuf/ .

### 2020-03-18 analysis/drake/sensitivity-analysis.R
Revise param dist fig and param table

@@ -188,11 +188,3 @@ plan <- bind_plans(plan, drake_plan(
     "analysis", "figures", "partial-variance-npp-end.png"
   )), sensitivity_period2_gg, width = 7.6, height = 8.4)
 ))
-
-plan <- bind_plans(plan, drake_plan(
-  ed2_param_table = readr::read_csv(file_in(
-    "analysis/data/derived-data/parameter-table.csv"
-  )) %>%
-    filter(Ran == "yes")
-))
-

### 2020-03-18 analysis/drake/trait-distribution.R

@@ -16,37 +16,71 @@ plan <- bind_plans(plan, drake_plan(
 
 ### Trait distribution figures
 plan <- bind_plans(plan, drake_plan(
+  ed2_param_table = read_csv(
+    file_in("analysis/data/derived-data/parameter-table.csv"),
+    col_types = "ffcccccc",
+    comment = "#"
+  ),
   trait_distribution = readRDS(file_in(!!trait_distribution_file)) %>%
-    mutate(pft = factor(pft, pfts("pft"))),
+    mutate(pft = factor(pft, pfts("pft")),
+           trait = factor(trait, ed2_param_table[["ED Name"]])) %>%
+    left_join(ed2_param_table, c("trait" = "ED Name")),
   ed2_default_params = ed_default_params() %>%
-    semi_join(trait_distribution, c("pft", "trait")),
+    mutate(trait = factor(trait, ed2_param_table[["ED Name"]])) %>%
+    semi_join(trait_distribution, c("pft", "trait")) %>%
+    left_join(ed2_param_table, c("trait" = "ED Name")),
+  ed2_default_plot = ed2_default_params %>%
+    mutate(
+      default_value = if_else(trait == "water_conductance",
+                              log10(default_value),
+                              default_value),
+      `Display name` = fct_recode(`Display name`, "log10(Water cond.)" = "Water cond.")
+    ),
   param_dist_gg = trait_distribution %>%
     unnest(draws) %>%
+    mutate(
+      draws = if_else(trait == "water_conductance",
+                      log10(draws),
+                      draws),
+      Median = if_else(trait == "water_conductance",
+                       log10(Median),
+                       Median),
+      `Display name` = fct_recode(`Display name`,
+                                  "log10(Water cond.)" = "Water cond.")
+    ) %>%
     ggplot() +
     aes(x = pft, y = draws, fill = pft) +
     geom_violin(alpha = 0.5) +
-    geom_point(aes(y = default_value, color = "default", shape = "default"),
-               data = ed2_default_params, size = 2) +
-    geom_point(aes(y = Median, color = "median", shape = "median"),
-               data = trait_distribution, size = 2) +
+    geom_point(aes(y = default_value,
+                   color = "ED-2.2 default", shape = "ED-2.2 default"),
+               data = ed2_default_plot, size = 2) +
+    geom_point(aes(y = Median,
+                   color = "Posterior median", shape = "Posterior median"),
+               size = 2) +
     # Parameter values
[…]

### 2020-03-27 analysis/drake/specific-parameters.R
Add pie graphs, first draft

@@ -83,3 +83,67 @@ plan <- bind_plans(plan, drake_plan(
     width = 9.7, height = 6.1, dpi = 300
   )
 ))
+
+scatter_pie <- function(indat, filterdat, filterdat_values,
+                        obslist, ...) {
+  pltdat <- indat %>%
+    semi_join(filterdat, "param_id") %>%
+    left_join(distinct(filterdat_values, param_id, label),
+              "param_id") %>%
+    mutate(x = as.numeric(model) + (runif(n(), -0.2, 0.2)),
+           pid = as.numeric(factor(param_id)))
+
+  colorvec <- pfts("color")
+  names(colorvec) <- pfts("pft")
+
+  ggplot(pltdat) +
+    scatterpie::geom_scatterpie(
+      aes(x = x, y = mmean_npp_py, group = pid),
+      cols = as.character(pfts("pft")),
+      data = pltdat,
+      ...
+    ) +
+    geom_label_repel(
+      aes(x = x, y = mmean_npp_py, group = pid, label = label)
+    ) +
+    geom_hline(yintercept = c(obslist$lo, obslist$hi), linetype = "dashed") +
+    scale_x_continuous(
+      breaks = seq_along(levels(pltdat$model)),
+      labels = gsub(" ", "\n", levels(pltdat$model))
+    ) +
+    scale_fill_manual(
+      breaks = pfts("pft"),
+      values = colorvec
+    ) +
+    labs(y = expression(NPP ~ (MgC ~ ha^-1)),
+         fill = "PFT") +
+    theme_bw() +
+    theme(axis.title.x = element_blank(),
+          legend.position = "bottom")
+}
+
+
+plan <- bind_plans(plan, drake_plan(
+  fitobs_pie = scatter_pie(
+    last_ten,
+    fit_observed,
+    fit_observed_param_values,
+    obs_npp
+  ),
+  fitobs_pie_png = ggsave(
+    file_out("analysis/figures/pie-fitobs.png"),
+    fitobs_pie,
+    width = 6.2, height = 5.0, dpi = 300
+  ),
+  diversity_pie = scatter_pie(
+    last_ten,
+    high_diversity,
+    high_diversity_param_values,
+    obs_npp
+  ),
+  diversity_pie_png = ggsave(
+    file_out("analysis/figures/pie-diversity.png"),
+    diversity_pie,
+    width = 6.2, height = 5.0, dpi = 300
+  )
+))

### 2020-03-27 analysis/drake/superlatives.R

@@ -39,9 +39,7 @@ plan <- bind_plans(plan, drake_plan(
     mutate(category = "Fit observations") %>%
     select(param_id, model, category),
   high_diversity = last_ten %>%
-    ## arrange(desc(npft_eff)) %>%
-    ## top_n(10) %>%
-    filter(npft_eff > 2) %>%
+    filter(npft_eff > 2, mmean_lai_py > 1) %>%
     mutate(category = "Most diverse") %>%
     select(param_id, model, category),
   superlatives = bind_rows(fit_observed, high_diversity)

### 2020-03-27 analysis/scratch/misc.R

@@ -2076,6 +2076,39 @@ last_ten %>%
   geom_text_repel(aes(label = factor(param_id))) +
   facet_wrap(vars(name), scales = "free_y")
 
+scatter_pie <- function(indat, filterdat, ...) {
+  pltdat <- indat %>%
+    semi_join(filterdat, c("param_id", "model")) %>%
+    mutate(x = as.numeric(model) + (runif(n(), -0.3, 0.3)),
+           pid = as.numeric(factor(param_id)))
+
+  colorvec <- pfts("color")
+  names(colorvec) <- pfts("pft")
+
+  ggplot(pltdat) +
+    scatterpie::geom_scatterpie(
+      aes(x = x, y = mmean_npp_py, group = pid),
+      cols = as.character(pfts("pft")),
+      data = pltdat,
+      ...
+    ) +
+    scale_x_continuous(
+      breaks = seq_along(levels(pltdat$model)),
+      labels = gsub(" ", "\n", levels(pltdat$model))
+    ) +
+    scale_fill_manual(
+      breaks = pfts("pft"),
+      values = colorvec
+    ) +
+    labs(y = expression(NPP ~ (MgC ~ ha^-1)),
+         fill = "PFT") +
+    theme_bw() +
+    theme(axis.title.x = element_blank(),
+          legend.position = "bottom")
+}
+
+scatter_pie(last_ten, high_diversity, pie_scale = 1.5)
+
 pltdat <- last_ten %>%
   semi_join(high_diversity, c("param_id", "model")) %>%
   mutate(x = as.numeric(model) + (runif(n(), -0.3, 0.3)),
@@ -2089,9 +2122,13 @@ ggplot(pltdat) +
   ) +
   scale_x_continuous(
     breaks = seq_along(levels(pltdat$model)),
-    labels = levels(pltdat$model)
+    labels = gsub(" ", "\n", levels(pltdat$model))
   ) +
-  scale_fill_manual(values = pfts("color"))
+  scale_fill_manual(values = pfts("color")) +
+  labs(y = expression(NPP ~ (MgC ~ ha^-1))) +
+  theme_bw() +
+  theme(axis.title.x = element_blank(),
+        legend.position = "bottom")
 
 last_ten %>%
   filter(is.na(model))
@@ -2111,3 +2148,8 @@ build_times(type = "build") %>%
 
 drake_cache_log() %>%
   filter(grepl("^942", hash))
+
+indat <- last_ten
[…]

### 2020-03-28 analysis/drake/other.R
Add UMBS map to supplementary figures

@@ -1,0 +1,16 @@
+plan <- bind_plans(plan, drake_plan(
+  umbs_map_gg = {
+    umbs <- sf::st_sfc(sf::st_point(c(-84.6975, 45.5625)), crs = 4326)
+    states <- rnaturalearth::ne_states(returnclass = "sf")
+    ggplot(states) +
+      geom_sf() +
+      geom_sf(data = umbs, size = 3) +
+      coord_sf(xlim = c(-90, -80), ylim = c(40, 48)) +
+      theme_bw()
+  },
+  umbs_map_png = ggsave(
+    file_out("analysis/figures/umbs-map.png"),
+    umbs_map_gg,
+    width = 6, height = 6.5, dpi = 300
+  )
+))

### 2020-03-28 analysis/scratch/misc.R

@@ -2183,15 +2183,3 @@ ggplot(d) +
     cols = c("A", "B"),
     data = d
   )
-
-##################################################
-umbs <- sf::st_sfc(sf::st_point(c(-84.6975, 45.5625)), crs = 4326)
-states <- rnaturalearth::ne_states(returnclass = "sf")
-
-library(ggplot2)
-
-ggplot(states) +
-  geom_sf() +
-  geom_sf(data = umbs, size = 3) +
-  coord_sf(xlim = c(-90, -80), ylim = c(40, 48)) +
-  theme_bw()
